### 2021-05-24 Manuscript/Jaconsville_manuscript.tex
initiate a manuscript file to be synced with overleaf project

\documentclass[11pt,letterpaper]{article}
\usepackage{xcolor}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage[left]{lineno}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage{pdfpages}
\usepackage{longtable}
\usepackage{url}

\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\doublespacing

\raggedright
\textwidth = 6.5 in
\textheight = 8.25 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = 0.0 in
\headheight = 0.0 in
\headsep = 0.5 in
\parskip = 0.1 in
\parindent = 0.2 in

\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{}
\rhead{\thepage}

\begin{document}

\begin{flushleft}

{\Large \textbf{Continuous motion from the Keweenawan track to the Grenville Loop: new paleomagnetic pole of the ca. 1 Ga Jacobsville sandstone}}
\\
\singlespacing
 Yiming Zhang\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, Blake Hodgin\textsuperscript{1}\\
\bigskip
\textsuperscript{1} Department of Earth and Planetary Science, University of California, Berkeley, CA, USA\\

\smallskip
\end{flushleft}

%\linenumbers
%\pagestyle{empty}

\section*{ABSTRACT}

\section*{INTRODUCTION}

The extensive paleomagnetic studies of the well-preserved late Mesoproterozoic to Neoproterozoic rocks of North America have provided the central record for global paleogeography reconstruction during the assembly and rearrangement of the postulated supercontinent Rodinia. Paleomagnetic study of rocks from Laurentia have led to a series of paleomagnetic poles that form an apparent polar wander path (APWP) known as the ``Logan Loop" for the older high­ latitude poles at its apex that continues into the ``Keweenawan Track" of younger lower ­latitude poles that form a progression as the APWP heads toward the ``Grenville Loop" \citep{Swanson-Hysell2019a}. In particular, recent advances in pairing high-precision zircon U-Pb geochronology with high-quality paleomagnetic poles significantly improved the resolution of the Keweenawan Track and revealed Laurentia's rapid motion ($>$20 cm/yr) during the Midcontinent Rift magmatic activity from ca. 1110 Ma to ca. 1085 Ma \citep{Swanson-Hysell2019a}.

As the Midcontinent Rift magmatic activity wanes its strength and evolved into a failed intracontinental rift, Laurentia experienced a long period of magmatic quiescence in its interior, where basin subsidence dominated the rift basin. This was followed by the Grenville Orogeny, causing the the inversion of the rift along thrust belts such as the Keweenaw Fault (Fig. xxx). The lack of extensive magmatism from ca.1080 Ma to ca. 980 Ma in the Laurentia interior resulted in a significant gap (for at least 50 myr) of paleomagnetic poles between the Keweenawan Track (developed from interior Laurentia igneous and sedimentary rocks) and the Grenville Loop (poles developed from the Grenville Province rocks) (Fig. xxx; \cite{Swanson-Hysell2019a}).

Due to the lack of pmag records from the interior Laurentia, various hypotheses for the spatial relationship between the end of the Keweenawan Track and the Grenville Loop have been debated. One model argues for that the difference in pole locations indicates that the Grenville Province was once thousands of kilometers away from Interior Laurentia and was progressively approaching the southern margin of Laurentia in the late Mesoproterozoic \citep{Halls2015}. A corollary of this hypothesis is that there had to be a 4000 km shortening during the Grenville orogeny - colliding Amazonia-Laurentia. The other model favors that the Grenville Loop is a continuation of the Keweenawan Track - that the southerly continental drift of Laurentia continued across the equator onto the southern hemisphere from ca. 1070 Ma to ca.1000 Ma. and  approached this problem by developing data from metamorphosed dike and anorthosite and argued for Grenville

Paleomagnetic poles from the interior of Laurentia during the quiescence magmatism period is thus crucial for reconcile the discordant paleomagnetic poles of the Keweenawan Track and the Grenville Loop. The
%Q: what are you using the poles from rocks that we know are from the Laurentia interior? (outside of the Grenville package?)

This new age constraint on the Jacobsville provides crucial insight into the tectonic evolution of the Midcontinent Rift and the inversion of the extensional region into a compressional region due to the Grenville orogen. It also makes the Jacobsville sandstone a unique opportunity to investigate Laurentia's paleogeographical position at ca. 1 Ga as a long period of magmatic quiescence for $\sim$50 myr preceeds Jacobsville deposition left a gap in Laurentia apparent polar wander path between the Keweenawan Track and the Grenville Loop (Fig. xxx).

Jacobsville sedimentation was preceded by a long period of volcanic and tectonic quiescence and cratonic stability so that bedrock surfaces became blanketed by paleosol and a surface of chemically resistant debris was dominated by quartz and iron-formation.

Previous paleomagnetic study on the red beds within the Jacobsville by \cite{Roy1978a} observed different components using AF, thermal, and chemical demagnetizations. Recognized dual polarities and recognized the complex remanence acquisition of red beds which lead to three isolated remanence components from the Jacobsville. However, that study did not recognize the problem of inclination shallowing and did not apply correction factor on the results. Moreover, the course demagnetization steps during thermal demagnetization would have had limited resolution for distinguishing the depositional remanent magnetization (DRM) and secondary chemical remanent magnetization (CRM) which have recently been shown to unblock over a narrow temperature window \cite{Swanson-Hysell2019b}.

In this study, we use fine-grained to silt-sized, hematite rich red beds from xxx sections along the Keweenaw Peninsula, Michigan with the goal of obtaining high-quality paleomagnetic pole from the jacobsville formation that is corrected for inclination shallowing problem and to investigate the connection between the Keweenawan Track and the Grenville Loop of the Laurentia APWP during the late Mesoproterozoic to the Neoproterozoic.

\section*{GEOLOGIC SETTINGS}

The Jacobsville sandstone is 
\cite{Kalliokoski1982a} In the Lake Superior syncline the Jacobsville Sandstone is a thick (+900m) fluvial sequence of feldspathic and quartzose sandstones, conglomerates, siltstones, and shales, completely devoid of lava flows or cross-cutting dikes. On the north and south sides of Lake Superior, most of the sandstone occurs as inward-dipping, fault-bounded wedges, separated by regional faults from the Oronto and Bayfield Groups of similar red sandstones, situated on the inner side of the syncline.

With its age being long debated through the past decades, recently detrital zircon work by \cite{Malone2016a} and \Hodgin{2021a} settled that the maximum deposition age of the Jacobsville sandstone  is 980 Ma, indicating the deposition of the Jacobsville predate or the upper Jacobsville overlaps with the Rigolet phase of the Grenville orogeny. The upper age of the Jacobsville is xxx Ma.

Keweenaw Fault motion post dating Jacobsville in the footwall
Youngest DZ from jacobsvill eis about 993 Ma (sandstone creek, about a few hundred meters from the fault). 
directly above pmag sampling at Agate falls grain B geochron date is (z2, s226, about 1000 Ma grain)

jacobsville (985.5 from calcite in the fault and <993 from zircons)

Blake's current tectonic model: after MCR volcanics ca. 1084 Ma emplacement follows the deposition of the sediments (Oronto Group) copper harbor conglomerate, Nonesuch formation and the Freda formation. Then the Grenville orogenesis took off with the Ottawan phase from 1020 Ma - 10?? Ma and the Rigolet phase from 995-980 Ma. Paleosol development and the unconformity between the Jacobsville and the volcanics indicate a prolonged period of gap during which the Oronto Group in the backbulge basin where they were deposited were lost, resulting in younger Jacobsville deposition directly on top of the volcanics. The fact that in the Jacobsville has conglomerate and sandstone and siltstone facies in them supports sediment input from the eroded Oronto Group. This indicate the older boudnary of the Jacobsville deposition might be closer to the Rigolet Phase than the Ottowan phase. The younger age boundary of the Jacobsville is well defined - given that the maximum deposition age from the zircons of the Jacobsville is 993 Ma and the calcite on the Keweenaw Fault is 985.5 Ma, the deposition of at least the upper portion of the Jacobsville is well constrained to be between 993 and 985 Ma. The deposition of Jacobsville in the back bulge in Laurentia interior is related to the far-field compression caused by the Grenville Orogeny.

This makes the Jacobsville sandstone one unique opportunity to acquire a paleomagnetic pole position within the gap between the

\cite{Roy1978a} sampled 37 sites across a lateral distance of

%\begin{figure}[!ht]
%\noindent\includegraphics[width=\textwidth]{}
%\centering
%\caption{\small{}}
%\label{fig:map}
%\end{figure}

\section*{METHODS and RESULTS}

\section*{DISCUSSION}

\subsection*{ACKNOWLEDGEMENTS}
%We thank James Pierce for his field assistance. 
%\footnotesize

\singlespacing
\clearpage
\bibliographystyle{gsabull}
\bibliography{YZ_ref}

\end{document}

### 2021-05-26 Manuscript/Jacobsville_manuscript.tex

--- 2021-05-24 Manuscript/Jaconsville_manuscript.tex
@@
 The extensive paleomagnetic studies of the well-preserved late Mesoproterozoic to Neoproterozoic rocks of North America have provided the central record for global paleogeography reconstruction during the assembly and rearrangement of the postulated supercontinent Rodinia.
-Paleomagnetic study of rocks from Laurentia have led to a series of paleomagnetic poles that form an apparent polar wander path (APWP) known as the ``Logan Loop" for the older high­ latitude poles at its apex that continues into the ``Keweenawan Track" of younger lower ­latitude poles that form a progression as the APWP heads toward the ``Grenville Loop" \citep{Swanson-Hysell2019a}.
+Paleomagnetic study of rocks from Laurentia have led to a series of paleomagnetic poles that form an apparent polar wander path (APWP) known as the ``Logan Loop" for the older high latitude poles at its apex that continues into the ``Keweenawan Track" of younger lower latitude poles that form a progression as the APWP heads toward the ``Grenville Loop" \citep{Swanson-Hysell2019a}.
 In particular, recent advances in pairing high-precision zircon U-Pb geochronology with high-quality paleomagnetic poles significantly improved the resolution of the Keweenawan Track and revealed Laurentia's rapid motion ($>$20 cm/yr) during the Midcontinent Rift magmatic activity from ca.
@@
 It also makes the Jacobsville sandstone a unique opportunity to investigate Laurentia's paleogeographical position at ca.
-1 Ga as a long period of magmatic quiescence for $\sim$50 myr preceeds Jacobsville deposition left a gap in Laurentia apparent polar wander path between the Keweenawan Track and the Grenville Loop (Fig.
+1 Ga as a long period of magmatic quiescence for $\sim$50 myr precedes Jacobsville deposition left a gap in Laurentia apparent polar wander path between the Keweenawan Track and the Grenville Loop (Fig.
 xxx).
@@
 The fact that in the Jacobsville has conglomerate and sandstone and siltstone facies in them supports sediment input from the eroded Oronto Group.
-This indicate the older boudnary of the Jacobsville deposition might be closer to the Rigolet Phase than the Ottowan phase.
+This indicate the older boundary of the Jacobsville deposition might be closer to the Rigolet Phase than the Ottawan phase.
 The younger age boundary of the Jacobsville is well defined - given that the maximum deposition age from the zircons of the Jacobsville is 993 Ma and the calcite on the Keweenaw Fault is 985.5 Ma, the deposition of at least the upper portion of the Jacobsville is well constrained to be between 993 and 985 Ma.
@@
 This makes the Jacobsville sandstone one unique opportunity to acquire a paleomagnetic pole position within the gap between the
+
+\subsection{Jacobsville stratigraphy}
+
+Agate Falls
+
+Natural Wall
+
+Snake Creek
+
+JK creek
 

### 2021-06-17 Manuscript/Jacobsville_manuscript.tex
Updates from Overleaf

@@
 
-{\Large \textbf{Continuous motion from the Keweenawan track to the Grenville Loop: new paleomagnetic pole of the ca. 1 Ga Jacobsville sandstone}}
+{\Large \textbf{Laurentia's continued drifting into the Neoproterozoic: new paleomagnetic pole from the upper Jacobsville sandstone}}
 \\
@@
 \section*{ABSTRACT}
+New paleomagnetic pole developed from red beds in the upper section of the ca.
+990 Ma Jacobsville Sandstone positions Laurentia in the southern hemisphere.
+The new data suggest Laurentia's continued southerly motion during post-Midcontinent Rift magmatism.
+This paleomagnetic pole from the intracontinental sedimentary package is inconsistent with Laurentia to have experienced 4000 km of shortening associated with the Grenville Orogeny.
 
@@
 The extensive paleomagnetic studies of the well-preserved late Mesoproterozoic to Neoproterozoic rocks of North America have provided the central record for global paleogeography reconstruction during the assembly and rearrangement of the postulated supercontinent Rodinia.
-Paleomagnetic study of rocks from Laurentia have led to a series of paleomagnetic poles that form an apparent polar wander path (APWP) known as the ``Logan Loop" for the older high latitude poles at its apex that continues into the ``Keweenawan Track" of younger lower latitude poles that form a progression as the APWP heads toward the ``Grenville Loop" \citep{Swanson-Hysell2019a}.
+Paleomagnetic studies of rocks from Laurentia have led to a series of paleomagnetic poles that form an apparent polar wander path (APWP) known as the ``Logan Loop" for the older high latitude poles at its apex that continues into the ``Keweenawan Track" of younger lower latitude poles that form a progression as the APWP heads toward the ``Grenville Loop" \citep{Swanson-Hysell2019a}.
 In particular, recent advances in pairing high-precision zircon U-Pb geochronology with high-quality paleomagnetic poles significantly improved the resolution of the Keweenawan Track and revealed Laurentia's rapid motion ($>$20 cm/yr) during the Midcontinent Rift magmatic activity from ca.
@@
 
+Along the south shore of Lake Superior, the Keweenawan Supergroup overlies Paleoproterozoic basement and consists of ~5 km of rift-related volcanic rocks that are conformably overlain by ~4 km of conglomerate, siltstone, and sandstone of the Oronto Group (Cannon et al., 1996).
+The uppermost unit of the Keweenawan Supergroup is the Jacobsville Sandstone, which consists of up to ~1 km of fluvial deposits that are more quartz-rich than the Oronto Group (Fig.
+1; Kalliokoski, 1982).
+The Jacobsville Sandstone unconformably overlies Archean and Paleoproterozoic basement as well as Midcontinent Rift volcanic and sedimentary rocks in northern Michigan (Fig.
+1; Cannon et al., 1993; Ojakangas et al., 2001).
+The overall extent of the Jacobsville Sandstone has been mapped along the southern shore of Lake Superior, from west to east - the Keweenaw Peninsula of northern Michigan to Sault Ste Marie, Canada (Fig.
+\ref{}).
+The Jacobsville Sandstone is unconformably overlain by the Cambrian Munising Formation (Hamblin, 1958).
+
+The Keweenaw Fault is a north to northwest-dipping thrust that extends ~250 km from the tip of the Keweenaw Peninsula in northern Michigan southeastward to a termination in northeastern Wisconsin (Fig.
+1; Cannon et al., 1996; Cannon and Nicholson, 2001).
+Thrust faults parallel to the Keweenaw Fault include the northwest-dipping Marenisco and Pelton Creek faults (Fig.
+1).
+Similarly-oriented thrusts continue southwestward through northern Wisconsin and eastern Minnesota (Ojakangas et al., 2001), and eastward below Lake Superior, where they are interpreted to join mapped thrust faults in southern Ontario (Fig.
+1; Manson and Halls, 1994; Ojakangas et al., 2001).
+Rock units adjacent to these regional thrusts have been moderately to intensely folded.
+In northern Michigan, the typically shallowly-dipping Jacobsville Sandstone has sub-vertical to overturned beds in the immediate footwall of the Keweenaw Fault (Irving and Chamberlin, 1885; Cannon and Nicholson, 2001).
+Near the Keweenaw Fault, Jacobsville Sandstone consists of quartz-rich, cross-bedded, medium-grained sandstone, locally abundant clast-supported conglomerate, and minor interbeds of red micaceous siltstone.
+The conglomerate contains volcanic clasts likely derived from uplifted Midcontinent Rift volcanics (Irving and Chamberlin, 1885; Brojanigo, 1984).
+Clast provenance in Jacobsville conglomerates includes Paleoproterozoic lithologies, such as vein quartz and iron formation clasts, which are abundant in southern exposures (Hamblin, 1958; Kalliokoski, 1982).
+Occurrence of conglomerates in proximity to the Keweenaw, Marenisco, and Pelton Creek thrusts is interpreted to represent syn-depositional development of local relief (Kalliokoski, 1982; Hedgman, 1992).
+
+It has been interpreted that the Jacobsville Sandstone is equivalent to the Bayfield Group in northern Wisconsin, the Fond Du Lac formation in Minnesota.
+Their shared lithofacies of conglomerate, sandstone, and siltstones as well as similar pattern of lithology transition from conglomerate dominated deposition to sandstone to occurrence of very fine grained siltstone facies  supports the interpretation that these sedimentary rocks formed during the period of thermal subsidence and inversion of the continental rifting after the major magmatism of the MCR.
+
+Red beds, particular fine-grained siltstone beds have been shown to be good paleomagnetic recorders, preserving magnetic remanence for up to billions of years.
+Occurrence of red siltstone beds have been reported in the Jacobsville Sandstone, particularly in the upper sections near Keweenaw Fault (e.g.
+Hamblin1958a, Irving1885a).
+They typically in the field are brick-red to dark-red, find-grained arenite to arkosic sandstone and siltstone.
+Minor grey reduction spots can be found.
+rounded detrital mica grains up to 5mm?
+size can be found commonly within very fine, fissile siltstone and shaly strata.
+
+\subsubsection{Paleomagnetic sample sites}
 Agate Falls
+A MDA of --- Ma was obtained from this site by Hodgin2021a.
 
 Natural Wall
+A MDA of --- Ma was obtained from this site by Hodgin2021a
 
 Snake Creek
+A MDA of --- Ma was obtained from this site by Hodgin2021a
 
 JK creek
+A MDA of --- Ma was obtained from this site by Hodgin2021a
 
@@
 
+paleomagnetic mean pole of the Jacobsville Sandstone and the age
+
+inclination shallowing correction of the Jacobsville Sandstone
+
+magnetic reversal during the Jacobsville and the Maya Superchron
+
+Implication on the crustal shortening of the Rigolet phase of the Grenville Orogeny
+
+Warnock on the age of magnetization on the Haliburton pole: based on the very coarse thermal demagnetization steps Warnock2000a assigned that all of the primary magnetite components unblocked above 550 \textdegree C, with an uncertainty of 555 \textdegree C $\pm$ 5\textdegree C.
+Using a thermal history model that combines U-Pb titanite ages, hornblende and biotite Ar/Ar ages and just by merely compare the thermal history model with the assigned narrow magnetite unblocking temperature range, the authors assigned the age of the magnetization to be 1015 $\pm$ 15 Ma.
+
+It is important to note that the result Ar ages of the hornblende and biotite are 983 $\pm$ 8 Ma, 996 $\pm$ 8 Ma, and 998 $\pm$ 8 for hornblende with unblocking temperature of 530 $\pm$ 40\textdegree C, consistent across metamorphic terranes (Bancroft).
+And the biotite Ar ages are 910 $\pm$ 7 Ma, 891 $\pm$ 7 Ma, corresponding to unblocking temperature of 280 $\pm$ 40\textdegree C.
+
+More, the thermal demagnetization plots shown in Warnock2000a suggest that the magnetite unblocking lower temperature boundary can be around 500\textdegree C, lower than 555 $\pm$ 5\textdegree C.
+In addition, given the case that the rocks underwent uniform rate of cooling due to post-formation uplifting, a slow cooling effect from 555\textdegree C to 300\textdegree C from 990 Ma to 900 Ma allows for a scenario where the magnetite remament magnetization that appear to have a 555\textdegree C unblocking temperature in lab was acquired through a period of --- myr cooling though 500\textdegree C during uplifting.
+In this case, the magnetite remanence acquisition can be interpreted to be closer to the biotite Ar age than previously thought.
+Therefore, the Haliburton intrusion pole could be interpreted to be a post Rigolet orogenic activity phase magnetization acquisition, and the apparent discrepancy between the Haliburton pole and the Jacobsville pole can be explained by Laurentia's whole plate motion after the Grenvillian orogeny from 990 Ma to 900 Ma.
+One corollary of such scenario is that Laurentia underwent plate motion at a rate of ---\textdegree/myr, which is similar to that of today's tectonic plate motion.
+
+Crucial link between the Keweenawan Track and the Grenville Loop
+
 \subsection*{ACKNOWLEDGEMENTS}

### 2022-02-01 Manuscript/Jacobsville_manuscript.tex

@@
 
-{\Large \textbf{Laurentia's continued drifting into the Neoproterozoic: new paleomagnetic pole from the upper Jacobsville sandstone}}
+{\Large \textbf{Laurentia's continued motion into the Neoproterozoic: new paleomagnetic pole from the Jacobsville sandstone}}
 \\
 \singlespacing
- Yiming Zhang\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, Blake Hodgin\textsuperscript{1}\\
+ Yiming Zhang\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, Blake Hodgin\textsuperscript{1}, James Pierce\textsuperscript{1}, Anthony Fuentes\textsuperscript{1}\\
 \bigskip
@@
 \section*{ABSTRACT}
-New paleomagnetic pole developed from red beds in the upper section of the ca.
-990 Ma Jacobsville Sandstone positions Laurentia in the southern hemisphere.
-The new data suggest Laurentia's continued southerly motion during post-Midcontinent Rift magmatism.
-This paleomagnetic pole from the intracontinental sedimentary package is inconsistent with Laurentia to have experienced 4000 km of shortening associated with the Grenville Orogeny.
+
+The ca.
+1.1 Ga North American Midcontinent Rift resulted in protracted magmatism and the associated deposition of sediments in the interior of Laurentia while the continent traveled rapidly from high to low latitudes.
+Coeval with the cessation of rifting and decrease in the speed of plate motion is the ca.
+1090-980 Ma Grenvillian Orogeny whose far-field stress is interpreted to have propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation.
+A lack of paleomagnetic record from the interior of Laurentia between the end of the Midcontinent Rift and the cooling of rocks associated with the Grenvillian Orogeny resulted in a gap of record for the position and orientation of Laurentia for nearly 100 Myr.
+We develop a new paleomagnetic pole from from red beds of the Jacobsville Formation along 5 stratigraphic sections in Northern Michigan, US.
+High-resolution thermal demagnetization experiments together with successful fold tests resolve primary magnetic remanence which show shallow inclinations despite being  flattened.
+Inclination shallowing corrected paleomagnetic pole position places Laurentia at low latitudes south to the Keweenawan Track ca.
+990 Ma, confirming continued plate motion of Laurentia across the equator in the Late Mesoproterozoic to early Neoproterozoic.
+The new pole from the interior of Laurentia indicate the Grenville Loop poles are younger than previously thought.
 
 \section*{INTRODUCTION}
+
+Paleomagnetic data for the Jacobsville have led to a paleomagnetic pole position that appears to be a continuation of the Keweenawan Track from the Oronto Group poles (Roy and Robertson, 1978; Fig.
+7; Table 4).
+As a result, its age has been interpolated to be latest Mesoproterozoic (e.g., 1020 Ma in Li et al., 2008) such that it was deposited prior to the ca.
+1000 Ma Grenville loop poles.
+However, this age assignment has recently been questioned on the basis of laser-ablation–inductively cou- pled plasma–mass spectrometry (LA-ICP-MS) U-Pb detrital zircon dates from Jacobsville Sandstone samples.
+Malone et al.
+(2016) interpreted a maximum depositional age of 959 ± 19 Ma based on the weighted mean of the four youngest grains from 2050 LA-ICP-MS zircon dates.
+A similar age was proposed on the basis of the youngest LA-ICP-MS U-Pb dates in the study of Craddock et al.
+(2013).
+This young maximum interpreted age is intriguing because it suggests the unconformity above the Freda Sandstone spans at least 100 m.y.
+and was followed by continued sedi- mentation in the region long after rift activity.
+Nevertheless, the paleo- magnetic pole of the Jacobsville Sandstone (Roy and Robertson, 1978) appears to extend the Keweenawan Track from the Oronto Group poles in a manner that is consistent with deposition of the Jacobsville Sand- stone in the late Mesoproterozoic.
+Given the conflict with this inferred age, the large uncertainty of individual LA-ICP-MS dates, and the pos- sibility of outlier dates in such a large data set, it would be beneficial for the youngest Jacobsville zircons to be analyzed using CA-ID-TIMS to obtain more precise constraints.
 

### 2022-02-01 Manuscript/Manuscript.tex
format initial manuscript file in AGU template

--- 2022-02-01 Manuscript/Jacobsville_manuscript.tex
@@
-\documentclass[11pt,letterpaper]{article}
-\usepackage{xcolor}
-\usepackage{textcomp,marvosym}
-\usepackage{amsmath,amssymb}
-\usepackage[left]{lineno}
-\usepackage{changepage}
-\usepackage{rotating}
-\usepackage{natbib}
-\usepackage{setspace}
-\usepackage{fancyhdr}
-\usepackage{graphicx}
-\usepackage{sidecap}
-\usepackage{pdfpages}
-\usepackage{longtable}
-\usepackage{url}
+\documentclass[draft]{agujournal2019}
+\usepackage{url} %this package should fix any errors with URLs in refs.
+\usepackage{lineno}
+\usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
+\usepackage{soul}
+\linenumbers
 
-\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
-\doublespacing
+\draftfalse
 
-\raggedright
-\textwidth = 6.5 in
-\textheight = 8.25 in
-\oddsidemargin = 0.0 in
-\evensidemargin = 0.0 in
-\topmargin = 0.0 in
-\headheight = 0.0 in
-\headsep = 0.5 in
-\parskip = 0.1 in
-\parindent = 0.2 in
+%% Enter journal name below.
+%% Choose from this list of Journals:
+%
+% JGR: Atmospheres
+% JGR: Biogeosciences
+% JGR: Earth Surface
+% JGR: Oceans
+% JGR: Planets
+% JGR: Solid Earth
+% JGR: Space Physics
+% Global Biogeochemical Cycles
+% Geophysical Research Letters
+% Paleoceanography and Paleoclimatology
+% Radio Science
+% Reviews of Geophysics
+% Tectonics
+% Space Weather
+% Water Resources Research
+% Geochemistry, Geophysics, Geosystems
+% Journal of Advances in Modeling Earth Systems (JAMES)
+% Earth's Future
+% Earth and Space Science
+% Geohealth
+%
 
-\pagestyle{myheadings}
-\pagestyle{fancy}
-\fancyhf{}
-\lhead{}
-\rhead{\thepage}
+\journalname{Geochemistry, Geophysics, Geosystems}
 
@@
 
-\begin{flushleft}
+\title{Laurentia's continued motion into the Neoproterozoic: new paleomagnetic pole from the Jacobsville sandstone}
 
-{\Large \textbf{Laurentia's continued motion into the Neoproterozoic: new paleomagnetic pole from the Jacobsville sandstone}}
-\\
-\singlespacing
- Yiming Zhang\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, Blake Hodgin\textsuperscript{1}, James Pierce\textsuperscript{1}, Anthony Fuentes\textsuperscript{1}\\
-\bigskip
-\textsuperscript{1} Department of Earth and Planetary Science, University of California, Berkeley, CA, USA\\
+\authors{Yiming Zhang\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, Blake Hodgin\textsuperscript{1}, James Pierce\textsuperscript{1}, Anthony Fuentes\textsuperscript{1}}
 
-\smallskip
-\end{flushleft}
+\affiliation{1}{Department of Earth and Planetary Science, University of California, Berkeley, CA, USA}
+% \affiliation{2}{Second Affiliation}
+% \affiliation{3}{Third Affiliation}
+% \affiliation{4}{Fourth Affiliation}
 
-%\linenumbers
-%\pagestyle{empty}
+\correspondingauthor{Yiming Zhang}{yimingzhang@berkeley.edu}
+
+\begin{keypoints}
+\item A new paleomagnetic pole has been developed from the ca.990 Ma Jacobsville Formation 
+\item Laurentia plate continued to move after 
+\item The new pole from the Jacobsville Formation indicates the Grenville Loop could be younger than thought
+\end{keypoints}
+
+\begin{abstract}
+[ enter your Abstract here ]
+\end{abstract}
+
+\section*{Plain Language Summary}
+[ enter your Plain Language Summary here or delete this section]
 
@@
 A lack of paleomagnetic record from the interior of Laurentia between the end of the Midcontinent Rift and the cooling of rocks associated with the Grenvillian Orogeny resulted in a gap of record for the position and orientation of Laurentia for nearly 100 Myr.
-We develop a new paleomagnetic pole from from red beds of the Jacobsville Formation along 5 stratigraphic sections in Northern Michigan, US.
+To fill this gap, we develop a new paleomagnetic pole from from red beds of the Jacobsville Formation along 5 stratigraphic sections in Northern Michigan, US.
 High-resolution thermal demagnetization experiments together with successful fold tests resolve primary magnetic remanence which show shallow inclinations despite being  flattened.
-Inclination shallowing corrected paleomagnetic pole position places Laurentia at low latitudes south to the Keweenawan Track ca.
+Inclination shallowing corrected paleomagnetic pole position places Laurentia at low latitudes south of the Keweenawan Track ca.
 990 Ma, confirming continued plate motion of Laurentia across the equator in the Late Mesoproterozoic to early Neoproterozoic.
-The new pole from the interior of Laurentia indicate the Grenville Loop poles are younger than previously thought.
+The new pole from the interior of Laurentia indicates the Grenville Loop poles are younger than previously thought.
+
+\begin{figure}
+\noindent\includegraphics[width=\textwidth]{../Figures/Geologic_map.pdf}
+\caption{caption}
+\label{pngfiguresample}
+\end{figure}
 
@@
 
-Paleomagnetic data for the Jacobsville have led to a paleomagnetic pole position that appears to be a continuation of the Keweenawan Track from the Oronto Group poles (Roy and Robertson, 1978; Fig.
+Paleomagnetic data for the Jacobsville Formation have led to a paleomagnetic pole position that appears to be a continuation of the Keweenawan Track from the Oronto Group poles (Roy and Robertson, 1978; Fig.
 7; Table 4).
@@
 This young maximum interpreted age is intriguing because it suggests the unconformity above the Freda Sandstone spans at least 100 m.y.
-and was followed by continued sedi- mentation in the region long after rift activity.
-Nevertheless, the paleo- magnetic pole of the Jacobsville Sandstone (Roy and Robertson, 1978) appears to extend the Keweenawan Track from the Oronto Group poles in a manner that is consistent with deposition of the Jacobsville Sand- stone in the late Mesoproterozoic.
-Given the conflict with this inferred age, the large uncertainty of individual LA-ICP-MS dates, and the pos- sibility of outlier dates in such a large data set, it would be beneficial for the youngest Jacobsville zircons to be analyzed using CA-ID-TIMS to obtain more precise constraints.
+and was followed by continued sedimentation in the region long after rift activity.
+Nevertheless, the paleo- magnetic pole of the Jacobsville Sandstone (Roy and Robertson, 1978) appears to extend the Keweenawan Track from the Oronto Group poles in a manner that may be more consistent with deposition of the Jacobsville Formation in the late Mesoproterozoic.
+Given the conflict with this inferred age, the large uncertainty of individual LA-ICP-MS dates, and the pos- sibility of outlier dates in such a large data set, it is beneficial to interpret the age of the youngest Jacobsville zircons dated CA-ID-TIMS to obtain more precise constraints.
+Using CA-ID-TIMS, a recent study by Hodgin et al (2022) redefines the maximum deposition age of the Jacobsville Formation as ca.
+993 Ma.
 
 The extensive paleomagnetic studies of the well-preserved late Mesoproterozoic to Neoproterozoic rocks of North America have provided the central record for global paleogeography reconstruction during the assembly and rearrangement of the postulated supercontinent Rodinia.
-Paleomagnetic studies of rocks from Laurentia have led to a series of paleomagnetic poles that form an apparent polar wander path (APWP) known as the ``Logan Loop" for the older high latitude poles at its apex that continues into the ``Keweenawan Track" of younger lower latitude poles that form a progression as the APWP heads toward the ``Grenville Loop" \citep{Swanson-Hysell2019a}.
+Paleomagnetic studies of rocks from Laurentia have led to a series of paleomagnetic poles that form an apparent polar wander path (APWP) known as the "Logan Loop" for the older high latitude poles at its apex that continues into the ``Keweenawan Track" of younger lower latitude poles that form a progression as the APWP heads toward the ``Grenville Loop" \citep{Swanson-Hysell2019a}.
 In particular, recent advances in pairing high-precision zircon U-Pb geochronology with high-quality paleomagnetic poles significantly improved the resolution of the Keweenawan Track and revealed Laurentia's rapid motion ($>$20 cm/yr) during the Midcontinent Rift magmatic activity from ca.
@@
 
-It has been interpreted that the Jacobsville Sandstone is equivalent to the Bayfield Group in northern Wisconsin, the Fond Du Lac formation in Minnesota.
-Their shared lithofacies of conglomerate, sandstone, and siltstones as well as similar pattern of lithology transition from conglomerate dominated deposition to sandstone to occurrence of very fine grained siltstone facies  supports the interpretation that these sedimentary rocks formed during the period of thermal subsidence and inversion of the continental rifting after the major magmatism of the MCR.
+It has been interpreted that the Jacobsville Sandstone is equivalent to the Bayfield Group in northern Wisconsin, and the Fond Du Lac and Hinckley formations in Minnesota.
+Their shared lithofacies of basal conglomerate, cross-bedded fluvial fine to medium grained sandstone, and interbedded red micaceous siltstone could  support the interpretation that these sedimentary rocks formed during a common period of post-rift basin formation following rift-related magmatism of the MCR.
 
@@
 \cite{Roy1978a} sampled 37 sites across a lateral distance of
-
-%\begin{figure}[!ht]
-%\noindent\includegraphics[width=\textwidth]{}
-%\centering
-%\caption{\small{}}
-%\label{fig:map}
-%\end{figure}
 
@@
 
-\subsection*{ACKNOWLEDGEMENTS}
-%We thank James Pierce for his field assistance. 
-%\footnotesize
+\acknowledgments
+Enter acknowledgments, including your data availability statement, here.
 
-\singlespacing
-\clearpage
-\bibliographystyle{gsabull}
 \bibliography{YZ_ref}

### 2022-03-25 Manuscript/Manuscript.tex

@@
 1.1 Ga North American Midcontinent Rift resulted in protracted magmatism and the associated deposition of sediments in the interior of Laurentia while the continent traveled rapidly from high to low latitudes.
-Coeval with the cessation of rifting and decrease in the speed of plate motion is the ca.
+Coeval with the cessation of rifting and decreasing in the speed of plate motion is the ca.
 1090-980 Ma Grenvillian Orogeny whose far-field stress is interpreted to have propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation.
-A lack of paleomagnetic record from the interior of Laurentia between the end of the Midcontinent Rift and the cooling of rocks associated with the Grenvillian Orogeny resulted in a gap of record for the position and orientation of Laurentia for nearly 100 Myr.
-To fill this gap, we develop a new paleomagnetic pole from from red beds of the Jacobsville Formation along 5 stratigraphic sections in Northern Michigan, US.
-High-resolution thermal demagnetization experiments together with successful fold tests resolve primary magnetic remanence which show shallow inclinations despite being  flattened.
+A lack of paleomagnetic record from the interior of Laurentia between the end of the Midcontinent Rift and the cooling of rocks associated with the Grenvillian Orogeny resulted in poorly constrained position and orientation of Laurentia for nearly 100 Myr.
+To fill this gap, we develop a new paleomagnetic pole from red beds of the Jacobsville Formation along 5 stratigraphic sections in Northern Michigan, US.
+Our high-resolution thermal demagnetization experiments successfully resolve primary magnetic remanence that pass a fold test and a conglomerate test.
+Our results agree with previous chemical demagnetization data but provide additional constraints on the shallowing of the inclinations and on the age of the pole once paired with max deposition ages developed from high-precision detrital zircon dating.
 Inclination shallowing corrected paleomagnetic pole position places Laurentia at low latitudes south of the Keweenawan Track ca.
@@
 The new pole from the interior of Laurentia indicates the Grenville Loop poles are younger than previously thought.
+
+\begin{keypoints}
+\item Laurentia  
+\item Inclination shallowing  
+\item Paleogeography  
+\item Grenville  
+\item Keweenawan Superchron  
+\item Red Bed 
+\end{keypoints}
 

### 2022-07-01 Manuscript/Manuscript.tex
Updates from Overleaf

--- 2022-02-01 Manuscript/Manuscript.tex
@@
 
-%% Enter journal name below.
-%% Choose from this list of Journals:
-%
-% JGR: Atmospheres
-% JGR: Biogeosciences
-% JGR: Earth Surface
-% JGR: Oceans
-% JGR: Planets
-% JGR: Solid Earth
-% JGR: Space Physics
-% Global Biogeochemical Cycles
-% Geophysical Research Letters
-% Paleoceanography and Paleoclimatology
-% Radio Science
-% Reviews of Geophysics
-% Tectonics
-% Space Weather
-% Water Resources Research
-% Geochemistry, Geophysics, Geosystems
-% Journal of Advances in Modeling Earth Systems (JAMES)
-% Earth's Future
-% Earth and Space Science
-% Geohealth
-%
-
-\journalname{Geochemistry, Geophysics, Geosystems}
+\journalname{Tectonics}
 
@@
 
-\title{Laurentia's continued motion into the Neoproterozoic: new paleomagnetic pole from the Jacobsville sandstone}
+\title{Laurentia's continued motion into the Neoproterozoic: new paleomagnetic constraint from the Jacobsville Formation}
 
-\authors{Yiming Zhang\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, Blake Hodgin\textsuperscript{1}, James Pierce\textsuperscript{1}, Anthony Fuentes\textsuperscript{1}}
+\authors{Yiming Zhang\textsuperscript{1}, Blake Hodgin\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, James Pierce\textsuperscript{1}, Anthony Fuentes\textsuperscript{1}}
 
 \affiliation{1}{Department of Earth and Planetary Science, University of California, Berkeley, CA, USA}
-% \affiliation{2}{Second Affiliation}
-% \affiliation{3}{Third Affiliation}
-% \affiliation{4}{Fourth Affiliation}
 
@@
 \begin{keypoints}
-\item A new paleomagnetic pole has been developed from the ca.990 Ma Jacobsville Formation 
-\item Laurentia plate continued to move after 
-\item The new pole from the Jacobsville Formation indicates the Grenville Loop could be younger than thought
+\item A new paleomagnetic pole from the ca.990 Ma Jacobsville Formation 
+\item Reversals put younger bounds on the normal Keweenawan superchron 
+\item The new pole indicates the Grenville Loop could be younger than thought
 \end{keypoints}
@@
 \begin{abstract}
-[ enter your Abstract here ]
+The ca.
+1.1 Ga North American Midcontinent Rift resulted in protracted magmatism and the associated deposition of sediments during rift-related extension and post-rift thermal subsidence.
+Coeval with the cessation of rifting and a decrease of the speed of plate motion is the onset of the Grenvillian Orogeny which lasted from ca.
+1090-980 Ma.
+The far-field stress of the orogeny is interpreted to have propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation whose maximum deposition age has been constrained to be ca.
+990 Ma.
+We develop a new paleomagnetic pole from red beds of the Jacobsville Formation along 5 stratigraphic sections in Northern Michigan, USA.
+High-resolution thermal demagnetization experiments resolve primary magnetic remanence that pass a fold test and two conglomerate tests.
+Our results confirm the occurrence of geomagnetic field reversals during the deposition of Jacobsville Formation, which provides constraints for the duration of the Keweenawan superchron.
+Inclination corrected paleomagnetic pole position from the Jacobsville Formation confirm continued plate motion of Laurentia across the equator in the Late Mesoproterozoic to early Neoproterozoic.
+The new pole from the interior of Laurentia indicates the paleomagnetic poles developed from the Grenville Province may record a coherent plate motion after major continental collision during the assemblage of supercontinent Rodinia rather than recording a significant crustal shortening.
 \end{abstract}
 
-\section*{Plain Language Summary}
-[ enter your Plain Language Summary here or delete this section]
+% \section*{Plain Language Summary}
+% [ enter your Plain Language Summary here or delete this section]
 
-\section*{ABSTRACT}
+\begin{keypoints}
+\item Neoproterozoic
+\item Laurentia  
+\item Paleogeography
+\item Inclination shallowing
+\item Grenville
+\item Keweenawan superchron
+\item Red bed
+\end{keypoints}
 
-The ca.
-1.1 Ga North American Midcontinent Rift resulted in protracted magmatism and the associated deposition of sediments in the interior of Laurentia while the continent traveled rapidly from high to low latitudes.
-Coeval with the cessation of rifting and decrease in the speed of plate motion is the ca.
-1090-980 Ma Grenvillian Orogeny whose far-field stress is interpreted to have propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation.
-A lack of paleomagnetic record from the interior of Laurentia between the end of the Midcontinent Rift and the cooling of rocks associated with the Grenvillian Orogeny resulted in a gap of record for the position and orientation of Laurentia for nearly 100 Myr.
-To fill this gap, we develop a new paleomagnetic pole from from red beds of the Jacobsville Formation along 5 stratigraphic sections in Northern Michigan, US.
-High-resolution thermal demagnetization experiments together with successful fold tests resolve primary magnetic remanence which show shallow inclinations despite being  flattened.
-Inclination shallowing corrected paleomagnetic pole position places Laurentia at low latitudes south of the Keweenawan Track ca.
-990 Ma, confirming continued plate motion of Laurentia across the equator in the Late Mesoproterozoic to early Neoproterozoic.
-The new pole from the interior of Laurentia indicates the Grenville Loop poles are younger than previously thought.
+\section*{Introduction}
 
-\begin{figure}
-\noindent\includegraphics[width=\textwidth]{../Figures/Geologic_map.pdf}
-\caption{caption}
-\label{pngfiguresample}
-\end{figure}
+The extensively studied paleomagnetic records of the ca.
+1109 Ma to 1084 Ma North American Midcontinent Rift (i.e.
+the Keweenawan Track) provides crucial constraints on global paleogeography in the Mesoproterozoic and has become the central record for reconstructing the development of ancient supercontinent Rodinia \cite{Swanson-Hysell2019a, Swanson-Hysell2021c}.
+As the rifting ceased and subsequently inverted following the onset of the Grenvillian Orogeny on Laurentia's eastern margin (Fig.
+\ref{fig:Geologic_map}), magmatism in the interior of Laurentia entered a quiescence period that could have lasted as much as $\sim$300 Myr until the emplacement of the ca.
+780 Ma Gunbarrel large igneous province \cite{Harlan2003a}.
+As a result, uncertainties exist in reconstructing Laurentia and its conjugate continents' configuration in the late Mesoproterozoic to early Neoproterozoic.
 
-\section*{INTRODUCTION}
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{Geologic_map.pdf}
+\caption{(A) Regional map showing location of the Grenville Front relative to the study area.
+(B) Geologic map of northern Michigan and Wisconsin showing Midcontinent Rift igneous and sedimentary rocks, the Jacobsville Formation, and major post-rift thrust faults.
+Jacobsville stratigraphic sections along which paleomagnetic samples were collected are shown as green diamond symbols.}
+\label{fig:Geologic_map}
+\end{figure*}
 
-Paleomagnetic data for the Jacobsville Formation have led to a paleomagnetic pole position that appears to be a continuation of the Keweenawan Track from the Oronto Group poles (Roy and Robertson, 1978; Fig.
+Although igneous activities are lacking, sedimentary rocks formed near the end of the Midcontinent Rift magmatism and afterwards, and can provide additional paleomagnetic data.
+In particular, hematite-bearing sedimentary rocks (i.e.
+red beds) can provide insights into the ancient geomagnetic field and the paleogeographic positions of sedimentary basins.
+In the Midcontinent Rift, the paleomagnetic poles developed from the ca.
+1075 Ma Nonesuch Formation and the ca.
+1070 Ma upper Freda Formation of the Oronto Group sediments that overlie the Midcontinent Rift volcanics in southern Lake Superior area (Fig.
+\ref{fig:Geologic_map}) provide crucial constraints for Laurentia's latitude, orientation, and plate motion in the late Mesoproterozoic \cite{Henry1977a}.
+
+In addition to the Oronto Group sediments, the overlying post-rift Jacobsville Formation which were deformed during the inversion of the Midcontinent Rift also contains abundant red beds which bear the opportunity to further extend the reconstruction of the apparent polar wander path of Laurentia toward more recent times \cite{Hamblin1958a, Hodgin2022a}.
+U-Pb detrital zircon ages and U-Pb calcite ages from late-kinematic veins constrain the maximum depositional age of the Jacobsville Formation to be ca.
+990 Ma \cite{Hodgin2022a}.
+Developing paleomagnetic data from the Jacobsville Formation will allow further development of Laurentia's paleogeographic evolution in the late Mesoproterozoic to early Neoproterozoic.
+
+Challenges exist in interpreting sedimentary paleomagnetic records as they acquire magnetic remanence not through cooling from above the magnetic ordering temperatures of magnetic minerals as igneous rocks do, but acquire detrital remanence as magnetic particles settle from the water column during deposition, and chemical remanence as magnetic minerals precipitate and grow within pore spaces of host sedimentary rocks syn- to post-deposition (i.e.
+pigmentary hematite).
+It has been found that the formation of the pigmentary hematite in red beds can significantly post-date the timing of sediment deposition, resulting in the chemical remanent magnetization being secondary, superimposed upon the primary remanent magnetization carried by detrital grains \cite{Butler1992a}.
+To isolate the remanence components of the red beds of the Jacobsville Formation, \citeA{Roy1974, Roy1978} adopted a method first applied by \citeA{Collinson1965a} to preferentially remove fine-grained pigmentary hematite through prolonged immersion in concentrated HCl acid whereas leaving the coarser detrital grains that are more resistant to dissolution in place.
+Paired acid etching demagnetization and thermal demagnetization experiments show that the pigmentary hematite grains tend to have lower unblocking temperatures than detrital ones \cite{Bilardello2010a, Tauxe1980a}.
+High-resolution thermal demagnetization experiments with temperature intervals as small as 1-2 \textdegree C have shown to be effective in isolating detrital remanence from chemical remanence as coarser hematite grains tend to have higher unblocking temperatures \cite{Swanson-Hysell2019b}.
+In particular, \cite{Swanson-Hysell2019b} conducted a conglomerate test on a suite of fluvial intraclasts of the red beds in the Freda Formation to show that via thermal demagnetization one can isolate the primary magnetization (which are reoriented in random directions in the clasts) from the secondary component carried by the co-occurring authigenic hematite (which yield similar directions amongst clasts).
+
+Paleomagnetism of red beds has also been challenged by the controversial issue of correcting for inclination shallowing \cite{King1955a, Taux2004b, Bilardello2016b}.
+Due to the rotation of detrital magnetic grains toward bedding-parallel caused by authigenic processes such as compaction during lithification, inclinations recorded by sediments often have shallower angles than what was during the time of deposition.
+To correct for such bias, \citeA{Tauxe2004b} developed a statistical model based on the geocentric axial dipole (GAD) hypothesis where the amount of inclination shallowing in a specific sediment package is evaluated based on the shape of the distribution of the sedimentary paleomagnetic directions is compared to that predicted by a paleosecular variation model developed for the past 5 Myr.
+\citeA{Pierce2022a} showed that the paleosecular variation model of \citeA{Tauxe2004b} is also suitable for application to sedimentary rocks formed during the Midcontinent Rift.
+That study also further extended the method to better represent uncertainties associated with the amount of inclination shallowing when reporting sedimentary paleomagnetic pole positions by using the spherical elliptical Kent distribution \cite{Kent1982a} instead of the commonly used circular Fisher distribution \cite{Fisher1953a}.
+
+In this study, we investigate the hematite-bearing, fine-grained sandstone to siltstone lithofacies along five stratigraphic sections of the Jacobsville Formation.
+High-resolution thermal demagnetization data on centimeter-scale siltstone intraclasts that were eroded by fluvial processes and redeposited at two locations constrain the timing of hematite acquisition by revealing a primary component that formed prior to the erosion of the clasts within the depositional environment and a secondary component that formed following their redeposition.
+Then we develop a new inclination-corrected paleomagnetic pole for the Jacobsville Formation based on a large number of samples and investigate the paleo-continental configuration of Laurentia and its association with Baltica during the development of the supercontinent Rodinia.
+
+% preservation of the Midcontinent Rift sedimentary rocks (Figure \ref{fig:Geologic_map}), complex magnetic remanence components that often form during significantly different times within the rocks and post-deposition modification of the remanence magnetization directions are common to sedimentary rocks, presenting challenges to the interpretation of their paleomagnetic records \cite{Beck2003a, Butler1992a, Vandervoo2012a}.
+In contrast to igneous rocks which acquire thermal remanent magnetizations through cooling of iron oxides, red beds often record mixtures of magnetic remanence components during primary hematite grain deposition and secondary hematite precipitation and growth that are often not coeval \cite{Tauxe1980a, Swanson-Hysell2019b}.
+Detrital hematite grains can record magnetic remanence at the time of deposition, whose age can be constrained by the maximum deposition age of the host sedimentary rock.
+The secondary source of hematite is the grains that precipitate and grow in situ within pore space of host sedimentary rocks that have been deposited (i.e.
+pigmentary hematite).
+These secondary hematite could form significantly after deposition of the sedimentary rocks and can keep growing through a protracted period of time.
+To isolate the primary magnetization component from the secondary one within red beds is crucial for accurately reconstructing paleomagnetic records at the time of red beds deposition.
+Another challenge to interpreting sedimentary paleomagnetic data is the problem of inclination shallowing \cite{King1955a}.
+Grain rotation toward being bedding plane parallel during processes associated with deposition and lithification can result in recorded inclinations being shallower than the local magnetic inclination.
+Such deviation may vary depending on the location where the sedimentary rock formed and can result in various degrees of uncertainties of paleomagnetic reconstructions.
+
+% The end of the ca.
+1109 Ma to 1084 Ma intracontinental Midcontinent Rift magmatism and extension was marked by a transition to deposition of sedimentary rocks of the Oronto Group \cite{Ojakangas2001a} as the rift basin thermally subsided prior to the rift undergoing contractional deformation in the later stages of the Grenvillian orogeny \cite{Cannon1992b, Hodgin2022a}.
+
+%  that has become central to the reconstruction of paleogeography of the assembly of supercontinent Rodinia.
+In addition to syn-rift rocks preservation, the cessation and later inversion of the rift allowed the development of syn-to post-rift basin development which accommodated the deposition of the ca.
+1080 - 1050 Ma Oronto Group sedimentary rocks that are in total about 5 km thick.
+Overlying the Oronto Group is the Jacobsville formation which has been interpreted to be deposited in a back-bulge setting associated with the far-field propagation of the crustal-scale deformation caused by the Grenvillian Orogeny.
+
+% Previous paleomagnetic data from the Midcontinent Rift rocks that are paired with high-precision geochronology data has led to the development of the Keweenawan Track, showing rapid plate motion (latitudinal motion of $\sim$ cm/yr) from high latitudes between ca.
+1105 Ma and ca.
+1092 Ma, followed by a decrease in plate motion (latitudinal motion of $\sim$) between ca.
+1092 Ma and ca.
+1080 Ma when Laurentia moved to the equator and ceased its rapid motion.
+In particular, paleomagnetic poles developed from the ca.
+1080 Ma Nonesuch Formation and the ca.
+1050 Ma upper Freda Formation of the Oronto Group sediments indicate that Laurentia's plate speed significantly reduced in the late Mesoproterozoic, resulting in little change in its paleocontinental configuration for tens of millions of years.
+
+% A leading tectonic model for the cessation of the Midcontinent Rift and extension is that it occurred due to far-field compressional stress associated with the ca.
+1090-980 Ma Grenvillian orogeny \cite{Cannon1992a}, when the collision of Laurentia with its conjugate margins led to the development of a back-bulge basin, accommodating the deposition of the Jacobsville Formation sediments as the inversion of the Midcontinent Rift progressed.
+Recent re-examination of the maximum deposition age of the Jacobsville Formation using zircon U-Pb whole grain chemical abrasion isotope dilution thermal ionization mass spectrometry has updated the maximum deposition age of the Jacobsville Formation to be ca.
+990 Ma \cite{Hodgin2022a}.
+Such age constraint is consistent with backbulge subsidence model \cite{DeCelles, 2012}, indicating that the Jacobsville Formation may have been deposited in a Grenvillian foreland basin system that resulted from lithospheric flexure induced by orogenic loading \cite{Rivers2012a}.
+The relatively young age of the Jacobsville Formation compared to other Midcontinent Rift rocks makes the Jacobsville Formation an crucial target for further reconstructing the global paleogeography in the late Proterozoic.
+For further refining the paleogeographic connection between the poles of the Grenville Loop and those of the Keweenawan Track.
+
+% \cite{Roy1978a} has done that but without the effort of distinguishing the detrital remanence and the chemical rmanence.
+% In addition, constraining the pole position of the
+
+% and shed new lights on reconciling the connection between the Keweenawan Track and the Grenville Loop paleomagnetic apparent polar wander paths.
+
+% Paleomagnetic data for the Jacobsville Formation have led to a paleomagnetic pole position that appears to be a continuation of the Keweenawan Track from the Oronto Group poles (Roy and Robertson, 1978; Fig.
 7; Table 4).
@@
 Nevertheless, the paleo- magnetic pole of the Jacobsville Sandstone (Roy and Robertson, 1978) appears to extend the Keweenawan Track from the Oronto Group poles in a manner that may be more consistent with deposition of the Jacobsville Formation in the late Mesoproterozoic.
-Given the conflict with this inferred age, the large uncertainty of individual LA-ICP-MS dates, and the pos- sibility of outlier dates in such a large data set, it is beneficial to interpret the age of the youngest Jacobsville zircons dated CA-ID-TIMS to obtain more precise constraints.
+Given the conflict with this inferred age, the large uncertainty of individual LA-ICP-MS dates, and the possibility of outlier dates in such a large data set, it is beneficial to interpret the age of the youngest Jacobsville zircons dated CA-ID-TIMS to obtain more precise constraints.
 Using CA-ID-TIMS, a recent study by Hodgin et al (2022) redefines the maximum deposition age of the Jacobsville Formation as ca.
@@
 
-The extensive paleomagnetic studies of the well-preserved late Mesoproterozoic to Neoproterozoic rocks of North America have provided the central record for global paleogeography reconstruction during the assembly and rearrangement of the postulated supercontinent Rodinia.
-Paleomagnetic studies of rocks from Laurentia have led to a series of paleomagnetic poles that form an apparent polar wander path (APWP) known as the "Logan Loop" for the older high latitude poles at its apex that continues into the ``Keweenawan Track" of younger lower latitude poles that form a progression as the APWP heads toward the ``Grenville Loop" \citep{Swanson-Hysell2019a}.
+% The extensive paleomagnetic studies of the well-preserved late Mesoproterozoic to Neoproterozoic rocks of North America have provided the central record for global paleogeography reconstruction during the assembly and rearrangement of the postulated supercontinent Rodinia.
+Paleomagnetic studies of rocks from Laurentia have led to a series of paleomagnetic poles that form an apparent polar wander path (APWP) known as the ``Logan Loop" for the older high latitude poles at its apex that continues into the ``Keweenawan Track" of younger lower latitude poles that form a progression as the APWP heads toward the ``Grenville Loop" \citep{Swanson-Hysell2019a}.
 In particular, recent advances in pairing high-precision zircon U-Pb geochronology with high-quality paleomagnetic poles significantly improved the resolution of the Keweenawan Track and revealed Laurentia's rapid motion ($>$20 cm/yr) during the Midcontinent Rift magmatic activity from ca.
@@
 
-As the Midcontinent Rift magmatic activity wanes its strength and evolved into a failed intracontinental rift, Laurentia experienced a long period of magmatic quiescence in its interior, where basin subsidence dominated the rift basin.
+% As the Midcontinent Rift magmatic activity wanes its strength and evolved into a failed intracontinental rift, Laurentia experienced a long period of magmatic quiescence in its interior, where basin subsidence dominated the rift basin.
 This was followed by the Grenville Orogeny, causing the the inversion of the rift along thrust belts such as the Keweenaw Fault (Fig.
@@
 
-Due to the lack of pmag records from the interior Laurentia, various hypotheses for the spatial relationship between the end of the Keweenawan Track and the Grenville Loop have been debated.
+% Due to the lack of pmag records from the interior Laurentia, various hypotheses for the spatial relationship between the end of the Keweenawan Track and the Grenville Loop have been debated.
 One model argues for that the difference in pole locations indicates that the Grenville Province was once thousands of kilometers away from Interior Laurentia and was progressively approaching the southern margin of Laurentia in the late Mesoproterozoic \citep{Halls2015}.
@@
 
-Paleomagnetic poles from the interior of Laurentia during the quiescence magmatism period is thus crucial for reconcile the discordant paleomagnetic poles of the Keweenawan Track and the Grenville Loop.
+% Paleomagnetic poles from the interior of Laurentia during the quiescence magmatism period is thus crucial for reconcile the discordant paleomagnetic poles of the Keweenawan Track and the Grenville Loop.
 The
-%Q: what are you using the poles from rocks that we know are from the Laurentia interior? (outside of the Grenville package?)
+% %Q: what are you using the poles from rocks that we know are from the Laurentia interior? (outside of the Grenville package?)
 
-This new age constraint on the Jacobsville provides crucial insight into the tectonic evolution of the Midcontinent Rift and the inversion of the extensional region into a compressional region due to the Grenville orogen.
+% This new age constraint on the Jacobsville provides crucial insight into the tectonic evolution of the Midcontinent Rift and the inversion of the extensional region into a compressional region due to the Grenville orogen.
 It also makes the Jacobsville sandstone a unique opportunity to investigate Laurentia's paleogeographical position at ca.
@@
 
-Jacobsville sedimentation was preceded by a long period of volcanic and tectonic quiescence and cratonic stability so that bedrock surfaces became blanketed by paleosol and a surface of chemically resistant debris was dominated by quartz and iron-formation.
+% Jacobsville sedimentation was preceded by a long period of volcanic and tectonic quiescence and cratonic stability so that bedrock surfaces became blanketed by paleosol and a surface of chemically resistant debris was dominated by quartz and iron-formation.
 
-Previous paleomagnetic study on the red beds within the Jacobsville by \cite{Roy1978a} observed different components using AF, thermal, and chemical demagnetizations.
+% Previous paleomagnetic study on the red beds within the Jacobsville by \cite{Roy1978a} observed different components using AF, thermal, and chemical demagnetizations.
 Recognized dual polarities and recognized the complex remanence acquisition of red beds which lead to three isolated remanence components from the Jacobsville.
@@
 
-In this study, we use fine-grained to silt-sized, hematite rich red beds from xxx sections along the Keweenaw Peninsula, Michigan with the goal of obtaining high-quality paleomagnetic pole from the jacobsville formation that is corrected for inclination shallowing problem and to investigate the connection between the Keweenawan Track and the Grenville Loop of the Laurentia APWP during the late Mesoproterozoic to the Neoproterozoic.
+% In this study, we use fine-grained to silt-sized, hematite rich red beds from xxx sections along the Keweenaw Peninsula, Michigan with the goal of obtaining high-quality paleomagnetic pole from the jacobsville formation that is corrected for inclination shallowing problem and to investigate the connection between the Keweenawan Track and the Grenville Loop of the Laurentia APWP during the late Mesoproterozoic to the Neoproterozoic.
 
-\section*{GEOLOGIC SETTINGS}
+\section*{Geologic settings}
 
-The Jacobsville sandstone is 
-\cite{Kalliokoski1982a} In the Lake Superior syncline the Jacobsville Sandstone is a thick (+900m) fluvial sequence of feldspathic and quartzose sandstones, conglomerates, siltstones, and shales, completely devoid of lava flows or cross-cutting dikes.
-On the north and south sides of Lake Superior, most of the sandstone occurs as inward-dipping, fault-bounded wedges, separated by regional faults from the Oronto and Bayfield Groups of similar red sandstones, situated on the inner side of the syncline.
+\subsection*{Structure}
+The ca.
+1109 Ma to 1084 Ma North American Midcontinent Rift is a major intracontinental rift system that extends over 2000 km through Laurentia craton (Figure \ref{fig:Geologic_map}).
+However, the rifting failed to separate Laurentia into two continents.
+As its magmatism waned and thermal subsidence progressed in the rift basin, the $>$4 km of ca.
+1080 Ma to 1050 Ma Oronto Group sediments \cite{Swanson-Hysell2019a, Fuentes2022a} were deposited conformably on top of the rift volcanics (Fig.
+\ref{fig:Geologic_map}; \cite{Cannon1992a}).
+Subsequently, the rift was inverted as the Grenvillian Orogeny commenced on the eastern margin of Laurentia (Figure \ref{fig:Geologic_map}; \citeA{Cannon1994a}), resulting in the propagation of far-field compressional stress deep into the interior of Laurentia.
+During the inversion, Paleoproterozoic and Archean crusts were uplifted via thrust faults such as the Marenisco fault and the Pelton Creek fault, forming the crustal-scale Montreal River monocline (Fig.
+\ref{fig:Geologic_map}; \cite{Cannon1993a}).
+The post-rift Jacobsville Formation overlies an erosional angular unconformity that developed on lithologies that were exhumed through this earlier reverse motion (Fig.
+\ref{fig:Geologic_map}; \citeA{Cannon1993a, Kalliokoski1982a}).
+The erosional contact and the development of paleosol between the Jacobsville and the underlying basement rocks is consistent with the interpretation that there was a prolonged period of sedimentation quiescence and net erosion between the formation of the Oronto Group sediments and the Jacobsville Formation.
+The overall extent of the Jacobsville Formation has been mapped along the lake shore of the Keweenaw Peninsula and is interpreted to extend eastward to Sault Ste Marie, Canada (Fig.
+\ref{fig:Geologic_map}; \citeA{Hamblin1958a}).
 
-With its age being long debated through the past decades, recently detrital zircon work by \cite{Malone2016a} and \Hodgin{2021a} settled that the maximum deposition age of the Jacobsville sandstone  is 980 Ma, indicating the deposition of the Jacobsville predate or the upper Jacobsville overlaps with the Rigolet phase of the Grenville orogeny.
-The upper age of the Jacobsville is xxx Ma.
+% The Jacobsville Sandstone is unconformably overlain by the Cambrian Munising Formation (Hamblin, 1958).
+
+As the far-field stress of the Grenvillian Orogeny continued to reach the interior of Laurentia, it thrusted and folded the Midcontinent Rift volcanics over the Jacobsville strata along the major Keweenaw fault which extends $\sim$250 km from the tip of the Keweenaw Peninsula in northern Michigan southeastward to a termination in northeastern Wisconsin (Fig.
+\ref{fig:Geologic_map}; \citeA{Cannon1996a, Cannon2001a, DeGraff2022a}).
+As a result, rift volcanic clasts where eroded into the Jacobsville basin at the same time when sediments from the older basement rocks along with distal input from the Grenville Province were deposited within the Jacobsville back-bulge basin system that in general wedges toward the northwest (Figure \ref{fig:Geologic_map}; \cite{Hedgman2014a, Hodgin2022a}.
+Immediately near the Keweenaw fault in the footwall, folding has also been recorded by strata of the upper Jacobsville Formation in contact with the volcanics \cite{Cannon1993a}.
+One famous remanence of that late-stage deformation is the Natural Wall ravine where sub-vertical to overturned sandstone beds are preserved (Figure \ref{fig:Geologic_map}; \cite{Hamblin1958a}).
+On shore of Bete Grise Bay near the northeastern edge of the Keweenaw Peninsula, the Keweenaw fault is interpreted to have accommodated compressional stresses by forming a series of segmented en echelon thrusts \cite{Tyrrell2019a}.
+The timing of such en echelon faulting could be coeval with the deposition of the upper Jacobsville Formation, leading to the nonconformable onlapping of the Jacobsville sediments over the volcanics in addition to the thrust faulting relationship between the two units.
+Late-kinematic calcite that precipitated within the brecciated Keweenaw fault zone and youngest detrital zircons of the Jacobsville Formation constrain the timing of the final inversion of the Keweenaw Fault to be during the Rigolet phase of the Grenvillian Orogeny \cite{Hodgin2022a}.
+
+\subsection{General stratigraphy}
+
+% It has been interpreted that the Jacobsville Sandstone is equivalent to the Bayfield Group in northern Wisconsin, and the Fond Du Lac and Hinckley formations in Minnesota.
+Their shared lithofacies of basal conglomerate, cross-bedded fluvial fine to medium grained sandstone, and interbedded red micaceous siltstone could support the interpretation that these sedimentary rocks formed during a common period of post-rift basin formation following rift-related magmatism of the Midcontinent Rift.
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=0.9\textwidth]{Field_photo.png}
+\caption{\scriptsize Field photos of red beds of the Jacobsville Formation sampled in this study.
+At Baby Snake Creek and the Natural Wall ravine in northern Keweenaw Peninsula, intervals of red fine-grained sandstone beds can be found through the steeply dipping to overturned basal strata near the Keweenaw Fault (A,C) and also the nearly horizontal top strata (B,D).
+At Hungarian Falls, relatively thick intervals of fine-grained red sandstone to siltstone beds separated by hematite-depleted arenite to arkosic sandstones or conglomerates are found along a $\sim$100-meter-tall waterfall (E).
+Abundant reoriented siltstone rip-up clasts often exist within the conglomeratic layers above siltstone beds (F).
+Drill cores or block samples of these rip-up clasts are sampled for the paleomagnetic conglomerate test presented in Fig.
+\ref{fig:Intraclast_pmag}.
+At Agate Falls, nearly flat-lying red fissile shale beds are exposed due to waterfall incision (G).
+Subrounded detrital mica grains deposited parallel to the bedding plane are often associated with the siltstone facies.
+Another set of siltstone rip-up clasts were sampled from a conglomeratic layer near the basal exposure for the paleomagnetic conglomerate test.
+}
+\label{fig:Field_photo}
+\end{figure*}
+
+The Jacobsville Formation is a fluvial sequence of feldspathic and quartzose sandstones, conglomerates, siltstones, and shales, completely devoid of lava flows or cross-cutting dikes.
+It is in general more quartz-rich than the Oronto Group sediments.
+Abundant near the bottom section in the southern exposures of the Jacobsville Formation, conglomerate facies typically have provenance sourced from Paleoproterozoic lithologies such as vein quartz and iron formation clasts \cite{Hamblin1958a, Kalliokoski1982a}.
+The occurrence of these conglomerates in proximity to the Keweenaw, Marenisco, and Pelton Creek thrust faults is interpreted to represent syn-depositional development of local relief \cite{Kalliokoski1982a, Hedgman1992a}).
+
+Near the Keweenaw Fault (Fig.
+\ref{fig:Geologic_map}), the Jacobsville Formation consists dominantly of quartz-rich, cross-bedded, medium-grained sandstone, locally abundant clast-supported conglomerate, with interbeds of hematite-bearing micaceous siltstone (Fig.
+\ref{fig:strat_column}).
+In this region, the conglomerate contains abundant volcanic clasts likely derived from uplifted Midcontinent Rift volcanics \cite{Irving1885a, Brojanigo1984a}.
+Occurrence of red siltstone beds have been reported in the Jacobsville Formation, particularly in the upper sections near Keweenaw Fault \cite<e.g.>{Hamblin1958a, Irving1885a}.
+In the field, they typically appear brick-red to dark-red, consisting of find-grained arenite to arkosic sandstone and siltstone sometimes with minor grey reduction spots (Fig.
+\ref{fig:Field_photo}).
+Rounded detrital mica grains up to $\sim$5mm in size can be found commonly within very fine, fissile siltstone and shaly strata (Figure \ref{fig:Field_photo}).
+The fact that in the Jacobsville has conglomerate and sandstone and siltstone facies in them supports sediment input from the eroded Oronto Group.
+This indicate the older boundary of the Jacobsville deposition might be closer to the Rigolet Phase than the Ottawan phase and the duration of Jacobsville sedimentation may be within a relatively constrained period of time.
+
+tyrrell 2019 fig. 10 recorded onlapping of the Jacobsville Formation on to the Portage Lake Volcanics which appears to have been faulted prior to the deposition of Jacobsville near Bete Grise Bay.
+
+% Hematite is one of the most stable form of iron oxides under ambient conditions, and magnetic remanence carried by the grains within red beds can be preserved for billions of years \cite{Butler1992a, Jiang2022a}.
+Previously, paleomagnetic records of the Jacobsville Formation have been investigated by \citeA{Roy1978a} who sampled a range of lithofacies of the Jacobsville Formation and obtained that a primary remanence component can be isolated via magnetic cleaning.
+In this study, we target the find-grained sandstone to siltstone facies in the upper Jacobsville Formation and segregate the primary and secondary remanent magnetizations using high-resolution thermal demagnetization method \cite{Swanson-Hysell2019b}.
+We update the paleomagnetic records of the Jacobsville Formation with additional evaluation for the degree inclination shallowing that \citeA{Roy1978a} did not take into account.
+
+% Along the south shore of Lake Superior, the Keweenawan Supergroup overlies Paleoproterozoic basement and consists of ~5 km of rift-related volcanic rocks that are conformably overlain by ~4 km of conglomerate, siltstone, and sandstone of the Oronto Group (Cannon et al., 1996).
+The uppermost unit of the Keweenawan Supergroup is the Jacobsville Sandstone, which consists of up to ~1 km of fluvial deposits that are more quartz-rich than the Oronto Group (Fig.
+1; Kalliokoski, 1982).
+The Jacobsville Sandstone unconformably overlies Archean and Paleoproterozoic basement as well as Midcontinent Rift volcanic and sedimentary rocks in northern Michigan (Fig.
+1; Cannon et al., 1993; Ojakangas et al., 2001).
+
+% The Keweenaw Fault is a north to northwest-dipping thrust that  Thrust faults parallel to the Keweenaw Fault include the northwest-dipping Marenisco and Pelton Creek faults (Fig.
+1).
+Similarly-oriented thrusts continue southwestward through northern Wisconsin and eastern Minnesota (Ojakangas et al., 2001), and eastward below Lake Superior, where they are interpreted to join mapped thrust faults in southern Ontario (Fig.
+1; Manson and Halls, 1994; Ojakangas et al., 2001).
+Rock units adjacent to these regional thrusts have been moderately to intensely folded.
+
+% In northern Michigan, the typically shallowly-dipping Jacobsville Sandstone has sub-vertical to overturned beds in the immediate footwall of the Keweenaw Fault (Irving and Chamberlin, 1885; Cannon and Nicholson, 2001).
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=0.9\textwidth]{strat_column.png}
+\caption{}
+\label{fig:strat_column}
+\end{figure*}
 
@@
 
-\subsection{Jacobsville stratigraphy}
+\section*{Paleomagnetic methods and results}
 
-Along the south shore of Lake Superior, the Keweenawan Supergroup overlies Paleoproterozoic basement and consists of ~5 km of rift-related volcanic rocks that are conformably overlain by ~4 km of conglomerate, siltstone, and sandstone of the Oronto Group (Cannon et al., 1996).
-The uppermost unit of the Keweenawan Supergroup is the Jacobsville Sandstone, which consists of up to ~1 km of fluvial deposits that are more quartz-rich than the Oronto Group (Fig.
-1; Kalliokoski, 1982).
-The Jacobsville Sandstone unconformably overlies Archean and Paleoproterozoic basement as well as Midcontinent Rift volcanic and sedimentary rocks in northern Michigan (Fig.
-1; Cannon et al., 1993; Ojakangas et al., 2001).
-The overall extent of the Jacobsville Sandstone has been mapped along the southern shore of Lake Superior, from west to east - the Keweenaw Peninsula of northern Michigan to Sault Ste Marie, Canada (Fig.
-\ref{}).
-The Jacobsville Sandstone is unconformably overlain by the Cambrian Munising Formation (Hamblin, 1958).
+Oriented paleomagnetic samples were collected with a portable electric drill along five stratigraphic sections of the Jacobsville Formation in norther Michigan (Fig.
+\ref{fig:Geologic_map}).
+Additional oriented cores and block samples of siltstone rip-up clasts within conglomerate facies were sampled at Hungarian Falls and at Agate Falls.
+In order to maximize sampling of paleosecular variation, we optimized for vertical stratigraphic coverage and collected one sample at each red bed horizon.
+As such, each sample constitutes a paleomagnetic site considering that a paleomagnetic site (which ideally captures a single snapshot of the local geomagnetic field) is a particular bed in a sedimentary sequence.
+Red fine-grained sandstone to shale layers were preferentially sampled as they have lower permeability and are less susceptible to diagenetic alteration through fluid flow than coarser grained sandstone.
+Care was taken to avoid samples containing reduction spots.
+Paleomagnetic samples were oriented using a magnetic compass and a sun compass whenever possible.
+Sun compass data were preferentially used when available.
+A total of 379 specimens including 30 intraclasts where collected for paleomagnetic study.
+The specimens underwent stepwise thermal demagnetization in the UC Berkeley Paleomagnetism Lab using an ASC demagnetizer (residual fields $<$10 nT) with measurements made on a 2G DC-SQUID magnetometer.
+The demagnetization protocol had high-resolution approaching the Ne\'el temperature of hematite (5 \textdegree C to 2 \textdegreeC) (Fig.
+\ref{fig:Intraclast_pmag}).
+All paleomagnetic data are available to the measurement level in the MagIC database (https://earthref.org/ MagIC/; THIS LINK IS AVAILABLE TO REVIEWERS AND WILL BE UPDATED WHEN A DOI IS AVAILABLE).
 
-The Keweenaw Fault is a north to northwest-dipping thrust that extends ~250 km from the tip of the Keweenaw Peninsula in northern Michigan southeastward to a termination in northeastern Wisconsin (Fig.
-1; Cannon et al., 1996; Cannon and Nicholson, 2001).
-Thrust faults parallel to the Keweenaw Fault include the northwest-dipping Marenisco and Pelton Creek faults (Fig.
-1).
-Similarly-oriented thrusts continue southwestward through northern Wisconsin and eastern Minnesota (Ojakangas et al., 2001), and eastward below Lake Superior, where they are interpreted to join mapped thrust faults in southern Ontario (Fig.
-1; Manson and Halls, 1994; Ojakangas et al., 2001).
-Rock units adjacent to these regional thrusts have been moderately to intensely folded.
-In northern Michigan, the typically shallowly-dipping Jacobsville Sandstone has sub-vertical to overturned beds in the immediate footwall of the Keweenaw Fault (Irving and Chamberlin, 1885; Cannon and Nicholson, 2001).
-Near the Keweenaw Fault, Jacobsville Sandstone consists of quartz-rich, cross-bedded, medium-grained sandstone, locally abundant clast-supported conglomerate, and minor interbeds of red micaceous siltstone.
-The conglomerate contains volcanic clasts likely derived from uplifted Midcontinent Rift volcanics (Irving and Chamberlin, 1885; Brojanigo, 1984).
-Clast provenance in Jacobsville conglomerates includes Paleoproterozoic lithologies, such as vein quartz and iron formation clasts, which are abundant in southern exposures (Hamblin, 1958; Kalliokoski, 1982).
-Occurrence of conglomerates in proximity to the Keweenaw, Marenisco, and Pelton Creek thrusts is interpreted to represent syn-depositional development of local relief (Kalliokoski, 1982; Hedgman, 1992).
+\subsection*{Fluvial intraclast conglomerate test}
 
-It has been interpreted that the Jacobsville Sandstone is equivalent to the Bayfield Group in northern Wisconsin, and the Fond Du Lac and Hinckley formations in Minnesota.
-Their shared lithofacies of basal conglomerate, cross-bedded fluvial fine to medium grained sandstone, and interbedded red micaceous siltstone could  support the interpretation that these sedimentary rocks formed during a common period of post-rift basin formation following rift-related magmatism of the MCR.
+Similar to the thermal demagnetization result of the hematite-bearing fluvial intraclasts of the Freda Formation \cite{Swanson-Hysell2019b}, the siltstone rip-up clasts of the Jacobsville Formation also typically reveal two distinct magnetization directions (Fig.
+\ref{fig:Intraclast_pmag}).
+One direction was similar throughout the intraclasts and was typically removed up to 640-655 \textdegree C (Fig.
+\ref{fig:Intraclast_pmag}).
+This component is directionally well-grouped amongst the reoriented intraclasts, indicating that it is likely a chemical remanent magnetization acquired through the growth of secondary hematite grains following redeposition of the clasts (Fig.
+\ref{fig:Intraclast_pmag}).
+After the removal of the mid-temperature component, further thermal demagnetization with small step increments at higher temperatures up to $\sim$689 \textdegree C often reveal an origin-trending component (Fig.
+\ref{fig:Intraclast_pmag}).
+In the data from the clasts, there is typically a significant directional change in the specimen magnetization between the mid-temperature component and the high-temperature component and as a result, 28 out of 30 intraclast specimens could be fit with distinct mid-temperature and high-temperature least squares lines \cite{Kirschvink1980a}.
+Distinct from the well-grouped mid-temperature component directions, the high-temperature directions of the intraclasts are dispersed and the null hypothesis of randomness cannot be rejected at the 95\% confidence level (n=6 for intraclasts at Agate Falls and n=22 for intraclasts at Hungarian Falls; Fig.
+\ref{fig:Intraclast_pmag}; \citeA{Watson1956a}).
+This indicates that the component was likely acquired by the detrital hematite grains during initial formation of the clasts prior to erosion and redeposition.
 
-Red beds, particular fine-grained siltstone beds have been shown to be good paleomagnetic recorders, preserving magnetic remanence for up to billions of years.
-Occurrence of red siltstone beds have been reported in the Jacobsville Sandstone, particularly in the upper sections near Keweenaw Fault (e.g.
-Hamblin1958a, Irving1885a).
-They typically in the field are brick-red to dark-red, find-grained arenite to arkosic sandstone and siltstone.
-Minor grey reduction spots can be found.
-rounded detrital mica grains up to 5mm?
-size can be found commonly within very fine, fissile siltstone and shaly strata.
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{Intraclast_pmag.pdf}
+\caption{Paleomagnetic data from intraclasts of the Jacobsville Formation reveal a mid-temperature component that typically unblocks prior to 655 \textdegree C and a high-temperature component that typically unblocks up to 687 \textdegree C.
+These components are present as varying fractions of the overall remanence as seen in the three individual clasts shown here on vector orthogonal plots and measurement-level equal area plots in tilt-corrected coordinates (developed using PmagPy; \citeA{Tauxe2016a}).
+The direction of the mid-temperature component is shown as green arrows on the vector component plots and green circles on the equal area plots, while the high-temperature component is shown with purple symbols.
+The mid-temperature component has a similar direction among the clasts as can be seen on the component directions equal area plots.
+In contrast, the high-temperature component directions (purple) are dispersed.
+NRM = natural remanent magnetization; AFC = Agate Falls clasts; HFC = Hungarian Falls clasts.}
+\label{fig:Intraclast_pmag}
+\end{figure*}
 
-\subsubsection{Paleomagnetic sample sites}
-Agate Falls
-A MDA of --- Ma was obtained from this site by Hodgin2021a.
+\subsection*{Tentative magnetostratigraphy}
 
-Natural Wall
-A MDA of --- Ma was obtained from this site by Hodgin2021a
+With the insights gained from the thermal demagnetization results of the Jacobsville fluvial intraclasts, we next isolate the high-temperature magnetization component from the mid-temperature component for all in situ specimens that were collected along the five localities in northern Michigan (Fig.
+\ref{fig:Geologic_map}).
+The resultant high-temperature component directions are plotted by locality in Fig.
+\ref{fig:in_situ_pmag}.
+These results show dual polarities recorded by the Jacobsville Formation (Fig.
+\ref{fig:in_situ_pmag}), consistent with the observation of \citeA{Roy1978a}.
+Furthermore, our new data show that the Jacobsville Formation recorded more than one geomagnetic field polarity reversal and a tentative magnetostratigraphy may be constructed (Fig.
+\ref{fig:in_situ_pmag}).
 
-Snake Creek
-A MDA of --- Ma was obtained from this site by Hodgin2021a
+At Agate Falls (Fig.
+\ref{fig:Geologic_map}), multiple hematite-rich, detrital mica-bearing siltstone layers occur through a $\sim$30-meter strata incised by the running Middle Branch Ontonagon River.
+68 paleomagnetic samples were collected on both sides of the river through overlapping stratigraphic sections (Fig.
+\ref{fig:strat_column}).
+A single transition from normal to reversed polarity is observed through the sections from bottom to top with a few strata in the middle of the section capturing transitional field directions (Fig.
+\ref{fig:strat_column}, \ref{fig:in_situ_pmag}).
 
-JK creek
-A MDA of --- Ma was obtained from this site by Hodgin2021a
+At Hungarian Falls (Fig.
+\ref{fig:Geologic_map}), high-energy fluvial environment led to the deposition of numerous hematite-rich siltstone to fine-grained sandstone layers, capturing richer information of the geomagnetic field variations than at other localities (Fig.
+\ref{fig:strat_column}).
+Throughout $\sim$240 meters of stratigraphy, a total of 158 samples were collected for paleomagnetic study.
+At stratigraphic height of $\sim$24 meters (Fig.
+\ref{fig:strat_column}), 8 samples from a $\sim$0.5-meter-thick siltstone layer record a reversed polarity (easterly declination and shallow inclination; horizon HF6).
+At stratigraphic level $\sim$42 meters (HF5; Fig.
+\ref{fig:strat_column}), 4 samples were collected through a $\sim$0.2 m siltstone horizon including one specimen from near the base of the layer and three specimens near the top of the layer.
+The high-temperature remanence component from the three specimens near the top all record a reversed polarity whereas the specimen collected near the base show a westerly declination.
+While more samples are needed to test whether the westerly declination recorded by the single sample is a result of the geomagnetic field having completely flipped polarity from that recorded by horizon HF6 or a result of transient unstable field, this anomalous direction indicates that the geomagnetic field did not stay consistent throughout the time of deposition of this horizon.
+As a result, we tentatively mark a period of transitional field near the base of horizon HF5 in Fig.
+\ref{fig:strat_column}.
+A reversed polarity field seems to have been maintained as 5 samples from horizon HF4 consistently record easterly directions similar to those recorded by samples near the top of HF5 (Fig.
+\ref{fig:strat_column}).
+However, a geomagnetic field polarity reversal happened between the deposition of siltstone layers of HF4 and HF3, as 7 samples from HF3 all record normal-polarity directions, antipodal to those of samples of HF4.
+Subsequently, the field appears to flip again back to reversed polarity, as is shown by 7 samples from HF2 higher above the section (Fig.
+\ref{fig:strat_column}).
+Through the $\sim$35-meter-thick horizon of HF1 characterized by numerous siltstone to fine-grained sandstone layers interbedded with coarser-grained sandstone and conglomerate, magnetic polarity reversals of R-N-R-N-R were recorded (Fig.
+\ref{fig:strat_column}).
 
-\cite{Roy1978a} sampled 37 sites across a lateral distance of
+The Natural Wall ravine and Hammel Creek are of close proximity from Hungarian Falls (Fig.
+\ref{fig:Geologic_map}) and the sedimentary facies characterized by trough cross-bedded sandstone and conglomerates are similar to the basal $\sim$120 meter of the section at Hungarian Falls (Fig.
+\ref{fig:strat_column}).
+Since only a couple of siltstone to fine-grained sandstone horizons were sampled at these two localities, we tentatively correlate the stratigraphic sections at these three localities based on their sedimentary facies and magnetic polarities (Fig.
+\ref{fig:strat_column}).
+At Hammel Creek, 14 samples from two siltstone layers near the base of the section are of reversed polarities, consistent with the reversed-polarity layers near the base of Hungarian Falls.
+Near the top of the Natural Wall section, consistent normal-polarity directions are recorded by 15 samples.
+This normal polarity could be coeval with one of the normal-polarity periods recorded at Hungarian Falls, or it could indicate an additional polarity reversal from reversed to normal polarity that happened after the deposition of the youngest reversed strata at Hungarian Falls (Fig.
+\ref{fig:strat_column}).
+Near the base of the stratigraphic section at Natural Wall (Fig.
+\ref{fig:strat_column}), the strata are steeply tilted or overturned (Fig.
+\ref{fig:Field_photo}).
+Tilt-corrected paleomagnetic directions from samples collected from these strata show dominantly southeasterly declinations that are distinct from either normal- or reversed-polarity directions higher along the strata or at Hungarian Falls.
+Possible causes of such distinct directions could be related to transitional fields associated with geomagnetic field polarity reversals or complications in the tilt correction such as noncylindrical folding or multiple tilting episodes.
 
-\section*{METHODS and RESULTS}
+A total of 70 specimens yielded interpretable high-temperature paleomagnetic remanence component throughout the stratigraphic section at Baby Snake Creek (Fig.
+\ref{fig:Geologic_map}, \ref{fig:in_situ_pmag}).
+The remanence directions recorded by red beds within the steeply folded basal section and within the flat-lying top portion of the section pass a bootstrap fold test (Supporting Information; \citeA{Tauxe1994a}).
+The remanence directions are all of normal polarity and show a dominantly southwest declination and upward-pointing inclination (Fig.
+\ref{fig:in_situ_pmag}).
 
-\section*{DISCUSSION}
+Given that multiple geomagnetic field reversals occurred during the deposition of the Jacobsville Formation, it is desirable that a magnetostratigraphy be developed.
+However, challenges exist in correlating the stratigraphic sections measured at the five localities in this study due to the lateral variability of the paleodepositional environment associated with the variable paleotopographic relief of the Precambrian surface upon which the Jacobsville Formation was deposited \cite{Hamblin1958a, Kalliokoski1982a}.
+In particular, there may not be a unique solution in reconstructing the relative stratigraphic positions of the strata at Baby Snake Creek and Agate Falls with respect to the other three sections in central Keweenaw Peninsula due to geographical separations (Fig.
+\ref{fig:Geologic_map}, \ref{fig:strat_column}).
+However, at both Baby Snake Creek and Agate Falls, the sedimentary strata are dominated by siltstone to cross-bedded sandstone and are lacking significant amount of conglomerate facies as seen near Hungarian Falls.
+This could be consistent with the interpretation that these strata were deposited as the Jacobsville basin of the Grenvillian foreland system resulted from lithospheric flexure induced by orogenic loading has been partially filled.
+Recent geologic mapping efforts near Baby Snake Creek have found evidence for the Jacobsville Formation onlapping the Midcontinent Rift volcanics that were previously faulted via en echelon segments\cite{Tyrrell2019a, Mueller2021a}.
+The unconformable contact between the sediments and the volcanics, instead of faulted contact typically observed near Hungarian Falls where the basal Jacobsville Formation is folded and faulted against the volcanics, indicate that the deposition of the Jacobsville Formation near the northern end of Keweenaw Peninsula likely post-date some episodes of motion along the Keweenaw Fault.
+Therefore, we interpret that the normal-polarity strata at Baby Snake Creek could be stratigraphically the youngest amongst the five localities, although a possibility that they instead are near the bottom of the stratigraphy and record the normal Keweenawan Superchron cannot be ruled out \cite{Driscoll2016b}.
+
+Overall, the multiple geomagnetic field polarity reversals recorded by the Jacobsville Formation constrain the maximum range of the Keweenawan normal superchron \cite{Driscoll2016b} to be between ca.
+1100 (Mamainse Point Volcanics; \cite{Swanson-Hysell2019a}) to ca.
+992 Ma (maximum deposition age of Jacobsville Formation; \cite{Hodgin2022a}).
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{in_situ_pmag.pdf}
+\caption{}
+\label{fig:in_situ_pmag}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{EI_results.pdf}
+\caption{}
+\label{fig:EI_results}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{Jacobsville_pole_plot.pdf}
+\caption{}
+\label{fig:pole_plot}
+\end{figure*}
+
+\section*{Discussion}
+\subsection*{Inclination shallowing correction and Jacobsville paleomagnetic pole}
+
+Although igneous and sedimentary rocks associated with the North America Midcontinent Rift provided high-resolution record of the ca.
+1110 to 1070 Ma Keweenawan Track which is the apparent polar wander path (APWP) for Laurentia in the Mesoproterozoic, high-quality paleogeographic constraints is lacking thereafter since tectonism in the interior of Laurentia entered a protracted period of quiescence until the emplacement of the ca.
+780 Ma Gunbarrel large igneous province \cite{Harlan2003a}.
+The new paleomagnetic data developed from the Jacobsville Formation presents an opportunity to extend the Keweenawan Track toward the early Neoproterozoic as its maximum deposition age has been recently tightly constrained to be ca.
+992 Ma \cite{Hodgin2022a}.
+
+However, inclination bias associated with sedimentary rocks need to be corrected when synthesizing a paleomagnetic pole.
+Rotation of ferromagnetic grains during sediment deposition and compaction during early diagenesis can result in the acquisition of a detrital remanent magnetization that is biased shallow relative to the local geomagnetic field in which it was acquired \cite{King1955a, Tauxe2005a, Kodama2012a}.
+This effect can be summarized as:
+\begin{equation*}
+\textup{tan}(I_o) = f\textup{tan}(I_f)
+\end{equation*}
+where $f$ represents the amount of inclination shallowing, $I_o$ represents the observed inclination of the specimen magnetization and $I_f$ represents the inclination of the field in which the magnetization was acquired.
+If uncorrected, shallower inclinations obtained from sedimentary rocks can result in erroneously low estimates of paleolatitudes, biasing the interpreted past positions of continents and hindering plate reconstructions \cite{Domeier2012a}.
+Methods for correcting sedimentary inclination bias have been developed.
+They include laboratory-based measurements of magnetic mineral fabrics \cite<e.g.>{Kodama1992a, Bilardello2015a}, as well as statistical methods based on the evaluation of the deviation of virtual geomagnetic pole (VGP) distribution patterns recorded by sedimentary rocks with an assumed paleosecular variation model (the ``elongation/inclination" [E/I] method of \citeA{Tauxe2004b}).
+When a large number of sedimentary paleomagnetic samples (typically n$>$100) capture an extensive period of time such that the paleosecular variation is sufficiently averaged, the E/I method can give accurate estimates and is more efficient than lab-based methods.
+In the ca.
+1.1 Ga Midcontinent Rift, \citeA{Pierce2022a} showed that the E/I method is effective in correcting the inclination bias recorded by the Cut Face Creek Sandstone---a lava flow-bracketed fluvial red bed sedimentary formation dominated by overbank deposition.
+
+Inclination shallowing is also present in the Jacobsville Formation.
+Scatter of the specimen high-temperature component directions in geographic coordinates show dispersion of declinations parallel to the bedding plane, indicating the directions are likely shallowed along the bedding-vertical plane (Fig.
+\ref{fig:in_situ_pmag}; \citeA{Tauxe2004b}).
+Since the deposition of the Jacobsville Formation span over geomagnetic field reversals, it is of interest to investigate the amount of inclination shallowing associated with the normal (n=130) and the reversed (n=148) polarity respectively.
+Applying the E/I method on the directions of normal polarity yields an estimated $f$ value of 0.67 with 95\% bootstrap uncertainty bounds from 0.91 to 0.55 (Fig.
+\ref{fig:in_situ_pmag}B).
+On the other hand, directions of reversed polarity yields an estimated $f$ factor of 0.52 with more uncertainties ranging from 0.97 to 0.29 (Fig.
+\ref{fig:in_situ_pmag}B).
+In addition, the dual-polarity directions do not pass a reversal test \cite{McFadden1990a, Tauxe2010a}, regardless of whether inclination correction is applied (Fig.
+\ref{fig:in_situ_pmag}).
+
+One interpretation of the polarity discrepancy could be that the deposition of the Jacobsville Formation was protracted such that the sediments at Hungarian Falls which dominantly record a reversed polarity was deposited prior to those at Baby Snake Creek which dominate the normal polarity records (Fig.
+\ref{fig:in_situ_pmag}).
+This is consistent with the magnetostratigraphic interpretation that Baby Snake Creek section is younger than the Hungarian Falls section (Fig.
+\ref{fig:strat_column}).
+Field observations of \citeA{Tyrrell2019a} and \citeA{Mueller2021a} near Bete Grise Bay, which is only $\sim$10 km east to Baby Snake Creek (Fig.
+\ref{fig:Geologic_map}), show that significant thrusting and deformation along the Keweenaw Fault and subsequent weathering of the Midcontinent Rift basalt occurred prior to the deposition of the Jacobsville Formation.
+In addition, U-Pb calcite geochronology and thermometry confirm that the final motion of $\sim$2 km of vertical displacement along the Keweenaw Fault occurred during the ca.
+980 Ma Rigolet phase of the Grenvillian Orogeny and the majority of the Keweenaw Fault motion likely occurred much earlier during the ca.
+1050 Ma Ottawan phase orogeny \cite{Hodgin2022b}.
+Therefore, we interpret that it is most likely that the Jacobsville Formation were deposited just prior to and during the Rigolet phase of the Grenvillian Orogeny with progressive sediment transportation to the northwest resulting in its onlapping on top of the volcanics in the northern tip of the Keweenaw Peninsula.
+As a result, the sediments at Baby Snake Creek records a slightly more southerly paleolatitude than those near Hungarian Falls as Laurentia's plate motion continues toward the south (Fig.
+\ref{fig:in_situ_pmag}, \ref{fig:pole_plot}).
+
+An alternative interpretation to the polarity discrepancy could be that there are insufficient samples such that either or both polarities are not average for secular variation.
+
+Overall, both inclination-corrected polarities recorded by the Jacobsville Formation show mean pole positions in the southern hemisphere with respect to today's continental configuration (normal polarity longitude=, latitude=, A95=; reversed polarity longitude= , latitude= , A95=).
+The two pole positions are close to each other, but are distinct from the youngest poles developed from rocks of the Midcontinent Rift (Fig.
+\ref{fig:pole_plot}).
+In addition, albeit that the Jacobsville sedimentation span over geomagnetic field reversals, the large arc distance between the poles of the Jacobsville Formation and that of the Freda Formation precludes the hypothesis that the Jacobsville Formation at the five studied localities (Fig.
+\ref{fig:Geologic_map}) is coeval with the Oronto Group and formed through a prolonged period of time.
+
+To succinctly summarize a paleomagnetic pole for the ca.
+992 Ma Jacobsville Formation, we combine the directions of both polarities and reassess the amount of inclination shallowing of the combined directional dataset (Fig.
+\ref{fig:in_situ_pmag}).
+The resultant estimate for the $f$ value is 0.61, and the inclination-corrected mean direction is dec= , inc= , $\alpha_{95}$= .
+The bootstrap 95\% uncertainty range for the $f$ value range from 0.80 to 0.51, which overlaps with the uncertainty ranges of directions associated with each polarity (Fig.
+\ref{fig:in_situ_pmag}).
+The Fisher mean paleomagnetic pole associated with the combined directions is shown in Fig.
+\ref{fig:pole_plot} (pole longitude= 186.3\textdegree E, pole latitude= 14.3\textdegree S, A95= 2.5\textdegree).
+
+\citeA{Pierce2022a} discussed that as sedimentary grains usually flatten perpendicular to the bedding plane but circularly symmetric about the bedding plane, the resultant shallow inclinations they record would typically result in asymmetric uncertainties in the recorded paleomagnetic pole position with more uncertainties in paleolatitues than paleolongitudes.
+The asymmetric uncertainties associated with sedimentary paleomagnetic poles violate the assumption of a Fisher distribution \cite{Fisher1953a} where uncertainties associated with a mean pole position is circularly distributed.
+Instead, the spherical asymmetric bivariate Kent distribution \cite{Kent1983a} is more appropriate for incorporating uncertainties associated with inclination shallowing in reported sedimentary paleomagnetic poles.
+We follow the approach of \citeA{Pierce2022a} who extended the E/I method of \citeA{Tauxe2004b}.
+First we take all estimated $f$ values from the bootstrap results of the E/I algorithm and apply each value to correcting all Jacobsville detrital remanence directions.
+Then we calculate the Fisher statistics for each of the inclination-corrected mean poles.
+Finally we obtain 100 Monte Carlo resamples of each Fisher mean pole.
+As a result, the distribution of the resampled mean pole positions have an elongated distribution which represents more uncertainty associated with paleolatitudes of Laurentia than paleolongitudes (Fig.
+\ref{fig:pole_plot}).
+The distribution of paleolatitudes of the resampled mean poles can be well-approximated by a normal distribution (Fig.
+\ref{fig:pole_plot}).
+We use the Kent distribution to parameterize the 95\% confidence ellipse of the distribution of the resampled mean poles (mean longitude=186.4\textdegree E, mean latitude=14.2\textdegree S, major axis longitude=262.5, major axis latitude=43.5, major axis magnitude=3.3\textdegree, minor axis longitude=110.0, minor axis latitude=43.0, minor axis magnitude=3.0\textdegree; Fig.
+\ref{fig:pole_plot}).
+As is shown in Fig.
+\ref{fig:pole_plot}, although the Kent ellipse uncertainty range of the Jacobsville Formation's paleomagnetic pole position has significant overlap with the circular uncertainty range represented by the Fisher mean, the Kent ellipse informs more uncertainty along the great circle path between the mean pole position and the locality of the Jacobsville Formation as it incorporates the uncertainties in the amount of inclination shallowing within the sediments.
+
+% \subsection*{Age of the Grenville Loop}
+
+\subsection*{Paleogeography of Laurentia during the late Mesoproterozoic to early Neoproterozoic}
+
+\citeA{McCabe1983a} developed paleomagnetic data from the Chequamegan Sandstone of the Bayfield Group and found reversed-polarity directions similar to that of the Jacobsville Formation poles developed by \citeA{Roy1978a}.
+
+Although igneous paleomagnetic records from the the interior of Laurentia has been lacking since the end of the Midcontinent Rift magmatism, data from metamorphosed rocks associated with the ca.
+1080 to 980 Ma Grenville continental collisional orogenesis has been developed to infer for Laurentia's plate configuration in the late Mesoproterozoic to early Neoproterozoic \cite<e.g.>{Weil1998a, Halls2015b}.
+However, in contrast to igneous rocks of the Midcontinent Rift which have relatively simple thermal history and straightforward paleomagnetic records \cite,e.g.>{Swanson-Hysell2019a}, rocks of the Grenville Province which experienced up to granulite facies metamorphism have much more complicated thermal history and the timing of their magnetic remanence acquisition is more difficult to constrain.
+As is shown in Fig.
+\ref{fig:pole_plot}, paleomagnetic poles developed from rocks of the Grenvillian Orogeny plots near Australia in today's continental reference frame, forming an arc distance of $\sim$30 degrees from the youngest paleomagnetic poles of the ca.
+1070 Ma Freda Formation \cite{Henry1977a}.
+Such great discrepancies between pole positions associated with uncertainties in the age of the magnetization of the Grenville rocks led to the ``Grenville Problem" which is still highly debated today - did the Grenville rocks acquire magnetic remanence prior to the continental collision and thus the pole discrepancies are a result of significant crustal shortening (4000 $\pm$ 1000 km; \citeA<e.g.>{Halls2015b}), or did they acquire remanence after the continental accretion and thus the southerly poles merely represent Laurentia's continued motion through more recent times \cite<e.g.>{Weil1998a}?
+
+As the ca.
+990 Ma Jacobsville Formation was deposited in Laurentia Interior (Fig.
+\ref{fig:Geologic_map}), its paleomagnetic pole position can help test the contrasting hypothesis regarding the ``Grenville Problem".
 
@@
 
+\section*{Conclusion}
+
 \acknowledgments

### 2022-07-17 Manuscript/Manuscript.tex

@@
 \usepackage{lineno}
+\usepackage{amsmath}
 \usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
@@
 \begin{keypoints}
-\item A new paleomagnetic pole from the ca.990 Ma Jacobsville Formation 
+\item A new paleomagnetic pole from the ca. 992 Ma Jacobsville Formation 
 \item Reversals put younger bounds on the normal Keweenawan superchron 
@@
 The ca.
-1.1 Ga North American Midcontinent Rift resulted in protracted magmatism and the associated deposition of sediments during rift-related extension and post-rift thermal subsidence.
-Coeval with the cessation of rifting and a decrease of the speed of plate motion is the onset of the Grenvillian Orogeny which lasted from ca.
-1090-980 Ma.
+1.1 Ga North American Midcontinent Rift resulted in protracted magmatism and associated deposition of sediments during rift-related extension and post-rift thermal subsidence.
+Coeval with the cessation of rifting and a decrease of Laurentia's plate speed is the onset of the ca.
+1090-980 Ma Grenvillian Orogeny along the eastern margin.
 The far-field stress of the orogeny is interpreted to have propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation whose maximum deposition age has been constrained to be ca.
-990 Ma.
-We develop a new paleomagnetic pole from red beds of the Jacobsville Formation along 5 stratigraphic sections in Northern Michigan, USA.
-High-resolution thermal demagnetization experiments resolve primary magnetic remanence that pass a fold test and two conglomerate tests.
-Our results confirm the occurrence of geomagnetic field reversals during the deposition of Jacobsville Formation, which provides constraints for the duration of the Keweenawan superchron.
-Inclination corrected paleomagnetic pole position from the Jacobsville Formation confirm continued plate motion of Laurentia across the equator in the Late Mesoproterozoic to early Neoproterozoic.
-The new pole from the interior of Laurentia indicates the paleomagnetic poles developed from the Grenville Province may record a coherent plate motion after major continental collision during the assemblage of supercontinent Rodinia rather than recording a significant crustal shortening.
+992 Ma.
+We develop a new paleomagnetic pole from red beds of the Jacobsville Formation along 5 stratigraphic sections in northern Michigan, USA.
+High-resolution thermal demagnetization experiments resolve detrital magnetic remanence that pass a fold test and two conglomerate tests.
+Our results confirm the occurrence of geomagnetic field reversals during the deposition of Jacobsville Formation, providing constraints for the duration of the Keweenawan superchron.
+Inclination corrected paleomagnetic pole position from the Jacobsville Formation confirm continued plate motion of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic.
+The new pole from the interior of Laurentia indicates the paleomagnetic poles developed from the Grenville Province most likely record coherent plate motion after major continental collision during the development of supercontinent Rodinia.
+% rather than recording a significant crustal shortening.
 \end{abstract}
@@
 \item Paleogeography
+\item apparent polar wander path
 \item Inclination shallowing
@@
 1109 Ma to 1084 Ma North American Midcontinent Rift (i.e.
-the Keweenawan Track) provides crucial constraints on global paleogeography in the Mesoproterozoic and has become the central record for reconstructing the development of ancient supercontinent Rodinia \cite{Swanson-Hysell2019a, Swanson-Hysell2021c}.
+the Keweenawan Track) provides crucial constraints on global paleogeography in the Mesoproterozoic and has become a central record for reconstructing the development of ancient supercontinent Rodinia \cite{Swanson-Hysell2019a, Swanson-Hysell2021c}.
 As the rifting ceased and subsequently inverted following the onset of the Grenvillian Orogeny on Laurentia's eastern margin (Fig.
@@
 
-\subsection*{Fluvial intraclast conglomerate test}
+\subsection*{Paleomagnetic field test}
+
+\subsubsection*{Fluvial intraclast conglomerate test}
 
@@
 \end{figure*}
+
+\subsubsection*{Fold test}
+
+A total of 70 specimens yielded interpretable high-temperature paleomagnetic remanence component throughout the stratigraphic section at Baby Snake Creek (Fig.
+\ref{fig:Geologic_map}, \ref{fig:in_situ_pmag}).
+The remanence directions recorded by red beds within the steeply folded basal section and within the flat-lying top portion of the section pass a bootstrap fold test (Supporting Information; \citeA{Tauxe1994a}).
+The remanence directions are all of normal polarity and show a dominantly southwest declination and upward-pointing inclination (Fig.
+\ref{fig:in_situ_pmag}).
 
@@
 \section*{Discussion}
-\subsection*{Inclination shallowing correction and Jacobsville paleomagnetic pole}
+\subsection*{Inclination shallowing correction}
 
-Although igneous and sedimentary rocks associated with the North America Midcontinent Rift provided high-resolution record of the ca.
-1110 to 1070 Ma Keweenawan Track which is the apparent polar wander path (APWP) for Laurentia in the Mesoproterozoic, high-quality paleogeographic constraints is lacking thereafter since tectonism in the interior of Laurentia entered a protracted period of quiescence until the emplacement of the ca.
+Although igneous and sedimentary rocks associated with the North American Midcontinent Rift provided high-resolution record of the ca.
+1110 to 1070 Ma Keweenawan Track which is the apparent polar wander path (APWP) for Laurentia in the late Mesoproterozoic, high-quality paleogeographic constraints is lacking thereafter since magmatism in the interior of Laurentia entered a protracted quiescence period until the emplacement of the ca.
 780 Ma Gunbarrel large igneous province \cite{Harlan2003a}.
-The new paleomagnetic data developed from the Jacobsville Formation presents an opportunity to extend the Keweenawan Track toward the early Neoproterozoic as its maximum deposition age has been recently tightly constrained to be ca.
-992 Ma \cite{Hodgin2022a}.
+The new paleomagnetic data developed from the Jacobsville Formation presents an opportunity to extend the Keweenawan Track toward the early Neoproterozoic as its maximum deposition age has been tightly constrained to be ca.
+992 Ma through high-precision U-Pb detrital zircon geochronology \cite{Hodgin2022a}.
 
-However, inclination bias associated with sedimentary rocks need to be corrected when synthesizing a paleomagnetic pole.
-Rotation of ferromagnetic grains during sediment deposition and compaction during early diagenesis can result in the acquisition of a detrital remanent magnetization that is biased shallow relative to the local geomagnetic field in which it was acquired \cite{King1955a, Tauxe2005a, Kodama2012a}.
+However, inclination bias associated with sedimentary paleomagnetic directions need to be corrected when synthesizing poles.
+Rotation of grains during sediment deposition and compaction during early diagenesis can result in the acquisition of a detrital remanent magnetization that is more shallowly inclined relative to the local geomagnetic field in which it was acquired \cite{King1955a, Tauxe2005a, Kodama2012a}.
 This effect can be summarized as:
@@
 \end{equation*}
-where $f$ represents the amount of inclination shallowing, $I_o$ represents the observed inclination of the specimen magnetization and $I_f$ represents the inclination of the field in which the magnetization was acquired.
-If uncorrected, shallower inclinations obtained from sedimentary rocks can result in erroneously low estimates of paleolatitudes, biasing the interpreted past positions of continents and hindering plate reconstructions \cite{Domeier2012a}.
-Methods for correcting sedimentary inclination bias have been developed.
-They include laboratory-based measurements of magnetic mineral fabrics \cite<e.g.>{Kodama1992a, Bilardello2015a}, as well as statistical methods based on the evaluation of the deviation of virtual geomagnetic pole (VGP) distribution patterns recorded by sedimentary rocks with an assumed paleosecular variation model (the ``elongation/inclination" [E/I] method of \citeA{Tauxe2004b}).
-When a large number of sedimentary paleomagnetic samples (typically n$>$100) capture an extensive period of time such that the paleosecular variation is sufficiently averaged, the E/I method can give accurate estimates and is more efficient than lab-based methods.
+where $f$ represents the amount of inclination shallowing, $I_o$ represents the observed inclination, and $I_f$ represents the inclination of the field in which the magnetization was acquired.
+If uncorrected, shallower inclinations obtained from sedimentary rocks can result in erroneously low estimates of paleolatitudes, biasing paleogeographic reconstructions \cite{Domeier2012a}.
+Methods for correcting sedimentary inclination bias include laboratory-based measurements of magnetic mineral fabrics \cite<e.g.>{Kodama1992a, Bilardello2015a}, as well as statistical methods based on the evaluation of the deviation of observed virtual geomagnetic pole (VGP) distribution patterns from an assumed paleosecular variation model (the ``elongation/inclination" [E/I] method of \citeA{Tauxe2004b}).
+When a large number of sedimentary paleomagnetic samples (typically n$>$100) capture an extensive period of time such that the paleosecular variation is sufficiently averaged, the E/I method can give accurate estimates and is less labor-intensive than methods based on mineral fabric measurements.
 In the ca.
-1.1 Ga Midcontinent Rift, \citeA{Pierce2022a} showed that the E/I method is effective in correcting the inclination bias recorded by the Cut Face Creek Sandstone---a lava flow-bracketed fluvial red bed sedimentary formation dominated by overbank deposition.
+1.1 Ga Midcontinent Rift, \citeA{Pierce2022a} showed that the E/I method is effective in correcting the inclination bias recorded by a package of lava flow-bracketed fluvial red bed sediments---the Cut Face Creek Sandstone.
 
-Inclination shallowing is also present in the Jacobsville Formation.
-Scatter of the specimen high-temperature component directions in geographic coordinates show dispersion of declinations parallel to the bedding plane, indicating the directions are likely shallowed along the bedding-vertical plane (Fig.
+Inclination shallowing is also present in the fluvial red beds of the Jacobsville Formation.
+Scatter of the specimen detrital remanence directions in geographic coordinates show dispersion of remanence vectors parallel to the bedding plane, indicating the directions are likely shallowed along the bedding-perpendicular plane (Fig.
 \ref{fig:in_situ_pmag}; \citeA{Tauxe2004b}).
 Since the deposition of the Jacobsville Formation span over geomagnetic field reversals, it is of interest to investigate the amount of inclination shallowing associated with the normal (n=130) and the reversed (n=148) polarity respectively.
-Applying the E/I method on the directions of normal polarity yields an estimated $f$ value of 0.67 with 95\% bootstrap uncertainty bounds from 0.91 to 0.55 (Fig.
+Applying the E/I method on the normal polarity directions yields an estimated $f$ value of 0.67 with 95\% bootstrap uncertainty bounds from 0.91 to 0.55 (Fig.
 \ref{fig:in_situ_pmag}B).
@@
 \ref{fig:in_situ_pmag}B).
-In addition, the dual-polarity directions do not pass a reversal test \cite{McFadden1990a, Tauxe2010a}, regardless of whether inclination correction is applied (Fig.
-\ref{fig:in_situ_pmag}).
+Although the estimated $f$ value for the reversed-polarity directions has large statistical uncertainty bounds, the uncertainties are likely associated with the overall shallow inclinations and distributions of directions in both upper and lower hemispheres (Fig.
+\ref{fig:in_situ_pmag}B).
+Therefore, inclination corrections do not result in significant changes in the mean reversed direction.
+In addition, directions of the two polarities do not pass a reversal test \cite{McFadden1990a, Tauxe2010a}, regardless of whether inclination correction is applied.
 
-One interpretation of the polarity discrepancy could be that the deposition of the Jacobsville Formation was protracted such that the sediments at Hungarian Falls which dominantly record a reversed polarity was deposited prior to those at Baby Snake Creek which dominate the normal polarity records (Fig.
+One interpretation of the polarity discrepancy could be that the deposition of the Jacobsville Formation was protracted such that the sediments at Hungarian Falls whose detrital remanence dominate reversed polarity records was deposited prior to those at Baby Snake Creek which dominate the normal polarity records (Fig.
 \ref{fig:in_situ_pmag}).
@@
 Field observations of \citeA{Tyrrell2019a} and \citeA{Mueller2021a} near Bete Grise Bay, which is only $\sim$10 km east to Baby Snake Creek (Fig.
-\ref{fig:Geologic_map}), show that significant thrusting and deformation along the Keweenaw Fault and subsequent weathering of the Midcontinent Rift basalt occurred prior to the deposition of the Jacobsville Formation.
-In addition, U-Pb calcite geochronology and thermometry confirm that the final motion of $\sim$2 km of vertical displacement along the Keweenaw Fault occurred during the ca.
-980 Ma Rigolet phase of the Grenvillian Orogeny and the majority of the Keweenaw Fault motion likely occurred much earlier during the ca.
-1050 Ma Ottawan phase orogeny \cite{Hodgin2022b}.
-Therefore, we interpret that it is most likely that the Jacobsville Formation were deposited just prior to and during the Rigolet phase of the Grenvillian Orogeny with progressive sediment transportation to the northwest resulting in its onlapping on top of the volcanics in the northern tip of the Keweenaw Peninsula.
-As a result, the sediments at Baby Snake Creek records a slightly more southerly paleolatitude than those near Hungarian Falls as Laurentia's plate motion continues toward the south (Fig.
-\ref{fig:in_situ_pmag}, \ref{fig:pole_plot}).
+\ref{fig:Geologic_map}), show unconformable onlap of the Jacobsville Formation on top of Midcontinent Rift volcanics that have experienced significant thrusting, deformation.
+At some locations, paleosol development has been found on top of the Midcontinent Rift basalt along the segmented Keweenaw Fault branches prior to the deposition of the Jacobsville sediments.
+This is in contrast with the faulted contact between the Jacobsville Formation and the volcanics to the south where the Jacobsville Formation is often folded against the volcanics (Fig.
+\ref{fig:Geologic_map}).
+U-Pb calcite geochronology and thermometry confirm that the final $\sim$2 km vertical displacement along the Keweenaw Fault occurred during the ca.
+980 Ma Rigolet phase of the Grenvillian Orogeny whereas the majority of the fault motion likely occurred during the ca.
+1050 Ma Ottawan phase \cite{Hodgin2022b}.
+We interpret that it is most likely that deposition of the Jacobsville Formation initiated close to the Rigolet phase of the Grenvillian Orogeny with progressive sediment transportation to the northwest as the Grenville back-bulge basin propagated toward the interior of Laurentia \cite{Hodgin2022b}.
+As a result, we hypothesize that the sediments at Baby Snake Creek likely records a slightly younger pole position with a more southerly paleolatitude than those near Hungarian Falls, indicating Laurentia's plate motion toward the south (Fig.
+\ref{fig:EI_results}).
+However, we cannot rule out an alternative interpretation that the sediments at Baby Snake Creek are coeval or older than some of the sediments to the south, as the basal Jacobsville Formation at Baby Snake Creek is folded against the Midcontinent Rift volcanics.
+Further field mapping and geochronological constraints are needed to test this hypothesis.
 
-An alternative interpretation to the polarity discrepancy could be that there are insufficient samples such that either or both polarities are not average for secular variation.
+% An alternative interpretation to the polarity discrepancy could be that there are insufficient samples such that either or both polarities are not average for secular variation. 
+Overall, both inclination-corrected polarities recorded by the Jacobsville Formation show mean pole positions in the southern hemisphere with respect to today's continental configuration (normal polarity longitude=175.1\textdegree E, latitude=19.1\textdegree S, A95=3.8\textdegree; reversed polarity longitude=191.1\textdegree E, latitude=14.5\textdegree S, A95=3.9\textdegree; both are corrected for inclination shallowing; Fig.
+\ref{fig:EI_results}).
+The two pole positions are distinct from the youngest poles developed from rocks of the Midcontinent Rift (Fig.
+\ref{fig:EI_results}).
+Despite that the Jacobsville sedimentation span over geomagnetic field reversals, the large arc distance between either polarity of the Jacobsville Formation and the pole of the ca.
+1070 Ma upper Freda Formation is consistent with the interpretation that the onset of deposition of the Jacobsville Formation at the five studied localities post-date the youngest member of the Oronto Group sediments by tens of millions of years.
 
-Overall, both inclination-corrected polarities recorded by the Jacobsville Formation show mean pole positions in the southern hemisphere with respect to today's continental configuration (normal polarity longitude=, latitude=, A95=; reversed polarity longitude= , latitude= , A95=).
-The two pole positions are close to each other, but are distinct from the youngest poles developed from rocks of the Midcontinent Rift (Fig.
-\ref{fig:pole_plot}).
-In addition, albeit that the Jacobsville sedimentation span over geomagnetic field reversals, the large arc distance between the poles of the Jacobsville Formation and that of the Freda Formation precludes the hypothesis that the Jacobsville Formation at the five studied localities (Fig.
-\ref{fig:Geologic_map}) is coeval with the Oronto Group and formed through a prolonged period of time.
-
+\subsection*{The Jacobsville paleomagnetic pole}
 To succinctly summarize a paleomagnetic pole for the ca.
@@
 \ref{fig:in_situ_pmag}).
-The resultant estimate for the $f$ value is 0.61, and the inclination-corrected mean direction is dec= , inc= , $\alpha_{95}$= .
-The bootstrap 95\% uncertainty range for the $f$ value range from 0.80 to 0.51, which overlaps with the uncertainty ranges of directions associated with each polarity (Fig.
+The resultant estimate for the $f$ value is 0.61 with a bootstrap 95\% uncertainty range for the $f$ value between 0.80 to 0.51, which overlaps with the uncertainty ranges of directions associated with either polarity (Fig.
 \ref{fig:in_situ_pmag}).
+The $f$ value of 0.61 is strikingly close to the value of 0.59 which is a calculated mean value of a compilation of inclination shallowing factor for hematite-bearing rocks derived by \cite{Bilardello2010b}.
 The Fisher mean paleomagnetic pole associated with the combined directions is shown in Fig.
-\ref{fig:pole_plot} (pole longitude= 186.3\textdegree E, pole latitude= 14.3\textdegree S, A95= 2.5\textdegree).
+\ref{fig:EI_results} (pole longitude= 186.3\textdegree E, pole latitude= 14.3\textdegree S, A95= 2.5\textdegree).
 
-\citeA{Pierce2022a} discussed that as sedimentary grains usually flatten perpendicular to the bedding plane but circularly symmetric about the bedding plane, the resultant shallow inclinations they record would typically result in asymmetric uncertainties in the recorded paleomagnetic pole position with more uncertainties in paleolatitues than paleolongitudes.
-The asymmetric uncertainties associated with sedimentary paleomagnetic poles violate the assumption of a Fisher distribution \cite{Fisher1953a} where uncertainties associated with a mean pole position is circularly distributed.
+\citeA{Pierce2022a} discussed that as inclination shallowing processes usually do not bias the recorded declination, it would typically result in asymmetric uncertainties in the recorded paleomagnetic pole position with more uncertainties in paleolatitues than paleolongitudes.
+The asymmetric uncertainties in the mean pole position thus violate the assumption of the Fisher distribution \cite{Fisher1953a} where uncertainties are circularly distributed about a mean vector.
 Instead, the spherical asymmetric bivariate Kent distribution \cite{Kent1983a} is more appropriate for incorporating uncertainties associated with inclination shallowing in reported sedimentary paleomagnetic poles.
 We follow the approach of \citeA{Pierce2022a} who extended the E/I method of \citeA{Tauxe2004b}.
-First we take all estimated $f$ values from the bootstrap results of the E/I algorithm and apply each value to correcting all Jacobsville detrital remanence directions.
-Then we calculate the Fisher statistics for each of the inclination-corrected mean poles.
-Finally we obtain 100 Monte Carlo resamples of each Fisher mean pole.
+First we take all estimated $f$ values from the bootstrap results of the E/I algorithm and apply each value to correct all Jacobsville detrital remanence directions (with the reversed directions flipped to antipodes).
+Then we calculate the Fisher statistics for each of the inclination-corrected mean pole.
+Finally, we perform a Monte Carlo approach to generate 100 resamples of each Fisher mean pole.
 As a result, the distribution of the resampled mean pole positions have an elongated distribution which represents more uncertainty associated with paleolatitudes of Laurentia than paleolongitudes (Fig.
+\ref{fig:EI_results}).
+The distribution of paleolatitudes of the resampled mean poles can be well-approximated by a normal distribution (Fig.
+\ref{fig:EI_results}).
+We use the Kent distribution to parameterize the 95\% confidence ellipse of the distribution of the resampled mean poles (mean longitude=186.4\textdegree E, mean latitude=14.2\textdegree S, major axis longitude=262.5, major axis latitude=43.5, major axis magnitude=3.3\textdegree, minor axis longitude=110.0, minor axis latitude=43.0, minor axis magnitude=3.0\textdegree; Fig.
+\ref{fig:EI_results}).
+As is shown in Fig.
+\ref{fig:EI_results}, although the Kent uncertainty ellipse has significant overlap with the circular uncertainty range calculated by the Fisher statistics due to the overall low paleolatitude of the Jacobsville basin, the Kent ellipse informs more uncertainty along the great circle path between the mean pole position and the locality of the basin as it incorporates the uncertainties in the amount of inclination shallowing within the sediments.
+
+\citeA{Roy1978a} isolated the detrital magnetization component via alternating-field, thermal, and chemical ``cleaning" demagnetization on a suite of Jacobsville red beds near Keweenaw Peninsula (their area A), the town of Marquette (area B), and Sault Ste Marie (area C).
+Although least-squares principal component analyses was not a routine approach in fitting paleomagnetic directions at the time, that study had success in removing the present day local field overprint and was able to resolve a magnetic polarity transition at one locality.
+The mean paleomagnetic pole position calculated from the interpreted primary magnetic remanence from rocks of the Keweenaw Peninsula and Marquette (J1 A+B pole; pole longitude=183\textdegree E, pole latitude=9\textdegree S, dp=3\textdegree, dm=6\textdegree; no inclination correction; Fig.
+\ref{fig:EI_results}; \citeA{Roy1978a}) is in the southern hemisphere and the uncertainty ellipse overlaps with that of the mean pole from this study (Fig.
+\ref{fig:EI_results}.
+However, their mean pole developed from red beds near Sault Ste Marie is in the northern hemisphere with a latitude of 12\textdegree N (Fig.
+\ref{fig:EI_results}).
+Despite the large uncertainty ellipse associated with this mean pole position, it is distinct from the mean pole position from Jacobsville rocks of the Keweenaw Peninsula and Marquette.
+Instead, it overlaps with the mean pole position of the ca.
+1070 Ma Upper Freda Formation (Fig.
+\ref{fig:EI_results}; \citeA{Henry1977a}).
+We hypothesize that the red beds collected by \citeA{Roy1978a} in Sault Ste Marie area are likely Freda-equivalent sediments, instead of being a part of the Jacobsville Formation.
+
+Deposition of the Bayfield Group sediments west to the Keweenaw Peninsula has been hypothesized to be coeval with the Jacobsville Formation but chronological constraints have been lacking.
+Our new paleomagnetic data from the Jacobsville Formation reaffirms that the pole positions of the the Chequamegon Sandstone of the Bayfield Group and the Jacobsville Formation are close albeit distinct (Fig.
+\ref{fig:EI_results}).
+Given that the paleomagnetic pole from the Chequamegon Sandstone of the Bayfield Group (pole longitude=177.7\textdegree E, pole latitude=12.3\textdegree S, A95=4.6\textdegree; \citeA{McCabe1983a}) was calculated without applying inclination correction, the two formations could have even closer a temporal association than the current paleomagnetic pole configuration appears (Fig.
+\ref{fig:EI_results}).
+In addition, the fact that only reversed polarity directions were recovered from the Chequamegon Sandstone could indicate that deposition of the sandstone collected by \citeA{McCabe1983a} could be coeval with one of the reversed polarity period recorded by the Jacobsville Formation (Fig.
+\ref{fig:strat_column}).
+To further test the temporal correlation between the sedimentary units, detailed paleomagnetic study and detrital zircon U-Pb geochronology data from the Chequamegon Sandstone are needed.
+
+\subsection*{Paleogeography of Laurentia in the late Proterozoic and the development of supercontinent Rodinia}
+
+Laurentia's apparent polar wander is well constrained in the Mesoproterozoic by the Keweenawan Track thanks to the large number of paleomagnetic poles paired with high-precision geochronology from rocks of the ca.
+1.1 Ga Midcontinent Rift \cite{Swanson-Hysell2019a}.
+In addition, the Midcontinent Rift have had simple a thermal history for their old ages, as rifting ceased prior to lithospheric separation, preserving rocks of the rift within the continental interior far from continental margin and subsequent orogenesis.
+Since the end of the Midcontinent Rift mamatism, igneous paleomagnetic records from the the interior of Laurentia is lacking, and data from rocks associated with the ca.
+1080 to 980 Ma Grenville Orogeny have been used to infer for Laurentia's plate configuration in the late Mesoproterozoic to early Neoproterozoic \cite<e.g.>{Weil1998a, Halls2015b}.
+However, in contrast to rocks of the Midcontinent Rift which usually have straightforward paleomagnetic records \cite<e.g.>{Swanson-Hysell2019a}, rocks of the Grenville Province which experienced up to granulite facies metamorphism have much more complicated thermal histories and the timing of their magnetic remanence acquisition and alteration is more difficult to constrain.
+As is shown in Fig.
+\ref{fig:pole_plot}, paleomagnetic poles developed from rocks of the Grenvillian Orogeny plots near Australia with respect to today's continental configuration, forming arc distances ranging from $\sim$35\textdegree to more than 50\textdegree away from the pole of the ca.
+1070 Ma Freda Formation \cite{Henry1977a}.
+Such large discrepancies between pole positions associated with uncertainties in the age of the magnetization of the Grenville rocks led to the controversial ``Grenville Problem"---did the Grenville rocks acquire magnetic remanence prior to the continental collision and thus the pole discrepancies are a result of significant crustal shortening (i.e.
+the ``two plate model"; \citeA<e.g.>{Irving1972a, Palmer1973a, Buchan1973a, Halls2015b}), or did they acquire remanence after the continental accretion and thus the southerly poles from the Grenville rocks represent Laurentia's continued motion through more recent times (i.e.
+the ``one plate model"; \citeA<e.g.>{Roy1978a, McWilliams1975a, Weil1998a})?
+
+The two plate hypothesis is based on the interpretation that timing of magnetic remanence acquisition of the Grenville rocks was close to that of the Midcontinent Rift rocks \cite<e.g.>{Warnock2000a, Halls2015b}, but the lack of southerly paleomagnetic poles from the interior of Laurentia is interpreted to indicate that the Grenville rocks acquired remanence on allochthonous terranes before the final accretion to Laurentia craton.
+Thus, up to 4000 $\pm$ 1000 km of crustal shortening was invoked by \citeA{Halls2015b} to explain the large arc distance between the poles.
+
+Given the new constraints on the age and position of the Jacobsville pole (Fig.
+\ref{fig:pole_plot}), the two plate model would require that the majority of the crustal shortening happened after the Rigolet phase of the Grenville Orogeny, as the ca.992 Ma Jacobsville pole is still $\sim$45\textdegree arc distance ($\sim$5000 km) away from the paleomagnetic pole of the Haliburton intrusions \cite{Buchan1976a} and $\sim$38\textdegree ($\sim$4000 km) from the pole of the Wilberforce intrusions \cite{Palmer1973a} (Fig, \ref{fig:pole_plot}).
+However, geologic and geochronologic evidence from both within the Grenville Province \cite{Aleinikoff2021a} and from interior Laurentia \cite{Hodgin2022a} indicate that majority of the deformation in the Grenville Province and the propagation of its associated deformation in the Midcontinent Rift happened during the ca.1090-1020 Ma Ottawan phase orogeny, whereas smaller amount of convergence happened during the ca.
+1010-980 Ma Rigolet phase.
+The lack of geological evidence for a striking amount of crustal shortening is supported by the slow average plate motion between the deposition of the upper Freda Formation and the deposition of the Jacobsville Formation (average $\sim$0.2\textdegree/Myr; Fig.
 \ref{fig:pole_plot}).
-The distribution of paleolatitudes of the resampled mean poles can be well-approximated by a normal distribution (Fig.
+Therefore, it is kinetically unlikely that the plate of Laurentia suddenly increased its motion during the waning phase of a collisional orogenesis in order to satisfy the Grenville pole positions.
+The plate kinetics interpretation based solely on paleomagnetic data and poorly constrained ages of magnetic remanence put forward by \citeA{Halls2015b} is implausible.
+Alternatively, the one plate model is much more straightforward in explaining the paleomagnetic pole positions of Grenville rocks as it interprets that the remanence was acquired during prolonged slow cooling of the orogeny that postdate the peak metamorphism.
+In contrast with the two plate model, the one plate model would predict that ages associated with the Grenville poles are much younger than ca.
+980 Ma as the rocks slowly cooled after the peak metamorphism associated with the collisional orogenesis (Fig.
 \ref{fig:pole_plot}).
-We use the Kent distribution to parameterize the 95\% confidence ellipse of the distribution of the resampled mean poles (mean longitude=186.4\textdegree E, mean latitude=14.2\textdegree S, major axis longitude=262.5, major axis latitude=43.5, major axis magnitude=3.3\textdegree, minor axis longitude=110.0, minor axis latitude=43.0, minor axis magnitude=3.0\textdegree; Fig.
+
+Adopting the one plate model to explain Laurentia's late Proterozoic apparent polar wander path necessitates that the ages assigned to many paleomagnetic poles derived from Grenville rocks need to be adjusted toward younger.
+Abundant petrological and geochronological evidence show large portion of the Grenville Province experienced up to granulite facies metamorphism (e.g.
+$>$700\textdegree C).
+Such high peak metamorphism temperatures would result in unblocking of any remanence the rocks had acquired earlier and complete overprint by remanence acquired during cooling after the orogenesis.
+Previous paleomagnetic studies have found that Grenville rocks commonly have magnetite and hematite being the remanence carriers \cite<e.g.>{Ueno1975a, Alvarez1998a, Buchan1973a, Buchan1976a, Brown2012a}.
+Theoretical calculations have shown that magnetite which formed above its Curie temperature (580\textdegree C) in Grenville rocks subsequently acquired thermoviscous remanence \cite{Pullaiah1975a, Dodson1980a} during prolonged cooling of the orogen, whereas hematite which often occur as exsolved lamellae interfingering with ilmenite acquired thermochemical remanence via magnetic ordering of the hemo-ilmenite below the Neel temperature of hematite \cite{Burton1985a, Harrison2006a}.
+
+Due to the long time scales involves with thermoviscous and thermochemical origin of the magnetic remanence, it is not straightforward to interpret the natural blocking temperatures of Grenville rocks based on observations of unblocking temperatures in the lab where rocks are heated and cooled rapidly.
+As a result, extrapolating equivalent natural blocking temperatures from lab unblocking temperatures using theoretical models \cite{Pullaiah1975a, Dodson1980a} and pairing the temperatures with thermochronology data is necessary to determine age of the remanence.
+\citeA{Brown2012a} recognized that the hemo-ilmenite exsolution occur at relatively low temperatures during cooling of the orogen and interpreted their new paleomagnetic data from rocks of the Adirondack Mountains, US, in context of recent understanding of the hemo-ilmenite phase diagram.
+That study assigned an age of ca.
+960 Ma to a paleomagnetic pole developed from hemo-ilmenite-bearing microcline gneiss from northwestern Adirondacks (Fig.
 \ref{fig:pole_plot}).
-As is shown in Fig.
-\ref{fig:pole_plot}, although the Kent ellipse uncertainty range of the Jacobsville Formation's paleomagnetic pole position has significant overlap with the circular uncertainty range represented by the Fisher mean, the Kent ellipse informs more uncertainty along the great circle path between the mean pole position and the locality of the Jacobsville Formation as it incorporates the uncertainties in the amount of inclination shallowing within the sediments.
+They assigned two additional paleomagnetic poles of ages ca.
+990 Ma and 970 Ma based on data from the Wanakena granite and anorthositic rocks associated with the Marcy Massif (Fig.
+\ref{fig:pole_plot}).
+Both units are interpreted to have magnetite as the dominant remanence carrier.
+However, in light of the ca.
+992 Ma Jacobsville pole position and the slow down of Laurentia's apparent polar wander since ca.
+1070 Ma, the ages assigned to the poles in \citeA{Brown2012a} could be still too old to be compatible with the magnetization being acquired after the major pulse of the orogenic collision.
 
-% \subsection*{Age of the Grenville Loop}
+That the estimated ages associated with the Adirondack paleomagnetic poles are biased old can be attributed to the outdated thermochronometer closure temperature calculations of \citeA{Mezger1991a, 1992a} whose data was used by \citeA{Brown2012a}.
+In particular, the cooling history for the Adirondack area estimated by \citeA{Mezger1991a, 1992a} show that the orogen had cooled to $\sim$400\textdegree C by the time of U-Pb system closure in rutile crystals.
+However, this estimated U-Pb closure temperatures range for rutile was not based on laboratory Pb diffusion experiment, but instead was an assumption that diffusion rate of Pb in rutile approximate that of Fe and Ti \cite{Mezger1989a}.
+Later experimental results \cite{Cherniak2000a} as well as analytical estimates \cite{Vry2006a, Kooijman2010a} both show that diffusion rates of Pb in rutile are in fact much slower, indicating that previous calculations significantly underestimated closure temperatures of the system.
+Therefore, timing of magnetic remanence acquisition for Grenville rocks in the Adirondack Mountains should be younger than previously thought.
 
-\subsection*{Paleogeography of Laurentia during the late Mesoproterozoic to early Neoproterozoic}
+The interpretation of young Grenville poles is supported by U-Pb apatite thermochronology data from iron oxide-apatite deposits of the eastern Adirondack Mountains \cite{Krestianinov2021a} and from the Wilberforce pyroxenite of the Wilberforce region, Canada \cite{Paul2021a}.
+Apatite from both regions yield ages of ca.
+900 Ma with associated closure temperatures $\sim$450\textdegree C \cite{Cherniak1991a, Chamberlain2001a}.
+Considering the thermoviscous remanence acquisition of magnetite and hematite (even after the exsolution of hemo-ilmenite lamellae) in the slowly cooled orogen \cite{Pullaiah1975a, Dodson1980a}, ca.
+900 Ma could be closer to the actual timing when rocks of the Grenville Province finally blocked in the remanence that is preserved today.
+However, the exact age of each paleomagnetic pole recorded by rocks of different domains of the Grenville Province remains difficult to constrain.
+Complications associated with the origin of the dated iron oxide-apatite ore, crystal-size dependence of closure temperature of accessory minerals such as rutile and apatite \cite{Smye2018a, Chamberlain2001a}, and the complex metamorphic histories in different domains of the Grenville Province require that more high-precision thermochronology data with detailed closure temperature calibration based on modern experimental results are needed.
 
-\citeA{McCabe1983a} developed paleomagnetic data from the Chequamegan Sandstone of the Bayfield Group and found reversed-polarity directions similar to that of the Jacobsville Formation poles developed by \citeA{Roy1978a}.
+Overall, albeit that numerous paleomagnetic data has been developed from rocks of the Grenville Province, few are paired with high-precision thermochronology data that are calibrated by modern diffusion paramters.
+More high-precision thermochronology data and statistical models that properly incorporates uncertainties associated with timing of closure of thermochronometers and with temperatures of magnetic remanence acquisition in slowly cooled orogen are needed in order to further calibrate the age of those poles.
+Refining the pole-age pairs for the Grenville rocks can further test the hypothesized ``Grenville Loop" of Laurentia's APWP \cite{Irving1974a} and refine the paleogeographic configuration during the development of super continent Rodinia in the Neoproterozoic.
 
-Although igneous paleomagnetic records from the the interior of Laurentia has been lacking since the end of the Midcontinent Rift magmatism, data from metamorphosed rocks associated with the ca.
-1080 to 980 Ma Grenville continental collisional orogenesis has been developed to infer for Laurentia's plate configuration in the late Mesoproterozoic to early Neoproterozoic \cite<e.g.>{Weil1998a, Halls2015b}.
-However, in contrast to igneous rocks of the Midcontinent Rift which have relatively simple thermal history and straightforward paleomagnetic records \cite,e.g.>{Swanson-Hysell2019a}, rocks of the Grenville Province which experienced up to granulite facies metamorphism have much more complicated thermal history and the timing of their magnetic remanence acquisition is more difficult to constrain.
-As is shown in Fig.
-\ref{fig:pole_plot}, paleomagnetic poles developed from rocks of the Grenvillian Orogeny plots near Australia in today's continental reference frame, forming an arc distance of $\sim$30 degrees from the youngest paleomagnetic poles of the ca.
-1070 Ma Freda Formation \cite{Henry1977a}.
-Such great discrepancies between pole positions associated with uncertainties in the age of the magnetization of the Grenville rocks led to the ``Grenville Problem" which is still highly debated today - did the Grenville rocks acquire magnetic remanence prior to the continental collision and thus the pole discrepancies are a result of significant crustal shortening (4000 $\pm$ 1000 km; \citeA<e.g.>{Halls2015b}), or did they acquire remanence after the continental accretion and thus the southerly poles merely represent Laurentia's continued motion through more recent times \cite<e.g.>{Weil1998a}?
-
-As the ca.
-990 Ma Jacobsville Formation was deposited in Laurentia Interior (Fig.
-\ref{fig:Geologic_map}), its paleomagnetic pole position can help test the contrasting hypothesis regarding the ``Grenville Problem".
-
-paleomagnetic mean pole of the Jacobsville Sandstone and the age
-
-inclination shallowing correction of the Jacobsville Sandstone
-
-magnetic reversal during the Jacobsville and the Maya Superchron
-
-Implication on the crustal shortening of the Rigolet phase of the Grenville Orogeny
-
-Warnock on the age of magnetization on the Haliburton pole: based on the very coarse thermal demagnetization steps Warnock2000a assigned that all of the primary magnetite components unblocked above 550 \textdegree C, with an uncertainty of 555 \textdegree C $\pm$ 5\textdegree C.
+% Warnock on the age of magnetization on the Haliburton pole: based on the very coarse thermal demagnetization steps Warnock2000a assigned that all of the primary magnetite components unblocked above 550 \textdegree C, with an uncertainty of 555 \textdegree C $\pm$ 5\textdegree C.
 Using a thermal history model that combines U-Pb titanite ages, hornblende and biotite Ar/Ar ages and just by merely compare the thermal history model with the assigned narrow magnetite unblocking temperature range, the authors assigned the age of the magnetization to be 1015 $\pm$ 15 Ma.
 
-It is important to note that the result Ar ages of the hornblende and biotite are 983 $\pm$ 8 Ma, 996 $\pm$ 8 Ma, and 998 $\pm$ 8 for hornblende with unblocking temperature of 530 $\pm$ 40\textdegree C, consistent across metamorphic terranes (Bancroft).
+% It is important to note that the result Ar ages of the hornblende and biotite are 983 $\pm$ 8 Ma, 996 $\pm$ 8 Ma, and 998 $\pm$ 8 for hornblende with unblocking temperature of 530 $\pm$ 40\textdegree C, consistent across metamorphic terranes (Bancroft).
 And the biotite Ar ages are 910 $\pm$ 7 Ma, 891 $\pm$ 7 Ma, corresponding to unblocking temperature of 280 $\pm$ 40\textdegree C.
 
-More, the thermal demagnetization plots shown in Warnock2000a suggest that the magnetite unblocking lower temperature boundary can be around 500\textdegree C, lower than 555 $\pm$ 5\textdegree C.
+% More, the thermal demagnetization plots shown in Warnock2000a suggest that the magnetite unblocking lower temperature boundary can be around 500\textdegree C, lower than 555 $\pm$ 5\textdegree C.
 In addition, given the case that the rocks underwent uniform rate of cooling due to post-formation uplifting, a slow cooling effect from 555\textdegree C to 300\textdegree C from 990 Ma to 900 Ma allows for a scenario where the magnetite remament magnetization that appear to have a 555\textdegree C unblocking temperature in lab was acquired through a period of --- myr cooling though 500\textdegree C during uplifting.
@@
 
-Crucial link between the Keweenawan Track and the Grenville Loop
-
 \section*{Conclusion}
 
+\section*{Data Availability}
+Paleomagnetic data associated with this study are available within the MagIC database (\url{XXX}; UPDATE WHEN DOI IS GENERATED) and all data are within a Github repository associated with this work (\url{https://github.com/Swanson-Hysell-Group/Jacobsville}) that is also archived on Zenodo (\url{https://doi.org/XXX}).
+This repository also contains Python code that implements all of the calculations, visualizations and statistical tests discussed herein.
+
 \acknowledgments
-Enter acknowledgments, including your data availability statement, here.
+This project is funded by NSF grant EAR-1847277 to Nicholas Swanson-Hysell.
 

### 2022-07-20 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 \usepackage{amsmath}
+\usepackage{sidecap}
 \usepackage[inline]{trackchanges} %for better track changes. finalnew option will compile document with changes incorporated.
@@
 1090-980 Ma Grenvillian Orogeny along the eastern margin.
-The far-field stress of the orogeny is interpreted to have propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation whose maximum deposition age has been constrained to be ca.
+The far-field stress of the orogeny is interpreted to have propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation whose maximum depositional age has been constrained to be ca.
 992 Ma.
-We develop a new paleomagnetic pole from red beds of the Jacobsville Formation along 5 stratigraphic sections in northern Michigan, USA.
+We develop a new paleomagnetic pole from red beds of the Jacobsville Formation along five stratigraphic sections in northern Michigan, USA.
 High-resolution thermal demagnetization experiments resolve detrital magnetic remanence that pass a fold test and two conglomerate tests.
-Our results confirm the occurrence of geomagnetic field reversals during the deposition of Jacobsville Formation, providing constraints for the duration of the Keweenawan superchron.
-Inclination corrected paleomagnetic pole position from the Jacobsville Formation confirm continued plate motion of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic.
-The new pole from the interior of Laurentia indicates the paleomagnetic poles developed from the Grenville Province most likely record coherent plate motion after major continental collision during the development of supercontinent Rodinia.
+Our results confirm the occurrence of geomagnetic field reversals during the deposition of Jacobsville Formation, providing constraints for the maximum duration of the Keweenawan superchron.
+An inclination corrected paleomagnetic pole position from the Jacobsville Formation confirm continued plate motion of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic.
+The new pole from the interior of Laurentia indicates the paleomagnetic poles developed from the Grenville Province most likely record coherent plate motion after major continental collision during the development of the supercontinent Rodinia.
 % rather than recording a significant crustal shortening.
@@
 
-\begin{keypoints}
-\item Neoproterozoic
-\item Laurentia  
-\item Paleogeography
-\item apparent polar wander path
-\item Inclination shallowing
-\item Grenville
-\item Keweenawan superchron
-\item Red bed
-\end{keypoints}
+\section*{Keywords}
+Neoproterozoic\\
+Laurentia  \\
+Paleogeography\\
+\
+
+apparent polar wander path\\
+Inclination shallowing\\
+Grenville\\
+Keweenawan superchron\\
+Red bed\\
 
@@
 1109 Ma to 1084 Ma North American Midcontinent Rift (i.e.
-the Keweenawan Track) provides crucial constraints on global paleogeography in the Mesoproterozoic and has become a central record for reconstructing the development of ancient supercontinent Rodinia \cite{Swanson-Hysell2019a, Swanson-Hysell2021c}.
+the Keweenawan Track) provide crucial constraints on global paleogeography in the Mesoproterozoic \cite{Swanson-Hysell2019a}.
+As a result, the Keweenawan Track has become a central record for reconstructing the development of the ancient supercontinent Rodinia \cite{2021a}.
 As the rifting ceased and subsequently inverted following the onset of the Grenvillian Orogeny on Laurentia's eastern margin (Fig.
-\ref{fig:Geologic_map}), magmatism in the interior of Laurentia entered a quiescence period that could have lasted as much as $\sim$300 Myr until the emplacement of the ca.
-780 Ma Gunbarrel large igneous province \cite{Harlan2003a}.
-As a result, uncertainties exist in reconstructing Laurentia and its conjugate continents' configuration in the late Mesoproterozoic to early Neoproterozoic.
+\ref{fig:Geologic_map}), magmatism in the interior of Laurentia entered a period of quiescence that could have lasted for as long as $\sim$300 Myr until the emplacement of the ca.
+780 Ma Gunbarrel large igneous province \cite{Harlan2003a, Mackinder2019a}.
+Due to this magmatic gap, Laurentia and its conjugate continents' paleogeographic configuration in the late Mesoproterozoic to mid-Neoproterozoic has remained largely uncertain \cite{Swanson-Hysell2021c}.
 
@@
 \caption{(A) Regional map showing location of the Grenville Front relative to the study area.
-(B) Geologic map of northern Michigan and Wisconsin showing Midcontinent Rift igneous and sedimentary rocks, the Jacobsville Formation, and major post-rift thrust faults.
-Jacobsville stratigraphic sections along which paleomagnetic samples were collected are shown as green diamond symbols.}
+The inset green box shows the extent of panel B.
+(B) Geologic map of northern Michigan and Wisconsin showing Midcontinent Rift igneous and sedimentary rocks, the Jacobsville Formation, other older Proterozoic rocks, and major post-rift thrust faults.
+Jacobsville stratigraphic sections along which paleomagnetic samples were collected in this study are shown as green diamond symbols.}
 \label{fig:Geologic_map}
@@
 
-Although igneous activities are lacking, sedimentary rocks formed near the end of the Midcontinent Rift magmatism and afterwards, and can provide additional paleomagnetic data.
-In particular, hematite-bearing sedimentary rocks (i.e.
-red beds) can provide insights into the ancient geomagnetic field and the paleogeographic positions of sedimentary basins.
+When igneous rock records are lacking, paleomagnetic data recorded by sedimentary rocks (e.g.
+red beds) can provide useful insights into the ancient geomagnetic field and record paleogeographic positions of sedimentary basins.
 In the Midcontinent Rift, the paleomagnetic poles developed from the ca.
@@
 1070 Ma upper Freda Formation of the Oronto Group sediments that overlie the Midcontinent Rift volcanics in southern Lake Superior area (Fig.
-\ref{fig:Geologic_map}) provide crucial constraints for Laurentia's latitude, orientation, and plate motion in the late Mesoproterozoic \cite{Henry1977a}.
+\ref{fig:Geologic_map}) provide crucial constraints for Laurentia's latitude, orientation, and plate motion in the late Mesoproterozoic \cite{Henry1977a, Swanson-Hysell2019a, Swanson-Hysell2019b}.
 
-In addition to the Oronto Group sediments, the overlying post-rift Jacobsville Formation which were deformed during the inversion of the Midcontinent Rift also contains abundant red beds which bear the opportunity to further extend the reconstruction of the apparent polar wander path of Laurentia toward more recent times \cite{Hamblin1958a, Hodgin2022a}.
+On top of the Oronto Group sediments overlies the post-rift Jacobsville Formation which was partly deformed during the inversion of the Midcontinent Rift (Fig.
+\ref{fig:Geologic_map}; \citeA{Hamblin1958a, Hodgin2022a, DeGraff2022a}).
+The Jacobsville Formation also contains abundant red beds which offer the opportunity to further extend the reconstruction of the apparent polar wander path of Laurentia toward more recent times \cite{Hamblin1958a, Roy1978a}.
 U-Pb detrital zircon ages and U-Pb calcite ages from late-kinematic veins constrain the maximum depositional age of the Jacobsville Formation to be ca.
-990 Ma \cite{Hodgin2022a}.
+992 Ma \cite{Hodgin2022a}.
 Developing paleomagnetic data from the Jacobsville Formation will allow further development of Laurentia's paleogeographic evolution in the late Mesoproterozoic to early Neoproterozoic.
 
-Challenges exist in interpreting sedimentary paleomagnetic records as they acquire magnetic remanence not through cooling from above the magnetic ordering temperatures of magnetic minerals as igneous rocks do, but acquire detrital remanence as magnetic particles settle from the water column during deposition, and chemical remanence as magnetic minerals precipitate and grow within pore spaces of host sedimentary rocks syn- to post-deposition (i.e.
-pigmentary hematite).
-It has been found that the formation of the pigmentary hematite in red beds can significantly post-date the timing of sediment deposition, resulting in the chemical remanent magnetization being secondary, superimposed upon the primary remanent magnetization carried by detrital grains \cite{Butler1992a}.
-To isolate the remanence components of the red beds of the Jacobsville Formation, \citeA{Roy1974, Roy1978} adopted a method first applied by \citeA{Collinson1965a} to preferentially remove fine-grained pigmentary hematite through prolonged immersion in concentrated HCl acid whereas leaving the coarser detrital grains that are more resistant to dissolution in place.
-Paired acid etching demagnetization and thermal demagnetization experiments show that the pigmentary hematite grains tend to have lower unblocking temperatures than detrital ones \cite{Bilardello2010a, Tauxe1980a}.
+Challenges exist in interpreting sedimentary paleomagnetic records.
+Primary detrital remanence acquired during settling of magnetic particles in the water column can be masked as secondary magnetic minerals (e.g.
+pigmentary hematite) that precipitate and grow within pore spaces of host sedimentary rocks syn- to post-deposition.
+In red beds, it has been found that formation of the pigmentary hematite can significantly post-date the timing of sediment deposition, resulting in the chemical remanent magnetization being secondary, superimposed upon the primary remanent magnetization carried by detrital grains \cite{Butler1992a}.
+To isolate the remanence components of the red beds of the Jacobsville Formation, \citeA{Roy1974a, Roy1978a} adopted a method first applied by \citeA{Collinson1965c} to preferentially remove fine-grained pigmentary hematite through prolonged immersion in concentrated HCl acid whereas leaving the coarser detrital grains that are more resistant to dissolution in place.
+Studies applying paired acid etching demagnetization and thermal demagnetization show that the pigmentary hematite grains tend to have lower unblocking temperatures than detrital ones \cite{Tauxe1980a, Bilardello2010c}.
 High-resolution thermal demagnetization experiments with temperature intervals as small as 1-2 \textdegree C have shown to be effective in isolating detrital remanence from chemical remanence as coarser hematite grains tend to have higher unblocking temperatures \cite{Swanson-Hysell2019b}.
-In particular, \cite{Swanson-Hysell2019b} conducted a conglomerate test on a suite of fluvial intraclasts of the red beds in the Freda Formation to show that via thermal demagnetization one can isolate the primary magnetization (which are reoriented in random directions in the clasts) from the secondary component carried by the co-occurring authigenic hematite (which yield similar directions amongst clasts).
+In particular, a conglomerate test on a suite of fluvial intraclasts of the red beds in the Freda Formation shows that via thermal demagnetization one can isolate detrital magnetization vectors (which are reoriented in random directions in the clasts) from the secondary components carried by co-occurring authigenic hematite (which yield similar directions amongst clasts) \cite{Swanson-Hysell2019b}.
 
-Paleomagnetism of red beds has also been challenged by the controversial issue of correcting for inclination shallowing \cite{King1955a, Taux2004b, Bilardello2016b}.
-Due to the rotation of detrital magnetic grains toward bedding-parallel caused by authigenic processes such as compaction during lithification, inclinations recorded by sediments often have shallower angles than what was during the time of deposition.
-To correct for such bias, \citeA{Tauxe2004b} developed a statistical model based on the geocentric axial dipole (GAD) hypothesis where the amount of inclination shallowing in a specific sediment package is evaluated based on the shape of the distribution of the sedimentary paleomagnetic directions is compared to that predicted by a paleosecular variation model developed for the past 5 Myr.
-\citeA{Pierce2022a} showed that the paleosecular variation model of \citeA{Tauxe2004b} is also suitable for application to sedimentary rocks formed during the Midcontinent Rift.
-That study also further extended the method to better represent uncertainties associated with the amount of inclination shallowing when reporting sedimentary paleomagnetic pole positions by using the spherical elliptical Kent distribution \cite{Kent1982a} instead of the commonly used circular Fisher distribution \cite{Fisher1953a}.
+Paleomagnetism of red beds has also been challenged by the controversial issue of correcting for inclination shallowing \cite{King1955a, Tauxe2004b, Bilardello2016b}, which is necessary due to the rotation of detrital magnetic grains to be parallel to bedding caused by authigenic processes such as compaction during lithification.
+As a result, inclinations recorded by sediments often have shallower angles than the local field inclinations during the time of deposition.
+To estimate the amount of inclination shallowing in a given package of sedimentary rocks, \citeA{Tauxe2004b} developed a statistical model (i.e.
+the elongation/inclination method) that evaluates the deviation of the distribution of a large number ($>$100) of observed sedimentary paleomagnetic directional data with the predicted distribution given by a paleosecular variation model developed for the past 5 Myr (TK03) under the the geocentric axial dipole (GAD) hypothesis.
+Although the TK03 paleosecular variation model is based on data of relatively recent geomagnetic field, it has been shown to be compatible with data from the ca.
+1.1 Ga lava flows of the Midcontinent Rift \cite{Tauxe2009a}.
+Further, \citeA{Pierce2022a} showed that the E/I method passes a field test for the 1.1 Ga Cut Face Creek Sandstone as the method successfully corrects the shallow paleomagnetic inclination of the sediments to that of the lava flows that bracket them.
+That study also further extended the method to better represent uncertainties associated with the amount of inclination shallowing when reporting sedimentary paleomagnetic pole positions by using the spherical elliptical Kent distribution \cite{Kent1982a} instead of the circular Fisher distribution \cite{Fisher1953a}.
 
 In this study, we investigate the hematite-bearing, fine-grained sandstone to siltstone lithofacies along five stratigraphic sections of the Jacobsville Formation.
-High-resolution thermal demagnetization data on centimeter-scale siltstone intraclasts that were eroded by fluvial processes and redeposited at two locations constrain the timing of hematite acquisition by revealing a primary component that formed prior to the erosion of the clasts within the depositional environment and a secondary component that formed following their redeposition.
-Then we develop a new inclination-corrected paleomagnetic pole for the Jacobsville Formation based on a large number of samples and investigate the paleo-continental configuration of Laurentia and its association with Baltica during the development of the supercontinent Rodinia.
-
-% preservation of the Midcontinent Rift sedimentary rocks (Figure \ref{fig:Geologic_map}), complex magnetic remanence components that often form during significantly different times within the rocks and post-deposition modification of the remanence magnetization directions are common to sedimentary rocks, presenting challenges to the interpretation of their paleomagnetic records \cite{Beck2003a, Butler1992a, Vandervoo2012a}.
-In contrast to igneous rocks which acquire thermal remanent magnetizations through cooling of iron oxides, red beds often record mixtures of magnetic remanence components during primary hematite grain deposition and secondary hematite precipitation and growth that are often not coeval \cite{Tauxe1980a, Swanson-Hysell2019b}.
-Detrital hematite grains can record magnetic remanence at the time of deposition, whose age can be constrained by the maximum deposition age of the host sedimentary rock.
-The secondary source of hematite is the grains that precipitate and grow in situ within pore space of host sedimentary rocks that have been deposited (i.e.
-pigmentary hematite).
-These secondary hematite could form significantly after deposition of the sedimentary rocks and can keep growing through a protracted period of time.
-To isolate the primary magnetization component from the secondary one within red beds is crucial for accurately reconstructing paleomagnetic records at the time of red beds deposition.
-Another challenge to interpreting sedimentary paleomagnetic data is the problem of inclination shallowing \cite{King1955a}.
-Grain rotation toward being bedding plane parallel during processes associated with deposition and lithification can result in recorded inclinations being shallower than the local magnetic inclination.
-Such deviation may vary depending on the location where the sedimentary rock formed and can result in various degrees of uncertainties of paleomagnetic reconstructions.
+High-resolution thermal demagnetization data on siltstone intraclasts that were eroded by fluvial processes and redeposited at two locations constrain the timing of hematite acquisition by revealing a primary component that formed prior to the erosion of the clasts within the depositional environment and a secondary component that formed following their redeposition.
+We then use a paleomagnetic fold test to further constrain the timing of detrital remanence acquisition in association with the structural inversion of the Midcontinent Rift.
+Finally, we develop a new inclination-corrected paleomagnetic pole for the Jacobsville Formation based on a large number of samples and investigate the paleo-continental configuration of Laurentia and its association with the development of the supercontinent Rodinia.
 
@@
 
-% Previous paleomagnetic data from the Midcontinent Rift rocks that are paired with high-precision geochronology data has led to the development of the Keweenawan Track, showing rapid plate motion (latitudinal motion of $\sim$ cm/yr) from high latitudes between ca.
-1105 Ma and ca.
-1092 Ma, followed by a decrease in plate motion (latitudinal motion of $\sim$) between ca.
-1092 Ma and ca.
-1080 Ma when Laurentia moved to the equator and ceased its rapid motion.
-In particular, paleomagnetic poles developed from the ca.
-1080 Ma Nonesuch Formation and the ca.
-1050 Ma upper Freda Formation of the Oronto Group sediments indicate that Laurentia's plate speed significantly reduced in the late Mesoproterozoic, resulting in little change in its paleocontinental configuration for tens of millions of years.
-
 % A leading tectonic model for the cessation of the Midcontinent Rift and extension is that it occurred due to far-field compressional stress associated with the ca.
@@
 For further refining the paleogeographic connection between the poles of the Grenville Loop and those of the Keweenawan Track.
-
-% \cite{Roy1978a} has done that but without the effort of distinguishing the detrital remanence and the chemical rmanence.
-% In addition, constraining the pole position of the
-
-% and shed new lights on reconciling the connection between the Keweenawan Track and the Grenville Loop paleomagnetic apparent polar wander paths.
 
@@
 
-% In this study, we use fine-grained to silt-sized, hematite rich red beds from xxx sections along the Keweenaw Peninsula, Michigan with the goal of obtaining high-quality paleomagnetic pole from the jacobsville formation that is corrected for inclination shallowing problem and to investigate the connection between the Keweenawan Track and the Grenville Loop of the Laurentia APWP during the late Mesoproterozoic to the Neoproterozoic.
-
 \section*{Geologic settings}
@@
 The ca.
-1109 Ma to 1084 Ma North American Midcontinent Rift is a major intracontinental rift system that extends over 2000 km through Laurentia craton (Figure \ref{fig:Geologic_map}).
+1109 Ma to 1084 Ma North American Midcontinent Rift is a major intracontinental rift system that extends over 2000 km through the Laurentia craton (Fig.
+\ref{fig:Geologic_map}).
 However, the rifting failed to separate Laurentia into two continents.
 As its magmatism waned and thermal subsidence progressed in the rift basin, the $>$4 km of ca.
-1080 Ma to 1050 Ma Oronto Group sediments \cite{Swanson-Hysell2019a, Fuentes2022a} were deposited conformably on top of the rift volcanics (Fig.
-\ref{fig:Geologic_map}; \cite{Cannon1992a}).
-Subsequently, the rift was inverted as the Grenvillian Orogeny commenced on the eastern margin of Laurentia (Figure \ref{fig:Geologic_map}; \citeA{Cannon1994a}), resulting in the propagation of far-field compressional stress deep into the interior of Laurentia.
-During the inversion, Paleoproterozoic and Archean crusts were uplifted via thrust faults such as the Marenisco fault and the Pelton Creek fault, forming the crustal-scale Montreal River monocline (Fig.
-\ref{fig:Geologic_map}; \cite{Cannon1993a}).
-The post-rift Jacobsville Formation overlies an erosional angular unconformity that developed on lithologies that were exhumed through this earlier reverse motion (Fig.
-\ref{fig:Geologic_map}; \citeA{Cannon1993a, Kalliokoski1982a}).
-The erosional contact and the development of paleosol between the Jacobsville and the underlying basement rocks is consistent with the interpretation that there was a prolonged period of sedimentation quiescence and net erosion between the formation of the Oronto Group sediments and the Jacobsville Formation.
-The overall extent of the Jacobsville Formation has been mapped along the lake shore of the Keweenaw Peninsula and is interpreted to extend eastward to Sault Ste Marie, Canada (Fig.
+1085-1050 Ma Oronto Group sediments \cite{Fairchild2017a, Swanson-Hysell2019a} were deposited conformably on top of the rift volcanics (Fig.
+\ref{fig:Geologic_map}; \citeA{Cannon1992a}).
+Subsequently, the rift was inverted as the ca.
+1090-980 Ma Grenvillian Orogeny commenced on the eastern margin of Laurentia (Fig.
+\ref{fig:Geologic_map}), eventually resulting in the propagation of far-field compressional stress deep into the interior of Laurentia \cite{Cannon1993a, Cannon1994a}.
+
+During the inversion, Paleoproterozoic and Archean crust was uplifted via thrust faults such as the Marenisco fault and the Pelton Creek fault, forming the crustal-scale Montreal River monocline (Fig.
+\ref{fig:Geologic_map}; \citeA{Cannon1993a}).
+The southern part of the post-rift Jacobsville Formation overlies an erosional angular unconformity that developed on lithologies that were exhumed through this earlier reverse motion (Fig.
+\ref{fig:Geologic_map}; \citeA{Hamblin1958a, Cannon1993a, Kalliokoski1982a}).
+More recent mapping of the Jacobsville Formation has also found that it can sometimes onlap the Midcontinent Rift volcanics within or adjacent to the Keweenaw Fault system around Bete Grise Bay on the Keweenaw Peninsula (Fig.
+\ref{fig:Geologic_map}; \citeA{Tyrrell2019a, Mueller2021a}).
+Field observations find erosional contact and development of paleosol between the Jacobsville Formation and underlying volcanic rocks \cite{Hamblin1958a, Kalliokoski1975a}, and interpretations of seismic reflection data show that sedimentary rocks interpreted to be Jacobsville Sandstone lie unconformably on the Oronto Group \cite{Cannon1989a}.
+These observations are consistent with the interpretation that there was a prolonged period of sedimentation quiescence and net erosion between the end of Oronto Group deposition and onset of Jacobsville Formation deposition.
+Based on dates from Cannon et al 1993 (and deformation of the Freda Fm within the Montreal River monocline) paired with new dates on the Jacobsville Fm (Hodgin et al., 2022), this time period of quiescence may have lasted up to ~60 Myrs from ca.
+1050 until ca.
+990 Ma.
+The overall extent of the Jacobsville Formation has been mapped along the lake shore of the Keweenaw Peninsula from northern Michigan, USA, to the north side of Michipicoten Island and perhaps to northern Sault Ste Marie, Canada (Fig.
 \ref{fig:Geologic_map}; \citeA{Hamblin1958a}).
 
-% The Jacobsville Sandstone is unconformably overlain by the Cambrian Munising Formation (Hamblin, 1958).
+As far-field stresses from the Grenvillian Orogeny propagated into the interior of Laurentia, the previously thinned crust of the Midcontinent Rift was structurally inverted by crustal-scale thrusts, such as the Keweenaw Fault which extends $\sim$250km from ...refs here.
+Final inversion of the Keweenaw Fault during the Grenvillian Orogeny thrust Midcontinent Rift volcanics over the Jacobsville strata, which were folded in the footwall of the fault.
+along much of the Keweenaw fault which extends $\sim$250 km from the tip of the Keweenaw Peninsula in northern Michigan southeastward to a termination in northeastern Wisconsin (Fig.
+\ref{fig:Geologic_map}; \citeA{Cannon1996a, Cannon2001a, DeGraff2022a}).
+Syndepositional faulting that took place during Jacobsville deposition is inferred from ...
+rift volcanic clasts where eroded into the Jacobsville basin at the same time when sediments from the older basement rocks along with distal input from the Grenville Province were deposited within the Jacobsville back-bulge basin system that in general wedges toward the northwest (Fig.
+\ref{fig:Geologic_map}; \cite{Hedgman2014a, Hodgin2022a}.
+Next to the footwall of the Keweenaw fault, folding has also been recorded by strata of the upper Jacobsville Formation in contact with the volcanics \cite{Cannon1993a}.
+One famous remnant of that late-stage deformation is the Natural Wall ravine where sub-vertical sandstone beds are preserved (Fig.
+\ref{fig:Geologic_map}; \cite{Hamblin1958a}).
+On the shore of Bete Grise Bay near the northeastern end of Keweenaw Peninsula, the Keweenaw fault is interpreted to have accommodated Grenville compressional stresses by forming a series of segmented en echelon thrusts \cite{Tyrrell2019a}.
+The timing of such faulting could be coeval with the deposition of the upper Jacobsville Formation, leading to the nonconformable onlapping of the Jacobsville sediments over the volcanics in addition to the thrust faulting relationship between the two units.
+Ca.
+985 Ma late-kinematic calcite that precipitated within the hanging wall of the Keweenaw Fault and ca.
+992 Ma maximum deposition age from detrital zircons of the Jacobsville Formation in the footwall constrain the timing of the final inversion of the Keweenaw Fault to be during the Rigolet phase of the Grenvillian Orogeny \cite{Hodgin2022a}.
 
-As the far-field stress of the Grenvillian Orogeny continued to reach the interior of Laurentia, it thrusted and folded the Midcontinent Rift volcanics over the Jacobsville strata along the major Keweenaw fault which extends $\sim$250 km from the tip of the Keweenaw Peninsula in northern Michigan southeastward to a termination in northeastern Wisconsin (Fig.
-\ref{fig:Geologic_map}; \citeA{Cannon1996a, Cannon2001a, DeGraff2022a}).
-As a result, rift volcanic clasts where eroded into the Jacobsville basin at the same time when sediments from the older basement rocks along with distal input from the Grenville Province were deposited within the Jacobsville back-bulge basin system that in general wedges toward the northwest (Figure \ref{fig:Geologic_map}; \cite{Hedgman2014a, Hodgin2022a}.
-Immediately near the Keweenaw fault in the footwall, folding has also been recorded by strata of the upper Jacobsville Formation in contact with the volcanics \cite{Cannon1993a}.
-One famous remanence of that late-stage deformation is the Natural Wall ravine where sub-vertical to overturned sandstone beds are preserved (Figure \ref{fig:Geologic_map}; \cite{Hamblin1958a}).
-On shore of Bete Grise Bay near the northeastern edge of the Keweenaw Peninsula, the Keweenaw fault is interpreted to have accommodated compressional stresses by forming a series of segmented en echelon thrusts \cite{Tyrrell2019a}.
-The timing of such en echelon faulting could be coeval with the deposition of the upper Jacobsville Formation, leading to the nonconformable onlapping of the Jacobsville sediments over the volcanics in addition to the thrust faulting relationship between the two units.
-Late-kinematic calcite that precipitated within the brecciated Keweenaw fault zone and youngest detrital zircons of the Jacobsville Formation constrain the timing of the final inversion of the Keweenaw Fault to be during the Rigolet phase of the Grenvillian Orogeny \cite{Hodgin2022a}.
-
-\subsection{General stratigraphy}
+\subsection{Sedimentology and general stratigraphy}
 
@@
 
-The Jacobsville Formation is a fluvial sequence of feldspathic and quartzose sandstones, conglomerates, siltstones, and shales, completely devoid of lava flows or cross-cutting dikes.
-It is in general more quartz-rich than the Oronto Group sediments.
+The Jacobsville Formation is a fluvial sequence of feldspathic and quartzose conglomerates, sandstones, siltstones, and shales completely devoid of lava flows or cross-cutting igneous dikes although cross-cutting clastic dikes have been observed.
 Abundant near the bottom section in the southern exposures of the Jacobsville Formation, conglomerate facies typically have provenance sourced from Paleoproterozoic lithologies such as vein quartz and iron formation clasts \cite{Hamblin1958a, Kalliokoski1982a}.
-The occurrence of these conglomerates in proximity to the Keweenaw, Marenisco, and Pelton Creek thrust faults is interpreted to represent syn-depositional development of local relief \cite{Kalliokoski1982a, Hedgman1992a}).
+The occurrence of these conglomerates in proximity to the Keweenaw, Marenisco, and Pelton Creek thrust faults is interpreted to represent syn-depositional development of local relief \cite{Kalliokoski1982a, Hedgman2014a}).
+It is in general more quartz-rich than the Oronto Group sediments \cite{Hamblin1958a}.
 
-Near the Keweenaw Fault (Fig.
-\ref{fig:Geologic_map}), the Jacobsville Formation consists dominantly of quartz-rich, cross-bedded, medium-grained sandstone, locally abundant clast-supported conglomerate, with interbeds of hematite-bearing micaceous siltstone (Fig.
+Close to the Keweenaw Fault in central Keweenaw Peninsula (Fig.
+\ref{fig:Geologic_map}), the Jacobsville Formation consists dominantly of quartz-rich, cross-bedded, medium-grained sandstone, with locally abundant clast-supported pebble to cobble conglomerate, and interbeds of red, hematite-bearing, micaceous shaly siltstone to fine-grained sandstone (Hungarian Falls, Hammel Creek, Natural Wall; Fig.
 \ref{fig:strat_column}).
 In this region, the conglomerate contains abundant volcanic clasts likely derived from uplifted Midcontinent Rift volcanics \cite{Irving1885a, Brojanigo1984a}.
-Occurrence of red siltstone beds have been reported in the Jacobsville Formation, particularly in the upper sections near Keweenaw Fault \cite<e.g.>{Hamblin1958a, Irving1885a}.
-In the field, they typically appear brick-red to dark-red, consisting of find-grained arenite to arkosic sandstone and siltstone sometimes with minor grey reduction spots (Fig.
+This clast-supported lithofacies can include siltstone rip-up clasts or channelized lenses of siltstones.
+Occurrence of red siltstone to sandstone beds are also commonly interbedded with sandstone and conglomerate \cite<e.g.>{Hamblin1958a, Irving1885a}.
+They typically appear brick-red to dark-red, consisting of find-grained arenite to arkosic sandstone and siltstone sometimes with minor grey reduction spots (Fig.
 \ref{fig:Field_photo}).
-Rounded detrital mica grains up to $\sim$5mm in size can be found commonly within very fine, fissile siltstone and shaly strata (Figure \ref{fig:Field_photo}).
-The fact that in the Jacobsville has conglomerate and sandstone and siltstone facies in them supports sediment input from the eroded Oronto Group.
+Rounded detrital mica grains up to $\sim$5mm in size can be found commonly within very fine, fissile siltstone and shaly strata (Fig.
+\ref{fig:Field_photo}).
+
+The exact stratigraphic association between sections of the Jacobsville Formation in the central Keweenaw Peninsula and those exposed at Agate Falls and Baby Snake Creek is difficult to constrain as the thickness of the Jacobsville strata can vary significantly through the peninsula \cite{Hamblin1958a, Kalliokoski1982a} and are not fully exposed at or in between the studied localities.
+As a result, various interpretations could exist for the relative stratigraphy at different localities.
+Based on the relative lack of conglomerate facies at Agate Falls and at Baby Snake Creek, and field observations of onlapping Jacobsville Formation on top of previously thrusted Midcontinent Rift volcanics, we interpret that Jacobsville Formation exposed at Agate Falls and Baby Snake Creek could be stratigraphically higher than the other three sections near central Keweenaw Peninsula (Fig.
+\ref{fig:Geologic_map}, \ref{fig:strat_column}).
+Further discussion regarding the stratigraphy is presented in section \ref{magstrat}.
+
+% Near the central part of Keweenaw Peninsula, we visited three adjacent Jacobsville stratigraphic sections at Hammel Creek, Hungarian Falls, and Natural Wall ravine (Fig.
+\ref{fig:Geologic_map}, \ref{fig:strat_column}).
+The exposed Jacobsville Formation share similarities in that they are all close to the Keweenaw Fault where the formation is steeply tilted in the footwall against the folded Midcontinent Rift volcanics in the hanging wall.
+They are also of similar thickness and are characterized by the interbedded layers of pebble conglomerate, massive cross-bedded sandstone and minor amount of siltstone to fine-grained sandstone (Fig.
+\ref{fig:strat_column}).
+
+% The fact that in the Jacobsville has conglomerate and sandstone and siltstone facies in them supports sediment input from the eroded Oronto Group.
 This indicate the older boundary of the Jacobsville deposition might be closer to the Rigolet Phase than the Ottawan phase and the duration of Jacobsville sedimentation may be within a relatively constrained period of time.
 
-tyrrell 2019 fig. 10 recorded onlapping of the Jacobsville Formation on to the Portage Lake Volcanics which appears to have been faulted prior to the deposition of Jacobsville near Bete Grise Bay.
-
-% Hematite is one of the most stable form of iron oxides under ambient conditions, and magnetic remanence carried by the grains within red beds can be preserved for billions of years \cite{Butler1992a, Jiang2022a}.
-Previously, paleomagnetic records of the Jacobsville Formation have been investigated by \citeA{Roy1978a} who sampled a range of lithofacies of the Jacobsville Formation and obtained that a primary remanence component can be isolated via magnetic cleaning.
-In this study, we target the find-grained sandstone to siltstone facies in the upper Jacobsville Formation and segregate the primary and secondary remanent magnetizations using high-resolution thermal demagnetization method \cite{Swanson-Hysell2019b}.
-We update the paleomagnetic records of the Jacobsville Formation with additional evaluation for the degree inclination shallowing that \citeA{Roy1978a} did not take into account.
+% tyrrell 2019 fig. 10 recorded onlapping of the Jacobsville Formation on to the Portage Lake Volcanics which appears to have been faulted prior to the deposition of Jacobsville near Bete Grise Bay.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{strat_column.png}
-\caption{}
+\includegraphics[width=0.9\textwidth]{Jacobsville_Sections_v3.pdf}
+\caption{Tentative sedimentary stratigraphy and magnetostratigraphy for the Jacobsville Formation in northern Michigan, USA.
+Stratigraphic locations of paleomagnetic samples are shown in context of Jacobsville lithofacies, sedimentary structures, as well as detrital zircon maximum deposition ages developed by \citeA{Hodgin2022a}.
+}
 \label{fig:strat_column}
@@
 
-Keweenaw Fault motion post dating Jacobsville in the footwall
-Youngest DZ from jacobsvill eis about 993 Ma (sandstone creek, about a few hundred meters from the fault). 
-directly above pmag sampling at Agate falls grain B geochron date is (z2, s226, about 1000 Ma grain)
+% Keweenaw Fault motion post dating Jacobsville in the footwall
+% Youngest DZ from jacobsvill eis about 993 Ma (sandstone creek, about a few hundred meters from the fault). 
+% directly above pmag sampling at Agate falls grain B geochron date is (z2, s226, about 1000 Ma grain)
 
-jacobsville (985.5 from calcite in the fault and <993 from zircons)
+% jacobsville (985.5 from calcite in the fault and <993 from zircons)
 
-Blake's current tectonic model: after MCR volcanics ca.
+% Blake's current tectonic model: after MCR volcanics ca.
 1084 Ma emplacement follows the deposition of the sediments (Oronto Group) copper harbor conglomerate, Nonesuch formation and the Freda formation.
@@
 
-This makes the Jacobsville sandstone one unique opportunity to acquire a paleomagnetic pole position within the gap between the
+% This makes the Jacobsville sandstone one unique opportunity to acquire a paleomagnetic pole position within the gap between the
 
@@
 \ref{fig:Geologic_map}).
-Additional oriented cores and block samples of siltstone rip-up clasts within conglomerate facies were sampled at Hungarian Falls and at Agate Falls.
-In order to maximize sampling of paleosecular variation, we optimized for vertical stratigraphic coverage and collected one sample at each red bed horizon.
-As such, each sample constitutes a paleomagnetic site considering that a paleomagnetic site (which ideally captures a single snapshot of the local geomagnetic field) is a particular bed in a sedimentary sequence.
-Red fine-grained sandstone to shale layers were preferentially sampled as they have lower permeability and are less susceptible to diagenetic alteration through fluid flow than coarser grained sandstone.
+Additional oriented cores and block samples of siltstone rip-up clasts within conglomerate lithofacies were sampled at Hungarian Falls and at Agate Falls (Fig.
+\ref{fig:Geologic_map}, \ref{fig:strat_column}).
+In order to maximize sampling of distinct time snapshots of the geomagnetic field at the time of Jacobsville deposition such that paleosecular variation can be sufficiently averaged out, we optimized for vertical stratigraphic coverage such that each sample constitutes a paleomagnetic site considering that a paleomagnetic site (which ideally captures a single snapshot of the local geomagnetic field) is a particular bed in a sedimentary sequence.
+Red fine-grained sandstone to shaly siltstone layers were preferentially sampled as they tend to have lower permeability and are less susceptible to diagenetic alteration through fluid flow than coarser grained sandstone.
 Care was taken to avoid samples containing reduction spots.
-Paleomagnetic samples were oriented using a magnetic compass and a sun compass whenever possible.
+Paleomagnetic cores and blocks were oriented using a magnetic compass and a sun compass whenever possible.
 Sun compass data were preferentially used when available.
 A total of 379 specimens including 30 intraclasts where collected for paleomagnetic study.
-The specimens underwent stepwise thermal demagnetization in the UC Berkeley Paleomagnetism Lab using an ASC demagnetizer (residual fields $<$10 nT) with measurements made on a 2G DC-SQUID magnetometer.
-The demagnetization protocol had high-resolution approaching the Ne\'el temperature of hematite (5 \textdegree C to 2 \textdegreeC) (Fig.
+The specimens underwent stepwise thermal demagnetization at the UC Berkeley Paleomagnetism Lab using an ASC demagnetizer (residual fields $<$10 nT) with measurements made on a 2G DC-SQUID magnetometer.
+The demagnetization protocol had high-resolution approaching the Ne\'el temperature of hematite (5 \textdegree C to 2 \textdegree C) (Fig.
 \ref{fig:Intraclast_pmag}).
+Demagnetization results are plotted in Figure \ref{fig:in_situ_pmag}A by locality.
 All paleomagnetic data are available to the measurement level in the MagIC database (https://earthref.org/ MagIC/; THIS LINK IS AVAILABLE TO REVIEWERS AND WILL BE UPDATED WHEN A DOI IS AVAILABLE).
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{in_situ_pmag.pdf}
+\caption{}
+\label{fig:in_situ_pmag}
+\end{figure*}
 
@@
 
-Similar to the thermal demagnetization result of the hematite-bearing fluvial intraclasts of the Freda Formation \cite{Swanson-Hysell2019b}, the siltstone rip-up clasts of the Jacobsville Formation also typically reveal two distinct magnetization directions (Fig.
+Similar to the thermal demagnetization result of the hematite-bearing fluvial intraclasts of the Freda Formation \cite{Swanson-Hysell2019b}, the siltstone rip-up clasts of the Jacobsville Formation typically reveal two distinct magnetization components (Fig.
 \ref{fig:Intraclast_pmag}).
-One direction was similar throughout the intraclasts and was typically removed up to 640-655 \textdegree C (Fig.
+One component show similar vector orientations amongst intraclasts and was typically removed up to 640-655 \textdegree C (Fig.
 \ref{fig:Intraclast_pmag}).
-This component is directionally well-grouped amongst the reoriented intraclasts, indicating that it is likely a chemical remanent magnetization acquired through the growth of secondary hematite grains following redeposition of the clasts (Fig.
+This indicates that it is likely a chemical remanent magnetization acquired through growth of secondary hematite in pore spaces following redeposition of the clasts.
+After removal of this component, further thermal demagnetization with small step increments at higher temperatures up to $\sim$689 \textdegree C often reveal an origin-trending component (Fig.
 \ref{fig:Intraclast_pmag}).
-After the removal of the mid-temperature component, further thermal demagnetization with small step increments at higher temperatures up to $\sim$689 \textdegree C often reveal an origin-trending component (Fig.
-\ref{fig:Intraclast_pmag}).
-In the data from the clasts, there is typically a significant directional change in the specimen magnetization between the mid-temperature component and the high-temperature component and as a result, 28 out of 30 intraclast specimens could be fit with distinct mid-temperature and high-temperature least squares lines \cite{Kirschvink1980a}.
+In the data from the clasts, there is typically a significant directional change in specimen magnetization between the mid-temperature component and the high-temperature component and as a result, 28 out of 30 intraclast specimens could be fit with distinct mid-temperature and high-temperature least squares lines \cite{Kirschvink1980a}.
 Distinct from the well-grouped mid-temperature component directions, the high-temperature directions of the intraclasts are dispersed and the null hypothesis of randomness cannot be rejected at the 95\% confidence level (n=6 for intraclasts at Agate Falls and n=22 for intraclasts at Hungarian Falls; Fig.
 \ref{fig:Intraclast_pmag}; \citeA{Watson1956a}).
-This indicates that the component was likely acquired by the detrital hematite grains during initial formation of the clasts prior to erosion and redeposition.
+This indicates that the component is held by detrital hematite grains during initial formation of the clasts prior to erosion and redeposition.
 
@@
 \includegraphics[width=\textwidth]{Intraclast_pmag.pdf}
-\caption{Paleomagnetic data from intraclasts of the Jacobsville Formation reveal a mid-temperature component that typically unblocks prior to 655 \textdegree C and a high-temperature component that typically unblocks up to 687 \textdegree C.
-These components are present as varying fractions of the overall remanence as seen in the three individual clasts shown here on vector orthogonal plots and measurement-level equal area plots in tilt-corrected coordinates (developed using PmagPy; \citeA{Tauxe2016a}).
+\caption{Paleomagnetic data from fluvial rip-up clasts of the Jacobsville Formation reveal a mid-temperature component that typically unblocks prior to 655 \textdegree C and a high-temperature component that typically unblocks up to 689 \textdegree C.
+The components are present as varying fractions of the overall remanence as seen in the three individual clasts shown on vector orthogonal plots and measurement-level equal area plots in tilt-corrected coordinates (developed using PmagPy; \citeA{Tauxe2016a}).
 The direction of the mid-temperature component is shown as green arrows on the vector component plots and green circles on the equal area plots, while the high-temperature component is shown with purple symbols.
-The mid-temperature component has a similar direction among the clasts as can be seen on the component directions equal area plots.
-In contrast, the high-temperature component directions (purple) are dispersed.
+The mid-temperature component has a similar direction among the clasts as can be seen on the summary equal area plots.
+In contrast, the high-temperature component directions (purple) are dispersed and pass a randomness test of \citeA{Watson1956a}.
 NRM = natural remanent magnetization; AFC = Agate Falls clasts; HFC = Hungarian Falls clasts.}
@@
 \ref{fig:Geologic_map}, \ref{fig:in_situ_pmag}).
-The remanence directions recorded by red beds within the steeply folded basal section and within the flat-lying top portion of the section pass a bootstrap fold test (Supporting Information; \citeA{Tauxe1994a}).
-The remanence directions are all of normal polarity and show a dominantly southwest declination and upward-pointing inclination (Fig.
+The tilt-corrected remanence directions are all of normal polarity and show a dominantly southwest declination and upward-pointing inclination (Fig.
 \ref{fig:in_situ_pmag}).
+The directions in geographic coordinates yield a bootstrap fold test \cite{Tauxe1994a} result that shows the 95\% confidence bounds for the best grouping of the directions from the two limbs of the fold is achieved between 55\% and 92\% unfolding.
+Although the 95\% confidence bounds exclude 100\% unfolding, that the grouping of the directions become much tighter than they were in geographic coordinates when significant unfolding is applied suggest the high-temperature component from Baby Snake Creek was likely acquired pre-folding.
+The non-ideal unfolding test could be a result of overlapping unblocking temperatures of detrital and pigmentary hematite in red beds of Baby Snake Creek or complex folding regime near the Keweenaw Fault or a combination of the two.
+
+\begin{SCfigure*}
+\centering
+\includegraphics[width=0.5\textwidth]{SC1_fold_test.pdf}
+\caption{Bootstrap paleomagnetic fold test \cite{Tauxe1994a} of the detrital remanent magnetization remanence directions recorded by specimens from Baby Snake Creek.
+Although the directional data set becomes more tightly grouped than in geographic coordinates when applying 55\%-92\% unfolding, both geographic and 100\% unfolded coordinate systems are excluded at the 95\% confidence level.
+}
+\label{fig:Intraclast_pmag}
+\end{SCfigure*}
 
 \subsection*{Tentative magnetostratigraphy}
-
+\label{magstrat}
 With the insights gained from the thermal demagnetization results of the Jacobsville fluvial intraclasts, we next isolate the high-temperature magnetization component from the mid-temperature component for all in situ specimens that were collected along the five localities in northern Michigan (Fig.
@@
 At Hungarian Falls (Fig.
-\ref{fig:Geologic_map}), high-energy fluvial environment led to the deposition of numerous hematite-rich siltstone to fine-grained sandstone layers, capturing richer information of the geomagnetic field variations than at other localities (Fig.
+\ref{fig:Geologic_map}), numerous hematite-rich siltstone to fine-grained silty sandstone layers were deposited.
+The siltstone and silty fine-grained sandstone layers can capture a relatively rich set of information from the paleomagnetic field variations compared to coarser-grained lithologies at other localities (Fig.
 \ref{fig:strat_column}).
@@
 
-The Natural Wall ravine and Hammel Creek are of close proximity from Hungarian Falls (Fig.
-\ref{fig:Geologic_map}) and the sedimentary facies characterized by trough cross-bedded sandstone and conglomerates are similar to the basal $\sim$120 meter of the section at Hungarian Falls (Fig.
+The Natural Wall ravine and Hammell Creek are in close proximity toHungarian Falls (Fig.
+\ref{fig:Geologic_map}) and the sedimentary facies characterized by trough cross-bedded sandstone and conglomerates are similar to the basal $\sim$120 meters of the section at Hungarian Falls (Fig.
 \ref{fig:strat_column}).
@@
 \ref{fig:strat_column}).
-At Hammel Creek, 14 samples from two siltstone layers near the base of the section are of reversed polarities, consistent with the reversed-polarity layers near the base of Hungarian Falls.
+At Hammell Creek, 14 samples from two siltstone layers near the base of the section are of reversed polarities, consistent with the reversed-polarity layers near the base of Hungarian Falls.
 Near the top of the Natural Wall section, consistent normal-polarity directions are recorded by 15 samples.
@@
 
-Given that multiple geomagnetic field reversals occurred during the deposition of the Jacobsville Formation, it is desirable that a magnetostratigraphy be developed.
+Given that multiple geomagnetic field reversals occurred during the deposition of the Jacobsville Formation, the development of amagnetostratigraphy should be possible and could be useful for future regional and global studies.
 However, challenges exist in correlating the stratigraphic sections measured at the five localities in this study due to the lateral variability of the paleodepositional environment associated with the variable paleotopographic relief of the Precambrian surface upon which the Jacobsville Formation was deposited \cite{Hamblin1958a, Kalliokoski1982a}.
-In particular, there may not be a unique solution in reconstructing the relative stratigraphic positions of the strata at Baby Snake Creek and Agate Falls with respect to the other three sections in central Keweenaw Peninsula due to geographical separations (Fig.
+In particular, there may not be a unique solution in reconstructing the relative stratigraphic positions of the strata at Baby Snake Creek and Agate Falls with respect to the other three sections in central Keweenaw Peninsula due to geographical separation (Fig.
 \ref{fig:Geologic_map}, \ref{fig:strat_column}).
@@
 992 Ma (maximum deposition age of Jacobsville Formation; \cite{Hodgin2022a}).
-
-\begin{figure*}[h!]
-\centering
-\includegraphics[width=\textwidth]{in_situ_pmag.pdf}
-\caption{}
-\label{fig:in_situ_pmag}
-\end{figure*}
 
@@
 \section*{Discussion}
+
 \subsection*{Inclination shallowing correction}
@@
 980 Ma Rigolet phase of the Grenvillian Orogeny whereas the majority of the fault motion likely occurred during the ca.
-1050 Ma Ottawan phase \cite{Hodgin2022b}.
-We interpret that it is most likely that deposition of the Jacobsville Formation initiated close to the Rigolet phase of the Grenvillian Orogeny with progressive sediment transportation to the northwest as the Grenville back-bulge basin propagated toward the interior of Laurentia \cite{Hodgin2022b}.
+1050 Ma Ottawan phase \cite{Hodgin2022a}.
+We interpret that it is most likely that deposition of the Jacobsville Formation initiated close to the Rigolet phase of the Grenvillian Orogeny with progressive sediment transportation to the northwest as the Grenville back-bulge basin propagated toward the interior of Laurentia \cite{Hodgin2022a}.
 As a result, we hypothesize that the sediments at Baby Snake Creek likely records a slightly younger pole position with a more southerly paleolatitude than those near Hungarian Falls, indicating Laurentia's plate motion toward the south (Fig.
@@
 The asymmetric uncertainties in the mean pole position thus violate the assumption of the Fisher distribution \cite{Fisher1953a} where uncertainties are circularly distributed about a mean vector.
-Instead, the spherical asymmetric bivariate Kent distribution \cite{Kent1983a} is more appropriate for incorporating uncertainties associated with inclination shallowing in reported sedimentary paleomagnetic poles.
+Instead, the spherical asymmetric bivariate Kent distribution \cite{Kent1982a} is more appropriate for incorporating uncertainties associated with inclination shallowing in reported sedimentary paleomagnetic poles.
 We follow the approach of \citeA{Pierce2022a} who extended the E/I method of \citeA{Tauxe2004b}.
@@
 Given the new constraints on the age and position of the Jacobsville pole (Fig.
-\ref{fig:pole_plot}), the two plate model would require that the majority of the crustal shortening happened after the Rigolet phase of the Grenville Orogeny, as the ca.992 Ma Jacobsville pole is still $\sim$45\textdegree arc distance ($\sim$5000 km) away from the paleomagnetic pole of the Haliburton intrusions \cite{Buchan1976a} and $\sim$38\textdegree ($\sim$4000 km) from the pole of the Wilberforce intrusions \cite{Palmer1973a} (Fig, \ref{fig:pole_plot}).
+\ref{fig:pole_plot}), the two plate model would require that the majority of the crustal shortening happened after the Rigolet phase of the Grenville Orogeny, as the ca.992 Ma Jacobsville pole is still $\sim$45\textdegree\ arc distance ($\sim$5000 km) away from the paleomagnetic pole of the Haliburton intrusions \cite{Buchan1976a} and $\sim$38\textdegree\ ($\sim$4000 km) from the pole of the Wilberforce intrusions \cite{Palmer1973a} (Fig, \ref{fig:pole_plot}).
 However, geologic and geochronologic evidence from both within the Grenville Province \cite{Aleinikoff2021a} and from interior Laurentia \cite{Hodgin2022a} indicate that majority of the deformation in the Grenville Province and the propagation of its associated deformation in the Midcontinent Rift happened during the ca.1090-1020 Ma Ottawan phase orogeny, whereas smaller amount of convergence happened during the ca.
@@
 Previous paleomagnetic studies have found that Grenville rocks commonly have magnetite and hematite being the remanence carriers \cite<e.g.>{Ueno1975a, Alvarez1998a, Buchan1973a, Buchan1976a, Brown2012a}.
-Theoretical calculations have shown that magnetite which formed above its Curie temperature (580\textdegree C) in Grenville rocks subsequently acquired thermoviscous remanence \cite{Pullaiah1975a, Dodson1980a} during prolonged cooling of the orogen, whereas hematite which often occur as exsolved lamellae interfingering with ilmenite acquired thermochemical remanence via magnetic ordering of the hemo-ilmenite below the Neel temperature of hematite \cite{Burton1985a, Harrison2006a}.
+Theoretical calculations have shown that magnetite which formed above its Curie temperature (580\textdegree C) in Grenville rocks subsequently acquired thermoviscous remanence \cite{Pullaiah1975a, Dodson1980a} during prolonged cooling of the orogen, whereas hematite which often occur as exsolved lamellae interfingering with ilmenite acquired thermochemical remanence via magnetic ordering of the hemo-ilmenite below the Neel temperature of hematite \cite{Burton1985a, Harrison2006b}.
 
@@
 
-That the estimated ages associated with the Adirondack paleomagnetic poles are biased old can be attributed to the outdated thermochronometer closure temperature calculations of \citeA{Mezger1991a, 1992a} whose data was used by \citeA{Brown2012a}.
-In particular, the cooling history for the Adirondack area estimated by \citeA{Mezger1991a, 1992a} show that the orogen had cooled to $\sim$400\textdegree C by the time of U-Pb system closure in rutile crystals.
+That the estimated ages associated with the Adirondack paleomagnetic poles are biased old can be attributed to the outdated thermochronometer closure temperature calculations of \citeA{Mezger1991a, Mezger1992a} whose data was used by \citeA{Brown2012a}.
+In particular, the cooling history for the Adirondack area estimated by \citeA{Mezger1991a, Mezger1992a} show that the orogen had cooled to $\sim$400\textdegree C by the time of U-Pb system closure in rutile crystals.
 However, this estimated U-Pb closure temperatures range for rutile was not based on laboratory Pb diffusion experiment, but instead was an assumption that diffusion rate of Pb in rutile approximate that of Fe and Ti \cite{Mezger1989a}.

### 2022-07-23 Manuscript/Manuscript.tex

@@
 
-\title{Laurentia's continued motion into the Neoproterozoic: new paleomagnetic constraint from the Jacobsville Formation}
+\title{Laurentia in the earliest Neoproterozoic: new paleomagnetic constraint from the Jacobsville Formation}
 
-\authors{Yiming Zhang\textsuperscript{1}, Blake Hodgin\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, James Pierce\textsuperscript{1}, Anthony Fuentes\textsuperscript{1}}
+\authors{Yiming Zhang\textsuperscript{1}, Blake Hodgin\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, James Pierce\textsuperscript{1,2}, Anthony Fuentes\textsuperscript{1}}
 
 \affiliation{1}{Department of Earth and Planetary Science, University of California, Berkeley, CA, USA}
+\affiliation{2}{Department of Earth and Planetary Sciences, Yale University, CT, USA}
 
@@
 \begin{keypoints}
-\item A new paleomagnetic pole from the ca. 992 Ma Jacobsville Formation 
-\item Reversals put younger bounds on the normal Keweenawan superchron 
-\item The new pole indicates the Grenville Loop could be younger than thought
+\item A new paleomagnetic pole from the ca. 990 Ma Jacobsville Formation \item Laurentia crossed the equator in the latest Mesoproterozoic to earliest Neoproterozoic
+\item The new pole indicates the Grenville Loop could be much younger than thought
 \end{keypoints}
@@
 \begin{abstract}
-The ca.
-1.1 Ga North American Midcontinent Rift resulted in protracted magmatism and associated deposition of sediments during rift-related extension and post-rift thermal subsidence.
-Coeval with the cessation of rifting and a decrease of Laurentia's plate speed is the onset of the ca.
-1090-980 Ma Grenvillian Orogeny along the eastern margin.
-The far-field stress of the orogeny is interpreted to have propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation whose maximum depositional age has been constrained to be ca.
-992 Ma.
-We develop a new paleomagnetic pole from red beds of the Jacobsville Formation along five stratigraphic sections in northern Michigan, USA.
-High-resolution thermal demagnetization experiments resolve detrital magnetic remanence that pass a fold test and two conglomerate tests.
-Our results confirm the occurrence of geomagnetic field reversals during the deposition of Jacobsville Formation, providing constraints for the maximum duration of the Keweenawan superchron.
-An inclination corrected paleomagnetic pole position from the Jacobsville Formation confirm continued plate motion of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic.
-The new pole from the interior of Laurentia indicates the paleomagnetic poles developed from the Grenville Province most likely record coherent plate motion after major continental collision during the development of the supercontinent Rodinia.
-% rather than recording a significant crustal shortening.
+Coeval with the cessation of the ca.
+1.1 Ga intracontinental Midcontinent Rift was the onset of the ca.
+1090-980 Ma Grenvillian Orogeny along Laurentia's eastern margin.
+Far-field compressional stresses of the orogeny propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation whose maximum depositional age has been constrained to be ca.
+990 Ma.
+We develop a new paleomagnetic pole from red beds of the Jacobsville Formation in northern Michigan, USA.
+High-resolution thermal demagnetization experiments resolve detrital remanent magnetizations that pass a conglomerate test.
+An inclination corrected paleomagnetic mean pole position from the Jacobsville Formation confirm continued plate motion of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic.
+The new pole from the interior of Laurentia indicates the paleomagnetic poles developed from the Grenville Province most likely record coherent plate motion after major continental collision during the initial assembly of the supercontinent Rodinia.
 \end{abstract}
@@
 \section*{Keywords}
-Neoproterozoic\\
+Rodinia\\
 Laurentia  \\
 Paleogeography\\
+Neoproterozoic\\
 \
@@
 1109 Ma to 1084 Ma North American Midcontinent Rift (i.e.
-the Keweenawan Track) provide crucial constraints on global paleogeography in the Mesoproterozoic \cite{Swanson-Hysell2019a}.
-As a result, the Keweenawan Track has become a central record for reconstructing the development of the ancient supercontinent Rodinia \cite{2021a}.
+the Keweenawan Track) provide crucial constraints on the paleogeography of Laurentia in the Mesoproterozoic \cite{Swanson-Hysell2019a}.
+As a result, the Keweenawan Track has become a central record for reconstructing the initial development of the ancient supercontinent Rodinia \cite{2021a}.
 As the rifting ceased and subsequently inverted following the onset of the Grenvillian Orogeny on Laurentia's eastern margin (Fig.
@@
 1070 Ma upper Freda Formation of the Oronto Group sediments that overlie the Midcontinent Rift volcanics in southern Lake Superior area (Fig.
-\ref{fig:Geologic_map}) provide crucial constraints for Laurentia's latitude, orientation, and plate motion in the late Mesoproterozoic \cite{Henry1977a, Swanson-Hysell2019a, Swanson-Hysell2019b}.
+\ref{fig:Geologic_map}) provided crucial constraints for Laurentia's paleolatitude, orientation, and plate speed in the late Mesoproterozoic \cite{Henry1977a, Swanson-Hysell2019a, Rose2022a}.
 
@@
 \ref{fig:Geologic_map}; \citeA{Hamblin1958a, Hodgin2022a, DeGraff2022a}).
-The Jacobsville Formation also contains abundant red beds which offer the opportunity to further extend the reconstruction of the apparent polar wander path of Laurentia toward more recent times \cite{Hamblin1958a, Roy1978a}.
-U-Pb detrital zircon ages and U-Pb calcite ages from late-kinematic veins constrain the maximum depositional age of the Jacobsville Formation to be ca.
-992 Ma \cite{Hodgin2022a}.
-Developing paleomagnetic data from the Jacobsville Formation will allow further development of Laurentia's paleogeographic evolution in the late Mesoproterozoic to early Neoproterozoic.
+U-Pb detrital zircon ages and U-Pb calcite ages from the Keweenaw fault where the Jacobsville Formation is deformed against thrusted rift volcanics constrain the maximum depositional age of the Jacobsville Formation to be ca.
+990 Ma \cite{Hodgin2022a}.
+The Jacobsville Formation contains abundant hematite-rich red beds which bear potential to further extend the apparent polar wander path (APWP) of Laurentia toward more recent times \cite{Hamblin1958a, Roy1978a}.
+Developing paleomagnetic data from the Jacobsville Formation will allow further development of Rodinia's paleogeographic evolution in the late Mesoproterozoic to early Neoproterozoic.
 
 Challenges exist in interpreting sedimentary paleomagnetic records.
-Primary detrital remanence acquired during settling of magnetic particles in the water column can be masked as secondary magnetic minerals (e.g.
-pigmentary hematite) that precipitate and grow within pore spaces of host sedimentary rocks syn- to post-deposition.
-In red beds, it has been found that formation of the pigmentary hematite can significantly post-date the timing of sediment deposition, resulting in the chemical remanent magnetization being secondary, superimposed upon the primary remanent magnetization carried by detrital grains \cite{Butler1992a}.
+Primary detrital remanence acquired through settling and deposition of magnetic particles in the water column can be masked by secondary remanence acquired through precipitation and growth of new minerals within pore spaces of host sedimentary rocks syn- to post-deposition (e.g.
+pigmentary hematite).
+In red beds, it has been found that formation of the pigmentary hematite can significantly post-date the timing of sediment deposition, resulting in a chemical remanent magnetization (CRM) superimposed upon primary detrital remanent magnetization (DRM) carried by detrital grains \cite{Butler1992a}.
 To isolate the remanence components of the red beds of the Jacobsville Formation, \citeA{Roy1974a, Roy1978a} adopted a method first applied by \citeA{Collinson1965c} to preferentially remove fine-grained pigmentary hematite through prolonged immersion in concentrated HCl acid whereas leaving the coarser detrital grains that are more resistant to dissolution in place.
 Studies applying paired acid etching demagnetization and thermal demagnetization show that the pigmentary hematite grains tend to have lower unblocking temperatures than detrital ones \cite{Tauxe1980a, Bilardello2010c}.
-High-resolution thermal demagnetization experiments with temperature intervals as small as 1-2 \textdegree C have shown to be effective in isolating detrital remanence from chemical remanence as coarser hematite grains tend to have higher unblocking temperatures \cite{Swanson-Hysell2019b}.
+High-resolution thermal demagnetization experiments with temperature intervals as small as 1-2 \textdegree C have shown to be effective in isolating detrital remanence from chemical remanence as coarser hematite grains tend to have higher unblocking temperatures closer to the Ne\'el temperature of hematite \cite{Swanson-Hysell2019b}.
 In particular, a conglomerate test on a suite of fluvial intraclasts of the red beds in the Freda Formation shows that via thermal demagnetization one can isolate detrital magnetization vectors (which are reoriented in random directions in the clasts) from the secondary components carried by co-occurring authigenic hematite (which yield similar directions amongst clasts) \cite{Swanson-Hysell2019b}.
 
-Paleomagnetism of red beds has also been challenged by the controversial issue of correcting for inclination shallowing \cite{King1955a, Tauxe2004b, Bilardello2016b}, which is necessary due to the rotation of detrital magnetic grains to be parallel to bedding caused by authigenic processes such as compaction during lithification.
-As a result, inclinations recorded by sediments often have shallower angles than the local field inclinations during the time of deposition.
+Paleomagnetism of red beds has also been challenged by the controversial issue of correcting for inclination shallowing \cite{King1955a, Tauxe2004b, Bilardello2016b}.
+Due to rotation of detrital magnetic grains caused by authigenic processes such as compaction during lithification, inclinations recorded by sediments often have shallower angles than the local field in which they are deposited.
 To estimate the amount of inclination shallowing in a given package of sedimentary rocks, \citeA{Tauxe2004b} developed a statistical model (i.e.
-the elongation/inclination method) that evaluates the deviation of the distribution of a large number ($>$100) of observed sedimentary paleomagnetic directional data with the predicted distribution given by a paleosecular variation model developed for the past 5 Myr (TK03) under the the geocentric axial dipole (GAD) hypothesis.
-Although the TK03 paleosecular variation model is based on data of relatively recent geomagnetic field, it has been shown to be compatible with data from the ca.
+the elongation/inclination [E/I]method) that evaluates the deviation of the distribution of a large number ($>$100) of observed sedimentary paleomagnetic directional data with the predicted distribution given by a paleosecular variation model developed for the past 5 Myr (TK03) under the the geocentric axial dipole (GAD) hypothesis.
+Although the TK03 paleosecular variation model is based on data of relatively recent geomagnetic field variations, it has been shown to be compatible with data from the ca.
 1.1 Ga lava flows of the Midcontinent Rift \cite{Tauxe2009a}.
-Further, \citeA{Pierce2022a} showed that the E/I method passes a field test for the 1.1 Ga Cut Face Creek Sandstone as the method successfully corrects the shallow paleomagnetic inclination of the sediments to that of the lava flows that bracket them.
+Further, \citeA{Pierce2022a} showed that the E/I method passes a field test for the 1.1 Ga Cut Face Creek Sandstone in the Midcontinent Rift as it successfully corrects the shallow paleomagnetic inclinations of the sediments to that of the North Shore Volcanic Group lava flows that bracket them.
 That study also further extended the method to better represent uncertainties associated with the amount of inclination shallowing when reporting sedimentary paleomagnetic pole positions by using the spherical elliptical Kent distribution \cite{Kent1982a} instead of the circular Fisher distribution \cite{Fisher1953a}.
 
-In this study, we investigate the hematite-bearing, fine-grained sandstone to siltstone lithofacies along five stratigraphic sections of the Jacobsville Formation.
-High-resolution thermal demagnetization data on siltstone intraclasts that were eroded by fluvial processes and redeposited at two locations constrain the timing of hematite acquisition by revealing a primary component that formed prior to the erosion of the clasts within the depositional environment and a secondary component that formed following their redeposition.
-We then use a paleomagnetic fold test to further constrain the timing of detrital remanence acquisition in association with the structural inversion of the Midcontinent Rift.
-Finally, we develop a new inclination-corrected paleomagnetic pole for the Jacobsville Formation based on a large number of samples and investigate the paleo-continental configuration of Laurentia and its association with the development of the supercontinent Rodinia.
+In this study, we investigate the hematite-bearing, fine-grained sandstone to siltstone lithofacies along five stratigraphic sections of the Jacobsville Formation on Keweenaw Peninsula, Michigan, USA.
+High-resolution thermal demagnetization isolate primary DRM from secondary CRM with the DRM passing a conglomerate test.
+We then use a paleomagnetic fold test to further constrain the timing of DRM acquisition in association with the structural inversion of the Midcontinent Rift.
+Finally, we develop a new inclination-corrected paleomagnetic pole for the Jacobsville Formation based on a large number of samples and investigate the paleo-continental configuration of Laurentia in the late Mesoproterozoic to early Neoproterozoic.
 
 % The end of the ca.
-1109 Ma to 1084 Ma intracontinental Midcontinent Rift magmatism and extension was marked by a transition to deposition of sedimentary rocks of the Oronto Group \cite{Ojakangas2001a} as the rift basin thermally subsided prior to the rift undergoing contractional deformation in the later stages of the Grenvillian orogeny \cite{Cannon1992b, Hodgin2022a}.
+1109 Ma to 1084 Ma intracontinental Midcontinent Rift magmatism and extension was marked by a transition to deposition of sedimentary rocks of the Oronto Group \cite{Ojakangas2001a} as the rift basin thermally subsided prior to the rift undergoing contractional deformation in the later stages of the Grenvillian orogeny \cite{Cannon1990b, Hodgin2022a}.
 
@@
 % A leading tectonic model for the cessation of the Midcontinent Rift and extension is that it occurred due to far-field compressional stress associated with the ca.
-1090-980 Ma Grenvillian orogeny \cite{Cannon1992a}, when the collision of Laurentia with its conjugate margins led to the development of a back-bulge basin, accommodating the deposition of the Jacobsville Formation sediments as the inversion of the Midcontinent Rift progressed.
+1090-980 Ma Grenvillian orogeny \cite{Cannon1990a}, when the collision of Laurentia with its conjugate margins led to the development of a back-bulge basin, accommodating the deposition of the Jacobsville Formation sediments as the inversion of the Midcontinent Rift progressed.
 Recent re-examination of the maximum deposition age of the Jacobsville Formation using zircon U-Pb whole grain chemical abrasion isotope dilution thermal ionization mass spectrometry has updated the maximum deposition age of the Jacobsville Formation to be ca.
@@
 
-% The extensive paleomagnetic studies of the well-preserved late Mesoproterozoic to Neoproterozoic rocks of North America have provided the central record for global paleogeography reconstruction during the assembly and rearrangement of the postulated supercontinent Rodinia.
-Paleomagnetic studies of rocks from Laurentia have led to a series of paleomagnetic poles that form an apparent polar wander path (APWP) known as the ``Logan Loop" for the older high latitude poles at its apex that continues into the ``Keweenawan Track" of younger lower latitude poles that form a progression as the APWP heads toward the ``Grenville Loop" \citep{Swanson-Hysell2019a}.
-In particular, recent advances in pairing high-precision zircon U-Pb geochronology with high-quality paleomagnetic poles significantly improved the resolution of the Keweenawan Track and revealed Laurentia's rapid motion ($>$20 cm/yr) during the Midcontinent Rift magmatic activity from ca.
-1110 Ma to ca.
-1085 Ma \citep{Swanson-Hysell2019a}.
-
 % As the Midcontinent Rift magmatic activity wanes its strength and evolved into a failed intracontinental rift, Laurentia experienced a long period of magmatic quiescence in its interior, where basin subsidence dominated the rift basin.
@@
 
-% Paleomagnetic poles from the interior of Laurentia during the quiescence magmatism period is thus crucial for reconcile the discordant paleomagnetic poles of the Keweenawan Track and the Grenville Loop.
-The
-% %Q: what are you using the poles from rocks that we know are from the Laurentia interior? (outside of the Grenville package?)
-
-% This new age constraint on the Jacobsville provides crucial insight into the tectonic evolution of the Midcontinent Rift and the inversion of the extensional region into a compressional region due to the Grenville orogen.
-It also makes the Jacobsville sandstone a unique opportunity to investigate Laurentia's paleogeographical position at ca.
-1 Ga as a long period of magmatic quiescence for $\sim$50 myr precedes Jacobsville deposition left a gap in Laurentia apparent polar wander path between the Keweenawan Track and the Grenville Loop (Fig.
-xxx).
-
-% Jacobsville sedimentation was preceded by a long period of volcanic and tectonic quiescence and cratonic stability so that bedrock surfaces became blanketed by paleosol and a surface of chemically resistant debris was dominated by quartz and iron-formation.
-
-% Previous paleomagnetic study on the red beds within the Jacobsville by \cite{Roy1978a} observed different components using AF, thermal, and chemical demagnetizations.
-Recognized dual polarities and recognized the complex remanence acquisition of red beds which lead to three isolated remanence components from the Jacobsville.
-However, that study did not recognize the problem of inclination shallowing and did not apply correction factor on the results.
-Moreover, the course demagnetization steps during thermal demagnetization would have had limited resolution for distinguishing the depositional remanent magnetization (DRM) and secondary chemical remanent magnetization (CRM) which have recently been shown to unblock over a narrow temperature window \cite{Swanson-Hysell2019b}.
-
 \section*{Geologic settings}
@@
 1085-1050 Ma Oronto Group sediments \cite{Fairchild2017a, Swanson-Hysell2019a} were deposited conformably on top of the rift volcanics (Fig.
-\ref{fig:Geologic_map}; \citeA{Cannon1992a}).
+\ref{fig:Geologic_map}; \citeA{Cannon1989a}).
 Subsequently, the rift was inverted as the ca.
 1090-980 Ma Grenvillian Orogeny commenced on the eastern margin of Laurentia (Fig.
-\ref{fig:Geologic_map}), eventually resulting in the propagation of far-field compressional stress deep into the interior of Laurentia \cite{Cannon1993a, Cannon1994a}.
+\ref{fig:Geologic_map}), eventually resulting in the propagation of far-field compressional stresses deep into the interior of Laurentia \cite{Cannon1993a, Cannon1994a}.
 
-During the inversion, Paleoproterozoic and Archean crust was uplifted via thrust faults such as the Marenisco fault and the Pelton Creek fault, forming the crustal-scale Montreal River monocline (Fig.
+During the rift inversion, Paleoproterozoic and Archean crust was uplifted via thrust faults such as the Marenisco fault and the Pelton Creek fault in the south, forming the crustal-scale Montreal River monocline (Fig.
 \ref{fig:Geologic_map}; \citeA{Cannon1993a}).
@@
 \ref{fig:Geologic_map}; \citeA{Hamblin1958a, Cannon1993a, Kalliokoski1982a}).
-More recent mapping of the Jacobsville Formation has also found that it can sometimes onlap the Midcontinent Rift volcanics within or adjacent to the Keweenaw Fault system around Bete Grise Bay on the Keweenaw Peninsula (Fig.
+North to these faults, Midcontinent Rift volcanic rocks were thrusted and folded along the Keweenaw fault which extends $\sim$250km from northern Wisconsin to northern Keweenaw Peninsula (Fig.
+\ref{fig:Geologic_map}; \citeA{DeGraff2022a}).
+Final inversion of the Keweenaw fault during the Rigolet phase of the Grenvillian Orogeny deformed part of the Jacobsville strata near the fault \cite{Hodgin2022a}.
+Recent mapping of the Jacobsville Formation also found that it can sometimes onlap the Midcontinent Rift volcanics within or adjacent to the Keweenaw Fault system near the northern end of the Keweenaw Peninsula (Fig.
 \ref{fig:Geologic_map}; \citeA{Tyrrell2019a, Mueller2021a}).
-Field observations find erosional contact and development of paleosol between the Jacobsville Formation and underlying volcanic rocks \cite{Hamblin1958a, Kalliokoski1975a}, and interpretations of seismic reflection data show that sedimentary rocks interpreted to be Jacobsville Sandstone lie unconformably on the Oronto Group \cite{Cannon1989a}.
-These observations are consistent with the interpretation that there was a prolonged period of sedimentation quiescence and net erosion between the end of Oronto Group deposition and onset of Jacobsville Formation deposition.
-Based on dates from Cannon et al 1993 (and deformation of the Freda Fm within the Montreal River monocline) paired with new dates on the Jacobsville Fm (Hodgin et al., 2022), this time period of quiescence may have lasted up to ~60 Myrs from ca.
-1050 until ca.
+
+Field observations find erosional contact and development of paleosol between the Jacobsville Formation and underlying volcanic rocks \cite{Hamblin1958a, Kalliokoski1975a}, and interpretations of seismic reflection data show that sedimentary rocks interpreted to be Jacobsville Formation lie unconformably on the Oronto Group \cite{Cannon1989a}.
+These data indicate that there was a prolonged period of sedimentation quiescence and net erosion between the end of Oronto Group deposition and onset of Jacobsville Formation deposition.
+Rb-Sr radiometric thermochronology data of \citeA{Cannon1993a} show the Montreal River monocline in which the Freda Formation and older basement rocks are deformed had been uplifted to mid to shallow crustal levels by ca.
+1060 Ma.
+U-Pb dates of late kinematic calcite crystals that precipitated at ($\sim$50\textdegree C) in the hanging wall of the Keweenaw fault \cite{Hodgin2022a} indicate that by ca.
+985 Ma the dominant inversion of Midcontinent Rift rocks along the Keweenaw fault has completed.
+Together, these data are consistent with the interpretation that the dominant inversion of the Midcontinent Rift basin happened during the ca.
+1090-1020 Ma Ottawan phase of the Grenvillian Orogeny.
+They also indicate that the period of sedimentary quiescence in the rift basin may have lasted up to $\sim$70 Myrs from ca.
+1060 to ca.
 990 Ma.
-The overall extent of the Jacobsville Formation has been mapped along the lake shore of the Keweenaw Peninsula from northern Michigan, USA, to the north side of Michipicoten Island and perhaps to northern Sault Ste Marie, Canada (Fig.
-\ref{fig:Geologic_map}; \citeA{Hamblin1958a}).
 
-As far-field stresses from the Grenvillian Orogeny propagated into the interior of Laurentia, the previously thinned crust of the Midcontinent Rift was structurally inverted by crustal-scale thrusts, such as the Keweenaw Fault which extends $\sim$250km from ...refs here.
-Final inversion of the Keweenaw Fault during the Grenvillian Orogeny thrust Midcontinent Rift volcanics over the Jacobsville strata, which were folded in the footwall of the fault.
-along much of the Keweenaw fault which extends $\sim$250 km from the tip of the Keweenaw Peninsula in northern Michigan southeastward to a termination in northeastern Wisconsin (Fig.
-\ref{fig:Geologic_map}; \citeA{Cannon1996a, Cannon2001a, DeGraff2022a}).
-Syndepositional faulting that took place during Jacobsville deposition is inferred from ...
-rift volcanic clasts where eroded into the Jacobsville basin at the same time when sediments from the older basement rocks along with distal input from the Grenville Province were deposited within the Jacobsville back-bulge basin system that in general wedges toward the northwest (Fig.
-\ref{fig:Geologic_map}; \cite{Hedgman2014a, Hodgin2022a}.
-Next to the footwall of the Keweenaw fault, folding has also been recorded by strata of the upper Jacobsville Formation in contact with the volcanics \cite{Cannon1993a}.
-One famous remnant of that late-stage deformation is the Natural Wall ravine where sub-vertical sandstone beds are preserved (Fig.
-\ref{fig:Geologic_map}; \cite{Hamblin1958a}).
+Syndepositional faulting that took place during Jacobsville deposition is inferred based on the observation of rift volcanic clasts within the Jacobsville basin.
+In the footwall of the Keweenaw fault near central Keweenaw Peninsula, strata of the Jacobsville Formation in contact with the Midcontinent volcanics are often steeply folded with the conglomeratic strata containing volcanic clasts \cite{Cannon1993a}.
 On the shore of Bete Grise Bay near the northeastern end of Keweenaw Peninsula, the Keweenaw fault is interpreted to have accommodated Grenville compressional stresses by forming a series of segmented en echelon thrusts \cite{Tyrrell2019a}.
 The timing of such faulting could be coeval with the deposition of the upper Jacobsville Formation, leading to the nonconformable onlapping of the Jacobsville sediments over the volcanics in addition to the thrust faulting relationship between the two units.
-Ca.
-985 Ma late-kinematic calcite that precipitated within the hanging wall of the Keweenaw Fault and ca.
-992 Ma maximum deposition age from detrital zircons of the Jacobsville Formation in the footwall constrain the timing of the final inversion of the Keweenaw Fault to be during the Rigolet phase of the Grenvillian Orogeny \cite{Hodgin2022a}.
 
-\subsection{Sedimentology and general stratigraphy}
+% at the same time when sediments from the older basement rocks along with distal input from the Grenville Province were deposited within the Jacobsville back-bulge basin system that in general wedges toward the northwest (Fig.
+\ref{fig:Geologic_map}; \cite{Hedgman2014a, Hodgin2022a}.
+
+\subsection{Sedimentary facies and general stratigraphy with interpretation}
 
@@
 \includegraphics[width=0.9\textwidth]{Field_photo.png}
-\caption{\scriptsize Field photos of red beds of the Jacobsville Formation sampled in this study.
-At Baby Snake Creek and the Natural Wall ravine in northern Keweenaw Peninsula, intervals of red fine-grained sandstone beds can be found through the steeply dipping to overturned basal strata near the Keweenaw Fault (A,C) and also the nearly horizontal top strata (B,D).
-At Hungarian Falls, relatively thick intervals of fine-grained red sandstone to siltstone beds separated by hematite-depleted arenite to arkosic sandstones or conglomerates are found along a $\sim$100-meter-tall waterfall (E).
+\caption{\scriptsize Field photos of red beds of the Jacobsville Formation.
+At Baby Snake Creek and the Natural Wall ravine in northern Keweenaw Peninsula, intervals of red fine-grained sandstone to siltstone beds can be found through the steeply dipping to overturned basal strata near the Keweenaw fault (A,C) and also the nearly horizontal top strata (B,D).
+At Hungarian Falls, relatively thick intervals of red beds separated by hematite-depleted arenite to arkosic sandstones or conglomerates are found along a $>$100-meter-tall waterfall (E).
 Abundant reoriented siltstone rip-up clasts often exist within the conglomeratic layers above siltstone beds (F).
-Drill cores or block samples of these rip-up clasts are sampled for the paleomagnetic conglomerate test presented in Fig.
-\ref{fig:Intraclast_pmag}.
-At Agate Falls, nearly flat-lying red fissile shale beds are exposed due to waterfall incision (G).
+Drill cores and block samples of these rip-up clasts were sampled for the paleomagnetic conglomerate test presented in Fig.
+\ref{fig:intraclast_pmag}.
+At Agate Falls, nearly flat-lying red fissile siltstone beds are exposed.
+(G).
 Subrounded detrital mica grains deposited parallel to the bedding plane are often associated with the siltstone facies.
@@
 
-The Jacobsville Formation is a fluvial sequence of feldspathic and quartzose conglomerates, sandstones, siltstones, and shales completely devoid of lava flows or cross-cutting igneous dikes although cross-cutting clastic dikes have been observed.
-Abundant near the bottom section in the southern exposures of the Jacobsville Formation, conglomerate facies typically have provenance sourced from Paleoproterozoic lithologies such as vein quartz and iron formation clasts \cite{Hamblin1958a, Kalliokoski1982a}.
-The occurrence of these conglomerates in proximity to the Keweenaw, Marenisco, and Pelton Creek thrust faults is interpreted to represent syn-depositional development of local relief \cite{Kalliokoski1982a, Hedgman2014a}).
-It is in general more quartz-rich than the Oronto Group sediments \cite{Hamblin1958a}.
+The Jacobsville Formation is a fluvial sequence of feldspathic and quartzose conglomerates, sandstones, siltstones, and shales completely devoid of lava flows or cross-cutting igneous dikes although numerous  cross-cutting clastic dikes are present \cite{Hamblin1958a}.
+Sediments of the Jacobsville Formation is interpreted to be dominantly derived from the Oronto Group sediments and older basement rocks that were uplifted and deformed in the Montreal River monocline.
+The Jacobsville sediments are in general more quartz-rich than the Oronto Group sediments \cite{Hamblin1958a}.
 
+In the southern exposures of the Jacobsville Formation, abundant conglomerate facies typically have provenance sourced from Paleoproterozoic lithologies such as vein quartz and iron formation clasts \cite{Hamblin1958a, Kalliokoski1982a}.
+The occurrence of these conglomerates in proximity to the Marenisco fault and Pelton Creek faults is interpreted to represent syn-depositional development of local relief \cite{Kalliokoski1982a, Hedgman2014a}.
 Close to the Keweenaw Fault in central Keweenaw Peninsula (Fig.
-\ref{fig:Geologic_map}), the Jacobsville Formation consists dominantly of quartz-rich, cross-bedded, medium-grained sandstone, with locally abundant clast-supported pebble to cobble conglomerate, and interbeds of red, hematite-bearing, micaceous shaly siltstone to fine-grained sandstone (Hungarian Falls, Hammel Creek, Natural Wall; Fig.
+\ref{fig:Geologic_map}), the Jacobsville Formation consists dominantly of quartz-rich, cross-bedded, medium-grained sandstone, with locally abundant clast-supported pebble to cobble conglomerate, and interbeds of red, hematite-bearing, micaceous siltstone to fine-grained sandstone (Hungarian Falls, Hammell Creek, Natural Wall; Fig.
 \ref{fig:strat_column}).
 In this region, the conglomerate contains abundant volcanic clasts likely derived from uplifted Midcontinent Rift volcanics \cite{Irving1885a, Brojanigo1984a}.
-This clast-supported lithofacies can include siltstone rip-up clasts or channelized lenses of siltstones.
-Occurrence of red siltstone to sandstone beds are also commonly interbedded with sandstone and conglomerate \cite<e.g.>{Hamblin1958a, Irving1885a}.
-They typically appear brick-red to dark-red, consisting of find-grained arenite to arkosic sandstone and siltstone sometimes with minor grey reduction spots (Fig.
-\ref{fig:Field_photo}).
-Rounded detrital mica grains up to $\sim$5mm in size can be found commonly within very fine, fissile siltstone and shaly strata (Fig.
+This clast-supported lithofacies can include siltstone rip-up clasts and sometimes channelized lenses of siltstones.
+More massive red beds also commonly interbed with hematite-depleted sandstone and conglomerate layers \cite<e.g.>{Hamblin1958a, Irving1885a}.
+The red beds typically appear brick-red to dark-red, consisting of fine-grained arenite to arkosic sandstone and siltstone sometimes with minor grey reduction spots (Fig.
 \ref{fig:Field_photo}).
 
-The exact stratigraphic association between sections of the Jacobsville Formation in the central Keweenaw Peninsula and those exposed at Agate Falls and Baby Snake Creek is difficult to constrain as the thickness of the Jacobsville strata can vary significantly through the peninsula \cite{Hamblin1958a, Kalliokoski1982a} and are not fully exposed at or in between the studied localities.
-As a result, various interpretations could exist for the relative stratigraphy at different localities.
-Based on the relative lack of conglomerate facies at Agate Falls and at Baby Snake Creek, and field observations of onlapping Jacobsville Formation on top of previously thrusted Midcontinent Rift volcanics, we interpret that Jacobsville Formation exposed at Agate Falls and Baby Snake Creek could be stratigraphically higher than the other three sections near central Keweenaw Peninsula (Fig.
+The occurrence of the rift volcanic-clast containing conglomeratic facies at Hungarian Falls, Hammell Creek, and Natural Wall indicates an interval of syn-depositional faulting in proximity to the Keweenaw fault in central part of the peninsula (Fig.
 \ref{fig:Geologic_map}, \ref{fig:strat_column}).
-Further discussion regarding the stratigraphy is presented in section \ref{magstrat}.
+Rapid weathering of the exhumed volcanics from the hanging wall also likely derived red clay and arkosic sandstone in the Jacobsville Formation which are most abundant near the fault but were also transported further away \cite{Hamblin1958a}.
+Rounded detrital mica grains commonly found within very fine, fissile siltstone strata are likely sourced from eroded Paleoproterozoic and Archean basement rocks in the Montreal River monocline highlands (Fig.
+\ref{fig:Geologic_map}).
+The mica grains can be up to $\sim$5mm in size (Fig.
+\ref{fig:Field_photo}).
 
-% Near the central part of Keweenaw Peninsula, we visited three adjacent Jacobsville stratigraphic sections at Hammel Creek, Hungarian Falls, and Natural Wall ravine (Fig.
+Thickness of the Jacobsville strata can vary significantly in northern Michigan \cite{Hamblin1958a, Kalliokoski1982a}, and complete stratigraphic sections are often not fully exposed at studied localities (Fig.
+\ref{fig:Geologic_map}).
+As a result, the exact stratigraphic association between sections of the Jacobsville Formation near central Keweenaw Peninsula and those exposed at Agate Falls and Baby Snake Creek is difficult to constrain.
+We propose one tentative stratigraphy model based on an interpreted Jacobsville sedimentary facies model.
+Given that deposition of conglomeratic facies was likely associated with syn-deposition faulting along the Keweenaw fault, this facies could have been deposited during the early stages of Jacobsville sedimentation, whereas siltstone and arkosic sandstone facies were probably deposited later with broader spatial distribution.
+Some early conglomeratic facies could have been reworked and contributed to the formation of Jacobsville red beds farther away from the fault.
+Based on this facies model, the relative lack of conglomerate facies at Agate Falls and at Baby Snake Creek, and field observations of onlapping Jacobsville Formation on top of previously deformed rift volcanics at Baby Snake Creek, could support an interpretation that Jacobsville sediments exposed at these two localities are slightly younger than the conglomerate-rich strata near central Keweenaw Peninsula (Fig.
 \ref{fig:Geologic_map}, \ref{fig:strat_column}).
-The exposed Jacobsville Formation share similarities in that they are all close to the Keweenaw Fault where the formation is steeply tilted in the footwall against the folded Midcontinent Rift volcanics in the hanging wall.
-They are also of similar thickness and are characterized by the interbedded layers of pebble conglomerate, massive cross-bedded sandstone and minor amount of siltstone to fine-grained sandstone (Fig.
-\ref{fig:strat_column}).
+Further discussion regarding the stratigraphy in light of the paleomagnetic data is presented in section \ref{magstrat}.
 
-% The fact that in the Jacobsville has conglomerate and sandstone and siltstone facies in them supports sediment input from the eroded Oronto Group.
-This indicate the older boundary of the Jacobsville deposition might be closer to the Rigolet Phase than the Ottawan phase and the duration of Jacobsville sedimentation may be within a relatively constrained period of time.
+%At some point earlier on, we may need to articulate a facies model that can be used to justify the correlation scheme by 1) demonstrating that the conglomeratic facies at the grouped sites in the lower Keweenaw Peninsula, which are associated with siltstone deposition and arkosic sands, were likely deposited during an interval of syn-depositional faulting in proximity to the Keweenaw Fault; next 2) it can be suggested that rapid weathering of the exhumed basaltic rock from the hanging wall resulted in red clay most abundant to the fault but also deposited further away, and abundant partially weathering arkosic sands but which also occur farther away to some degree; 3) This interpretation is supported by the facies association being volcanic conglomerate, red siltstone, and arkosic sands (along with some other typical Jacobsville facies) and is the basis for suggesting that farther travelled silt and arkosic sands could be correlated with the conglomerates.
+Tentatively this stratigraphicaly significant and unique facies association may serve as a basis for correlation across the basin during a time of more significant syn-depositional fault uplift.
 
-% tyrrell 2019 fig. 10 recorded onlapping of the Jacobsville Formation on to the Portage Lake Volcanics which appears to have been faulted prior to the deposition of Jacobsville near Bete Grise Bay.
-
-% Along the south shore of Lake Superior, the Keweenawan Supergroup overlies Paleoproterozoic basement and consists of ~5 km of rift-related volcanic rocks that are conformably overlain by ~4 km of conglomerate, siltstone, and sandstone of the Oronto Group (Cannon et al., 1996).
-The uppermost unit of the Keweenawan Supergroup is the Jacobsville Sandstone, which consists of up to ~1 km of fluvial deposits that are more quartz-rich than the Oronto Group (Fig.
-1; Kalliokoski, 1982).
-The Jacobsville Sandstone unconformably overlies Archean and Paleoproterozoic basement as well as Midcontinent Rift volcanic and sedimentary rocks in northern Michigan (Fig.
-1; Cannon et al., 1993; Ojakangas et al., 2001).
-
-% The Keweenaw Fault is a north to northwest-dipping thrust that  Thrust faults parallel to the Keweenaw Fault include the northwest-dipping Marenisco and Pelton Creek faults (Fig.
-1).
-Similarly-oriented thrusts continue southwestward through northern Wisconsin and eastern Minnesota (Ojakangas et al., 2001), and eastward below Lake Superior, where they are interpreted to join mapped thrust faults in southern Ontario (Fig.
-1; Manson and Halls, 1994; Ojakangas et al., 2001).
-Rock units adjacent to these regional thrusts have been moderately to intensely folded.
-
-% In northern Michigan, the typically shallowly-dipping Jacobsville Sandstone has sub-vertical to overturned beds in the immediate footwall of the Keweenaw Fault (Irving and Chamberlin, 1885; Cannon and Nicholson, 2001).
+Overall, the extent of the Jacobsville Formation has been mapped along the lake shore of the Keweenaw Peninsula from northern Michigan, USA, to the north side of Michipicoten Island and perhaps to northern Sault Ste Marie, Canada (Fig.
+\ref{fig:Geologic_map}; \citeA{Hamblin1958a}).
+It has also been interpreted that the Jacobsville Formation could be temporally and lithologically correlated with the Bayfield Group sediments in northern Wisconsin and the Fond du Lac and Hinckley Formation in Minnesota \cite{Hamblin1958a, Kalliokoski1982a, Malone2016a}.
 
@@
 \end{figure*}
-
-% Keweenaw Fault motion post dating Jacobsville in the footwall
-% Youngest DZ from jacobsvill eis about 993 Ma (sandstone creek, about a few hundred meters from the fault). 
-% directly above pmag sampling at Agate falls grain B geochron date is (z2, s226, about 1000 Ma grain)
-
-% jacobsville (985.5 from calcite in the fault and <993 from zircons)
 
@@
 
-% This makes the Jacobsville sandstone one unique opportunity to acquire a paleomagnetic pole position within the gap between the
+\section*{Paleomagnetic results and inerpretation}
 
-\section*{Paleomagnetic methods and results}
-
-Oriented paleomagnetic samples were collected with a portable electric drill along five stratigraphic sections of the Jacobsville Formation in norther Michigan (Fig.
-\ref{fig:Geologic_map}).
+Oriented paleomagnetic samples were collected with a portable electric drill along five stratigraphic sections of the Jacobsville Formation in northern Michigan (Fig.
+\ref{fig:Geologic_map}, \ref{fig:strat_column}).
 Additional oriented cores and block samples of siltstone rip-up clasts within conglomerate lithofacies were sampled at Hungarian Falls and at Agate Falls (Fig.
 \ref{fig:Geologic_map}, \ref{fig:strat_column}).
-In order to maximize sampling of distinct time snapshots of the geomagnetic field at the time of Jacobsville deposition such that paleosecular variation can be sufficiently averaged out, we optimized for vertical stratigraphic coverage such that each sample constitutes a paleomagnetic site considering that a paleomagnetic site (which ideally captures a single snapshot of the local geomagnetic field) is a particular bed in a sedimentary sequence.
+In order to maximize sampling of distinct time snapshots of the geomagnetic field at the time of Jacobsville deposition such that paleosecular variation can be sufficiently averaged, we optimized for vertical stratigraphic coverage such that each sample constitutes a paleomagnetic site considering that a paleomagnetic site (which ideally captures a single snapshot of the local geomagnetic field) is a particular bed in a sedimentary sequence.
 Red fine-grained sandstone to shaly siltstone layers were preferentially sampled as they tend to have lower permeability and are less susceptible to diagenetic alteration through fluid flow than coarser grained sandstone.
@@
 The specimens underwent stepwise thermal demagnetization at the UC Berkeley Paleomagnetism Lab using an ASC demagnetizer (residual fields $<$10 nT) with measurements made on a 2G DC-SQUID magnetometer.
-The demagnetization protocol had high-resolution approaching the Ne\'el temperature of hematite (5 \textdegree C to 2 \textdegree C) (Fig.
-\ref{fig:Intraclast_pmag}).
+The demagnetization protocol had high-resolution approaching the Ne\'el temperature of hematite.
+As a result, 356 specimens yielded detrital remanent magnetizations that can be fitted with least-square lines.
+(5 \textdegree C to 2 \textdegree C) (Fig.
+\ref{fig:intraclast_pmag}).
 Demagnetization results are plotted in Figure \ref{fig:in_situ_pmag}A by locality.
@@
 \includegraphics[width=\textwidth]{in_situ_pmag.pdf}
-\caption{}
+\caption{A) Specimen detrital remanence directions plotted by locality on equal area plots.
+During sample collection, we optimized for vertical stratigraphic coverage such that each sample constitutes a paleomagnetic site considering that a paleomagnetic site.
+B) Summary equal area plots at the study-level combining directions from all localities.
+The elongation/inclination (E/I) method \cite{Tauxe2004b} for estimating the amount of inclination shallowing in Jacobsville red beds are applied separately to specimen remanence directions of each polarity and to all specimen directions after flipping the reversed (orange) directions to their antipodes (blue).
+Details on inclination shallowing corrections are shown in Figure \ref{fig:EI_results}.}
 \label{fig:in_situ_pmag}
@@
 Similar to the thermal demagnetization result of the hematite-bearing fluvial intraclasts of the Freda Formation \cite{Swanson-Hysell2019b}, the siltstone rip-up clasts of the Jacobsville Formation typically reveal two distinct magnetization components (Fig.
-\ref{fig:Intraclast_pmag}).
+\ref{fig:intraclast_pmag}).
 One component show similar vector orientations amongst intraclasts and was typically removed up to 640-655 \textdegree C (Fig.
-\ref{fig:Intraclast_pmag}).
+\ref{fig:intraclast_pmag}).
 This indicates that it is likely a chemical remanent magnetization acquired through growth of secondary hematite in pore spaces following redeposition of the clasts.
 After removal of this component, further thermal demagnetization with small step increments at higher temperatures up to $\sim$689 \textdegree C often reveal an origin-trending component (Fig.
-\ref{fig:Intraclast_pmag}).
-In the data from the clasts, there is typically a significant directional change in specimen magnetization between the mid-temperature component and the high-temperature component and as a result, 28 out of 30 intraclast specimens could be fit with distinct mid-temperature and high-temperature least squares lines \cite{Kirschvink1980a}.
+\ref{fig:intraclast_pmag}).
+In the data from the clasts, there is typically a significant directional change in specimen magnetization between the mid-temperature component and the high-temperature component.
+As a result, 28 out of 30 intraclast specimens could be fit with distinct mid-temperature and high-temperature least squares lines \cite{Kirschvink1980a}.
 Distinct from the well-grouped mid-temperature component directions, the high-temperature directions of the intraclasts are dispersed and the null hypothesis of randomness cannot be rejected at the 95\% confidence level (n=6 for intraclasts at Agate Falls and n=22 for intraclasts at Hungarian Falls; Fig.
-\ref{fig:Intraclast_pmag}; \citeA{Watson1956a}).
-This indicates that the component is held by detrital hematite grains during initial formation of the clasts prior to erosion and redeposition.
+\ref{fig:intraclast_pmag}; \citeA{Watson1956a}).
+This indicates that the remanence was acquired by detrital hematite grains during initial deposition of the grains.
 
@@
 \centering
-\includegraphics[width=\textwidth]{Intraclast_pmag.pdf}
-\caption{Paleomagnetic data from fluvial rip-up clasts of the Jacobsville Formation reveal a mid-temperature component that typically unblocks prior to 655 \textdegree C and a high-temperature component that typically unblocks up to 689 \textdegree C.
+\includegraphics[width=\textwidth]{intraclast_pmag.pdf}
+\caption{Paleomagnetic data from fluvial rip-up clasts of the Jacobsville Formation reveal a mid-temperature component that typically unblocks up to 655 \textdegree C and a high-temperature component that typically unblocks up to 689 \textdegree C.
 The components are present as varying fractions of the overall remanence as seen in the three individual clasts shown on vector orthogonal plots and measurement-level equal area plots in tilt-corrected coordinates (developed using PmagPy; \citeA{Tauxe2016a}).
-The direction of the mid-temperature component is shown as green arrows on the vector component plots and green circles on the equal area plots, while the high-temperature component is shown with purple symbols.
+The direction of the mid-temperature component (interpreted as CRM) is shown as green arrows on the vector component plots and green circles on the equal area plots, while the high-temperature component (interpreted as DRM) is shown with purple symbols.
 The mid-temperature component has a similar direction among the clasts as can be seen on the summary equal area plots.
@@
 NRM = natural remanent magnetization; AFC = Agate Falls clasts; HFC = Hungarian Falls clasts.}
-\label{fig:Intraclast_pmag}
+\label{fig:intraclast_pmag}
 \end{figure*}
@@
 
-A total of 70 specimens yielded interpretable high-temperature paleomagnetic remanence component throughout the stratigraphic section at Baby Snake Creek (Fig.
+A total of 70 specimens yielded interpretable DRM directions throughout the stratigraphic section at Baby Snake Creek (Fig.
 \ref{fig:Geologic_map}, \ref{fig:in_situ_pmag}).
-The tilt-corrected remanence directions are all of normal polarity and show a dominantly southwest declination and upward-pointing inclination (Fig.
-\ref{fig:in_situ_pmag}).
-The directions in geographic coordinates yield a bootstrap fold test \cite{Tauxe1994a} result that shows the 95\% confidence bounds for the best grouping of the directions from the two limbs of the fold is achieved between 55\% and 92\% unfolding.
-Although the 95\% confidence bounds exclude 100\% unfolding, that the grouping of the directions become much tighter than they were in geographic coordinates when significant unfolding is applied suggest the high-temperature component from Baby Snake Creek was likely acquired pre-folding.
-The non-ideal unfolding test could be a result of overlapping unblocking temperatures of detrital and pigmentary hematite in red beds of Baby Snake Creek or complex folding regime near the Keweenaw Fault or a combination of the two.
+46 specimens were collected from the steeply tilted beds close to the Keweenaw fault and the rest 24 were from nearly horizontal beds.
+This allows us to conduct the paleomagnetic fold test on these specimens to investigate the timing of detrital remanence acquisition.
+A bootstrap fold test \cite{Tauxe1994a} result shows 95\% confidence bounds for the best grouping of the directions from the various beds of the fold is achieved between 55\% and 92\% unfolding (Fig.
+\ref{fig:fold_test}).
+Although the 95\% confidence bounds exclude both 100\% unfolding and geographic coordinates, the results indicate the directions are better grouped with a significant amount of tilt correction is applied complex folding regime such as non-cylindrical folding or multiple folding episodes about different axes could contributed to this fold test result not including 100\% unfolding within the 95\% confidence bounds.
 
@@
 \includegraphics[width=0.5\textwidth]{SC1_fold_test.pdf}
-\caption{Bootstrap paleomagnetic fold test \cite{Tauxe1994a} of the detrital remanent magnetization remanence directions recorded by specimens from Baby Snake Creek.
-Although the directional data set becomes more tightly grouped than in geographic coordinates when applying 55\%-92\% unfolding, both geographic and 100\% unfolded coordinate systems are excluded at the 95\% confidence level.
+\caption{Bootstrap paleomagnetic fold test \cite{Tauxe1994a} of the detrital remanent magnetization directions recorded by specimens from Baby Snake Creek.
+The directional data set becomes more tightly grouped than in geographic coordinates when applying 55\%-92\% unfolding, but both geographic and 100\% unfolded coordinate systems are excluded at the 95\% confidence level.
 }
-\label{fig:Intraclast_pmag}
+\label{fig:fold_test}
 \end{SCfigure*}
@@
 \label{magstrat}
-With the insights gained from the thermal demagnetization results of the Jacobsville fluvial intraclasts, we next isolate the high-temperature magnetization component from the mid-temperature component for all in situ specimens that were collected along the five localities in northern Michigan (Fig.
+With the insights gained from the thermal demagnetization results of the Jacobsville fluvial intraclasts (Fig.
+\ref{fig:intraclast_pmag}), we next investigate the detrital remanent magnetizations for all in situ specimens that were collected along the five localities in northern Michigan (Fig.
 \ref{fig:Geologic_map}).
-The resultant high-temperature component directions are plotted by locality in Fig.
+The resultant DRM directions are plotted by locality in Fig.
 \ref{fig:in_situ_pmag}.
-These results show dual polarities recorded by the Jacobsville Formation (Fig.
+These results show dual polarities with some specimens recording transitional field directions (Fig.
 \ref{fig:in_situ_pmag}), consistent with the observation of \citeA{Roy1978a}.
-Furthermore, our new data show that the Jacobsville Formation recorded more than one geomagnetic field polarity reversal and a tentative magnetostratigraphy may be constructed (Fig.
+While \citeA{Roy1978a} observed only one polarity reversal, our new data show that the Jacobsville Formation recorded multiple geomagnetic field polarity reversals.
+A tentative magnetostratigraphy is proposed in Fig.
+\ref{fig:strat_column}.
+In order to aid later discussions regarding the paleogeography of Laurentia, we refer to the westerly DRM directions as of normal polarity and the easterly DRM directions of reversed polarity (Fig.
 \ref{fig:in_situ_pmag}).
 
-At Agate Falls (Fig.
-\ref{fig:Geologic_map}), multiple hematite-rich, detrital mica-bearing siltstone layers occur through a $\sim$30-meter strata incised by the running Middle Branch Ontonagon River.
-68 paleomagnetic samples were collected on both sides of the river through overlapping stratigraphic sections (Fig.
+At Hungarian Falls (Fig.
+\ref{fig:Geologic_map}), numerous hematite-rich siltstone to fine-grained sandstone layers were deposited (Fig.
 \ref{fig:strat_column}).
-A single transition from normal to reversed polarity is observed through the sections from bottom to top with a few strata in the middle of the section capturing transitional field directions (Fig.
-\ref{fig:strat_column}, \ref{fig:in_situ_pmag}).
-
-At Hungarian Falls (Fig.
-\ref{fig:Geologic_map}), numerous hematite-rich siltstone to fine-grained silty sandstone layers were deposited.
-The siltstone and silty fine-grained sandstone layers can capture a relatively rich set of information from the paleomagnetic field variations compared to coarser-grained lithologies at other localities (Fig.
+Throughout $\sim$240 meters of strata, a total of 158 samples were collected for paleomagnetic study.
+At stratigraphic height of $\sim$24 meters (Fig.
+\ref{fig:strat_column}), 8 samples from a $\sim$0.5-meter-thick siltstone layer record reversed polarity directions (easterly declination and shallow inclination; horizon HF6).
+At stratigraphic level $\sim$42 meters (horizon HF5; Fig.
+\ref{fig:strat_column}), 4 samples were collected through a $\sim$0.2 m siltstone horizon including one specimen from near the base of the layer and three specimens near the top of the layer.
+The DRM component from the three specimens near the top all record a reversed polarity whereas the specimen collected near the base show a westerly direction.
+The westerly direction indicates that the geomagnetic field could have gone through a reversal between strata at 24 and 42 m, although more samples are needed to test the length and stability of the potential period of normal polarity.
+Therefore, we tentatively mark a period of transitional field near the base of horizon HF5 in Fig.
+\ref{fig:strat_column}.
+A reversed polarity field seems to have been maintained as 5 samples from horizon HF4 consistently record easterly directions just above the top of HF5 (Fig.
 \ref{fig:strat_column}).
-Throughout $\sim$240 meters of stratigraphy, a total of 158 samples were collected for paleomagnetic study.
-At stratigraphic height of $\sim$24 meters (Fig.
-\ref{fig:strat_column}), 8 samples from a $\sim$0.5-meter-thick siltstone layer record a reversed polarity (easterly declination and shallow inclination; horizon HF6).
-At stratigraphic level $\sim$42 meters (HF5; Fig.
-\ref{fig:strat_column}), 4 samples were collected through a $\sim$0.2 m siltstone horizon including one specimen from near the base of the layer and three specimens near the top of the layer.
-The high-temperature remanence component from the three specimens near the top all record a reversed polarity whereas the specimen collected near the base show a westerly declination.
-While more samples are needed to test whether the westerly declination recorded by the single sample is a result of the geomagnetic field having completely flipped polarity from that recorded by horizon HF6 or a result of transient unstable field, this anomalous direction indicates that the geomagnetic field did not stay consistent throughout the time of deposition of this horizon.
-As a result, we tentatively mark a period of transitional field near the base of horizon HF5 in Fig.
-\ref{fig:strat_column}.
-A reversed polarity field seems to have been maintained as 5 samples from horizon HF4 consistently record easterly directions similar to those recorded by samples near the top of HF5 (Fig.
-\ref{fig:strat_column}).
-However, a geomagnetic field polarity reversal happened between the deposition of siltstone layers of HF4 and HF3, as 7 samples from HF3 all record normal-polarity directions, antipodal to those of samples of HF4.
-Subsequently, the field appears to flip again back to reversed polarity, as is shown by 7 samples from HF2 higher above the section (Fig.
+However, at least one geomagnetic field polarity reversal happened between the deposition of siltstone layers of HF4 and HF3, as 7 samples from HF3 all record normal-polarity directions.
+Subsequently, the field appears to flip again back to reversed polarity, as is shown by 7 samples from HF2 higher in the section (Fig.
 \ref{fig:strat_column}).
@@
 
-The Natural Wall ravine and Hammell Creek are in close proximity toHungarian Falls (Fig.
+The Natural Wall ravine and Hammell Creek are in close proximity to Hungarian Falls (Fig.
 \ref{fig:Geologic_map}) and the sedimentary facies characterized by trough cross-bedded sandstone and conglomerates are similar to the basal $\sim$120 meters of the section at Hungarian Falls (Fig.
 \ref{fig:strat_column}).
-Since only a couple of siltstone to fine-grained sandstone horizons were sampled at these two localities, we tentatively correlate the stratigraphic sections at these three localities based on their sedimentary facies and magnetic polarities (Fig.
+Since only a few siltstone to fine-grained sandstone horizons were sampled at these two localities, we tentatively correlate the stratigraphic sections at these three localities based on their sedimentary facies (Fig.
 \ref{fig:strat_column}).
@@
 Near the top of the Natural Wall section, consistent normal-polarity directions are recorded by 15 samples.
-This normal polarity could be coeval with one of the normal-polarity periods recorded at Hungarian Falls, or it could indicate an additional polarity reversal from reversed to normal polarity that happened after the deposition of the youngest reversed strata at Hungarian Falls (Fig.
+This normal polarity is likely coeval with one of the normal-polarity periods recorded at Hungarian Falls, although it could indicate an additional polarity reversal from reversed to normal polarity that happened after the deposition of the youngest reversed strata at Hungarian Falls (Fig.
 \ref{fig:strat_column}).
-Near the base of the stratigraphic section at Natural Wall (Fig.
+Near the base of the exposed stratigraphic section at Natural Wall (Fig.
 \ref{fig:strat_column}), the strata are steeply tilted or overturned (Fig.
@@
 Tilt-corrected paleomagnetic directions from samples collected from these strata show dominantly southeasterly declinations that are distinct from either normal- or reversed-polarity directions higher along the strata or at Hungarian Falls.
-Possible causes of such distinct directions could be related to transitional fields associated with geomagnetic field polarity reversals or complications in the tilt correction such as noncylindrical folding or multiple tilting episodes.
+It is most likely that these directions resulted from noncylindrical or multiple episodes of complex folding.
+We hypothesize that these DRM directions were acquired during a reverse polarity episode similar to that recorded near the base of Hammell Creek (Fig.
+\ref{fig:strat_column}), but we will not include these directions in further paleomagnetic pole analyses.
 
-A total of 70 specimens yielded interpretable high-temperature paleomagnetic remanence component throughout the stratigraphic section at Baby Snake Creek (Fig.
-\ref{fig:Geologic_map}, \ref{fig:in_situ_pmag}).
-The remanence directions recorded by red beds within the steeply folded basal section and within the flat-lying top portion of the section pass a bootstrap fold test (Supporting Information; \citeA{Tauxe1994a}).
-The remanence directions are all of normal polarity and show a dominantly southwest declination and upward-pointing inclination (Fig.
-\ref{fig:in_situ_pmag}).
+At Baby Snake Creek (Fig.
+\ref{fig:Geologic_map}), all 70 specimens that yielded interpretable high-temperature remanence have normal-polarity DRM directions \ref{fig:in_situ_pmag}).
+At Agate Falls (Fig.
+\ref{fig:Geologic_map}), multiple hematite-rich, detrital mica-bearing siltstone layers occur through a $\sim$30-meter strata incised by the running Middle Branch Ontonagon River.
+68 paleomagnetic samples were collected on both sides of the waterfall (Fig.
+\ref{fig:strat_column}).
+A single transition from normal to reversed polarity is observed through the strata from bottom to top with a few strata in the middle of the section capturing transitional directions (Fig.
+\ref{fig:strat_column}, \ref{fig:in_situ_pmag}).
 
-Given that multiple geomagnetic field reversals occurred during the deposition of the Jacobsville Formation, the development of amagnetostratigraphy should be possible and could be useful for future regional and global studies.
-However, challenges exist in correlating the stratigraphic sections measured at the five localities in this study due to the lateral variability of the paleodepositional environment associated with the variable paleotopographic relief of the Precambrian surface upon which the Jacobsville Formation was deposited \cite{Hamblin1958a, Kalliokoski1982a}.
+Challenges exist in correlating the stratigraphic sections measured at the five localities in this study due to the lateral variability of the thickness of the formation associated with the variable paleotopographic relief of the Precambrian surface upon which the Jacobsville Formation was deposited \cite{Hamblin1958a, Kalliokoski1982a}.
 In particular, there may not be a unique solution in reconstructing the relative stratigraphic positions of the strata at Baby Snake Creek and Agate Falls with respect to the other three sections in central Keweenaw Peninsula due to geographical separation (Fig.
@@
 However, at both Baby Snake Creek and Agate Falls, the sedimentary strata are dominated by siltstone to cross-bedded sandstone and are lacking significant amount of conglomerate facies as seen near Hungarian Falls.
-This could be consistent with the interpretation that these strata were deposited as the Jacobsville basin of the Grenvillian foreland system resulted from lithospheric flexure induced by orogenic loading has been partially filled.
-Recent geologic mapping efforts near Baby Snake Creek have found evidence for the Jacobsville Formation onlapping the Midcontinent Rift volcanics that were previously faulted via en echelon segments\cite{Tyrrell2019a, Mueller2021a}.
-The unconformable contact between the sediments and the volcanics, instead of faulted contact typically observed near Hungarian Falls where the basal Jacobsville Formation is folded and faulted against the volcanics, indicate that the deposition of the Jacobsville Formation near the northern end of Keweenaw Peninsula likely post-date some episodes of motion along the Keweenaw Fault.
-Therefore, we interpret that the normal-polarity strata at Baby Snake Creek could be stratigraphically the youngest amongst the five localities, although a possibility that they instead are near the bottom of the stratigraphy and record the normal Keweenawan Superchron cannot be ruled out \cite{Driscoll2016b}.
+Given the relative facies abundance difference at these localities, an interpreted stratigraphic correlation could exist based on an interpreted sedimentary facies model.
+The conglomerates, hematite-bearing siltstones, and arkosic sandstones deposited near the Keweenaw fault could be associated with syn-depositional faulting that uplifted the Midcontinent Rift volcanics that subsequently eroded and shedded sediments near the fault.
+These facies, particularly the conglomerate facies, could have then been reworked and transported farther, resulting in the deposition of more sandstone-rich younger strata (Fig.
+\ref{fig:strat_column}).
+The relative lack of conglomerates at Agate Falls which is $\sim$25 km away from the Keweenaw fault could indicate that the Jacobsville sediments there are derived from far-travelled sediments.
+Recent field mapping near Baby Snake Creek have found evidence for the Jacobsville Formation onlapping the Midcontinent Rift volcanics that were previously faulted via en echelon segments\cite{Tyrrell2019a, Mueller2021a}.
+The unconformable contact between the sediments and the volcanics, instead of faulted contact typically observed near Hungarian could indicate that the deposition of the Jacobsville Formation near the northern end of Keweenaw Peninsula likely post-date some episodes of fault motion, and are thus younger than the sediments near the fault.
+Therefore, we interpret that the normal-polarity strata at Agate Falls and at Baby Snake Creek could be stratigraphically the youngest amongst the five localities, although a possibility that they instead are older strata deposited during the Keweenawan superchron \cite{Driscoll2016b} cannot be ruled out.
 
-Overall, the multiple geomagnetic field polarity reversals recorded by the Jacobsville Formation constrain the maximum range of the Keweenawan normal superchron \cite{Driscoll2016b} to be between ca.
-1100 (Mamainse Point Volcanics; \cite{Swanson-Hysell2019a}) to ca.
-992 Ma (maximum deposition age of Jacobsville Formation; \cite{Hodgin2022a}).
-
-\begin{figure*}[h!]
-\centering
-\includegraphics[width=\textwidth]{EI_results.pdf}
-\caption{}
-\label{fig:EI_results}
-\end{figure*}
-
-\begin{figure*}[h!]
-\centering
-\includegraphics[width=\textwidth]{Jacobsville_pole_plot.pdf}
-\caption{}
-\label{fig:pole_plot}
-\end{figure*}
+Overall, the multiple geomagnetic field polarity reversals recorded by the Jacobsville Formation constrain the maximum range of the Keweenawan superchron \cite{Driscoll2016b} to be between ca.
+1100 Ma (Mamainse Point Volcanics; \cite{Swanson-Hysell2019a}) to ca.
+990 Ma (maximum deposition age of Jacobsville Formation; \cite{Hodgin2022a}).
 
@@
 
-Although igneous and sedimentary rocks associated with the North American Midcontinent Rift provided high-resolution record of the ca.
-1110 to 1070 Ma Keweenawan Track which is the apparent polar wander path (APWP) for Laurentia in the late Mesoproterozoic, high-quality paleogeographic constraints is lacking thereafter since magmatism in the interior of Laurentia entered a protracted quiescence period until the emplacement of the ca.
+\begin{figure*}[H!]
+\centering
+\includegraphics[width=\textwidth]{EI_results.pdf}
+\caption{\footnotesize Results of the estimated amount of inclination shallowing of the detrital remanent magnetization of the Jacobsville Formation using the elongation/inclination (E/I) method \cite{Tauxe2004b}.
+Directions from both polarities are used with the reversed-polarity directions flipped to their antipodes (Fig.
+\ref{fig:in_situ_pmag}).
+A) The E/I method results in an estimated flattening factor $f$=0.61 (red curve) based on where elongation/inclination curve of the dataset intersects that predicted by the TK03 paleosecular variation model (black curve; \citeA{Tauxe2004b}).
+The grey lines show 1,000 bootstrap resamples of the DRM directions which provide an estimate of the uncertainty associated with the $f$ factor estimate.
+B) The distribution of the paleolatitudes implied from the resultant corrected inclinations from the E/I bootstrap resamples.
+That the distribution is drawn from a normal distribution with a mean of 11.0 and standard deviation of 1.2 cannot be rejected in the Kolmogorov-Smirnov test.
+The 95\% confidence bounds shown by dashed lnes span a range of paleolatitudes that need to be incorporated into the uncertainty of the reported paleomagnetic pole.
+C) The cumulative distribution of all plausible inclinations based on the E/I bootstrap results with the 95\% confidence bounds shown in inclination and $f$ value axes.
+D) Kent mean pole and Fisher mean pole of the Jacobsville Formation in context of previous poles from \citeA{Roy1978a} and the pole position derived from the Chequamegon Formation of the Bayfield Group developed by \citeA{McCabe1983a}.
+The mean pole position of the Kent distribution calculated from the distribution of resampled Jacobsville mean poles following the method of \citeA{Pierce2022a} overlaps with the Fisher mean pole position corrected by a single inclination shallowing value of 0.61.
+Both mean poles lie in between the Fisher mean poles calculated using the inclination-corrected normal and reversed pole groups respectfully.
+The close proximity between Jacobsville mean poles with the Chequamegon pole indicate their deposition could be largely coeval.
+The J1 (C) pole of \citeA{Roy1978a} is recorded by hematite-bearing sediments in Sault Ste Marie area.
+Its position in the northern hemisphere close to the Upper Freda Formation pole indicate that those sediment could be Freda-equivalent instead of Jacobsville-equivalent.
+$^*$---Fisher mean pole positions calculated with inclination shallowing correction given by the E/I results shown in Fig.
+\ref{fig:in_situ_pmag}.
+$^+$---the Freda Formation Fisher mean pole is corrected for inclination shallowing assuming $f$=0.6.
+}
+\label{fig:EI_results}
+\end{figure*}
+
+Igneous and sedimentary rocks associated with the North American Midcontinent Rift provide high-resolution record of the ca.
+1109 to 1070 Ma Keweenawan Track which tightly constrains the apparent polar wander path for Laurentia in the late Mesoproterozoic (Fig.
+\ref{fig:pole_plot}; \cite{Swanson-Hysell2019a}).
+However, high-quality paleogeographic constraints is lacking thereafter since magmatism in the interior of Laurentia entered a protracted quiescence period until the emplacement of the ca.
 780 Ma Gunbarrel large igneous province \cite{Harlan2003a}.
-The new paleomagnetic data developed from the Jacobsville Formation presents an opportunity to extend the Keweenawan Track toward the early Neoproterozoic as its maximum deposition age has been tightly constrained to be ca.
-992 Ma through high-precision U-Pb detrital zircon geochronology \cite{Hodgin2022a}.
+The new paleomagnetic pole developed from the Jacobsville Formation presents an opportunity to extend the APWP to the early Neoproterozoic as its maximum deposition age has been tightly constrained to be ca.
+990 Ma through high-precision U-Pb detrital zircon geochronology and U-Pb calcite geochronology \cite{Hodgin2022a}.
 
 However, inclination bias associated with sedimentary paleomagnetic directions need to be corrected when synthesizing poles.
-Rotation of grains during sediment deposition and compaction during early diagenesis can result in the acquisition of a detrital remanent magnetization that is more shallowly inclined relative to the local geomagnetic field in which it was acquired \cite{King1955a, Tauxe2005a, Kodama2012a}.
+Rotation of grains during sediment deposition and compaction during early diagenesis can result in the acquisition of a detrital remanent magnetization that is shallower than the local geomagnetic field in which it was acquired \cite{King1955a, Tauxe2005a, Kodama2012a}.
 This effect can be summarized as:
@@
 If uncorrected, shallower inclinations obtained from sedimentary rocks can result in erroneously low estimates of paleolatitudes, biasing paleogeographic reconstructions \cite{Domeier2012a}.
-Methods for correcting sedimentary inclination bias include laboratory-based measurements of magnetic mineral fabrics \cite<e.g.>{Kodama1992a, Bilardello2015a}, as well as statistical methods based on the evaluation of the deviation of observed virtual geomagnetic pole (VGP) distribution patterns from an assumed paleosecular variation model (the ``elongation/inclination" [E/I] method of \citeA{Tauxe2004b}).
-When a large number of sedimentary paleomagnetic samples (typically n$>$100) capture an extensive period of time such that the paleosecular variation is sufficiently averaged, the E/I method can give accurate estimates and is less labor-intensive than methods based on mineral fabric measurements.
-In the ca.
-1.1 Ga Midcontinent Rift, \citeA{Pierce2022a} showed that the E/I method is effective in correcting the inclination bias recorded by a package of lava flow-bracketed fluvial red bed sediments---the Cut Face Creek Sandstone.
+Methods for correcting sedimentary inclination bias include laboratory-based measurements of magnetic mineral fabrics \cite<e.g.>{Kodama1990a, Bilardello2015a}, as well as statistical methods based on the evaluation of the deviation of an observed virtual geomagnetic pole (VGP) distribution from that of a paleosecular variation model (the ``elongation/inclination" [E/I] method of \citeA{Tauxe2004b}).
+When a large number of sedimentary paleomagnetic sites (typically n$>$100) capture an extensive period of time such that the paleosecular variation is sufficiently averaged, the E/I method can give efficient estimates with reasonable uncertainty bounds via bootstrapping and is less labor-intensive than methods based on mineral fabric measurements.
+\citeA{Tauxe2009a} based on paleomagnetic data from lava flows of the North Shore Volcanic Group in the Midcontinent Rift to show that the ca.
+1.1 Ga geomagnetic field had a behavior consistent with the prediction given by the TK03 paleosecular variation model of \citeA{Tauxe2004b}.
+\citeA{Pierce2022a} showed that the E/I method is effective in correcting the inclination bias recorded by a package of lava flow-bracketed fluvial red bed sediments---the Cut Face Creek Sandstone.
 
 Inclination shallowing is also present in the fluvial red beds of the Jacobsville Formation.
-Scatter of the specimen detrital remanence directions in geographic coordinates show dispersion of remanence vectors parallel to the bedding plane, indicating the directions are likely shallowed along the bedding-perpendicular plane (Fig.
+Scatter of the specimen detrital remanence directions in geographic coordinates show elongation of DRM directions parallel to the bedding plane, indicating the directions are shallowed (Fig.
 \ref{fig:in_situ_pmag}; \citeA{Tauxe2004b}).
-Since the deposition of the Jacobsville Formation span over geomagnetic field reversals, it is of interest to investigate the amount of inclination shallowing associated with the normal (n=130) and the reversed (n=148) polarity respectively.
+Since the deposition of the Jacobsville Formation span over geomagnetic field reversals, it is of interest to investigate the amount of inclination shallowing associated with the normal (n=130) and the reversed (n=148) polarity respectively (transitional fields and specimens affected by complex tilting regime are excluded).
 Applying the E/I method on the normal polarity directions yields an estimated $f$ value of 0.67 with 95\% bootstrap uncertainty bounds from 0.91 to 0.55 (Fig.
 \ref{fig:in_situ_pmag}B).
-On the other hand, directions of reversed polarity yields an estimated $f$ factor of 0.52 with more uncertainties ranging from 0.97 to 0.29 (Fig.
+Directions of reversed polarity yields an estimated $f$ factor of 0.52 with uncertainty bounds between 0.97 and 0.29 (Fig.
 \ref{fig:in_situ_pmag}B).
@@
 Therefore, inclination corrections do not result in significant changes in the mean reversed direction.
+
 In addition, directions of the two polarities do not pass a reversal test \cite{McFadden1990a, Tauxe2010a}, regardless of whether inclination correction is applied.
+One interpretation of the asymmetric polarities could be that the deposition of the Jacobsville Formation was protracted such that the sediments at Hungarian Falls which dominate the reversed polarity records were deposited prior to those at Baby Snake Creek which dominate the normal polarity records (Fig.
+\ref{fig:in_situ_pmag}).
+This is consistent with our stratigraphic interpretation that Baby Snake Creek sediments could be younger than those near central Keweenaw Peninsula (i.e.
+Hungarian Falls, Natural Wall, Hammell Creek; Fig.
+\ref{fig:strat_column}).
+Other alternative interpretations could include that the mean poles are biased by transient non-dipolar fields due to insufficient averaging of paleosecular variation or transitional fields during field reversals.
+More data are needed to test these hypotheses.
 
-One interpretation of the polarity discrepancy could be that the deposition of the Jacobsville Formation was protracted such that the sediments at Hungarian Falls whose detrital remanence dominate reversed polarity records was deposited prior to those at Baby Snake Creek which dominate the normal polarity records (Fig.
-\ref{fig:in_situ_pmag}).
-This is consistent with the magnetostratigraphic interpretation that Baby Snake Creek section is younger than the Hungarian Falls section (Fig.
-\ref{fig:strat_column}).
-Field observations of \citeA{Tyrrell2019a} and \citeA{Mueller2021a} near Bete Grise Bay, which is only $\sim$10 km east to Baby Snake Creek (Fig.
+Overall, both inclination-corrected polarities recorded by the Jacobsville Formation show mean pole positions in the southern hemisphere with respect to today's continental configuration (normal polarity longitude=175.1\textdegree E, latitude=19.1\textdegree S, A95=3.8\textdegree; reversed polarity longitude=191.1\textdegree E, latitude=14.5\textdegree S, A95=3.9\textdegree; Fig.
+\ref{fig:EI_results}).
+The two pole positions are distinct from the youngest poles developed from rocks of the Midcontinent Rift (Fig.
+\ref{fig:EI_results}).
+Despite that the Jacobsville sedimentation span over geomagnetic field reversals, the large arc distance between either polarity of the Jacobsville Formation and the pole of the ca.
+1070 Ma upper Freda Formation \cite{Henry1977a} is consistent with the interpretation that the onset of deposition of the Jacobsville Formation at the five studied localities post-date the youngest member of the Oronto Group sediments by tens of millions of years.
+
+% Field observations of \citeA{Tyrrell2019a} and \citeA{Mueller2021a} near Bete Grise Bay, which is only $\sim$10 km east to Baby Snake Creek (Fig.
 \ref{fig:Geologic_map}), show unconformable onlap of the Jacobsville Formation on top of Midcontinent Rift volcanics that have experienced significant thrusting, deformation.
@@
 
-% An alternative interpretation to the polarity discrepancy could be that there are insufficient samples such that either or both polarities are not average for secular variation. 
-Overall, both inclination-corrected polarities recorded by the Jacobsville Formation show mean pole positions in the southern hemisphere with respect to today's continental configuration (normal polarity longitude=175.1\textdegree E, latitude=19.1\textdegree S, A95=3.8\textdegree; reversed polarity longitude=191.1\textdegree E, latitude=14.5\textdegree S, A95=3.9\textdegree; both are corrected for inclination shallowing; Fig.
-\ref{fig:EI_results}).
-The two pole positions are distinct from the youngest poles developed from rocks of the Midcontinent Rift (Fig.
-\ref{fig:EI_results}).
-Despite that the Jacobsville sedimentation span over geomagnetic field reversals, the large arc distance between either polarity of the Jacobsville Formation and the pole of the ca.
-1070 Ma upper Freda Formation is consistent with the interpretation that the onset of deposition of the Jacobsville Formation at the five studied localities post-date the youngest member of the Oronto Group sediments by tens of millions of years.
+\subsection*{The Jacobsville paleomagnetic pole}
 
-\subsection*{The Jacobsville paleomagnetic pole}
-To succinctly summarize a paleomagnetic pole for the ca.
-992 Ma Jacobsville Formation, we combine the directions of both polarities and reassess the amount of inclination shallowing of the combined directional dataset (Fig.
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{Jacobsville_pole_plot.pdf}
+\caption{The Jacobsville Formation Kent mean pole position is plotted in context of the ca.
+1109-1070 Ma Keweenawan Track (poles color-coded by age shown in the colorbar) and selected poles from the Grenville Province.
+The southerly pole position of the ca.
+990 Ma Jacobsville Formation indicates Laurentia crossed the equator in the late Mesoproterozoic to early Neoproterozoic.
+The large arc distances between the Jacobsville pole and the poles of the Haliburton intrusions \cite{Buchan1976a} and the Wilberforce pyroxenite \cite{Palmer1973a} (black) indicate previous estimates of them being $>$1000 Ma \cite{Halls2015b} based on poor age and temperature calibration of their paleomagnetic and thermochronologic data are too old.
+\citeA{Brown2012a} developed poles (blue) from rocks of the Adirondack Mountains and estimated their ages to be ca.
+960 Ma to 990 Ma, with the gneiss pole (lightest blue) being the youngest and the granites pole (darkest blue) being the oldest.
+However, the estimated closure temperatures for rutile was too low, biasing the estimated pole age to be too old.
+The true age of the Grenville poles could be much older than previously thought in light of paleomagnetic and chronological constraints from the Jacobsville Formation.
+}
+\label{fig:pole_plot}
+\end{figure*}
+
+In light of the lithofacies model and detrital remanent magnetization data of the Jacobsville Formation, we prefer the interpretation that the deposition of the Jacobsville Formation initiated long after the the Oronto Group sedimentation and likely spanned a short time period close to the end of the ca.
+1010 to 980 Ma Rigolet phase of the Grenville Orogeny \cite{Rivers2012a, Hodgin2022a}.
+Therefore, we combine DRM directions of both polarities of the ca.
+990 Ma Jacobsville Formation to succinctly report one paleomagnetic pole for the formation.
+We reassess the amount of inclination shallowing of the combined directional dataset (Fig.
 \ref{fig:in_situ_pmag}).
@@
 \ref{fig:EI_results}.
-However, their mean pole developed from red beds near Sault Ste Marie is in the northern hemisphere with a latitude of 12\textdegree N (Fig.
+However, their mean pole developed from red beds near Sault Ste Marie lies in the northern hemisphere with a latitude of 12\textdegree N (Fig.
 \ref{fig:EI_results}).
-Despite the large uncertainty ellipse associated with this mean pole position, it is distinct from the mean pole position from Jacobsville rocks of the Keweenaw Peninsula and Marquette.
-Instead, it overlaps with the mean pole position of the ca.
-1070 Ma Upper Freda Formation (Fig.
+Despite the large uncertainty ellipse associated with this mean pole position, it is distinct from the mean pole position from Jacobsville rocks in the Keweenaw Peninsula and Marquette.
+Instead, it overlaps with the mean poles of the ca.
+1070 Ma Upper Freda Formation and the ca.
+1075 Ma Nonesuch Formation (Fig.
 \ref{fig:EI_results}; \citeA{Henry1977a}).
@@
 
-Deposition of the Bayfield Group sediments west to the Keweenaw Peninsula has been hypothesized to be coeval with the Jacobsville Formation but chronological constraints have been lacking.
-Our new paleomagnetic data from the Jacobsville Formation reaffirms that the pole positions of the the Chequamegon Sandstone of the Bayfield Group and the Jacobsville Formation are close albeit distinct (Fig.
+Deposition of the Bayfield Group sediments west to the Keweenaw Peninsula has been hypothesized to be coeval with the Jacobsville Formation but chronological constraints have been lacking \cite{Hamblin1958a, Kalliokoski1982a, Malone2016a}.
+Our new paleomagnetic data from the Jacobsville Formation confirm that the pole positions of the the Chequamegon Sandstone of the Bayfield Group and the Jacobsville Formation are close albeit distinct (Fig.
 \ref{fig:EI_results}).
-Given that the paleomagnetic pole from the Chequamegon Sandstone of the Bayfield Group (pole longitude=177.7\textdegree E, pole latitude=12.3\textdegree S, A95=4.6\textdegree; \citeA{McCabe1983a}) was calculated without applying inclination correction, the two formations could have even closer a temporal association than the current paleomagnetic pole configuration appears (Fig.
+Given that the paleomagnetic pole from the Chequamegon Sandstone of the Bayfield Group (pole longitude=177.7\textdegree E, pole latitude=12.3\textdegree S, A95=4.6\textdegree; \citeA{McCabe1983a}) was calculated without applying inclination correction, the two formations could have an even closer temporal association than what appears in the current paleomagnetic pole configuration (Fig.
 \ref{fig:EI_results}).
-In addition, the fact that only reversed polarity directions were recovered from the Chequamegon Sandstone could indicate that deposition of the sandstone collected by \citeA{McCabe1983a} could be coeval with one of the reversed polarity period recorded by the Jacobsville Formation (Fig.
+In addition, the fact that only reversed polarity directions were observed in the Chequamegon Sandstone indicate that deposition of the sandstone collected by \citeA{McCabe1983a} could be coeval with one of the reversed polarity periods recorded by the Jacobsville Formation (Fig.
 \ref{fig:strat_column}).
@@
 
-Laurentia's apparent polar wander is well constrained in the Mesoproterozoic by the Keweenawan Track thanks to the large number of paleomagnetic poles paired with high-precision geochronology from rocks of the ca.
-1.1 Ga Midcontinent Rift \cite{Swanson-Hysell2019a}.
-In addition, the Midcontinent Rift have had simple a thermal history for their old ages, as rifting ceased prior to lithospheric separation, preserving rocks of the rift within the continental interior far from continental margin and subsequent orogenesis.
-Since the end of the Midcontinent Rift mamatism, igneous paleomagnetic records from the the interior of Laurentia is lacking, and data from rocks associated with the ca.
-1080 to 980 Ma Grenville Orogeny have been used to infer for Laurentia's plate configuration in the late Mesoproterozoic to early Neoproterozoic \cite<e.g.>{Weil1998a, Halls2015b}.
-However, in contrast to rocks of the Midcontinent Rift which usually have straightforward paleomagnetic records \cite<e.g.>{Swanson-Hysell2019a}, rocks of the Grenville Province which experienced up to granulite facies metamorphism have much more complicated thermal histories and the timing of their magnetic remanence acquisition and alteration is more difficult to constrain.
+Constraining the paleogeography of Laurentia in the late Proterozoic is central to global paleogeography due to its central position in the ancient supercontinent Rodinia.
+During the initial assembly of Rodinia in the late Mesoproterozoic, Laurentia's apparent polar wander is well constrained by the ca.
+1.1 Ga Keweenawan Track thanks to the large number of paleomagnetic poles paired with high-precision geochronology from rocks of the Midcontinent Rift \cite{Swanson-Hysell2019a}.
+Since the end of the Midcontinent Rift magmatism, well-dated paleomagnetic records from the the interior of Laurentia is lacking, and data from rocks associated with the ca.
+1090 to 980 Ma Grenville Orogeny have been used to infer for Laurentia's plate configuration in the late Mesoproterozoic to early Neoproterozoic \cite<e.g.>{Weil1998a, Halls2015b}.
+However, in contrast to rocks of the Midcontinent Rift which usually have straightforward paleomagnetic records thanks to their simple thermal history \cite<e.g.>{Swanson-Hysell2019a}, rocks of the Grenville Province which experienced up to granulite facies metamorphism have much more complicated thermal histories and the timing of their magnetic remanence acquisition and alteration is more difficult to constrain.
 As is shown in Fig.
@@
 1070 Ma Freda Formation \cite{Henry1977a}.
-Such large discrepancies between pole positions associated with uncertainties in the age of the magnetization of the Grenville rocks led to the controversial ``Grenville Problem"---did the Grenville rocks acquire magnetic remanence prior to the continental collision and thus the pole discrepancies are a result of significant crustal shortening (i.e.
-the ``two plate model"; \citeA<e.g.>{Irving1972a, Palmer1973a, Buchan1973a, Halls2015b}), or did they acquire remanence after the continental accretion and thus the southerly poles from the Grenville rocks represent Laurentia's continued motion through more recent times (i.e.
+Such large distance between pole positions associated with uncertainties in ages of Grenville magnetizations led to the controversial ``Grenville Problem"---did the Grenville rocks acquire magnetic remanence prior to the continental collision and thus the pole discrepancies are a result of significant crustal shortening (i.e.
+the ``two plate model"; \citeA<e.g.>{Irving1972a, Palmer1973a, Buchan1973a, Halls2015b}), or did they acquire remanence after the continental accretion and thus the southerly poles from the Grenville rocks represent a coherent plate motion through more recent times (i.e.
 the ``one plate model"; \citeA<e.g.>{Roy1978a, McWilliams1975a, Weil1998a})?
 
-The two plate hypothesis is based on the interpretation that timing of magnetic remanence acquisition of the Grenville rocks was close to that of the Midcontinent Rift rocks \cite<e.g.>{Warnock2000a, Halls2015b}, but the lack of southerly paleomagnetic poles from the interior of Laurentia is interpreted to indicate that the Grenville rocks acquired remanence on allochthonous terranes before the final accretion to Laurentia craton.
-Thus, up to 4000 $\pm$ 1000 km of crustal shortening was invoked by \citeA{Halls2015b} to explain the large arc distance between the poles.
+The two plate model is based on the interpretation that the Grenville rocks acquired magnetizations on allochthonous terranes in the late Mesoproterozoic, prior to the final accretion to Laurentia craton, resulting in their apparent southerly paleomagnetic pole positions today (Fig.
+\ref{fig:pole_plot}).
+Further, the lack of paleomagnetic poles in the southern hemisphere from interior Laurentia at this time (i.e.
+the Keweenawan Track) is further interpreted to support such allochthonous magnetic remanence hypothesis \cite<e.g.>{Warnock2000a, Halls2015b}.
+A corollary of such interpretation is that the Grenville paleomagnetic poles could record a large amount of crustal shortening associated with the Grenvillian Orogeny.
+\citeA{Halls2015b} suggested the lost crust must be up to 4000 $\pm$ 1000 km long.
 
 Given the new constraints on the age and position of the Jacobsville pole (Fig.
-\ref{fig:pole_plot}), the two plate model would require that the majority of the crustal shortening happened after the Rigolet phase of the Grenville Orogeny, as the ca.992 Ma Jacobsville pole is still $\sim$45\textdegree\ arc distance ($\sim$5000 km) away from the paleomagnetic pole of the Haliburton intrusions \cite{Buchan1976a} and $\sim$38\textdegree\ ($\sim$4000 km) from the pole of the Wilberforce intrusions \cite{Palmer1973a} (Fig, \ref{fig:pole_plot}).
-However, geologic and geochronologic evidence from both within the Grenville Province \cite{Aleinikoff2021a} and from interior Laurentia \cite{Hodgin2022a} indicate that majority of the deformation in the Grenville Province and the propagation of its associated deformation in the Midcontinent Rift happened during the ca.1090-1020 Ma Ottawan phase orogeny, whereas smaller amount of convergence happened during the ca.
+\ref{fig:pole_plot}), the two plate model would require that the majority of the crustal shortening happened during or after the Rigolet phase of the Grenville Orogeny, as the ca.
+990 Ma Jacobsville pole is still $\sim$45\textdegree\ arc distance ($\sim$5000 km) away from the paleomagnetic pole of the Haliburton intrusions \cite{Buchan1976a} and $\sim$38\textdegree\ ($\sim$4000 km) from the pole of the Wilberforce intrusions \cite{Palmer1973a} (Fig, \ref{fig:pole_plot}).
+However, geologic and geochronologic evidence from both within the Grenville Province \cite{Aleinikoff2021a} and from interior Laurentia \cite{Hodgin2022a} indicate that majority of the deformation in the Grenville Province and the propagation of its compressional stresses into interior Laurentia happened during the ca.
+1090-1020 Ma Ottawan phase orogeny, whereas smaller amount of convergence happened during the ca.
 1010-980 Ma Rigolet phase.
-The lack of geological evidence for a striking amount of crustal shortening is supported by the slow average plate motion between the deposition of the upper Freda Formation and the deposition of the Jacobsville Formation (average $\sim$0.2\textdegree/Myr; Fig.
+The lack of geological evidence for a striking amount of crustal shortening is consistent with the slow average plate motion between the deposition of the upper Freda Formation and the deposition of the Jacobsville Formation (average $\sim$0.2\textdegree/Myr; Fig.
 \ref{fig:pole_plot}).
-Therefore, it is kinetically unlikely that the plate of Laurentia suddenly increased its motion during the waning phase of a collisional orogenesis in order to satisfy the Grenville pole positions.
+Therefore, it is kinetically unlikely that the plate of Laurentia suddenly increased its motion near the end of a collisional orogenesis in order to satisfy the Grenville pole positions.
 The plate kinetics interpretation based solely on paleomagnetic data and poorly constrained ages of magnetic remanence put forward by \citeA{Halls2015b} is implausible.
-Alternatively, the one plate model is much more straightforward in explaining the paleomagnetic pole positions of Grenville rocks as it interprets that the remanence was acquired during prolonged slow cooling of the orogeny that postdate the peak metamorphism.
-In contrast with the two plate model, the one plate model would predict that ages associated with the Grenville poles are much younger than ca.
-980 Ma as the rocks slowly cooled after the peak metamorphism associated with the collisional orogenesis (Fig.
+
+Alternatively, the one plate model is much more straightforward in explaining the paleomagnetic pole positions of Grenville rocks as it interprets that those rocks acquired remanence during prolonged exhumation of the orogeny that postdate the peak metamorphism.
+In contrast to the two plate model, the one plate model would predict that ages associated with the Grenville poles are much younger than ca.
+980 Ma (Fig.
 \ref{fig:pole_plot}).
+Adopting this model necessitates that the ages previously assigned to many Grenville poles be adjusted to much younger.
+Many previous paleomagnetic studies recognized that a large portion of the Grenville Province experienced up to granulite facies metamorphism (e.g.
+$>$700\textdegree C).
+The high metamorphism temperatures resulted in complete unblocking and overwriting of any previous remanence regardless of whether magnetite or hematite being the remanence carriers \cite<e.g.>{Ueno1975a, Alvarez1998a, Buchan1973a, Buchan1976a, Brown2012a}.
+However, what has not been fully taken into account in previous studies is the effective lowering of magnetic blocking temperatures of the new remanence.
+Theoretical calculations have shown that during prolonged cooling, magnetite which form above its Curie temperature (580\textdegree C) will subsequently acquire thermoviscous remanence \cite{Pullaiah1975a, Dodson1980a}, whereas hematite which often occur as exsolved lamellae in ilmenite in Grenville rocks acquired thermochemical remanence via magnetic ordering of the hemo-ilmenite below the Ne\'el temperature of hematite \cite{Burton1985a, Harrison2006b}.
 
-Adopting the one plate model to explain Laurentia's late Proterozoic apparent polar wander path necessitates that the ages assigned to many paleomagnetic poles derived from Grenville rocks need to be adjusted toward younger.
-Abundant petrological and geochronological evidence show large portion of the Grenville Province experienced up to granulite facies metamorphism (e.g.
-$>$700\textdegree C).
-Such high peak metamorphism temperatures would result in unblocking of any remanence the rocks had acquired earlier and complete overprint by remanence acquired during cooling after the orogenesis.
-Previous paleomagnetic studies have found that Grenville rocks commonly have magnetite and hematite being the remanence carriers \cite<e.g.>{Ueno1975a, Alvarez1998a, Buchan1973a, Buchan1976a, Brown2012a}.
-Theoretical calculations have shown that magnetite which formed above its Curie temperature (580\textdegree C) in Grenville rocks subsequently acquired thermoviscous remanence \cite{Pullaiah1975a, Dodson1980a} during prolonged cooling of the orogen, whereas hematite which often occur as exsolved lamellae interfingering with ilmenite acquired thermochemical remanence via magnetic ordering of the hemo-ilmenite below the Neel temperature of hematite \cite{Burton1985a, Harrison2006b}.
-
-Due to the long time scales involves with thermoviscous and thermochemical origin of the magnetic remanence, it is not straightforward to interpret the natural blocking temperatures of Grenville rocks based on observations of unblocking temperatures in the lab where rocks are heated and cooled rapidly.
-As a result, extrapolating equivalent natural blocking temperatures from lab unblocking temperatures using theoretical models \cite{Pullaiah1975a, Dodson1980a} and pairing the temperatures with thermochronology data is necessary to determine age of the remanence.
-\citeA{Brown2012a} recognized that the hemo-ilmenite exsolution occur at relatively low temperatures during cooling of the orogen and interpreted their new paleomagnetic data from rocks of the Adirondack Mountains, US, in context of recent understanding of the hemo-ilmenite phase diagram.
-That study assigned an age of ca.
+Due to the long time scales involved with thermoviscous and thermochemical origin of the magnetic remanence, it is not straightforward to interpret the natural blocking temperatures of Grenville rocks based on observations of unblocking temperatures in the lab where rocks are heated and cooled rapidly.
+\citeA{Brown2012a} studied Grenville rocks from the Adirondack Mountains.
+They recognized that the hemo-ilmenite exsolution occur at relatively low temperatures and acquire thermochemical remanence during cooling of the orogen and assigned an age of ca.
 960 Ma to a paleomagnetic pole developed from hemo-ilmenite-bearing microcline gneiss from northwestern Adirondacks (Fig.
 \ref{fig:pole_plot}).
-They assigned two additional paleomagnetic poles of ages ca.
-990 Ma and 970 Ma based on data from the Wanakena granite and anorthositic rocks associated with the Marcy Massif (Fig.
+Using the theoretical models of \citeA{Pullaiah1975a}, that study also extrapolated remanence ages of ca.
+990 and ca.
+970 Ma for the Wanakena granite and Marcy Massif anorthositic rocks whose dominant remanence carriers are magnetite grains that acquired thermoviscous remanence (Fig.
 \ref{fig:pole_plot}).
-Both units are interpreted to have magnetite as the dominant remanence carrier.
 However, in light of the ca.
-992 Ma Jacobsville pole position and the slow down of Laurentia's apparent polar wander since ca.
+990 Ma Jacobsville pole position and the slow down of Laurentia's apparent polar wander since ca.
 1070 Ma, the ages assigned to the poles in \citeA{Brown2012a} could be still too old to be compatible with the magnetization being acquired after the major pulse of the orogenic collision.
 
-That the estimated ages associated with the Adirondack paleomagnetic poles are biased old can be attributed to the outdated thermochronometer closure temperature calculations of \citeA{Mezger1991a, Mezger1992a} whose data was used by \citeA{Brown2012a}.
-In particular, the cooling history for the Adirondack area estimated by \citeA{Mezger1991a, Mezger1992a} show that the orogen had cooled to $\sim$400\textdegree C by the time of U-Pb system closure in rutile crystals.
-However, this estimated U-Pb closure temperatures range for rutile was not based on laboratory Pb diffusion experiment, but instead was an assumption that diffusion rate of Pb in rutile approximate that of Fe and Ti \cite{Mezger1989a}.
-Later experimental results \cite{Cherniak2000a} as well as analytical estimates \cite{Vry2006a, Kooijman2010a} both show that diffusion rates of Pb in rutile are in fact much slower, indicating that previous calculations significantly underestimated closure temperatures of the system.
+That the estimated ages associated with the Adirondack paleomagnetic poles are biased old can be attributed to the outdated thermochronometer closure temperature calculations of \citeA{Mezger1991a, Mezger1990a} whose data was used by \citeA{Brown2012a}.
+In particular, the cooling history for the Adirondack area estimated by \citeA{Mezger1991a, Mezger1990a} show that the orogen had cooled to $\sim$400\textdegree C by the time of U-Pb system closure in rutile crystals.
+However, this estimated U-Pb closure temperatures range for rutile was not based on laboratory Pb diffusion experiment, but instead was based on an assumption that thediffusion rate of Pb in rutile approximate that of Fe and Ti \cite{Mezger1989a}.
+Later experimental results \cite{Cherniak2000a} as well as analytical estimates \cite{Vry2006a, Kooijman2010a} show that diffusion rates of Pb in rutile are in fact much slower, indicating that previous calculations significantly underestimated closure temperatures of the system.
 Therefore, timing of magnetic remanence acquisition for Grenville rocks in the Adirondack Mountains should be younger than previously thought.
-
-The interpretation of young Grenville poles is supported by U-Pb apatite thermochronology data from iron oxide-apatite deposits of the eastern Adirondack Mountains \cite{Krestianinov2021a} and from the Wilberforce pyroxenite of the Wilberforce region, Canada \cite{Paul2021a}.
+The interpretation of young Grenville poles is supported by high-precision U-Pb apatite thermochronology data from iron oxide-apatite deposits of the eastern Adirondack Mountains \cite{Krestianinov2021a} and from the Wilberforce pyroxenite of the Wilberforce region, Canada \cite{Paul2021a}.
 Apatite from both regions yield ages of ca.
 900 Ma with associated closure temperatures $\sim$450\textdegree C \cite{Cherniak1991a, Chamberlain2001a}.
-Considering the thermoviscous remanence acquisition of magnetite and hematite (even after the exsolution of hemo-ilmenite lamellae) in the slowly cooled orogen \cite{Pullaiah1975a, Dodson1980a}, ca.
-900 Ma could be closer to the actual timing when rocks of the Grenville Province finally blocked in the remanence that is preserved today.
-However, the exact age of each paleomagnetic pole recorded by rocks of different domains of the Grenville Province remains difficult to constrain.
-Complications associated with the origin of the dated iron oxide-apatite ore, crystal-size dependence of closure temperature of accessory minerals such as rutile and apatite \cite{Smye2018a, Chamberlain2001a}, and the complex metamorphic histories in different domains of the Grenville Province require that more high-precision thermochronology data with detailed closure temperature calibration based on modern experimental results are needed.
+Considering delayed remanence acquisition of magnetite and hematite, ca.
+900 Ma should be closer to the actual timing when rocks of the Grenville Province finally blocked in their remanence that is preserved today.
+However, the exact age of each paleomagnetic pole recorded by rocks of different domains of the Grenville Province remains largely uncertain.
+Complications such as the origin of the dated iron oxide-apatite ore, crystal-size dependence of closure temperature of accessory minerals including rutile and apatite \cite{Smye2018a, Chamberlain2001a}, and the complex metamorphic histories in different domains of the Grenville Province require that more high-precision thermochronology data with modern closure temperature calibration are developed.
 
 Overall, albeit that numerous paleomagnetic data has been developed from rocks of the Grenville Province, few are paired with high-precision thermochronology data that are calibrated by modern diffusion paramters.
-More high-precision thermochronology data and statistical models that properly incorporates uncertainties associated with timing of closure of thermochronometers and with temperatures of magnetic remanence acquisition in slowly cooled orogen are needed in order to further calibrate the age of those poles.
-Refining the pole-age pairs for the Grenville rocks can further test the hypothesized ``Grenville Loop" of Laurentia's APWP \cite{Irving1974a} and refine the paleogeographic configuration during the development of super continent Rodinia in the Neoproterozoic.
-
-% Warnock on the age of magnetization on the Haliburton pole: based on the very coarse thermal demagnetization steps Warnock2000a assigned that all of the primary magnetite components unblocked above 550 \textdegree C, with an uncertainty of 555 \textdegree C $\pm$ 5\textdegree C.
-Using a thermal history model that combines U-Pb titanite ages, hornblende and biotite Ar/Ar ages and just by merely compare the thermal history model with the assigned narrow magnetite unblocking temperature range, the authors assigned the age of the magnetization to be 1015 $\pm$ 15 Ma.
-
-% It is important to note that the result Ar ages of the hornblende and biotite are 983 $\pm$ 8 Ma, 996 $\pm$ 8 Ma, and 998 $\pm$ 8 for hornblende with unblocking temperature of 530 $\pm$ 40\textdegree C, consistent across metamorphic terranes (Bancroft).
-And the biotite Ar ages are 910 $\pm$ 7 Ma, 891 $\pm$ 7 Ma, corresponding to unblocking temperature of 280 $\pm$ 40\textdegree C.
-
-% More, the thermal demagnetization plots shown in Warnock2000a suggest that the magnetite unblocking lower temperature boundary can be around 500\textdegree C, lower than 555 $\pm$ 5\textdegree C.
-In addition, given the case that the rocks underwent uniform rate of cooling due to post-formation uplifting, a slow cooling effect from 555\textdegree C to 300\textdegree C from 990 Ma to 900 Ma allows for a scenario where the magnetite remament magnetization that appear to have a 555\textdegree C unblocking temperature in lab was acquired through a period of --- myr cooling though 500\textdegree C during uplifting.
-In this case, the magnetite remanence acquisition can be interpreted to be closer to the biotite Ar age than previously thought.
-Therefore, the Haliburton intrusion pole could be interpreted to be a post Rigolet orogenic activity phase magnetization acquisition, and the apparent discrepancy between the Haliburton pole and the Jacobsville pole can be explained by Laurentia's whole plate motion after the Grenvillian orogeny from 990 Ma to 900 Ma.
-One corollary of such scenario is that Laurentia underwent plate motion at a rate of ---\textdegree/myr, which is similar to that of today's tectonic plate motion.
+More high-precision thermochronology data and statistical models that properly incorporates uncertainties associated with timing of closure of thermochronometers and with magnetic remanence acquisition temperatures in the slowly cooled orogen are needed in order to further calibrate the age of those poles.
+Further development of paired pole-age data from the rich suite of metamorphic rocks in the Grenville Province can help update Laurentia's APWP \cite{Irving1974a} and the paleogeographic configuration of super continent Rodinia in the Neoproterozoic.
 
 \section*{Conclusion}
+
+Our new paleomagnetic pole from the ca.
+990 Ma Jacobsville Formation passes paleomagnetic field tests and has robust radiometric age constraints.
+This high-quality pole is a crucial addition to Laurentia's apparent polar wander path in the late Proterozoic.
+It extends the Mesoproterozoic Keweenawan Track (Fig.
+\ref{fig:pole_plot}) by showing that the plate slowed down significantly as the Grenville collisional orogenesis commenced on Laurentia's margin.
+It also sets an anchor for the position of supercontinent Rodinia in early Neoproterozoic due to Laurentia's central position in the configuration.
+Being a robust record from rocks in interior Laurentia, the Jacobsville pole implies that ages associated with the paleomagnetic poles recorded by metamorphic rocks of the Grenville Province might need to be adjusted significantly.
+Further paired studies of high-quality paleomagnetism and high-precision thermochronolgy are needed to illuminate Laurentia's development toward mid Neoproterozoic and its connection with conjugate continents such as Baltica.
 

### 2022-07-23 Manuscript/Manuscript.tex
Updates from Overleaf

@@
 \begin{abstract}
-Coeval with the cessation of the ca.
-1.1 Ga intracontinental Midcontinent Rift was the onset of the ca.
-1090-980 Ma Grenvillian Orogeny along Laurentia's eastern margin.
-Far-field compressional stresses of the orogeny propagated into the interior of Laurentia and inverted the Midcontinent Rift, deforming the post-rift Jacobsville Formation whose maximum depositional age has been constrained to be ca.
-990 Ma.
-We develop a new paleomagnetic pole from red beds of the Jacobsville Formation in northern Michigan, USA.
+The paleogeography of Laurentia in the late Proterozoic is central to global paleogeography due to its central position in the supercontinent Rodinia.
+We develop a new paleomagnetic pole from red beds of the ca.
+990 Ma Jacobsville Formation which serves as a crucial anchor point for Laurentia's position in the earliest Neoproterozoic.
 High-resolution thermal demagnetization experiments resolve detrital remanent magnetizations that pass a conglomerate test.
 An inclination corrected paleomagnetic mean pole position from the Jacobsville Formation confirm continued plate motion of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic.
-The new pole from the interior of Laurentia indicates the paleomagnetic poles developed from the Grenville Province most likely record coherent plate motion after major continental collision during the initial assembly of the supercontinent Rodinia.
+The new pole from the interior of Laurentia indicates the plate slowed down significantly following the onset of the Grenvillian orogenesis on its eastern margin.
+This pole's position and age indicate the paleomagnetic poles developed from the Grenville Province most likely record coherent plate motion after major continental collision during the initial assembly of the supercontinent Rodinia.
 \end{abstract}

### 2023-03-29 Manuscript/Manuscript.tex
update manuscript tex files

@@
 \begin{document}
+\title{Tracking Rodinia into the Neoproterozoic: new paleomagnetic constraints from the Jacobsville Formation}
 
-\title{Laurentia in the earliest Neoproterozoic: new paleomagnetic constraint from the Jacobsville Formation}
-
-\authors{Yiming Zhang\textsuperscript{1}, Blake Hodgin\textsuperscript{1}, Nicholas Swanson-Hysell\textsuperscript{1}, James Pierce\textsuperscript{1,2}, Anthony Fuentes\textsuperscript{1}}
+\authors{Yiming Zhang\textsuperscript{1}, Eben B.
+Hodgin\textsuperscript{1,2}, Tadesse Alemu\textsuperscript{1}, James Pierce\textsuperscript{1,3}, Anthony Fuentes\textsuperscript{1}, Nicholas L.
+Swanson-Hysell\textsuperscript{1}}
 
 \affiliation{1}{Department of Earth and Planetary Science, University of California, Berkeley, CA, USA}
-\affiliation{2}{Department of Earth and Planetary Sciences, Yale University, CT, USA}
+\affiliation{2}{Department of Earth, Environmental, and Planetary Science, Brown University, Providence, RI, USA}
+\affiliation{3}{Department of Earth and Planetary Sciences, Yale University, CT, USA}
 
@@
 \begin{keypoints}
-\item A new paleomagnetic pole from the ca. 990 Ma Jacobsville Formation \item Laurentia crossed the equator in the latest Mesoproterozoic to earliest Neoproterozoic
-\item The new pole indicates the Grenville Loop could be much younger than thought
+\item Siltstones and fine-grained sandstones of the ca. 990 Ma Jacobsville Formation have primary magnetization held by detrital hematite.
+\item Following closure of the Unimos Ocean and onset of the Grenvillian orogeny in the late Mesoproterozoic, Laurentia's motion significantly slowed.
+\item The pole position and age suggests that the Grenville Loop paleomagnetic poles are younger than current interpretations.
 \end{keypoints}
@@
 \begin{abstract}
-The paleogeography of Laurentia in the late Proterozoic is central to global paleogeography due to its central position in the supercontinent Rodinia.
-We develop a new paleomagnetic pole from red beds of the ca.
-990 Ma Jacobsville Formation which serves as a crucial anchor point for Laurentia's position in the earliest Neoproterozoic.
-High-resolution thermal demagnetization experiments resolve detrital remanent magnetizations that pass a conglomerate test.
-An inclination corrected paleomagnetic mean pole position from the Jacobsville Formation confirm continued plate motion of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic.
-The new pole from the interior of Laurentia indicates the plate slowed down significantly following the onset of the Grenvillian orogenesis on its eastern margin.
-This pole's position and age indicate the paleomagnetic poles developed from the Grenville Province most likely record coherent plate motion after major continental collision during the initial assembly of the supercontinent Rodinia.
+The paleogeography of Laurentia throughout the Neoproterozoic is critical for reconstructing global paleogeography due to its central position in the supercontinent Rodinia.
+We develop a new paleomagnetic pole from red siltstones and fine-grained sandstones of the early Neoproterozoic Jacobsville Formation which is now constrained to be ca.
+990 Ma in age.
+High-resolution thermal demagnetization experiments resolve detrital remanent magnetizations held by hematite.
+These directions were reoriented within siltstone rip-up clasts and pass an intraformational conglomerate test---giving confidence that the magnetization is detrital and primary.
+An inclination-corrected mean paleomagnetic pole position for the Jacobsville Formation indicates that Laurentia's motion slowed down significantly following the onset of the Grenvillian orogeny.
+Prior rapid plate motion associated with closure of the Unimos Ocean transitioned to slow drift of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic.
+We interpret the distinct position of this well-dated pole from those in the Grenville orogen that have been assigned a similar age to indicate that the ages of the poles associated with the Grenville Loop likely need to be revised to be younger due to prolonged exhumation.
 \end{abstract}
 
-% \section*{Plain Language Summary}
-% [ enter your Plain Language Summary here or delete this section]
+\section*{Plain Language Summary}
 
-\section*{Keywords}
-Rodinia\\
-Laurentia  \\
-Paleogeography\\
-Neoproterozoic\\
-\
-
-apparent polar wander path\\
-Inclination shallowing\\
-Grenville\\
-Keweenawan superchron\\
-Red bed\\
+There have been times in Earth history when many of the continents are joined together into a big continent that is called a supercontinent.
+The most recent supercontinent is Pangea which had Africa at its center.
+An older supercontinent called Rodinia formed about 1 billion years ago and had North America at its center.
+The old parts of North America that were in the supercontinent are called Laurentia.
+The position of Laurentia is key to understanding where Rodinia was and how different continents were connected in the supercontinent.
+We now know the age of sedimentary rocks called the Jacobsville Formation that were deposited in ancient rivers and can use these rocks to determine the past position of Laurentia (and therefore Rodinia) 990 million years ago using paleomagnetism.
+This position is quite different in current interpretations which rely on data from rocks that cooled slowly within an ancient mountain belt that formed when the supercontinent assembled---they likely show where the supercontinent was after 990 Ma.
+As a result, we have a new understanding of where the supercontinent was that can enable further progress in putting all the pieces together to further understand the ancient Earth.
 
@@
 
-The extensively studied paleomagnetic records of the ca.
-1109 Ma to 1084 Ma North American Midcontinent Rift (i.e.
-the Keweenawan Track) provide crucial constraints on the paleogeography of Laurentia in the Mesoproterozoic \cite{Swanson-Hysell2019a}.
-As a result, the Keweenawan Track has become a central record for reconstructing the initial development of the ancient supercontinent Rodinia \cite{2021a}.
-As the rifting ceased and subsequently inverted following the onset of the Grenvillian Orogeny on Laurentia's eastern margin (Fig.
-\ref{fig:Geologic_map}), magmatism in the interior of Laurentia entered a period of quiescence that could have lasted for as long as $\sim$300 Myr until the emplacement of the ca.
-780 Ma Gunbarrel large igneous province \cite{Harlan2003a, Mackinder2019a}.
-Due to this magmatic gap, Laurentia and its conjugate continents' paleogeographic configuration in the late Mesoproterozoic to mid-Neoproterozoic has remained largely uncertain \cite{Swanson-Hysell2021c}.
+The extensively studied paleomagnetic records of ca.
+1109 to 1084 Ma volcanics and intrusions associated with the North American Midcontinent Rift provide crucial constraints on the paleogeography of Laurentia in the late Mesoproterozoic Era \cite{Halls1982a, Fairchild2017a, Swanson-Hysell2019a}.
+The resulting sequence of poles known as the Keweenawan Track is a central record for reconstructing the assembly of the ancient supercontinent Rodinia \cite{Evans2021b}.
+An advantage of intracratonic magmatic events, such as those preserved in the Midcontinent Rift, is that they lead to emplacement of igneous rocks that can readily retain primary thermal remanent magnetization (TRM) that can be confidently associated with the emplacement age of the unit.
+However, following the development of the Midcontinent Rift, there was a $\sim$300 Myr quiescence in known intracratonic magmatic activity in Laurentia that lasted until the emplacement of the ca.
+775 Ma Gunbarrel large igneous province \cite{Harlan2003a, Mackinder2019a, Swanson-Hysell2021c}.
+The lack of Laurentian rocks with primary TRM in the early Neoproterozoic has led paleogeography reconstructions to be reliant on magnetizations acquired by rocks that were metamorphosed during the ca.
+1090 to 980 Ma Grenvillian orogeny \cite{Rivers2012a, Swanson-Hysell2023a}.
+However, rocks within the orogen acquired their magnetizations during exhumation long after peak metamorphic conditions \cite{McWilliams1975a}.
+As a result, the timing of remanence acquisition needs to be constrained through the more difficult task of dating cooling associated with exhumation.
+This gap in well-dated paleomagnetic poles for Laurentia in the late Mesoproterozoic to mid-Neoproterozoic contributes to uncertainty in paleogeography at that time including the configuration of the supercontinent Rodinia.
 
@@
 \includegraphics[width=\textwidth]{Geologic_map.pdf}
-\caption{(A) Regional map showing location of the Grenville Front relative to the study area.
+\caption{(A) Regional map showing the extent of the Midcontinent Rift and the location of the Grenville Front relative to the study area.
 The inset green box shows the extent of panel B.
-(B) Geologic map of northern Michigan and Wisconsin showing Midcontinent Rift igneous and sedimentary rocks, the Jacobsville Formation, other older Proterozoic rocks, and major post-rift thrust faults.
-Jacobsville stratigraphic sections along which paleomagnetic samples were collected in this study are shown as green diamond symbols.}
+(B) Geologic map of northern Michigan and Wisconsin showing Midcontinent Rift igneous and sedimentary rocks, the Jacobsville Formation, sedimentary rocks of the Bayfield Group, other older Proterozoic rocks, and major post-rift thrust faults.
+The location of the Jacobsville stratigraphic sections in this study are shown with green diamonds.
+Map modified from \citeA{Nicholson2004a}.}
 \label{fig:Geologic_map}
@@
 
-When igneous rock records are lacking, paleomagnetic data recorded by sedimentary rocks (e.g.
-red beds) can provide useful insights into the ancient geomagnetic field and record paleogeographic positions of sedimentary basins.
-In the Midcontinent Rift, the paleomagnetic poles developed from the ca.
-1075 Ma Nonesuch Formation and the ca.
-1070 Ma upper Freda Formation of the Oronto Group sediments that overlie the Midcontinent Rift volcanics in southern Lake Superior area (Fig.
-\ref{fig:Geologic_map}) provided crucial constraints for Laurentia's paleolatitude, orientation, and plate speed in the late Mesoproterozoic \cite{Henry1977a, Swanson-Hysell2019a, Rose2022a}.
+Paleomagnetic poles can also be developed from sedimentary rocks deposited in basins.
+Sedimentary rocks of the Oronto Group deposited during post-rift thermal subsidence following the bulk of Midcontinent Rift magmatic activity provide paleomagnetic poles that extend the Keweenawan Track to ca.
+1070 Ma \cite{Henry1977a, Slotznick2023a}.
+While there are some contemporaneous basins in northern Laurentia \cite{Greenman2021a}, there is minimal basin development within Laurentia following the Grenvillian orogeny until the Rodinia supercontinent began to rift apart ca.
+770 Ma \cite{Macdonald2023a}.
 
-On top of the Oronto Group sediments overlies the post-rift Jacobsville Formation which was partly deformed during the inversion of the Midcontinent Rift (Fig.
+One rare preservation of an early Neoproterozoic sedimentary succession in Laurentia's interior is the Jacobsville Formation (Fig.
 \ref{fig:Geologic_map}; \citeA{Hamblin1958a, Hodgin2022a, DeGraff2022a}).
-U-Pb detrital zircon ages and U-Pb calcite ages from the Keweenaw fault where the Jacobsville Formation is deformed against thrusted rift volcanics constrain the maximum depositional age of the Jacobsville Formation to be ca.
+A 992.51$\pm$0.64 Ma (2$\sigma$ analytical uncertainty) U-Pb detrital zircon date from the Jacobsville Formation constrains its maximum deposition age \cite{Hodgin2022a}.
+A 985.5$\pm$35.8 Ma (2$\sigma$ analytical uncertainty) U-Pb date from calcite veins within the Keweenaw fault provides a minimum age constraint on Jacobsville deposition, as the Jacobsville Formation is folded in the footwall against Midcontinent rift volcanics in the hanging wall.
+Together, these dates bracket the depositional age of the Jacobsville Formation and constrain it to be ca.
 990 Ma \cite{Hodgin2022a}.
-The Jacobsville Formation contains abundant hematite-rich red beds which bear potential to further extend the apparent polar wander path (APWP) of Laurentia toward more recent times \cite{Hamblin1958a, Roy1978a}.
-Developing paleomagnetic data from the Jacobsville Formation will allow further development of Rodinia's paleogeographic evolution in the late Mesoproterozoic to early Neoproterozoic.
+These results constrain the formation to have been deposited within a syn-orogenic basin during the Rigolet phase of the Grenvillian orogen and to have been deformed during the same orogenic phase \cite{Hodgin2022a, Swanson-Hysell2023a}.
 
-Challenges exist in interpreting sedimentary paleomagnetic records.
-Primary detrital remanence acquired through settling and deposition of magnetic particles in the water column can be masked by secondary remanence acquired through precipitation and growth of new minerals within pore spaces of host sedimentary rocks syn- to post-deposition (e.g.
-pigmentary hematite).
-In red beds, it has been found that formation of the pigmentary hematite can significantly post-date the timing of sediment deposition, resulting in a chemical remanent magnetization (CRM) superimposed upon primary detrital remanent magnetization (DRM) carried by detrital grains \cite{Butler1992a}.
-To isolate the remanence components of the red beds of the Jacobsville Formation, \citeA{Roy1974a, Roy1978a} adopted a method first applied by \citeA{Collinson1965c} to preferentially remove fine-grained pigmentary hematite through prolonged immersion in concentrated HCl acid whereas leaving the coarser detrital grains that are more resistant to dissolution in place.
-Studies applying paired acid etching demagnetization and thermal demagnetization show that the pigmentary hematite grains tend to have lower unblocking temperatures than detrital ones \cite{Tauxe1980a, Bilardello2010c}.
-High-resolution thermal demagnetization experiments with temperature intervals as small as 1-2 \textdegree C have shown to be effective in isolating detrital remanence from chemical remanence as coarser hematite grains tend to have higher unblocking temperatures closer to the Ne\'el temperature of hematite \cite{Swanson-Hysell2019b}.
-In particular, a conglomerate test on a suite of fluvial intraclasts of the red beds in the Freda Formation shows that via thermal demagnetization one can isolate detrital magnetization vectors (which are reoriented in random directions in the clasts) from the secondary components carried by co-occurring authigenic hematite (which yield similar directions amongst clasts) \cite{Swanson-Hysell2019b}.
+The Jacobsville Formation contains abundant hematite-bearing sandstone and siltstone which have the potential to record a key pole in Laurentia's apparent polar wander path (APWP) \cite{Dubois1962a, Roy1978a}.
+\citeA{Roy1978a} conducted a paleomagnetic study that developed a pole which has been used in paleogeographic syntheses where it has been assigned variable ages including an age of ca.
+1020 Ma based on APWP extrapolation \cite<e.g.>{Li2008a}.
+However, the uncertainty in its age combined with a lack of field tests, and no application of inclination correction have resulted in the pole not being included in recent curated compilations \cite<e.g.>{Evans2021a}.
+Obtaining high-quality paleomagnetic data, including robust field tests, motivates the development of new data from the Jacobsville Formation.
 
-Paleomagnetism of red beds has also been challenged by the controversial issue of correcting for inclination shallowing \cite{King1955a, Tauxe2004b, Bilardello2016b}.
-Due to rotation of detrital magnetic grains caused by authigenic processes such as compaction during lithification, inclinations recorded by sediments often have shallower angles than the local field in which they are deposited.
-To estimate the amount of inclination shallowing in a given package of sedimentary rocks, \citeA{Tauxe2004b} developed a statistical model (i.e.
-the elongation/inclination [E/I]method) that evaluates the deviation of the distribution of a large number ($>$100) of observed sedimentary paleomagnetic directional data with the predicted distribution given by a paleosecular variation model developed for the past 5 Myr (TK03) under the the geocentric axial dipole (GAD) hypothesis.
-Although the TK03 paleosecular variation model is based on data of relatively recent geomagnetic field variations, it has been shown to be compatible with data from the ca.
-1.1 Ga lava flows of the Midcontinent Rift \cite{Tauxe2009a}.
-Further, \citeA{Pierce2022a} showed that the E/I method passes a field test for the 1.1 Ga Cut Face Creek Sandstone in the Midcontinent Rift as it successfully corrects the shallow paleomagnetic inclinations of the sediments to that of the North Shore Volcanic Group lava flows that bracket them.
-That study also further extended the method to better represent uncertainties associated with the amount of inclination shallowing when reporting sedimentary paleomagnetic pole positions by using the spherical elliptical Kent distribution \cite{Kent1982a} instead of the circular Fisher distribution \cite{Fisher1953a}.
+In this study, we investigate five stratigraphic sections of the Jacobsville Formation on and near the Keweenaw Peninsula, Michigan, USA (Figs.
+\ref{fig:Geologic_map} and \ref{fig:strat_column}).
+High-resolution thermal demagnetization isolates primary detrital remanent magnetization (DRM) from secondary chemical remanent magnetization (CRM) within analyzed samples.
+This interpretation is supported by the DRM directions passing an intraformational conglomerate test on siltstone intraclasts that were ripped-up and remobilized within the fluvial depositional environment.
+A paleomagnetic fold test further constrains the timing of DRM acquisition to predate the folding of the Jacobsville Formation in the footwall of the Keweenaw Fault soon after deposition.
+With these data, we develop a new inclination-corrected paleomagnetic pole for the Jacobsville Formation based on a large number of samples that gives new constraints on the paleogeographic position of Laurentia in the earliest Neoproterozoic.
 
-In this study, we investigate the hematite-bearing, fine-grained sandstone to siltstone lithofacies along five stratigraphic sections of the Jacobsville Formation on Keweenaw Peninsula, Michigan, USA.
-High-resolution thermal demagnetization isolate primary DRM from secondary CRM with the DRM passing a conglomerate test.
-We then use a paleomagnetic fold test to further constrain the timing of DRM acquisition in association with the structural inversion of the Midcontinent Rift.
-Finally, we develop a new inclination-corrected paleomagnetic pole for the Jacobsville Formation based on a large number of samples and investigate the paleo-continental configuration of Laurentia in the late Mesoproterozoic to early Neoproterozoic.
+\section*{Geologic setting}
 
-% The end of the ca.
-1109 Ma to 1084 Ma intracontinental Midcontinent Rift magmatism and extension was marked by a transition to deposition of sedimentary rocks of the Oronto Group \cite{Ojakangas2001a} as the rift basin thermally subsided prior to the rift undergoing contractional deformation in the later stages of the Grenvillian orogeny \cite{Cannon1990b, Hodgin2022a}.
-
-%  that has become central to the reconstruction of paleogeography of the assembly of supercontinent Rodinia.
-In addition to syn-rift rocks preservation, the cessation and later inversion of the rift allowed the development of syn-to post-rift basin development which accommodated the deposition of the ca.
-1080 - 1050 Ma Oronto Group sedimentary rocks that are in total about 5 km thick.
-Overlying the Oronto Group is the Jacobsville formation which has been interpreted to be deposited in a back-bulge setting associated with the far-field propagation of the crustal-scale deformation caused by the Grenvillian Orogeny.
-
-% A leading tectonic model for the cessation of the Midcontinent Rift and extension is that it occurred due to far-field compressional stress associated with the ca.
-1090-980 Ma Grenvillian orogeny \cite{Cannon1990a}, when the collision of Laurentia with its conjugate margins led to the development of a back-bulge basin, accommodating the deposition of the Jacobsville Formation sediments as the inversion of the Midcontinent Rift progressed.
-Recent re-examination of the maximum deposition age of the Jacobsville Formation using zircon U-Pb whole grain chemical abrasion isotope dilution thermal ionization mass spectrometry has updated the maximum deposition age of the Jacobsville Formation to be ca.
-990 Ma \cite{Hodgin2022a}.
-Such age constraint is consistent with backbulge subsidence model \cite{DeCelles, 2012}, indicating that the Jacobsville Formation may have been deposited in a Grenvillian foreland basin system that resulted from lithospheric flexure induced by orogenic loading \cite{Rivers2012a}.
-The relatively young age of the Jacobsville Formation compared to other Midcontinent Rift rocks makes the Jacobsville Formation an crucial target for further reconstructing the global paleogeography in the late Proterozoic.
-For further refining the paleogeographic connection between the poles of the Grenville Loop and those of the Keweenawan Track.
-
-% Paleomagnetic data for the Jacobsville Formation have led to a paleomagnetic pole position that appears to be a continuation of the Keweenawan Track from the Oronto Group poles (Roy and Robertson, 1978; Fig.
-7; Table 4).
-As a result, its age has been interpolated to be latest Mesoproterozoic (e.g., 1020 Ma in Li et al., 2008) such that it was deposited prior to the ca.
-1000 Ma Grenville loop poles.
-However, this age assignment has recently been questioned on the basis of laser-ablation–inductively cou- pled plasma–mass spectrometry (LA-ICP-MS) U-Pb detrital zircon dates from Jacobsville Sandstone samples.
-Malone et al.
-(2016) interpreted a maximum depositional age of 959 ± 19 Ma based on the weighted mean of the four youngest grains from 2050 LA-ICP-MS zircon dates.
-A similar age was proposed on the basis of the youngest LA-ICP-MS U-Pb dates in the study of Craddock et al.
-(2013).
-This young maximum interpreted age is intriguing because it suggests the unconformity above the Freda Sandstone spans at least 100 m.y.
-and was followed by continued sedimentation in the region long after rift activity.
-Nevertheless, the paleo- magnetic pole of the Jacobsville Sandstone (Roy and Robertson, 1978) appears to extend the Keweenawan Track from the Oronto Group poles in a manner that may be more consistent with deposition of the Jacobsville Formation in the late Mesoproterozoic.
-Given the conflict with this inferred age, the large uncertainty of individual LA-ICP-MS dates, and the possibility of outlier dates in such a large data set, it is beneficial to interpret the age of the youngest Jacobsville zircons dated CA-ID-TIMS to obtain more precise constraints.
-Using CA-ID-TIMS, a recent study by Hodgin et al (2022) redefines the maximum deposition age of the Jacobsville Formation as ca.
-993 Ma.
-
-% As the Midcontinent Rift magmatic activity wanes its strength and evolved into a failed intracontinental rift, Laurentia experienced a long period of magmatic quiescence in its interior, where basin subsidence dominated the rift basin.
-This was followed by the Grenville Orogeny, causing the the inversion of the rift along thrust belts such as the Keweenaw Fault (Fig.
-xxx).
-The lack of extensive magmatism from ca.1080 Ma to ca.
-980 Ma in the Laurentia interior resulted in a significant gap (for at least 50 myr) of paleomagnetic poles between the Keweenawan Track (developed from interior Laurentia igneous and sedimentary rocks) and the Grenville Loop (poles developed from the Grenville Province rocks) (Fig.
-xxx; \cite{Swanson-Hysell2019a}).
-
-% Due to the lack of pmag records from the interior Laurentia, various hypotheses for the spatial relationship between the end of the Keweenawan Track and the Grenville Loop have been debated.
-One model argues for that the difference in pole locations indicates that the Grenville Province was once thousands of kilometers away from Interior Laurentia and was progressively approaching the southern margin of Laurentia in the late Mesoproterozoic \citep{Halls2015}.
-A corollary of this hypothesis is that there had to be a 4000 km shortening during the Grenville orogeny - colliding Amazonia-Laurentia.
-The other model favors that the Grenville Loop is a continuation of the Keweenawan Track - that the southerly continental drift of Laurentia continued across the equator onto the southern hemisphere from ca.
-1070 Ma to ca.1000 Ma.
-and  approached this problem by developing data from metamorphosed dike and anorthosite and argued for Grenville
-
-\section*{Geologic settings}
-
-\subsection*{Structure}
 The ca.
@@
 \ref{fig:Geologic_map}).
-However, the rifting failed to separate Laurentia into two continents.
-As its magmatism waned and thermal subsidence progressed in the rift basin, the $>$4 km of ca.
-1085-1050 Ma Oronto Group sediments \cite{Fairchild2017a, Swanson-Hysell2019a} were deposited conformably on top of the rift volcanics (Fig.
-\ref{fig:Geologic_map}; \citeA{Cannon1989a}).
-Subsequently, the rift was inverted as the ca.
-1090-980 Ma Grenvillian Orogeny commenced on the eastern margin of Laurentia (Fig.
-\ref{fig:Geologic_map}), eventually resulting in the propagation of far-field compressional stresses deep into the interior of Laurentia \cite{Cannon1993a, Cannon1994a}.
+Following the end of active extension in the rift, thermal subsidence resulted in the accumulation of $>$4 km of ca.
+1085-1050 Ma Oronto Group sedimentary rocks that are both intercalated with the youngest volcanics and deposited atop them (Fig.
+\ref{fig:Geologic_map}; \citeA{Daniels1982a, Cannon1989a}).
+Subsequently, the rift was inverted as contraction associated with the ca.
+1090-980 Ma Grenvillian Orogeny on the eastern margin of Laurentia (Fig.
+\ref{fig:Geologic_map}) propagated into Laurentia's interior \cite{Cannon1993a, Cannon1994a, Hodgin2022a, Swanson-Hysell2023a}.
 
-During the rift inversion, Paleoproterozoic and Archean crust was uplifted via thrust faults such as the Marenisco fault and the Pelton Creek fault in the south, forming the crustal-scale Montreal River monocline (Fig.
+During rift inversion, Midcontinent Rift volcanic and sedimentary rocks were uplifted along with Paleoproterozoic and Archean lithologies via thrust faults such as the Marenisco fault, forming the crustal-scale Montreal River monocline (Fig.
 \ref{fig:Geologic_map}; \citeA{Cannon1993a}).
-The southern part of the post-rift Jacobsville Formation overlies an erosional angular unconformity that developed on lithologies that were exhumed through this earlier reverse motion (Fig.
+Rb-Sr radiometric thermochronology data of \citeA{Cannon1993a} show Archean lithologies within the Montreal River monocline to have been exhumed to mid to shallow crustal temperatures ($\sim$270\textdegree C) by ca.
+1050 Ma.
+The Jacobsville Formation overlies an erosional angular unconformity that developed on lithologies that were exhumed through this earlier episode of contractional deformation associated with Grenvillian orogenesis (Fig.
 \ref{fig:Geologic_map}; \citeA{Hamblin1958a, Cannon1993a, Kalliokoski1982a}).
-North to these faults, Midcontinent Rift volcanic rocks were thrusted and folded along the Keweenaw fault which extends $\sim$250km from northern Wisconsin to northern Keweenaw Peninsula (Fig.
-\ref{fig:Geologic_map}; \citeA{DeGraff2022a}).
-Final inversion of the Keweenaw fault during the Rigolet phase of the Grenvillian Orogeny deformed part of the Jacobsville strata near the fault \cite{Hodgin2022a}.
-Recent mapping of the Jacobsville Formation also found that it can sometimes onlap the Midcontinent Rift volcanics within or adjacent to the Keweenaw Fault system near the northern end of the Keweenaw Peninsula (Fig.
-\ref{fig:Geologic_map}; \citeA{Tyrrell2019a, Mueller2021a}).
+Conglomeratic facies can occur in the basal Jacobsville Formation, with clasts derived from locally uplifted basement lithologies \cite{Irving1885a, Hamblin1958a, Kalliokoski1982a, Alemu2023a}.
 
-Field observations find erosional contact and development of paleosol between the Jacobsville Formation and underlying volcanic rocks \cite{Hamblin1958a, Kalliokoski1975a}, and interpretations of seismic reflection data show that sedimentary rocks interpreted to be Jacobsville Formation lie unconformably on the Oronto Group \cite{Cannon1989a}.
-These data indicate that there was a prolonged period of sedimentation quiescence and net erosion between the end of Oronto Group deposition and onset of Jacobsville Formation deposition.
-Rb-Sr radiometric thermochronology data of \citeA{Cannon1993a} show the Montreal River monocline in which the Freda Formation and older basement rocks are deformed had been uplifted to mid to shallow crustal levels by ca.
-1060 Ma.
-U-Pb dates of late kinematic calcite crystals that precipitated at ($\sim$50\textdegree C) in the hanging wall of the Keweenaw fault \cite{Hodgin2022a} indicate that by ca.
-985 Ma the dominant inversion of Midcontinent Rift rocks along the Keweenaw fault has completed.
-Together, these data are consistent with the interpretation that the dominant inversion of the Midcontinent Rift basin happened during the ca.
-1090-1020 Ma Ottawan phase of the Grenvillian Orogeny.
-They also indicate that the period of sedimentary quiescence in the rift basin may have lasted up to $\sim$70 Myrs from ca.
-1060 to ca.
-990 Ma.
+Bedding planes in the Jacobsville Formation typically have shallow dips.
+The exception to these near-horizontal orientations is proximal to the Keweenaw fault (Fig.
+\ref{fig:Geologic_map}) where the Jacobsville Formation was occasionally deformed into monoclinal drag folds resulting in steeply tilted to overturned beds (Fig.
+\ref{fig:Geologic_map}; \citeA{Irving1885a, Cannon2001a}).
+Recent mapping of the Jacobsville Formation also found that it sometimes onlaps the Midcontinent Rift volcanics within or adjacent to en echelon fault branches of the Keweenaw fault system near the northern end of the Keweenaw Peninsula \cite{Tyrrell2019a, Mueller2021a}.
+Clumped isotope data for vein calcite with a U-Pb date of 985.5$\pm$35.8 Ma in the Keweenaw fault breccia within the hanging wall give temperatures of $\sim$50\textdegree C  \cite{Hodgin2022a}.
+The low temperature indicates that the calcite precipitated at shallow depth and that the majority of contractional inversion along the fault was complete by ca.
+985 Ma.
+Combined with the 992.51$\pm$0.64 Ma maximum depositional age, these data are interpreted to indicate that the Jacobsville is a syn-orogenic unit deposited during the ca.
+1010-980 Ma Rigolet phase of the Grenvillian orogeny \cite{Hodgin2022a, Swanson-Hysell2023a}.
 
-Syndepositional faulting that took place during Jacobsville deposition is inferred based on the observation of rift volcanic clasts within the Jacobsville basin.
-In the footwall of the Keweenaw fault near central Keweenaw Peninsula, strata of the Jacobsville Formation in contact with the Midcontinent volcanics are often steeply folded with the conglomeratic strata containing volcanic clasts \cite{Cannon1993a}.
-On the shore of Bete Grise Bay near the northeastern end of Keweenaw Peninsula, the Keweenaw fault is interpreted to have accommodated Grenville compressional stresses by forming a series of segmented en echelon thrusts \cite{Tyrrell2019a}.
-The timing of such faulting could be coeval with the deposition of the upper Jacobsville Formation, leading to the nonconformable onlapping of the Jacobsville sediments over the volcanics in addition to the thrust faulting relationship between the two units.
+In addition, field observations reveal erosional contacts and development of paleosols between the Jacobsville Formation and underlying volcanic rocks \cite{Hamblin1958a, Kalliokoski1975a}, and seismic reflection data beneath Lake Superior have been interpreted to indicate that the Jacobsville Formation lies in angular unconformity atop sedimentary rocks of the Oronto Group \cite{Cannon1989a}.
+These data indicate that there was an interval of erosion following ca.
+1050 Ma prior to the onset of Jacobsville Formation deposition.
 
-% at the same time when sediments from the older basement rocks along with distal input from the Grenville Province were deposited within the Jacobsville back-bulge basin system that in general wedges toward the northwest (Fig.
-\ref{fig:Geologic_map}; \cite{Hedgman2014a, Hodgin2022a}.
+\subsection*{Overview of Jacobsville sedimentology and stratigraphy}
 
-\subsection{Sedimentary facies and general stratigraphy with interpretation}
+The Jacobsville Formation is a fluvial sequence of feldspathic and quartzose conglomerates, sandstones, siltstones, and shales devoid of lava flows or cross-cutting igneous dikes although internal clastic dikes are present \cite{Hamblin1958a}.
+Sandstones of the Jacobsville Formation are more quartz-rich with fewer lithic fragments than sandstones of the Oronto Group such that they can be classified as subarkose to quartz arenites (Fig.
+\ref{fig:Field_photo}; \citeA{Hamblin1958a, Ojakangas2002a}).
+While much of the Jacobsville Formation is dominated by sandstone channel deposits, there are clay-rich siltstones that formed in overbank settings that are a major focus of this study.
 
-% It has been interpreted that the Jacobsville Sandstone is equivalent to the Bayfield Group in northern Wisconsin, and the Fond Du Lac and Hinckley formations in Minnesota.
-Their shared lithofacies of basal conglomerate, cross-bedded fluvial fine to medium grained sandstone, and interbedded red micaceous siltstone could support the interpretation that these sedimentary rocks formed during a common period of post-rift basin formation following rift-related magmatism of the Midcontinent Rift.
 \begin{figure*}[h!]
 \centering
-\includegraphics[width=0.9\textwidth]{Field_photo.png}
-\caption{\scriptsize Field photos of red beds of the Jacobsville Formation.
-At Baby Snake Creek and the Natural Wall ravine in northern Keweenaw Peninsula, intervals of red fine-grained sandstone to siltstone beds can be found through the steeply dipping to overturned basal strata near the Keweenaw fault (A,C) and also the nearly horizontal top strata (B,D).
-At Hungarian Falls, relatively thick intervals of red beds separated by hematite-depleted arenite to arkosic sandstones or conglomerates are found along a $>$100-meter-tall waterfall (E).
-Abundant reoriented siltstone rip-up clasts often exist within the conglomeratic layers above siltstone beds (F).
-Drill cores and block samples of these rip-up clasts were sampled for the paleomagnetic conglomerate test presented in Fig.
-\ref{fig:intraclast_pmag}.
-At Agate Falls, nearly flat-lying red fissile siltstone beds are exposed.
-(G).
-Subrounded detrital mica grains deposited parallel to the bedding plane are often associated with the siltstone facies.
-Another set of siltstone rip-up clasts were sampled from a conglomeratic layer near the basal exposure for the paleomagnetic conglomerate test.
-}
+\includegraphics[width=0.8\textwidth]{Jacobsville_Sections_v6.pdf}
+\caption{Lithostratigraphy and magnetostratigraphy for studied sections of the Jacobsville Formation in northern Michigan, USA.
+Paleomagnetic sampling focused on the dark red siltstone to fine-grained sandstone lithofacies.
+Paleomagnetic specimen declinations are plotted with blue circles corresponding to normal polarity and orange circles to reverse polarity.
+The maximum depositional age of the Agate Falls section is constrained by a CA-ID-TIMS detrital zircon U-Pb date of 1003.2 $\pm$ 2.2 Ma (sample AF1-29.3 of \citeA{Hodgin2022a}).
+In the Snake Creek tributary and Natural Wall sections, the position of the detrital zircon samples BSC1-92.5 and NW1–61.5 from \citeA{Hodgin2022a} are shown.
+No U-Pb zircon dates from these samples were younger than the age of Midcontinent Rift volcanics.
+A correlation of a facies association consisting of volcanic-clast conglomerate, coarse arkosic sandstone, and dark red clay-rich siltstone is inferred between the sections at Hungarian Falls (along Dover Creek and a tributary falls), Hammel Creek, and Natural Wall \cite{Brojanigo1984a}.
+The stratigraphic correlation between Jacobsville sections at Agate Falls and Snake Creek tributary with these correlated sections from the central Keweenaw Peninsula is relatively uncertain.}
+\label{fig:strat_column}
+\end{figure*}
+
+In southern exposures of the Jacobsville Formation, conglomerate facies typically have provenance sourced from Paleoproterozoic lithologies such as vein quartz and iron formation clasts \cite{Hamblin1958a, Kalliokoski1982a}.
+The occurrence of these conglomerates in proximity to the Marenisco fault and Pelton Creek fault is interpreted to represent syn-depositional development of local relief (Fig.
+\ref{fig:Geologic_map}; \citeA{Kalliokoski1982a, Hedgman2014a}).
+Close to the Keweenaw fault in the central Keweenaw Peninsula (Fig.
+\ref{fig:Geologic_map}), the Jacobsville Formation consists dominantly of quartz-rich, trough cross-bedded, medium-grained sandstone, with locally abundant clast-supported pebble to cobble conglomerate, and interbeds of red, hematite-bearing, micaceous siltstone to fine-grained sandstone (Hungarian Falls, Hammel Creek, Natural Wall; Figs.
+\ref{fig:strat_column} and \ref{fig:Field_photo}).
+In this region, the conglomerate contains abundant rounded to angular volcanic clasts as large as boulders that are likely derived from uplifted Midcontinent Rift volcanics \cite{Irving1885a, Brojanigo1984a}.
+Thicker intervals of dark red to brick red siltstone and very fine-grained sandstone are interbeded with coarser-grained sandstone and conglomerate at Agate Falls and Hungarian Falls (Figs.
+\ref{fig:strat_column} and \ref{fig:Field_photo}).
+Rip-up clasts of the siltstones are found within some channelized interbeds of sandstone and conglomerate (Fig.
+\ref{fig:intraclast_pmag}).
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=\textwidth]{Field_photo.png}
+\caption{\scriptsize Field photos of the Jacobsville Formation.
+(A) The red/white medium-grained sandstone with decimeter-scale trough cross-stratification in this image is a wide-spread lithology in the formation, but was not targeted for paleomagnetic sampling in this study.
+(B, C, D) At the Snake Creek tributary and the Natural Wall ravine in the northeastern Keweenaw Peninsula, intervals of red fine-grained sandstone to siltstone beds can be found through the steeply dipping to overturned basal strata near the Keweenaw fault and also the nearly horizontal upper strata.
+At the Hungarian Falls section (E), relatively thick red siltstone horizons interbedded or interfingering with sandstones and conglomerates are found along a waterfall associated with a stream that is a tributary to Dover Creek.
+Abundant reoriented siltstone rip-up clasts can be present within the conglomeratic layers above siltstone beds (Fig.
+\ref{fig:intraclast_pmag}).
+(F) At Agate Falls, nearly flat-lying red fissile siltstone beds are exposed.
+Subrounded detrital mica grains deposited parallel to the bedding plane are often present within the siltstone facies.
+Another set of siltstone rip-up clasts were sampled from a conglomeratic layer at Agate Falls for a paleomagnetic conglomerate test (Fig.
+\ref{fig:intraclast_pmag}).}
 \label{fig:Field_photo}
@@
 
-The Jacobsville Formation is a fluvial sequence of feldspathic and quartzose conglomerates, sandstones, siltstones, and shales completely devoid of lava flows or cross-cutting igneous dikes although numerous  cross-cutting clastic dikes are present \cite{Hamblin1958a}.
-Sediments of the Jacobsville Formation is interpreted to be dominantly derived from the Oronto Group sediments and older basement rocks that were uplifted and deformed in the Montreal River monocline.
-The Jacobsville sediments are in general more quartz-rich than the Oronto Group sediments \cite{Hamblin1958a}.
+The abundance of rift volcanic-clasts in the conglomeratic facies at Hungarian Falls, Hammel Creek, and Natural Wall indicates the presence of uplifted volcanics in proximity to the Keweenaw fault in the central part of the peninsula during Jacobsville deposition (Figs.
+\ref{fig:Geologic_map} and \ref{fig:strat_column}).
+Coarse-grained arkosic to lithic sandstone is a facies associated with the volcanic-clast conglomerates and likely derived from mechanical weathering of the exhumed volcanics.
+Chemical weathering of exhumed volcanics from the hanging wall is a likely source of clay within the abundant fine-grained lithologies that are associated with the volcanic-clast conglomeratic facies \cite{Hamblin1958a}.
+Detrital mica grains up to $\sim$5 mm in size commonly found within siltstone to fine-grained sandstone are likely sourced from Paleoproterozoic and Archean lithologies in the region (Fig.
+\ref{fig:Geologic_map}).
 
-In the southern exposures of the Jacobsville Formation, abundant conglomerate facies typically have provenance sourced from Paleoproterozoic lithologies such as vein quartz and iron formation clasts \cite{Hamblin1958a, Kalliokoski1982a}.
-The occurrence of these conglomerates in proximity to the Marenisco fault and Pelton Creek faults is interpreted to represent syn-depositional development of local relief \cite{Kalliokoski1982a, Hedgman2014a}.
-Close to the Keweenaw Fault in central Keweenaw Peninsula (Fig.
-\ref{fig:Geologic_map}), the Jacobsville Formation consists dominantly of quartz-rich, cross-bedded, medium-grained sandstone, with locally abundant clast-supported pebble to cobble conglomerate, and interbeds of red, hematite-bearing, micaceous siltstone to fine-grained sandstone (Hungarian Falls, Hammell Creek, Natural Wall; Fig.
+The thickness of the Jacobsville strata can vary significantly in northern Michigan \cite{Hamblin1958a, Kalliokoski1982a}, and complete stratigraphic sections are not exposed at the studied localities (Figs.
+\ref{fig:Geologic_map} and \ref{fig:strat_column}).
+Clay-rich red siltstone is more abundant in conjunction with conglomerates (e.g.
+20-120 m in the Hungarian Falls section; Figs.
+\ref{fig:strat_column} and \ref{fig:Field_photo}) than with the cross-stratified sandstone which dominates much of the formation.
+Relatively thin lithic to arkosic sandstone beds, which are otherwise uncommon in the Jacobsville Formation, are commonly associated with the volcanic conglomerates and red siltstones near the Keweenaw Fault (Fig.
 \ref{fig:strat_column}).
-In this region, the conglomerate contains abundant volcanic clasts likely derived from uplifted Midcontinent Rift volcanics \cite{Irving1885a, Brojanigo1984a}.
-This clast-supported lithofacies can include siltstone rip-up clasts and sometimes channelized lenses of siltstones.
-More massive red beds also commonly interbed with hematite-depleted sandstone and conglomerate layers \cite<e.g.>{Hamblin1958a, Irving1885a}.
-The red beds typically appear brick-red to dark-red, consisting of fine-grained arenite to arkosic sandstone and siltstone sometimes with minor grey reduction spots (Fig.
-\ref{fig:Field_photo}).
+Together, the volcanic conglomerates, red siltstones, and arkosic sandstones make up a unique facies association.
+As previously interpreted by \citeA{Brojanigo1984a}, this facies association was likely deposited proximal to actively uplifting volcanics near the Hungarian Falls, Hammel Creek, and Natural Wall sections (Fig.
+\ref{fig:strat_column}).
+The relative stratigraphic positions of the Agate Falls and Snake Creek tributary sections are less certain (Figs.
+\ref{fig:Geologic_map} and \ref{fig:strat_column}).
 
-The occurrence of the rift volcanic-clast containing conglomeratic facies at Hungarian Falls, Hammell Creek, and Natural Wall indicates an interval of syn-depositional faulting in proximity to the Keweenaw fault in central part of the peninsula (Fig.
-\ref{fig:Geologic_map}, \ref{fig:strat_column}).
-Rapid weathering of the exhumed volcanics from the hanging wall also likely derived red clay and arkosic sandstone in the Jacobsville Formation which are most abundant near the fault but were also transported further away \cite{Hamblin1958a}.
-Rounded detrital mica grains commonly found within very fine, fissile siltstone strata are likely sourced from eroded Paleoproterozoic and Archean basement rocks in the Montreal River monocline highlands (Fig.
-\ref{fig:Geologic_map}).
-The mica grains can be up to $\sim$5mm in size (Fig.
-\ref{fig:Field_photo}).
+The Jacobsville Formation has been mapped to have a continuous extent from the Wisconsin and Michigan border northeast to the Keweenaw Peninsula and east to the Marquette region (Fig.
+\ref{fig:Geologic_map}; \citeA{Hamblin1958a, Cannon1995a, Cannon1996a, Cannon2001a}).
+In the eastern Lake Superior region near Sault Ste Marie, Ontario, there are similar exposures of sedimentary rocks that are also mapped as the Jacobsville Formation \cite{Hamblin1958a}.
+As observed on the Keweenaw Peninsula, these sandstones are within the footwall of faults where they are overthrusted by Midcontinent Rift volcanics near Mamainse Point \cite{Manson1994a}.
+To the west, the Jacobsville Formation has been interpreted to be correlative with sedimentary rocks of the Bayfield Group in northern Wisconsin and the Fond du Lac Formation and Hinckley Formation in Minnesota \cite{Hamblin1958a, Kalliokoski1982a, Ojakangas2001b}.
+A challenge associated with these correlations is that the ca.
+1075 to 1050 Ma Freda Formation of the Oronto Group and the ca.
+990 Ma Jacobsville Formation have quite similar fluvial lithofacies despite being deposited in distinct basinal settings at different times.
 
-Thickness of the Jacobsville strata can vary significantly in northern Michigan \cite{Hamblin1958a, Kalliokoski1982a}, and complete stratigraphic sections are often not fully exposed at studied localities (Fig.
-\ref{fig:Geologic_map}).
-As a result, the exact stratigraphic association between sections of the Jacobsville Formation near central Keweenaw Peninsula and those exposed at Agate Falls and Baby Snake Creek is difficult to constrain.
-We propose one tentative stratigraphy model based on an interpreted Jacobsville sedimentary facies model.
-Given that deposition of conglomeratic facies was likely associated with syn-deposition faulting along the Keweenaw fault, this facies could have been deposited during the early stages of Jacobsville sedimentation, whereas siltstone and arkosic sandstone facies were probably deposited later with broader spatial distribution.
-Some early conglomeratic facies could have been reworked and contributed to the formation of Jacobsville red beds farther away from the fault.
-Based on this facies model, the relative lack of conglomerate facies at Agate Falls and at Baby Snake Creek, and field observations of onlapping Jacobsville Formation on top of previously deformed rift volcanics at Baby Snake Creek, could support an interpretation that Jacobsville sediments exposed at these two localities are slightly younger than the conglomerate-rich strata near central Keweenaw Peninsula (Fig.
-\ref{fig:Geologic_map}, \ref{fig:strat_column}).
-Further discussion regarding the stratigraphy in light of the paleomagnetic data is presented in section \ref{magstrat}.
+\section*{Paleomagnetic results and interpretation}
 
-%At some point earlier on, we may need to articulate a facies model that can be used to justify the correlation scheme by 1) demonstrating that the conglomeratic facies at the grouped sites in the lower Keweenaw Peninsula, which are associated with siltstone deposition and arkosic sands, were likely deposited during an interval of syn-depositional faulting in proximity to the Keweenaw Fault; next 2) it can be suggested that rapid weathering of the exhumed basaltic rock from the hanging wall resulted in red clay most abundant to the fault but also deposited further away, and abundant partially weathering arkosic sands but which also occur farther away to some degree; 3) This interpretation is supported by the facies association being volcanic conglomerate, red siltstone, and arkosic sands (along with some other typical Jacobsville facies) and is the basis for suggesting that farther travelled silt and arkosic sands could be correlated with the conglomerates.
-Tentatively this stratigraphicaly significant and unique facies association may serve as a basis for correlation across the basin during a time of more significant syn-depositional fault uplift.
+Oriented paleomagnetic samples were collected with a portable electric drill from five stratigraphic sections of the Jacobsville Formation in northern Michigan (Figs.
+\ref{fig:Geologic_map} and \ref{fig:strat_column}).
+Additional oriented cores and block samples of siltstone rip-up clasts within conglomerates and conglomeratic sandstones were sampled within the Hungarian Falls and Agate Falls sections (Figs.
+\ref{fig:Geologic_map} and \ref{fig:strat_column}, and \ref{fig:intraclast_pmag}).
+To maximize sampling of distinct time snapshots of the geomagnetic field at the time of Jacobsville deposition and average out paleosecular variation, we optimized for vertical stratigraphic coverage.
+Each sample is a distinct stratigraphic horizon and therefore a paleomagnetic site, representing a unique record of the geomagnetic field.
+Red fine-grained sandstone to shaly siltstone layers were preferentially sampled as they have lower permeability and are less susceptible to diagenetic alteration through fluid flow than coarser-grained sandstone.
+These fine-grained lithologies are typically of deep red color in contrast to coarse-grained lithologies that have splotchy tan, red, and light green coloration associated with secondary reduction (Fig.
+\ref{fig:Field_photo}A).
+Care was taken to avoid collecting samples with reduction spots.
+Paleomagnetic cores and blocks were oriented using a magnetic compass and a sun compass when possible.
+Sun compass data were preferentially used when available.
+A total of 379 specimens including 30 rip-up clasts were collected for paleomagnetic study.
+To minimize the visual impact of sampling, rock containing core holes was knocked out of outcrops following samples---readily done for the friable siltstones.
 
-Overall, the extent of the Jacobsville Formation has been mapped along the lake shore of the Keweenaw Peninsula from northern Michigan, USA, to the north side of Michipicoten Island and perhaps to northern Sault Ste Marie, Canada (Fig.
-\ref{fig:Geologic_map}; \citeA{Hamblin1958a}).
-It has also been interpreted that the Jacobsville Formation could be temporally and lithologically correlated with the Bayfield Group sediments in northern Wisconsin and the Fond du Lac and Hinckley Formation in Minnesota \cite{Hamblin1958a, Kalliokoski1982a, Malone2016a}.
+Challenges exist in isolating the primary magnetization in sedimentary rocks.
+Primary detrital remanence acquired during deposition can be masked by secondary remanence acquired through precipitation and growth of new minerals within sedimentary rocks syn- to post-deposition (e.g.
+pigmentary hematite formed from precursor phases such as ferrihydrite; \citeA{Jiang2018a}).
+In red beds, it has been found that formation of pigmentary hematite can post-date the timing of deposition, resulting in a chemical remanent magnetization superimposed upon primary detrital remanent magnetization carried by detrital grains \cite{Collinson1974a, Tauxe1980a}.
+To isolate the remanence components of sandstones of various grain sizes of the Jacobsville Formation, \citeA{Roy1978a} adopted a method first applied by \citeA{Collinson1965c} to preferentially remove fine-grained pigmentary hematite through prolonged immersion in concentrated HCl acid.
+Progressively longer immersion times first dissolve the nanometer-scale pigmentary hematite prior to dissolving the coarser micrometer-scale detrital grains.
+Studies applying paired acid etching demagnetization and thermal demagnetization show that the pigmentary hematite grains tend to have lower unblocking temperatures than detrital ones \cite{Tauxe1980a, Bilardello2010c}.
+High-resolution thermal demagnetization experiments with temperature intervals as small as 1-2 \textdegree C have also been shown to be effective in isolating detrital remanence from chemical remanence as coarser hematite grains tend to have higher unblocking temperatures closer to the N\'eel temperature of hematite \cite{Jiang2015a,Swanson-Hysell2019b}.
+
+We adopted a thermal demagnetization protocol with increasingly higher resolution approaching the N\'eel temperature of hematite (5 \textdegree C to 2 \textdegree C; Fig.
+\ref{fig:intraclast_pmag}).
+The specimens underwent stepwise thermal demagnetization at the UC Berkeley Paleomagnetism Lab using an ASC demagnetizer (residual fields $<$10 nT) with measurements made on a 2G DC-SQUID magnetometer.
+While the very fine-grained sandstone and siltstone lithologies typically do not have a significant present-day local field overprint component, such component can be present in fine-grained sandstones and is removed by $\sim$300\textdegree C during thermal demagnetization (Fig.
+S1).
+A total of 356 specimens yielded detrital remanent magnetizations that can be fit with least-square lines \cite{Kirschvink1980a}.
+These fits were made using PmagPy \cite{Tauxe2016a} and all paleomagnetic data are available to the measurement level in the MagIC database (https://earthref.org/MagIC/19780/df1b7c0c-b1cc-4e1e-b40c-2585f007178a; \textit{this link provides access to reviewers and will be updated with a final version of the link when the manuscript is accepted and assigned a DOI}).
+
+\subsection*{Paleomagnetic field tests}
+
+\subsubsection*{Fluvial intraclast conglomerate tests}
+
+Similar to the thermal demagnetization results of the hematite-bearing fluvial intraclasts of the Freda Formation \cite{Swanson-Hysell2019b}, the siltstone rip-up clasts of the Jacobsville Formation typically reveal two distinct magnetization components (Fig.
+\ref{fig:intraclast_pmag}).
+One component shows similar vector orientations amongst intraclasts and was typically removed up to 640-655 \textdegree C (Fig.
+\ref{fig:intraclast_pmag}).
+The relatively low unblocking temperatures and similarity in directions among the reoriented rip-up clasts indicate that it is chemical remanent magnetization acquired through crystallization of secondary pigmentary hematite following clast redeposition.
+After removal of this component, further thermal demagnetization with small step increments at higher temperatures up to $\sim$689 \textdegree C often reveal an origin-trending component (Fig.
+\ref{fig:intraclast_pmag}).
+In the data from the clasts, there is typically a significant directional change in specimen magnetization between the mid-temperature component and the high-temperature component.
+As a result, 27 out of 30 intraclast specimens could be fit with high-temperature least squares lines.
+Distinct from the well-grouped mid-temperature component directions, the high-temperature directions of the intraclasts are dispersed and the null hypothesis of randomness cannot be rejected at the 95\% confidence level (n=5 for intraclasts at Agate Falls and n=22 for intraclasts at Hungarian Falls; Fig.
+\ref{fig:intraclast_pmag}; \citeA{Watson1956a}).
+This result indicates that the high unblocking temperature remanence is primary and was acquired by detrital hematite grains during initial deposition prior to the clasts being ripped up and reoriented within the depositional environment.
+This conglomerate test provides strong support for interpreting the high unblocking temperature remanence in the \textit{in situ} beds as a primary DRM.
 
@@
 \centering
-\includegraphics[width=0.9\textwidth]{Jacobsville_Sections_v3.pdf}
-\caption{Tentative sedimentary stratigraphy and magnetostratigraphy for the Jacobsville Formation in northern Michigan, USA.
-Stratigraphic locations of paleomagnetic samples are shown in context of Jacobsville lithofacies, sedimentary structures, as well as detrital zircon maximum deposition ages developed by \citeA{Hodgin2022a}.
-}
-\label{fig:strat_column}
-\end{figure*}
-
-% Blake's current tectonic model: after MCR volcanics ca.
-1084 Ma emplacement follows the deposition of the sediments (Oronto Group) copper harbor conglomerate, Nonesuch formation and the Freda formation.
-Then the Grenville orogenesis took off with the Ottawan phase from 1020 Ma - 10??
-Ma and the Rigolet phase from 995-980 Ma.
-Paleosol development and the unconformity between the Jacobsville and the volcanics indicate a prolonged period of gap during which the Oronto Group in the backbulge basin where they were deposited were lost, resulting in younger Jacobsville deposition directly on top of the volcanics.
-The fact that in the Jacobsville has conglomerate and sandstone and siltstone facies in them supports sediment input from the eroded Oronto Group.
-This indicate the older boundary of the Jacobsville deposition might be closer to the Rigolet Phase than the Ottawan phase.
-The younger age boundary of the Jacobsville is well defined - given that the maximum deposition age from the zircons of the Jacobsville is 993 Ma and the calcite on the Keweenaw Fault is 985.5 Ma, the deposition of at least the upper portion of the Jacobsville is well constrained to be between 993 and 985 Ma.
-The deposition of Jacobsville in the back bulge in Laurentia interior is related to the far-field compression caused by the Grenville Orogeny.
-
-\section*{Paleomagnetic results and inerpretation}
-
-Oriented paleomagnetic samples were collected with a portable electric drill along five stratigraphic sections of the Jacobsville Formation in northern Michigan (Fig.
-\ref{fig:Geologic_map}, \ref{fig:strat_column}).
-Additional oriented cores and block samples of siltstone rip-up clasts within conglomerate lithofacies were sampled at Hungarian Falls and at Agate Falls (Fig.
-\ref{fig:Geologic_map}, \ref{fig:strat_column}).
-In order to maximize sampling of distinct time snapshots of the geomagnetic field at the time of Jacobsville deposition such that paleosecular variation can be sufficiently averaged, we optimized for vertical stratigraphic coverage such that each sample constitutes a paleomagnetic site considering that a paleomagnetic site (which ideally captures a single snapshot of the local geomagnetic field) is a particular bed in a sedimentary sequence.
-Red fine-grained sandstone to shaly siltstone layers were preferentially sampled as they tend to have lower permeability and are less susceptible to diagenetic alteration through fluid flow than coarser grained sandstone.
-Care was taken to avoid samples containing reduction spots.
-Paleomagnetic cores and blocks were oriented using a magnetic compass and a sun compass whenever possible.
-Sun compass data were preferentially used when available.
-A total of 379 specimens including 30 intraclasts where collected for paleomagnetic study.
-The specimens underwent stepwise thermal demagnetization at the UC Berkeley Paleomagnetism Lab using an ASC demagnetizer (residual fields $<$10 nT) with measurements made on a 2G DC-SQUID magnetometer.
-The demagnetization protocol had high-resolution approaching the Ne\'el temperature of hematite.
-As a result, 356 specimens yielded detrital remanent magnetizations that can be fitted with least-square lines.
-(5 \textdegree C to 2 \textdegree C) (Fig.
-\ref{fig:intraclast_pmag}).
-Demagnetization results are plotted in Figure \ref{fig:in_situ_pmag}A by locality.
-All paleomagnetic data are available to the measurement level in the MagIC database (https://earthref.org/ MagIC/; THIS LINK IS AVAILABLE TO REVIEWERS AND WILL BE UPDATED WHEN A DOI IS AVAILABLE).
-
-\begin{figure*}[h!]
-\centering
-\includegraphics[width=\textwidth]{in_situ_pmag.pdf}
-\caption{A) Specimen detrital remanence directions plotted by locality on equal area plots.
-During sample collection, we optimized for vertical stratigraphic coverage such that each sample constitutes a paleomagnetic site considering that a paleomagnetic site.
-B) Summary equal area plots at the study-level combining directions from all localities.
-The elongation/inclination (E/I) method \cite{Tauxe2004b} for estimating the amount of inclination shallowing in Jacobsville red beds are applied separately to specimen remanence directions of each polarity and to all specimen directions after flipping the reversed (orange) directions to their antipodes (blue).
-Details on inclination shallowing corrections are shown in Figure \ref{fig:EI_results}.}
-\label{fig:in_situ_pmag}
-\end{figure*}
-
-\subsection*{Paleomagnetic field test}
-
-\subsubsection*{Fluvial intraclast conglomerate test}
-
-Similar to the thermal demagnetization result of the hematite-bearing fluvial intraclasts of the Freda Formation \cite{Swanson-Hysell2019b}, the siltstone rip-up clasts of the Jacobsville Formation typically reveal two distinct magnetization components (Fig.
-\ref{fig:intraclast_pmag}).
-One component show similar vector orientations amongst intraclasts and was typically removed up to 640-655 \textdegree C (Fig.
-\ref{fig:intraclast_pmag}).
-This indicates that it is likely a chemical remanent magnetization acquired through growth of secondary hematite in pore spaces following redeposition of the clasts.
-After removal of this component, further thermal demagnetization with small step increments at higher temperatures up to $\sim$689 \textdegree C often reveal an origin-trending component (Fig.
-\ref{fig:intraclast_pmag}).
-In the data from the clasts, there is typically a significant directional change in specimen magnetization between the mid-temperature component and the high-temperature component.
-As a result, 28 out of 30 intraclast specimens could be fit with distinct mid-temperature and high-temperature least squares lines \cite{Kirschvink1980a}.
-Distinct from the well-grouped mid-temperature component directions, the high-temperature directions of the intraclasts are dispersed and the null hypothesis of randomness cannot be rejected at the 95\% confidence level (n=6 for intraclasts at Agate Falls and n=22 for intraclasts at Hungarian Falls; Fig.
-\ref{fig:intraclast_pmag}; \citeA{Watson1956a}).
-This indicates that the remanence was acquired by detrital hematite grains during initial deposition of the grains.
-
-\begin{figure*}[h!]
-\centering
-\includegraphics[width=\textwidth]{intraclast_pmag.pdf}
-\caption{Paleomagnetic data from fluvial rip-up clasts of the Jacobsville Formation reveal a mid-temperature component that typically unblocks up to 655 \textdegree C and a high-temperature component that typically unblocks up to 689 \textdegree C.
-The components are present as varying fractions of the overall remanence as seen in the three individual clasts shown on vector orthogonal plots and measurement-level equal area plots in tilt-corrected coordinates (developed using PmagPy; \citeA{Tauxe2016a}).
-The direction of the mid-temperature component (interpreted as CRM) is shown as green arrows on the vector component plots and green circles on the equal area plots, while the high-temperature component (interpreted as DRM) is shown with purple symbols.
-The mid-temperature component has a similar direction among the clasts as can be seen on the summary equal area plots.
-In contrast, the high-temperature component directions (purple) are dispersed and pass a randomness test of \citeA{Watson1956a}.
-NRM = natural remanent magnetization; AFC = Agate Falls clasts; HFC = Hungarian Falls clasts.}
+\includegraphics[width=.7\textwidth]{intraclast_pmag.png}
+\caption{\scriptsize(A) Field photos of siltstone to fine-grained sandstone fluvial rip-up clasts within conglomerate and conglomeratic sandstone beds in the Jacobsville Formation.
+(B) Representative step-wise thermal demagnetization data from a rip-up clast from Hungarian Falls (HFC-23) plotted on an orthogonal vector diagram.
+The plot reveals a mid-temperature component and an origin-trending high-temperature component.
+The components are present as varying fractions of the overall remanence in different specimens.
+The direction of the mid-temperature component (interpreted as secondary CRM) is shown as dark red arrows on the orthogonal vector plots and dark red circles on the equal area plots, while the high-temperature component (interpreted as primary DRM) is shown in grey.
+(C, D) The mid-temperature component has a similar direction among the clasts as can be seen on the summary equal area plots.
+In contrast, the high-temperature component directions are dispersed and pass the randomness test of \citeA{Watson1956a}.
+The DRM directions are shown in tilt-corrected bedding coordinates and CRM directions are shown in geographic coordinates.
+NRM = natural remanent magnetization.}
 \label{fig:intraclast_pmag}
@@
 
-A total of 70 specimens yielded interpretable DRM directions throughout the stratigraphic section at Baby Snake Creek (Fig.
-\ref{fig:Geologic_map}, \ref{fig:in_situ_pmag}).
-46 specimens were collected from the steeply tilted beds close to the Keweenaw fault and the rest 24 were from nearly horizontal beds.
-This allows us to conduct the paleomagnetic fold test on these specimens to investigate the timing of detrital remanence acquisition.
-A bootstrap fold test \cite{Tauxe1994a} result shows 95\% confidence bounds for the best grouping of the directions from the various beds of the fold is achieved between 55\% and 92\% unfolding (Fig.
+A total of 71 specimens yielded interpretable DRM directions throughout the stratigraphic section at the Snake Creek tributary (Figs.
+\ref{fig:Geologic_map} and \ref{fig:in_situ_pmag}).
+The exposures of steeply tilted, moderately tilted, as well as nearly horizontal beds along this section allow us to conduct a paleomagnetic fold test to investigate the timing of detrital remanence acquisition.
+The folding in this section is the result of deformation in the immediate footwall of the Keweenaw fault.
+The chronological constraints on the timing of Keweenaw fault motion indicate that it occurred within $\sim$10 Myr of deposition---making this fold test a more informative constraint on the age of the remanence than typical for such tests.
+Given that the more steeply tilted beds accommodated significant deformation during shortening associated with the Keweenaw fault motion, they could be more prone to non-cylindrical folding and complicate tilt-correction of the paleomagnetic directional data \cite<e.g.>{Pueyo2003a, Nabavi2021a}.
+Therefore, we conduct a bootstrap fold test \cite{Tauxe1994a} using specimens from the nearly horizontal beds and the moderately tilted beds.
+The result shows that the degrees of unfolding that result in the best grouping of the DRM directions have 95\% confidence bounds that overlap with 100\%, consistent with the specimens having acquired their magnetization prior to folding (Fig.
 \ref{fig:fold_test}).
-Although the 95\% confidence bounds exclude both 100\% unfolding and geographic coordinates, the results indicate the directions are better grouped with a significant amount of tilt correction is applied complex folding regime such as non-cylindrical folding or multiple folding episodes about different axes could contributed to this fold test result not including 100\% unfolding within the 95\% confidence bounds.
+This positive fold test further supports that the Jacobsville red beds acquired a primary detrital remanence during deposition.
+In contrast, the low-temperature component directions from the Snake Creek tributary fails a fold test (Fig.
+S2), consistent with them being acquired as chemical remanence magnetizations that postdate the folding.
 
@@
 \includegraphics[width=0.5\textwidth]{SC1_fold_test.pdf}
-\caption{Bootstrap paleomagnetic fold test \cite{Tauxe1994a} of the detrital remanent magnetization directions recorded by specimens from Baby Snake Creek.
-The directional data set becomes more tightly grouped than in geographic coordinates when applying 55\%-92\% unfolding, but both geographic and 100\% unfolded coordinate systems are excluded at the 95\% confidence level.
-}
+\caption{Bootstrap paleomagnetic fold test \cite{Tauxe1994a} of the detrital remanent magnetization directions recorded by specimens from the nearly horizontal beds and moderately tilted beds at the Snake Creek tributary.
+Complete unfolding lies within the 95\% confidence limits of the test, consistent with the magnetization having been acquired prior to folding.}
 \label{fig:fold_test}
@@
 
-\subsection*{Tentative magnetostratigraphy}
-\label{magstrat}
-With the insights gained from the thermal demagnetization results of the Jacobsville fluvial intraclasts (Fig.
-\ref{fig:intraclast_pmag}), we next investigate the detrital remanent magnetizations for all in situ specimens that were collected along the five localities in northern Michigan (Fig.
-\ref{fig:Geologic_map}).
-The resultant DRM directions are plotted by locality in Fig.
-\ref{fig:in_situ_pmag}.
-These results show dual polarities with some specimens recording transitional field directions (Fig.
-\ref{fig:in_situ_pmag}), consistent with the observation of \citeA{Roy1978a}.
-While \citeA{Roy1978a} observed only one polarity reversal, our new data show that the Jacobsville Formation recorded multiple geomagnetic field polarity reversals.
-A tentative magnetostratigraphy is proposed in Fig.
-\ref{fig:strat_column}.
-In order to aid later discussions regarding the paleogeography of Laurentia, we refer to the westerly DRM directions as of normal polarity and the easterly DRM directions of reversed polarity (Fig.
+\subsection*{Paleomagnetic reversals}
+
+With the insights gained from the thermal demagnetization results of the Jacobsville fluvial intraclasts and the fold test (Figs.
+\ref{fig:intraclast_pmag} and \ref{fig:fold_test}), we next investigate the detrital remanent magnetizations for all \textit{in situ} specimens from the five stratigraphic sections (Figs.
+\ref{fig:Geologic_map} and \ref{fig:strat_column}).
+The DRM directions are plotted by stratigraphic section location in Figure \ref{fig:in_situ_pmag}.
+As observed in the sparser data of \citeA{Roy1978a}, there are dual polarity magnetizations (Fig.
+\ref{fig:strat_column}).
+The paleomagnetic polarity reversals are shown in stratigraphic context in Figure \ref{fig:strat_column} with specimen magnetization declinations plotted against stratigraphic heights.
+The paleogeographic context of Laurentia within Rodinia and into the Paleozoic enables us to interpret the westerly DRM directions as normal geomagnetic polarity and the easterly DRM directions as reversed polarity (Fig.
 \ref{fig:in_situ_pmag}).
 
-At Hungarian Falls (Fig.
-\ref{fig:Geologic_map}), numerous hematite-rich siltstone to fine-grained sandstone layers were deposited (Fig.
-\ref{fig:strat_column}).
-Throughout $\sim$240 meters of strata, a total of 158 samples were collected for paleomagnetic study.
-At stratigraphic height of $\sim$24 meters (Fig.
-\ref{fig:strat_column}), 8 samples from a $\sim$0.5-meter-thick siltstone layer record reversed polarity directions (easterly declination and shallow inclination; horizon HF6).
-At stratigraphic level $\sim$42 meters (horizon HF5; Fig.
-\ref{fig:strat_column}), 4 samples were collected through a $\sim$0.2 m siltstone horizon including one specimen from near the base of the layer and three specimens near the top of the layer.
-The DRM component from the three specimens near the top all record a reversed polarity whereas the specimen collected near the base show a westerly direction.
-The westerly direction indicates that the geomagnetic field could have gone through a reversal between strata at 24 and 42 m, although more samples are needed to test the length and stability of the potential period of normal polarity.
-Therefore, we tentatively mark a period of transitional field near the base of horizon HF5 in Fig.
-\ref{fig:strat_column}.
-A reversed polarity field seems to have been maintained as 5 samples from horizon HF4 consistently record easterly directions just above the top of HF5 (Fig.
-\ref{fig:strat_column}).
-However, at least one geomagnetic field polarity reversal happened between the deposition of siltstone layers of HF4 and HF3, as 7 samples from HF3 all record normal-polarity directions.
-Subsequently, the field appears to flip again back to reversed polarity, as is shown by 7 samples from HF2 higher in the section (Fig.
-\ref{fig:strat_column}).
-Through the $\sim$35-meter-thick horizon of HF1 characterized by numerous siltstone to fine-grained sandstone layers interbedded with coarser-grained sandstone and conglomerate, magnetic polarity reversals of R-N-R-N-R were recorded (Fig.
+At the Hungarian Falls section along Dover Creek, numerous hematite-rich siltstone to fine sandstone layers commonly occur in association with trough cross-bedded sandstone and conglomeratic facies.
+These lithofacies are particularly well-exposed along a tributary waterfall to Dover Creek such that 158 samples were collected for paleomagnetic study (Fig.
+\ref{fig:Geologic_map}, \ref{fig:strat_column}), making this section the most densely sampled amongst all sections.
+The specimen DRMs show that the geomagnetic field reversed at least nine times during deposition within the section (Fig.
 \ref{fig:strat_column}).
 
-The Natural Wall ravine and Hammell Creek are in close proximity to Hungarian Falls (Fig.
-\ref{fig:Geologic_map}) and the sedimentary facies characterized by trough cross-bedded sandstone and conglomerates are similar to the basal $\sim$120 meters of the section at Hungarian Falls (Fig.
+The Natural Wall and Hammel Creek sections are in close proximity to the Hungarian Falls section (Fig.
+\ref{fig:Geologic_map}) and the unique sedimentary facies association of volcanic-clast conglomerate, red siltstone, and coarse arkosic sandstone is similar to the basal $\sim$120 meters of the section at Hungarian Falls (Fig.
 \ref{fig:strat_column}).
-Since only a few siltstone to fine-grained sandstone horizons were sampled at these two localities, we tentatively correlate the stratigraphic sections at these three localities based on their sedimentary facies (Fig.
+Our sampling at these two localities is more limited due to fewer occurrences of exposed siltstone and fine-grained sandstone beds due to more limited exposure throughout the sections (Fig.
 \ref{fig:strat_column}).
-At Hammell Creek, 14 samples from two siltstone layers near the base of the section are of reversed polarities, consistent with the reversed-polarity layers near the base of Hungarian Falls.
-Near the top of the Natural Wall section, consistent normal-polarity directions are recorded by 15 samples.
-This normal polarity is likely coeval with one of the normal-polarity periods recorded at Hungarian Falls, although it could indicate an additional polarity reversal from reversed to normal polarity that happened after the deposition of the youngest reversed strata at Hungarian Falls (Fig.
+At Hammel Creek, 16 samples from two siltstone layers near the base of the section are of reversed polarity (Fig.
 \ref{fig:strat_column}).
-Near the base of the exposed stratigraphic section at Natural Wall (Fig.
-\ref{fig:strat_column}), the strata are steeply tilted or overturned (Fig.
+Near the top of the Natural Wall section, consistent normal-polarity directions are recorded by 15 samples (Fig.
+\ref{fig:strat_column}).
+
+At the Snake Creek tributary (Fig.
+\ref{fig:Geologic_map}), 71 specimens yielded interpretable DRMs through multiple levels along the $\sim$95-meter-thick section and all show normal-polarity directions (Figs.
+\ref{fig:in_situ_pmag} and \ref{fig:strat_column}).
+At Agate Falls (Fig.
+\ref{fig:Geologic_map}), clay-rich detrital mica-bearing siltstone layers occur through $\sim$30 meters of strata that are incised by the Middle Branch Ontonagon River (Fig.
 \ref{fig:Field_photo}).
-Tilt-corrected paleomagnetic directions from samples collected from these strata show dominantly southeasterly declinations that are distinct from either normal- or reversed-polarity directions higher along the strata or at Hungarian Falls.
-It is most likely that these directions resulted from noncylindrical or multiple episodes of complex folding.
-We hypothesize that these DRM directions were acquired during a reverse polarity episode similar to that recorded near the base of Hammell Creek (Fig.
-\ref{fig:strat_column}), but we will not include these directions in further paleomagnetic pole analyses.
+67 total paleomagnetic samples were collected on both sides of the waterfall (Fig.
+\ref{fig:strat_column}).
+At least three geomagnetic polarity reversals occurred during deposition within the Agate Falls section (Fig.
+\ref{fig:strat_column}).
 
-At Baby Snake Creek (Fig.
-\ref{fig:Geologic_map}), all 70 specimens that yielded interpretable high-temperature remanence have normal-polarity DRM directions \ref{fig:in_situ_pmag}).
-At Agate Falls (Fig.
-\ref{fig:Geologic_map}), multiple hematite-rich, detrital mica-bearing siltstone layers occur through a $\sim$30-meter strata incised by the running Middle Branch Ontonagon River.
-68 paleomagnetic samples were collected on both sides of the waterfall (Fig.
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=0.95\textwidth]{in_situ_pmag.pdf}
+\caption{(A) Specimen detrital remanence directions (DRM) plotted by locality on equal area plots.
+During sample collection, we optimized for vertical stratigraphic coverage such that each sample constitutes a single horizon and therefore a paleomagnetic site.
+The measurement-level data are available in the MagIC database (https://earthref.org/MagIC/19780/df1b7c0c-b1cc-4e1e-b40c-2585f007178a; \textit{this link provides access to reviewers and will be updated with a final version of the link when the manuscript is accepted}).
+(B) Summary equal area plots at the study-level combining directions from all localities.
+The elongation/inclination (E/I) method \cite{Tauxe2004b} was used to estimate the amount of inclination shallowing in Jacobsville specimen DRMs after unifying the polarities (flipping the normal [blue] directions to their antipodes).
+Details on inclination shallowing corrections are shown in Figure \ref{fig:EI_results} with the value of $f$=0.64 used for the directions in the middle panel.
+All DRM directions are shown in tilt-corrected bedding coordinates.
+The summary plot of all specimen chemical remanence directions (CRM) are shown in geographic coordinates show well-grouped reversed-polarity directions whose directions are similar to the DRMs.
+}
+\label{fig:in_situ_pmag}
+\end{figure*}
+
+The paleomagnetic reversal test assesses whether two directional data sets share a common mean when one of the polarity directions is flipped to its antipode.
+Although the occurrence of multiple geomagnetic field reversals during deposition of the Jacobsville Formation is evident (Figs.
+\ref{fig:strat_column} and \ref{fig:in_situ_pmag}), the limited and often discrete records of many polarity chrons are insufficient for pair-wise reversal tests within single sections (Fig.
 \ref{fig:strat_column}).
-A single transition from normal to reversed polarity is observed through the strata from bottom to top with a few strata in the middle of the section capturing transitional directions (Fig.
-\ref{fig:strat_column}, \ref{fig:in_situ_pmag}).
+Sedimentation within individual siltstone horizons was likely quite rapid given their deposition within fluvial floodplains potentially without sufficient time to average out secular variation within discrete siltstone intervals.
+We combine all normal and reversed directions from the five sections and perform a reversal test.
+The \citeA{McFadden1990a} reversal test results show the angle between the two mean directions (15.9\textdegree; no inclination shallowing correction) exceeds the critical angle (7.5\textdegree) (Fig.
+S3), thereby failing the test.
+This difference largely arises due to normal (west directed) directions in the Snake Creek tributary section being slightly steeper in their upward inclination.
+There are at least three possible explanations for this behavior: 1) the lower number of normal directions which dominantly come from the Snake Creek tributary section may not have averaged out secular variation; 2) there could have been plate motion during Jacobsville deposition with the normal polarity in the Snake Creek tributary section being acquired when Laurentia was at a slightly higher latitude; 3) there could be a bias from the reversed CRM superimposed on the normal DRM in the Snake Creek tributary specimens given that removal of the CRM is associated with progressively shallowing the inclination in some specimens \cite{Cox1960a, Scott1996a}.
+Regardless, the positive intraclast conglomerate test gives confidence that the Jacobsville Formation has a primary detrital remanent magnetization that can be isolated through thermal demagnetization.
+This result is strengthened through the positive fold test and the broadly antipodal dual polarity data.
+Additionally, taking the large number of samples together from both polarities to calculate a pole increases the likelihood that the dual polarity pole has averaged out secular variation.
 
-Challenges exist in correlating the stratigraphic sections measured at the five localities in this study due to the lateral variability of the thickness of the formation associated with the variable paleotopographic relief of the Precambrian surface upon which the Jacobsville Formation was deposited \cite{Hamblin1958a, Kalliokoski1982a}.
-In particular, there may not be a unique solution in reconstructing the relative stratigraphic positions of the strata at Baby Snake Creek and Agate Falls with respect to the other three sections in central Keweenaw Peninsula due to geographical separation (Fig.
-\ref{fig:Geologic_map}, \ref{fig:strat_column}).
-However, at both Baby Snake Creek and Agate Falls, the sedimentary strata are dominated by siltstone to cross-bedded sandstone and are lacking significant amount of conglomerate facies as seen near Hungarian Falls.
-Given the relative facies abundance difference at these localities, an interpreted stratigraphic correlation could exist based on an interpreted sedimentary facies model.
-The conglomerates, hematite-bearing siltstones, and arkosic sandstones deposited near the Keweenaw fault could be associated with syn-depositional faulting that uplifted the Midcontinent Rift volcanics that subsequently eroded and shedded sediments near the fault.
-These facies, particularly the conglomerate facies, could have then been reworked and transported farther, resulting in the deposition of more sandstone-rich younger strata (Fig.
-\ref{fig:strat_column}).
-The relative lack of conglomerates at Agate Falls which is $\sim$25 km away from the Keweenaw fault could indicate that the Jacobsville sediments there are derived from far-travelled sediments.
-Recent field mapping near Baby Snake Creek have found evidence for the Jacobsville Formation onlapping the Midcontinent Rift volcanics that were previously faulted via en echelon segments\cite{Tyrrell2019a, Mueller2021a}.
-The unconformable contact between the sediments and the volcanics, instead of faulted contact typically observed near Hungarian could indicate that the deposition of the Jacobsville Formation near the northern end of Keweenaw Peninsula likely post-date some episodes of fault motion, and are thus younger than the sediments near the fault.
-Therefore, we interpret that the normal-polarity strata at Agate Falls and at Baby Snake Creek could be stratigraphically the youngest amongst the five localities, although a possibility that they instead are older strata deposited during the Keweenawan superchron \cite{Driscoll2016b} cannot be ruled out.
+A total of 186 specimens yielded interpretable chemical remanence directions that are dominantly reversed polarity.
+These directions are plotted by stratigraphic section location in Figure S4 and are all shown in Figure 6B.
+That the CRM directions are a secondary remanence component that postdate the deposition of the Jacobsville Formation is supported by the negative intraformational conglomerate test results and failure in the fold test (Figs.
+\ref{fig:intraclast_pmag}, S2).
+However, the similarity of the CRM directions (declination=96.2\textdegree, inclination=15.1\textdegree, $\alpha_{95}$=4.2\textdegree, n=186) with the primary DRM directions (declination=103.1\textdegree, inclination=13.4\textdegree, $\alpha_{95}$=3.3\textdegree, n=307, no inclination shallowing correction) indicate that growth of the secondary pigmentary hematite which carries the CRM likely occurred relatively soon after deposition when Laurentia was in a similar paleogeographic position.
 
-Overall, the multiple geomagnetic field polarity reversals recorded by the Jacobsville Formation constrain the maximum range of the Keweenawan superchron \cite{Driscoll2016b} to be between ca.
-1100 Ma (Mamainse Point Volcanics; \cite{Swanson-Hysell2019a}) to ca.
-990 Ma (maximum deposition age of Jacobsville Formation; \cite{Hodgin2022a}).
+Overall, the multiple geomagnetic field polarity reversals recorded by the Jacobsville Formation constrain that the end of the Keweenawan normal superchron \cite{Driscoll2016b} (which started ca.
+1099 Ma; \cite{Swanson-Hysell2019a}) had to have ended by the onset of ca.
+990 Ma Jacobsville deposition \cite{Hodgin2022a}.
 
 \section*{Discussion}
+\subsection*{Averaging paleosecular variation and correcting for inclination shallowing}
 
-\subsection*{Inclination shallowing correction}
+Igneous and sedimentary rocks associated with the North American Midcontinent Rift provide a high-resolution record of the ca.
+1109 to 1070 Ma Keweenawan Track which tightly constrains the apparent polar wander path for Laurentia in the late Mesoproterozoic (Fig.
+\ref{fig:pole_plot}; \citeA{Swanson-Hysell2019a}).
+However, high-quality paleogeographic constraints thereafter are sparser and more uncertain in their age until the ca.
+775 Ma Gunbarrel large igneous province \cite{Harlan2003a} and subsequent ca.
+775-719 Ma poles from western Laurentia extensional basins \cite{Weil2006a, Eyster2019a}.
+Developing a new paleomagnetic pole from the Jacobsville Formation presents an opportunity for a well-constrained early Neoproterozoic pole given that the depositional age of the formation has recently been constrained to ca.
+990 Ma \cite{Hodgin2022a}.
 
-\begin{figure*}[H!]
+Our high-resolution thermal demagnetization successfully isolated detrital magnetization from the secondary chemical magnetization carried by pigmentary hematite that grew after deposition of the Jacobsville Formation.
+The scatter in the DRM directions can be interpreted to reflect variations in the geomagnetic field during the deposition of the sediments \cite{Steiner1983a, Tauxe1984a}.
+The multiple geomagnetic field reversals captured by specimens collected from the Hungarian Falls and Agate Falls sections support that a prolonged period of time is represented by the specimens given that the duration of geomagnetic polarity chrons are typically on the order of tens of Kyr to Myr timescales (Fig.
+\ref{fig:strat_column}).
+Therefore, we combine all specimen detrital remanent magnetizations to calculate a paleomagnetic pole for the Jacobsville Formation.
+
+A challenge in interpreting detrital magnetizations is the issue of correcting for inclination shallowing \cite{King1955a, Tauxe2004b, Bilardello2016b}.
+Scatter of the specimen detrital remanence directions in tilt-corrected coordinates show elongation parallel to the bedding plane, consistent with them being shallowed (Fig.
+\ref{fig:in_situ_pmag}; \citeA{Tauxe2004b}).
+Such a directional distribution is due to rotation of detrital hematite grains during deposition and subsequent compaction, resulting in inclinations recorded by sedimentary rocks having shallower angles than the local field in which they are deposited.
+If uncorrected, shallower inclinations obtained from sedimentary rocks can result in erroneously low estimates of paleolatitudes, biasing paleogeographic reconstructions \cite{Muttoni2003a, Domeier2012a}.
+To estimate the amount of inclination shallowing, we apply the statistical elongation/inclination (E/I) method that evaluates the deviation of the distribution of a large number ($>$100) of observed sedimentary paleomagnetic directional data with the predicted distribution given by a statistical paleosecular variation model \cite{Tauxe2004b}.
+Although the TK03 paleosecular variation model is based on data of relatively recent geomagnetic field variations, it has been shown to be compatible with data from the ca.
+1.1 Ga lava flows of the Midcontinent Rift \cite{Tauxe2009a}.
+Furthermore, \citeA{Pierce2022a} showed that applying the E/I method to detrital hematite magnetizations within the ca.
+1093 Ma Cut Face Creek Sandstone in the Midcontinent Rift successfully corrects the shallow paleomagnetic inclinations of the sediments to that of the North Shore Volcanic Group lava flows that bracket them.
+\citeA{Pierce2022a} also developed a method to represent uncertainties associated with the amount of inclination shallowing into sedimentary paleomagnetic pole positions which can then be summarized with an elliptical Kent distribution \cite{Kent1982a} rather than a circular Fisher distribution \cite{Fisher1953a}.
+
+Applying the E/I method to the Jacobsville DRM directions results in an estimate for the flattening factor ($f$ factor) of 0.64 with a 95\% uncertainty range of 0.83 to 0.51 (Figs.
+\ref{fig:EI_results}).
+The $f$ factor of 0.64 is close to the value of 0.58 which is a the mean of inclination shallowing factors compiled for hematite-bearing rocks in \citeA{Pierce2022a} (building on compilations of \citeA{Bilardello2016b} and \citeA{Vaes2021a}).
+The Fisher mean inclination-corrected paleomagnetic pole associated with this nominal $f$ factor is Plon=183.4\textdegree E, Plat=16.9\textdegree S, A$_{95}$=3.1\textdegree.
+We apply the method of \citeA{Pierce2022a} to incorporate the uncertainty in the E/I estimate of the $f$ factor into the uncertainty of the paleomagnetic pole (Fig.
+\ref{fig:EI_results}).
+This uncertainty results in more uncertainty associated with paleolatitude which for the pole is along the great circle path between the mean pole position and the locality of the Jacobsville sections.
+Following \citeA{Pierce2022a}, the pole can be represented with a Kent distribution 95\% confidence ellipse which is: mean longitude=183.4\textdegree E, mean latitude=16.9\textdegree S, major axis longitude=255.5\textdegree E, major axis latitude=45.2\textdegree N, major axis magnitude=4.1\textdegree, minor axis longitude=108.1\textdegree E, minor axis latitude=39.9\textdegree N, minor axis magnitude=3.1\textdegree (Fig.
+\ref{fig:EI_results}; Table \ref{tab:Kent_means})
+
+% To further represent the uncertainty associated with the $f$ factor, and take all estimated $f$ values from the bootstrap results of the E/I algorithm and apply each value to correct all Jacobsville DRMs (with the reversed directions flipped to antipodes).
+Then we calculate the Fisher statistics for each of the inclination-corrected mean poles.
+Finally, we perform a Monte Carlo approach to generate 100 resamples of each Fisher mean pole.
+As a result, the distribution of the resampled mean pole positions have an elongated distribution which represents more uncertainty associated with paleolatitudes of Laurentia than the orientation of the continent (Fig.
+\ref{fig:EI_results}).
+The distribution of paleolatitudes of the resampled mean poles can be approximated by a normal distribution (Fig.
+\ref{fig:EI_results}).
+Following the \citeA{Pierce2022a} methodology, we then use the Kent distribution to determine the 95\% confidence ellipse of the distribution of the resampled mean poles (mean longitude=183.4\textdegree E, mean latitude=17.1\textdegree S, major axis longitude=255.3\textdegree E, major axis latitude=45.2\textdegree N, major axis magnitude=4.2\textdegree, minor axis longitude=108.2\textdegree E, minor axis latitude=39.8\textdegree N, minor axis magnitude=3.1\textdegree; Fig.
+\ref{fig:EI_results}).
+As compared to the paleomagnetic pole calculated with Fisher statistics using the nominal inclination shallowing correction factor of $f$=0.66, this Kent ellipse indicates that there is more uncertainty along the great circle path between the mean pole position and the locality of the Jacobsville basin (Fig.
+\ref{fig:EI_results}; Table \ref{tab:Kent_means}).
+
+\begin{figure*}[h!]
 \centering
 \includegraphics[width=\textwidth]{EI_results.pdf}
-\caption{\footnotesize Results of the estimated amount of inclination shallowing of the detrital remanent magnetization of the Jacobsville Formation using the elongation/inclination (E/I) method \cite{Tauxe2004b}.
-Directions from both polarities are used with the reversed-polarity directions flipped to their antipodes (Fig.
-\ref{fig:in_situ_pmag}).
-A) The E/I method results in an estimated flattening factor $f$=0.61 (red curve) based on where elongation/inclination curve of the dataset intersects that predicted by the TK03 paleosecular variation model (black curve; \citeA{Tauxe2004b}).
-The grey lines show 1,000 bootstrap resamples of the DRM directions which provide an estimate of the uncertainty associated with the $f$ factor estimate.
-B) The distribution of the paleolatitudes implied from the resultant corrected inclinations from the E/I bootstrap resamples.
-That the distribution is drawn from a normal distribution with a mean of 11.0 and standard deviation of 1.2 cannot be rejected in the Kolmogorov-Smirnov test.
-The 95\% confidence bounds shown by dashed lnes span a range of paleolatitudes that need to be incorporated into the uncertainty of the reported paleomagnetic pole.
-C) The cumulative distribution of all plausible inclinations based on the E/I bootstrap results with the 95\% confidence bounds shown in inclination and $f$ value axes.
-D) Kent mean pole and Fisher mean pole of the Jacobsville Formation in context of previous poles from \citeA{Roy1978a} and the pole position derived from the Chequamegon Formation of the Bayfield Group developed by \citeA{McCabe1983a}.
-The mean pole position of the Kent distribution calculated from the distribution of resampled Jacobsville mean poles following the method of \citeA{Pierce2022a} overlaps with the Fisher mean pole position corrected by a single inclination shallowing value of 0.61.
-Both mean poles lie in between the Fisher mean poles calculated using the inclination-corrected normal and reversed pole groups respectfully.
-The close proximity between Jacobsville mean poles with the Chequamegon pole indicate their deposition could be largely coeval.
-The J1 (C) pole of \citeA{Roy1978a} is recorded by hematite-bearing sediments in Sault Ste Marie area.
-Its position in the northern hemisphere close to the Upper Freda Formation pole indicate that those sediment could be Freda-equivalent instead of Jacobsville-equivalent.
-$^*$---Fisher mean pole positions calculated with inclination shallowing correction given by the E/I results shown in Fig.
-\ref{fig:in_situ_pmag}.
-$^+$---the Freda Formation Fisher mean pole is corrected for inclination shallowing assuming $f$=0.6.
+\caption{\scriptsize Results of the estimated amount of inclination shallowing of the detrital remanent magnetization of the Jacobsville Formation using the elongation/inclination (E/I) method \cite{Tauxe2004b}.
+(A) The E/I method givens an estimated flattening factor $f$=0.64 (red curve) based on where the elongation/inclination curve of the dataset intersects that predicted by the TK03 paleosecular variation model (black curve; \citeA{Tauxe2004b}).
+The grey lines show 1,000 bootstrap resamples of the DRM directions which provide an estimate of the uncertainty associated with the $f$ factor estimate (95\% interval from 0.83 to 0.51).
+(B) The distribution of the paleolatitudes resulting from the corrected inclinations of the E/I bootstrap resamples.
+That the distribution is drawn from a normal distribution with a mean of -11.5\textdegree\ and standard deviation of 1.4 cannot be rejected in the Kolmogorov-Smirnov test.
+The 95\% confidence bounds shown by dashed lines span a range of paleolatitudes that need to be incorporated into the uncertainty of the reported paleomagnetic pole.
+(C) The cumulative distribution of inclinations based on the E/I bootstrap results with the 95\% confidence bounds shown in terms of inclination and $f$ factor.
+(D) The new Kent mean inclination corrected DRM pole (red square) of the Jacobsville Formation is shown in the context of the pole calculated for the CRM (purple triangle) as well as the Jacobsville pole developed by \citeA{Roy1978a}, the Chequamegon Formation pole of \citeA{McCabe1983a}, and poles from the Oronto Group (Nonesuch Formation, \citeA{Slotznick2023a}, Freda Formation, \citeA{Henry1977a}).
+The mean pole position of the Kent distribution was calculated following the method of \citeA{Pierce2022a}.
+The close proximity between Jacobsville poles with the Chequamegon pole is consistent with their deposition being coeval.
+In contrast, the J1 (C) pole of \citeA{Roy1978a} developed from hematite-bearing sedimentary rocks in the Sault Ste Marie region plots in the northern hemisphere close to the Freda Formation pole.
 }
@@
 
-Igneous and sedimentary rocks associated with the North American Midcontinent Rift provide high-resolution record of the ca.
-1109 to 1070 Ma Keweenawan Track which tightly constrains the apparent polar wander path for Laurentia in the late Mesoproterozoic (Fig.
-\ref{fig:pole_plot}; \cite{Swanson-Hysell2019a}).
-However, high-quality paleogeographic constraints is lacking thereafter since magmatism in the interior of Laurentia entered a protracted quiescence period until the emplacement of the ca.
-780 Ma Gunbarrel large igneous province \cite{Harlan2003a}.
-The new paleomagnetic pole developed from the Jacobsville Formation presents an opportunity to extend the APWP to the early Neoproterozoic as its maximum deposition age has been tightly constrained to be ca.
-990 Ma through high-precision U-Pb detrital zircon geochronology and U-Pb calcite geochronology \cite{Hodgin2022a}.
+\begin{table}[h!]
+   \caption{Kent mean paleomagnetic pole for the Jacobsville Formation} 
+   \label{tab:Kent_means}
+   \footnotesize
+   \singlespacing
+   \centering
+   \begin{tabular}{p{0.8 in}p{0.8 in}p{0.8 in}p{0.8 in}p{0.8 in}p{0.8 in}}
+   \hline
+   pole & mean pole position (Plon/Plat) & major axis & major axis 95\% confidence angle & minor axis & minor axis 95\% confidence angle \\
+   & $\gamma_{1}$ & $\gamma_{2}$ & $\zeta_{95}$ & $\gamma_{3}$ & $\eta_{95}$ \\
+   \hline
+   Jacobsville $E/I$ corrected & 183.4\textdegree E / 16.9\textdegree S &  255.5\textdegree E / 45.2\textdegree N & 4.1\textdegree &
+   108.1\textdegree E / 39.9\textdegree N & 3.1\textdegree \\
+   \hline
+ \multicolumn{6}{p{5.5 in}}{Notes: The Fisher mean of the Jacobsville detrital remanence paleomagnetic pole without an inclination shallowing correction is Plon=185.5\textdegree E, Plat=14.0\textdegree S, A$_{95}$=2.7\textdegree; the Fisher mean of the Jacobsville detrital remanence paleomagnetic pole with an inclination shallowing correction of $f$=0.64 is Plon=183.4\textdegree E, Plat=16.9\textdegree S, A$_{95}$=3.1\textdegree; the Fisher mean of the Jacobsville chemical remanence paleomagnetic pole is Plon=179.6\textdegree E, Plat=10.2\textdegree S, A$_{95}$=3.6\textdegree}
+   \end{tabular}
+\end{table}
 
-However, inclination bias associated with sedimentary paleomagnetic directions need to be corrected when synthesizing poles.
-Rotation of grains during sediment deposition and compaction during early diagenesis can result in the acquisition of a detrital remanent magnetization that is shallower than the local geomagnetic field in which it was acquired \cite{King1955a, Tauxe2005a, Kodama2012a}.
-This effect can be summarized as:
-\begin{equation*}
-\textup{tan}(I_o) = f\textup{tan}(I_f)
-\end{equation*}
-where $f$ represents the amount of inclination shallowing, $I_o$ represents the observed inclination, and $I_f$ represents the inclination of the field in which the magnetization was acquired.
-If uncorrected, shallower inclinations obtained from sedimentary rocks can result in erroneously low estimates of paleolatitudes, biasing paleogeographic reconstructions \cite{Domeier2012a}.
-Methods for correcting sedimentary inclination bias include laboratory-based measurements of magnetic mineral fabrics \cite<e.g.>{Kodama1990a, Bilardello2015a}, as well as statistical methods based on the evaluation of the deviation of an observed virtual geomagnetic pole (VGP) distribution from that of a paleosecular variation model (the ``elongation/inclination" [E/I] method of \citeA{Tauxe2004b}).
-When a large number of sedimentary paleomagnetic sites (typically n$>$100) capture an extensive period of time such that the paleosecular variation is sufficiently averaged, the E/I method can give efficient estimates with reasonable uncertainty bounds via bootstrapping and is less labor-intensive than methods based on mineral fabric measurements.
-\citeA{Tauxe2009a} based on paleomagnetic data from lava flows of the North Shore Volcanic Group in the Midcontinent Rift to show that the ca.
-1.1 Ga geomagnetic field had a behavior consistent with the prediction given by the TK03 paleosecular variation model of \citeA{Tauxe2004b}.
-\citeA{Pierce2022a} showed that the E/I method is effective in correcting the inclination bias recorded by a package of lava flow-bracketed fluvial red bed sediments---the Cut Face Creek Sandstone.
+\citeA{Roy1978a} also studied the paleomagnetism of sedimentary rocks grouped as the Jacobsville Formation and isolated a characteristic magnetization component via alternating-field, thermal, and chemical demagnetization on a suite of Jacobsville red beds near Keweenaw Peninsula (their area A), the town of Marquette (their area B), and Sault Ste Marie (their area C).
+Although least-squares principal component analyses was not a routine approach in fitting paleomagnetic directions at the time, that study had success in removing the present day local field overprint and was able to resolve dual magnetic polarities at one locality.
+The mean paleomagnetic pole position calculated from the interpreted primary magnetic remanence from rocks of the Keweenaw Peninsula and Marquette area (J1 A+B pole; pole longitude=183\textdegree E, pole latitude=9\textdegree S, dp=3\textdegree, dm=6\textdegree; no inclination correction; Fig.
+\ref{fig:EI_results}; \citeA{Roy1978a}) is in the southern hemisphere and lies close to the mean pole from this study (Fig.
+\ref{fig:EI_results}).
+However, their mean pole developed from fine- and medium-grained sandstone near Sault Ste Marie lies in the northern hemisphere with a pole latitude of 12\textdegree N (Fig.
+\ref{fig:EI_results}).
+Despite the large uncertainty ellipse associated with this mean pole position, it is distinct from the mean pole position from the Jacobsville Formation in northern Michigan.
+Instead, it overlaps with the Oronto Group paleomagnetic poles of the ca.
+1070 Ma lower Freda Formation and the ca.
+1075 Ma Nonesuch Formation (Fig.
+\ref{fig:EI_results}; \citeA{Henry1977a, Slotznick2023a}).
+We hypothesize that the red beds studied by \citeA{Roy1978a} from the Sault Ste Marie area are likely time equivalent to sedimentary rocks of the Oronto Group, instead of being correlative with the Jacobsville Formation.
 
-Inclination shallowing is also present in the fluvial red beds of the Jacobsville Formation.
-Scatter of the specimen detrital remanence directions in geographic coordinates show elongation of DRM directions parallel to the bedding plane, indicating the directions are shallowed (Fig.
-\ref{fig:in_situ_pmag}; \citeA{Tauxe2004b}).
-Since the deposition of the Jacobsville Formation span over geomagnetic field reversals, it is of interest to investigate the amount of inclination shallowing associated with the normal (n=130) and the reversed (n=148) polarity respectively (transitional fields and specimens affected by complex tilting regime are excluded).
-Applying the E/I method on the normal polarity directions yields an estimated $f$ value of 0.67 with 95\% bootstrap uncertainty bounds from 0.91 to 0.55 (Fig.
-\ref{fig:in_situ_pmag}B).
-Directions of reversed polarity yields an estimated $f$ factor of 0.52 with uncertainty bounds between 0.97 and 0.29 (Fig.
-\ref{fig:in_situ_pmag}B).
-Although the estimated $f$ value for the reversed-polarity directions has large statistical uncertainty bounds, the uncertainties are likely associated with the overall shallow inclinations and distributions of directions in both upper and lower hemispheres (Fig.
-\ref{fig:in_situ_pmag}B).
-Therefore, inclination corrections do not result in significant changes in the mean reversed direction.
+Deposition of Bayfield Group sedimentary rocks in northern Wisconsin (west of the studied exposures; Fig.
+\ref{fig:Geologic_map}) has been hypothesized to be coeval with the Jacobsville Formation \cite{Hamblin1958a, Kalliokoski1982a, Malone2016a}.
+Published low precision detrital zircon U-Pb dates developed by LA-ICP-MS dates are consistent with this interpretation with a maximum depositional age of ca.
+1035 Ma for the Chequamegon Formation of the Bayfield Group \cite{Craddock2013a}.
+Our updated Jacobsville pole position is close to a pole developed from the Chequamegon Sandstone of the Bayfield Group (Fig.
+\ref{fig:EI_results}; pole longitude=177.7\textdegree E, pole latitude=12.3\textdegree S, A$_95$=4.6\textdegree; no inclination correction; \citeA{McCabe1983a}).
+These data are consistent with the Bayfield Group being deposited within a contiguous syn-orogenic basin with the Jacobsville Formation during the Rigolet phase of the Grenvillian orogeny.
+This interpretation can be evaluated with further high-precision geochronology focused on the Bayfield Group.
 
-In addition, directions of the two polarities do not pass a reversal test \cite{McFadden1990a, Tauxe2010a}, regardless of whether inclination correction is applied.
-One interpretation of the asymmetric polarities could be that the deposition of the Jacobsville Formation was protracted such that the sediments at Hungarian Falls which dominate the reversed polarity records were deposited prior to those at Baby Snake Creek which dominate the normal polarity records (Fig.
-\ref{fig:in_situ_pmag}).
-This is consistent with our stratigraphic interpretation that Baby Snake Creek sediments could be younger than those near central Keweenaw Peninsula (i.e.
-Hungarian Falls, Natural Wall, Hammell Creek; Fig.
-\ref{fig:strat_column}).
-Other alternative interpretations could include that the mean poles are biased by transient non-dipolar fields due to insufficient averaging of paleosecular variation or transitional fields during field reversals.
-More data are needed to test these hypotheses.
+\subsection*{Slowdown of Laurentia's plate motion due to the Grenvillian orogeny}
 
-Overall, both inclination-corrected polarities recorded by the Jacobsville Formation show mean pole positions in the southern hemisphere with respect to today's continental configuration (normal polarity longitude=175.1\textdegree E, latitude=19.1\textdegree S, A95=3.8\textdegree; reversed polarity longitude=191.1\textdegree E, latitude=14.5\textdegree S, A95=3.9\textdegree; Fig.
-\ref{fig:EI_results}).
-The two pole positions are distinct from the youngest poles developed from rocks of the Midcontinent Rift (Fig.
-\ref{fig:EI_results}).
-Despite that the Jacobsville sedimentation span over geomagnetic field reversals, the large arc distance between either polarity of the Jacobsville Formation and the pole of the ca.
-1070 Ma upper Freda Formation \cite{Henry1977a} is consistent with the interpretation that the onset of deposition of the Jacobsville Formation at the five studied localities post-date the youngest member of the Oronto Group sediments by tens of millions of years.
-
-% Field observations of \citeA{Tyrrell2019a} and \citeA{Mueller2021a} near Bete Grise Bay, which is only $\sim$10 km east to Baby Snake Creek (Fig.
-\ref{fig:Geologic_map}), show unconformable onlap of the Jacobsville Formation on top of Midcontinent Rift volcanics that have experienced significant thrusting, deformation.
-At some locations, paleosol development has been found on top of the Midcontinent Rift basalt along the segmented Keweenaw Fault branches prior to the deposition of the Jacobsville sediments.
-This is in contrast with the faulted contact between the Jacobsville Formation and the volcanics to the south where the Jacobsville Formation is often folded against the volcanics (Fig.
-\ref{fig:Geologic_map}).
-U-Pb calcite geochronology and thermometry confirm that the final $\sim$2 km vertical displacement along the Keweenaw Fault occurred during the ca.
-980 Ma Rigolet phase of the Grenvillian Orogeny whereas the majority of the fault motion likely occurred during the ca.
-1050 Ma Ottawan phase \cite{Hodgin2022a}.
-We interpret that it is most likely that deposition of the Jacobsville Formation initiated close to the Rigolet phase of the Grenvillian Orogeny with progressive sediment transportation to the northwest as the Grenville back-bulge basin propagated toward the interior of Laurentia \cite{Hodgin2022a}.
-As a result, we hypothesize that the sediments at Baby Snake Creek likely records a slightly younger pole position with a more southerly paleolatitude than those near Hungarian Falls, indicating Laurentia's plate motion toward the south (Fig.
-\ref{fig:EI_results}).
-However, we cannot rule out an alternative interpretation that the sediments at Baby Snake Creek are coeval or older than some of the sediments to the south, as the basal Jacobsville Formation at Baby Snake Creek is folded against the Midcontinent Rift volcanics.
-Further field mapping and geochronological constraints are needed to test this hypothesis.
-
-\subsection*{The Jacobsville paleomagnetic pole}
+The ca.
+1109-1070 Ma Keweenawan Track constrains rapid motion of Laurentia from high latitudes toward the equator (Fig.
+\ref{fig:paleogeography}; \citeA{Davis1997a, Swanson-Hysell2019a}).
+Apparent polar wander path inversion results are consistent with the pole path being dominated by plate tectonic motion that could have reached a rate of 30 cm/yr (95\% interval of 27-34 cm/yr; one Euler pole scenario; \citeA{Swanson-Hysell2019a, Rose2022a}); even faster than the rate of Indian plate motion during the closure of the Neotethys Ocean \cite{Hinsbergen2022a, Jagoutz2015a}.
+Laurentia's rapid motion preceded and coincided with the onset of the collisional Grenvillian orogeny.
+In contrast to the rapid changes in pole position associated with the Keweenawan Track, the relatively close proximity between the ca.
+990 Ma Jacobsville pole and ca.
+1085 to 1070 Ma poles indicate that Laurentia's motion significantly slowed following the onset of the Grenvillian orogeny (Figs.
+\ref{fig:pole_plot} and \ref{fig:paleogeography}).
+Lacking chronostratigraphic constraints, previous treatments have assigned an age to the Jacobsville pole of \citeA{Roy1978a} through extrapolation of the motion of the Keweenawan Track \cite<e.g.>{Li2008a}.
+With the improved age constraints of \citeA{Hodgin2022a} and our new paleomagnetic pole, we can instead constrain that Laurentia's motion decreased by an order of magnitude.
+Incorporating temporal and spatial uncertainty on the poles gives APWP rate of $\sim$3.2 cm/yr (95\% interval from 2.5 cm/yr to 4.3 cm/yr; Fig.
+S5) between ca.
+1070 to 990 Ma.
 
@@
 \includegraphics[width=\textwidth]{Jacobsville_pole_plot.pdf}
-\caption{The Jacobsville Formation Kent mean pole position is plotted in context of the ca.
-1109-1070 Ma Keweenawan Track (poles color-coded by age shown in the colorbar) and selected poles from the Grenville Province.
+\caption{The inclination-corrected Jacobsville Formation mean pole position with its Kent uncertainty ellipse is plotted in context of the ca.
+1109-1070 Ma Keweenawan Track (poles color-coded by age shown in the colorbar), selected poles from the Grenville Province, and late Tonian poles of Laurentia as compiled by \citeA{Eyster2019a} (orange colored poles).
 The southerly pole position of the ca.
-990 Ma Jacobsville Formation indicates Laurentia crossed the equator in the late Mesoproterozoic to early Neoproterozoic.
-The large arc distances between the Jacobsville pole and the poles of the Haliburton intrusions \cite{Buchan1976a} and the Wilberforce pyroxenite \cite{Palmer1973a} (black) indicate previous estimates of them being $>$1000 Ma \cite{Halls2015b} based on poor age and temperature calibration of their paleomagnetic and thermochronologic data are too old.
-\citeA{Brown2012a} developed poles (blue) from rocks of the Adirondack Mountains and estimated their ages to be ca.
-960 Ma to 990 Ma, with the gneiss pole (lightest blue) being the youngest and the granites pole (darkest blue) being the oldest.
-However, the estimated closure temperatures for rutile was too low, biasing the estimated pole age to be too old.
-The true age of the Grenville poles could be much older than previously thought in light of paleomagnetic and chronological constraints from the Jacobsville Formation.
+990 Ma Jacobsville Formation indicates that the study area was crossing the equator in the late Mesoproterozoic to early Neoproterozoic.
+Given the large arc distance between the ca.
+990 Ma Jacobsville pole and the poles of the Grenville loop, such as those of the the Haliburton intrusions, we suggest that estimates of their age as $>$1000 Ma \cite{Warnock2000a, Halls2015b, Evans2021b} are likely too old.
+The pole compilation used in this figure is included as Table S1.
 }
@@
 
-In light of the lithofacies model and detrital remanent magnetization data of the Jacobsville Formation, we prefer the interpretation that the deposition of the Jacobsville Formation initiated long after the the Oronto Group sedimentation and likely spanned a short time period close to the end of the ca.
-1010 to 980 Ma Rigolet phase of the Grenville Orogeny \cite{Rivers2012a, Hodgin2022a}.
-Therefore, we combine DRM directions of both polarities of the ca.
-990 Ma Jacobsville Formation to succinctly report one paleomagnetic pole for the formation.
-We reassess the amount of inclination shallowing of the combined directional dataset (Fig.
-\ref{fig:in_situ_pmag}).
-The resultant estimate for the $f$ value is 0.61 with a bootstrap 95\% uncertainty range for the $f$ value between 0.80 to 0.51, which overlaps with the uncertainty ranges of directions associated with either polarity (Fig.
-\ref{fig:in_situ_pmag}).
-The $f$ value of 0.61 is strikingly close to the value of 0.59 which is a calculated mean value of a compilation of inclination shallowing factor for hematite-bearing rocks derived by \cite{Bilardello2010b}.
-The Fisher mean paleomagnetic pole associated with the combined directions is shown in Fig.
-\ref{fig:EI_results} (pole longitude= 186.3\textdegree E, pole latitude= 14.3\textdegree S, A95= 2.5\textdegree).
+This significant slow down in plate motion is geodynamically consistent with the onset of Grenvillian orogenesis.
+The earliest records of orogenesis on the leading margin of Laurentia occur ca.
+1090 Ma with estimates of peak orogenesis associated with the Ottawan stage of the orogeny occurring ca.
+1050 Ma \cite{Rivers2012a, Swanson-Hysell2023a}.
+Inversions of the Keweenawan Track that incorporate multiple tectonic Euler poles indicate a slowdown from rates exceeding 20 cm/yr prior to 1095 Ma to rates below 20 cm/yr after 1095 Ma \cite{Swanson-Hysell2019a, Rose2022a}.
+This initial slowdown could be associated with soft collision as contractional deformation initiated on the leading edge of Laurentia \cite{Staal2020a}.
+Continued contractional collision led to substantially thickened crust recorded by granulite-facies metamorphic rocks by the time of peak Ottawan phase metamorphism ca.
+1050 Ma \cite{Rivers2012a}.
+The large slowdown in plate speed as now constrained by the Jacobsville paleomagnetic pole is associated with this progression to hard continent-continent collision involving increased crustal thickening and progressive transmission of contractional stress that eventually developed into the interior of Laurentia \cite{Cannon1994a}.
 
-\citeA{Pierce2022a} discussed that as inclination shallowing processes usually do not bias the recorded declination, it would typically result in asymmetric uncertainties in the recorded paleomagnetic pole position with more uncertainties in paleolatitues than paleolongitudes.
-The asymmetric uncertainties in the mean pole position thus violate the assumption of the Fisher distribution \cite{Fisher1953a} where uncertainties are circularly distributed about a mean vector.
-Instead, the spherical asymmetric bivariate Kent distribution \cite{Kent1982a} is more appropriate for incorporating uncertainties associated with inclination shallowing in reported sedimentary paleomagnetic poles.
-We follow the approach of \citeA{Pierce2022a} who extended the E/I method of \citeA{Tauxe2004b}.
-First we take all estimated $f$ values from the bootstrap results of the E/I algorithm and apply each value to correct all Jacobsville detrital remanence directions (with the reversed directions flipped to antipodes).
-Then we calculate the Fisher statistics for each of the inclination-corrected mean pole.
-Finally, we perform a Monte Carlo approach to generate 100 resamples of each Fisher mean pole.
-As a result, the distribution of the resampled mean pole positions have an elongated distribution which represents more uncertainty associated with paleolatitudes of Laurentia than paleolongitudes (Fig.
-\ref{fig:EI_results}).
-The distribution of paleolatitudes of the resampled mean poles can be well-approximated by a normal distribution (Fig.
-\ref{fig:EI_results}).
-We use the Kent distribution to parameterize the 95\% confidence ellipse of the distribution of the resampled mean poles (mean longitude=186.4\textdegree E, mean latitude=14.2\textdegree S, major axis longitude=262.5, major axis latitude=43.5, major axis magnitude=3.3\textdegree, minor axis longitude=110.0, minor axis latitude=43.0, minor axis magnitude=3.0\textdegree; Fig.
-\ref{fig:EI_results}).
-As is shown in Fig.
-\ref{fig:EI_results}, although the Kent uncertainty ellipse has significant overlap with the circular uncertainty range calculated by the Fisher statistics due to the overall low paleolatitude of the Jacobsville basin, the Kent ellipse informs more uncertainty along the great circle path between the mean pole position and the locality of the basin as it incorporates the uncertainties in the amount of inclination shallowing within the sediments.
+Laurentia's rapid motion was associated with closure of the Unimos Ocean and is consistent with it being on the lower plate that subducted under a conjugate continent (e.g.
+Amazonia) that collided on its margin (\citeA{Swanson-Hysell2023a}; Fig.
+\ref{fig:paleogeography}).
+The Jacobsville paleomagnetic pole constrains this motion to have slowed to below average plate speed as orogeny progressed to be a large-scale continent-continent collision.
+As a result, Laurentia maintained a low-latitude position throughout the duration of the Grenvillian orogeny which sutured continents together into the supercontinent Rodinia (Fig.
+\ref{fig:paleogeography}).
 
-\citeA{Roy1978a} isolated the detrital magnetization component via alternating-field, thermal, and chemical ``cleaning" demagnetization on a suite of Jacobsville red beds near Keweenaw Peninsula (their area A), the town of Marquette (area B), and Sault Ste Marie (area C).
-Although least-squares principal component analyses was not a routine approach in fitting paleomagnetic directions at the time, that study had success in removing the present day local field overprint and was able to resolve a magnetic polarity transition at one locality.
-The mean paleomagnetic pole position calculated from the interpreted primary magnetic remanence from rocks of the Keweenaw Peninsula and Marquette (J1 A+B pole; pole longitude=183\textdegree E, pole latitude=9\textdegree S, dp=3\textdegree, dm=6\textdegree; no inclination correction; Fig.
-\ref{fig:EI_results}; \citeA{Roy1978a}) is in the southern hemisphere and the uncertainty ellipse overlaps with that of the mean pole from this study (Fig.
-\ref{fig:EI_results}.
-However, their mean pole developed from red beds near Sault Ste Marie lies in the northern hemisphere with a latitude of 12\textdegree N (Fig.
-\ref{fig:EI_results}).
-Despite the large uncertainty ellipse associated with this mean pole position, it is distinct from the mean pole position from Jacobsville rocks in the Keweenaw Peninsula and Marquette.
-Instead, it overlaps with the mean poles of the ca.
-1070 Ma Upper Freda Formation and the ca.
-1075 Ma Nonesuch Formation (Fig.
-\ref{fig:EI_results}; \citeA{Henry1977a}).
-We hypothesize that the red beds collected by \citeA{Roy1978a} in Sault Ste Marie area are likely Freda-equivalent sediments, instead of being a part of the Jacobsville Formation.
+\subsection*{Implications for the age of the Grenville Loop}
 
-Deposition of the Bayfield Group sediments west to the Keweenaw Peninsula has been hypothesized to be coeval with the Jacobsville Formation but chronological constraints have been lacking \cite{Hamblin1958a, Kalliokoski1982a, Malone2016a}.
-Our new paleomagnetic data from the Jacobsville Formation confirm that the pole positions of the the Chequamegon Sandstone of the Bayfield Group and the Jacobsville Formation are close albeit distinct (Fig.
-\ref{fig:EI_results}).
-Given that the paleomagnetic pole from the Chequamegon Sandstone of the Bayfield Group (pole longitude=177.7\textdegree E, pole latitude=12.3\textdegree S, A95=4.6\textdegree; \citeA{McCabe1983a}) was calculated without applying inclination correction, the two formations could have an even closer temporal association than what appears in the current paleomagnetic pole configuration (Fig.
-\ref{fig:EI_results}).
-In addition, the fact that only reversed polarity directions were observed in the Chequamegon Sandstone indicate that deposition of the sandstone collected by \citeA{McCabe1983a} could be coeval with one of the reversed polarity periods recorded by the Jacobsville Formation (Fig.
-\ref{fig:strat_column}).
-To further test the temporal correlation between the sedimentary units, detailed paleomagnetic study and detrital zircon U-Pb geochronology data from the Chequamegon Sandstone are needed.
+Given that the age and pole position of the Jacobsville Formation had been uncertain and lacking other constraints from sedimentary or volcanic rocks, reconstruction of the paleogeography of Laurentia in the early Neoproterozoic has been reliant on paleomagnetic poles developed from metamorphic rocks within the Grenville Province \cite<e.g.>{Weil1998a}.
+As is shown in Figure \ref{fig:pole_plot}, paleomagnetic poles of the Grenville Loop plot near Australia in present-day coordinates; forming arc distances ranging from $\sim$35\textdegree\ to more than 50\textdegree\ away from poles at the end of the ca.
+1110 to 1070 Ma Keweenawan Track (Fig.
+\ref{fig:pole_plot}).
+Determining ages associated with those Grenville Loop poles is crucial for constraining the motion of Laurentia at this time and the configuration between Laurentia and hypothesized conjugate continents such as Baltica within Rodinia \cite{Cawood2017a, Gong2018a, Swanson-Hysell2021c}.
 
-\subsection*{Paleogeography of Laurentia in the late Proterozoic and the development of supercontinent Rodinia}
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=0.8\textwidth]{Jacobsville_paleogeography.pdf}
+\caption{Paleogeographic position of Laurentia through the late Mesoproterozoic to early Neoproterozoic.
+The color-coded reconstructions show snapshots of Laurentia's position and orientation at 5 Myr intervals from 1110 to 1075 Ma.
+These reconstructions implement the two-stage tectonic Euler pole rotation inversion of \citeA{Swanson-Hysell2019a}.
+The paleogeographic snapshot colored in red shows the position and orientation of Laurentia ca.
+990 Ma as constrained by the Jacobsville paleomagnetic pole developed in this study.
+Interpreted positions of Laurentia's conjugate continents along its margin are shown in grey.
+This reconstruction shows that while Laurentia experienced rapid latitudinal changes through the Keweenawan Track that its motion was much slower between ca.
+1075 and 990 Ma as the Grenvillian orogeny progressed.
+The inset figure compares Laurentia's position predicted by the new Jacobsville paleomagnetic pole (red) and that constrained by the Haliburton intrusion of the Grenvillian orogeny (light blue) which has been assigned an age of 1015 Ma \cite{Warnock2000a}).
+We suggest that this and other Grenville loop poles are younger and that Laurentia traveled from low latitudes toward higher latitudes following ca.
+990 Ma while the Grenville orogen was slowly cooling and exhuming.}
+\label{fig:paleogeography}
+\end{figure*}
 
-Constraining the paleogeography of Laurentia in the late Proterozoic is central to global paleogeography due to its central position in the ancient supercontinent Rodinia.
-During the initial assembly of Rodinia in the late Mesoproterozoic, Laurentia's apparent polar wander is well constrained by the ca.
-1.1 Ga Keweenawan Track thanks to the large number of paleomagnetic poles paired with high-precision geochronology from rocks of the Midcontinent Rift \cite{Swanson-Hysell2019a}.
-Since the end of the Midcontinent Rift magmatism, well-dated paleomagnetic records from the the interior of Laurentia is lacking, and data from rocks associated with the ca.
-1090 to 980 Ma Grenville Orogeny have been used to infer for Laurentia's plate configuration in the late Mesoproterozoic to early Neoproterozoic \cite<e.g.>{Weil1998a, Halls2015b}.
-However, in contrast to rocks of the Midcontinent Rift which usually have straightforward paleomagnetic records thanks to their simple thermal history \cite<e.g.>{Swanson-Hysell2019a}, rocks of the Grenville Province which experienced up to granulite facies metamorphism have much more complicated thermal histories and the timing of their magnetic remanence acquisition and alteration is more difficult to constrain.
-As is shown in Fig.
-\ref{fig:pole_plot}, paleomagnetic poles developed from rocks of the Grenvillian Orogeny plots near Australia with respect to today's continental configuration, forming arc distances ranging from $\sim$35\textdegree to more than 50\textdegree away from the pole of the ca.
-1070 Ma Freda Formation \cite{Henry1977a}.
-Such large distance between pole positions associated with uncertainties in ages of Grenville magnetizations led to the controversial ``Grenville Problem"---did the Grenville rocks acquire magnetic remanence prior to the continental collision and thus the pole discrepancies are a result of significant crustal shortening (i.e.
+In contrast to rocks of the Midcontinent Rift where magnetizations can be confidently assigned to be the same age as the crystallization ages of the rocks \cite<e.g.>{Davis1997a,Fairchild2017a,Swanson-Hysell2019a}, rocks of the Grenville orogen experienced up to granulite facies metamorphism (temperatures $>$900\textdegree C; \citeA<e.g.>{Shinevar2021a, Metzger2021a}) and acquired magnetic remanence during subsequent exhumation \cite{McWilliams1975a, Dunlop1985a, Dodson1985a}.
+During slow post-orogenic exhumation (1-3\textdegree C/Myr; \citeA<e.g.>{Rivers2023a}), magnetic minerals acquired remanence by cooling through an extended temperature range over millions to tens of millions of years \cite{Pullaiah1975a, Dodson1980a} or by crystallizing through exsolution \cite{McEnroe2007a}.
+Determining the age of magnetization in such rocks requires reconstructing cooling histories using isotopic thermochronometers which experience radioisotopic accumulation on mineral-specific closure of elemental diffusion \cite{Dodson1973a, Dodson1985a}.
+
+Based on the interpretation that blocking of magnetic remanence of the magnetite-bearing Haliburton intrusions of the Grenville Central Metamorphic Belt occurred during relatively narrow temperature ranges close to the Curie temperature of magnetite (i.e.
+580\textdegree C), \citeA{Warnock2000a} considered that the timing of remanence acquisition during cooling of the intrusions is bracketed between U-Pb titanite and $^{40}$Ar/$^{39}$Ar hornblende dates.
+As a result, that study interpreted the age associated with the Haliburton paleomagnetic pole to be ca.
+1015 Ma.
+This age has been assigned to the pole in subsequent compilations \cite<e.g.>{Evans2021a}.
+In the central Adirondack Highlands of the Grenville orogen, magnetite-bearing granitic rocks and anorthositic rocks record pole positions that overlap with the Haliburton pole while their ages have been interpreted by \citeA{Brown2012a} to be ca.
+990-970 Ma, based on U-Pb and $^{40}$Ar/$^{39}$Ar thermochronology data developed by \citeA{Mezger1991a}.
+The ca.
+990 Ma Jacobsville pole position is both distinct from the Grenville Loop poles and is consistent with a slowdown of Laurentia's plate motion associated with the Grenvillian orogeny (Fig.
+\ref{fig:pole_plot}).
+If ages that have been assigned to Grenville Loop poles are taken at face value, those poles represent the position of Laurentia both before and after the position constrained by the 990 Ma Jacobsville pole.
+Such an interpretation would imply either rapid plate tectonic motion of Laurentia coeval with the Rigolet phase of the Grenvillian orogeny or rapid oscillatory true polar wander \cite<e.g.>{Evans2003a}.
+Alternatively, a more straightforward model that could explain the distinct pole positions between the Jacobsville pole and the Grenville Loop poles is that the poles from the Grenville orogen are younger than currently interpreted.
+Instead of poles such as that from the Haliburton intrusions having remanence that was acquired during the Grenvillian orogeny, the magnetic remanence could instead have been acquired during protracted exhumation that followed the ca.
+980 Ma cessation of contractional deformation associated with the Rigolet stage of the orogeny.
+In this scenario, the migration of Rodinia to the higher latitude position represented by the Grenville loop poles would have occurred further into the Neoproterozoic after the Grenvillian orogeny had ceased.
+
+While numerous paleomagnetic data sets have been developed from rocks of the Grenville Province, few are paired with well-calibrated, high-precision thermochronology data.
+Recent data sets, such as new apatite U-Pb dates of 923$\pm$14 Ma (2$\sigma$ analytical uncertainty) from the Wilberforce pyroxenite \cite{Paul2021a}, give tantalizing hints that temperatures that blocked magnetizations in portions of the orogen were achieved later than the ages assigned to paleomagnetic poles would suggest; given that U-Pb apatite closure temperatures are close to magnetite blocking temperatures.
+To test the hypothesis that the Grenville Loop is younger than in current interpretations, geochronology studies are needed to exploit the resolving power of high-precision whole grain analytical methods (i.e.
+ID-TIMS) and high-spatial-resolution in-situ methods (e.g.
+laser ablation depth profiling; \citeA{Chew2021a}) for reconstructing the cooling rate and time-temperature history of the Grenville Province.
+Testing this hypothesis is critical for reconstructions of the paleogeography of Rodinia in the early Neoproterozoic.
+
+\section*{Conclusion}
+
+The magnetization of hematite-bearing fine-grained siliciclastic sedimentary rocks from the Jacobsville Formation can be shown to be detrital and primary through positive intraformational conglomerate tests.
+We use these data to develop a new inclination-corrected paleomagnetic pole that can be constrained with recently published radiometric age constraints from \citeA{Hodgin2022a} to be ca.
+990 Ma.
+This high-quality pole is a crucial addition to Laurentia's apparent polar wander path in the earliest Neoproterozoic.
+Its position indicates that Laurentia's plate motion significantly slowed during Grenvillian collisional orogenesis.
+The new pole establishes a well-calibrated constraint for the position of the supercontinent Rodinia in early Neoproterozoic with Laurentia being at the center.
+The Jacobsville pole implies that ages associated with the paleomagnetic poles recorded by metamorphic rocks of the Grenville Province are likely younger than in current interpretations.
+Further paired studies of high-quality paleomagnetism and high-precision thermochronology are needed to illuminate Rodinia's motion and configuration in the early Neoproterozoic.
+
+\section*{Data Availability}
+Paleomagnetic data associated with this study are available within the MagIC database (\url{10.7288/V4/MAGIC/19780}) and all data are within a Github repository associated with this work (\url{https://github.com/Swanson-Hysell-Group/Jacobsville}) that is also archived on Zenodo (\url{https://doi.org/XXX}; \textit{will be archived in Zenodo upon acceptance}).
+This repository also contains Python code that implements all of the calculations, visualizations and statistical tests discussed herein.
+
+\acknowledgments
+Project research was funded by NSF CAREER grant EAR-1847277.
+Jim DeGraff provided helpful guidance in identifying exposures of the Jacobsville Formation.
+We gratefully acknowledge the Michigan Department of Natural Resources for sampling permitting.
+We thank Tim Lyons for additional private land access.
+We thank Madeline Swanson-Hysell for her assistance in the field.
+
+\nocite{Swanson-Hysell2014a, Swanson-Hysell2014b, Books1972a, Davis1997a, Davis1985a, Denyszyn2009a, Diehl1994a, Eyster2019a, Fairchild2017a, Halls1974a, Hnat2006a, Palmer1986a, Tauxe2009a, Tauxe1991a, Weil2003a}
+\bibliography{YZ_ref}
+
+\end{document}
+
+%%%EXTRA TEXT%%%
+
+% Evidence for the Grenville Loop being younger than previously estimated is emerging in recent advances of thermochronological studies in the Grenville Province.
+\citeA{Paul2021a} developed a new apatite U-Pb age of 923$\pm$14 Ma (2$\sigma$ analytical uncertainty) from the Wilberforce pyroxenite using isotope dilution thermal ionization mass spectrometry (ID-TIMS).
+In the Adirondack Highlands, \citeA{Krestianinov2021a} also developed U-Pb ID-TIMS ages of 907$\pm$14 Ma and 924$\pm$13 Ma (2$\sigma$ analytical uncertainties) from apatite associated with low-Ti iron ores that are likely cogenetic with Adirondack AMCG units \cite{Postel1952a, Krestianinov2021a}.
+These ages represent the timing of closure of Pb diffusion within apatite grains during exhumation \cite{Dodson1973a}.
+The temperature range that corresponds to this process for apatite is 420-540\textdegree C \cite{Cherniak1991a} which overlaps with magnetic blocking temperatures of these magnetite-bearing rocks, taking into account the effective lowering in blocking temperatures during slow cooling (400-580\textdegree C; \citeA{Palmer1973a, Brown2012a, Halgedahl1980a, Dodson1980a}).
+Therefore, the Grenville Loop pole ages are better assigned to be close to the dates given by apatite thermochronology.
+Given that K-rich minerals are known to be prone to obtain excess Ar \cite{McDougall1999a}, the discrepancy between the younger apatite U-Pb ages and generally older hornblende $^{40}$Ar/$^{39}$Ar ages indicate that potentially widespread excess Ar could have biased previous cooling history interpretations of the Grenville orogen \cite<e.g.>{Warnock2000a, Mezger1991a}.
+
+% Previously, the large-magnitude shifts in Laurentia's apparent polar wander path through the Keweenawan Track and Grenville Loop (Fig.
+\ref{fig:pole_plot}; \citeA{Irving1981a, Weil1998a, Swanson-Hysell2019a}) were hypothesized to represent rapid true polar wander which involves the rotation of the entire solid Earth about the spin axis \cite{Evans2003a, Mitchell2014b}.
+However, The slowdown in Laurentia's plate motion implied by the new Jacobsville paleomagnetic pole is consistent with the interpretation that the onset of the Grenvillian orogeny had dominant kinematic control on plate speeds as it involved continental-scale collisional orogenesis that likely led to the cessation and subsequent inversion of the Midcontinent Rift \cite{Cannon1994a}.
+
+% More high-precision thermochronology data and statistical models that properly incorporates uncertainties associated with timing of closure of thermochronometers and with magnetic remanence acquisition temperatures in the slowly cooled orogen are needed in order to further calibrate the age of those poles.
+Further development of paired pole-age data from the rich suite of metamorphic rocks in the Grenville Province can help update Laurentia's APWP \cite{Irving1974a} and the paleogeographic configuration of super continent Rodinia in the Neoproterozoic.
+
+% The interpretation of young Grenville poles is supported by high-precision U-Pb apatite thermochronology data from iron oxide-apatite deposits of the eastern Adirondack Mountains \cite{Krestianinov2021a} and from the Wilberforce pyroxenite of the Wilberforce region, Canada \cite{Paul2021a}.
+Apatite from both regions yield ages of ca.
+900 Ma with associated closure temperatures $\sim$450\textdegree C \cite{Cherniak1991a, Chamberlain2001a}.
+Considering delayed remanence acquisition of magnetite and hematite, ca.
+900 Ma should be closer to the actual timing when rocks of the Grenville Province finally blocked in their remanence that is preserved today.
+However, the exact age of each paleomagnetic pole recorded by rocks of different domains of the Grenville Province remains largely uncertain.
+Complications such as the origin of the dated iron oxide-apatite ore, crystal-size dependence of closure temperature of accessory minerals including rutile and apatite \cite{Smye2018a, Chamberlain2001a}, and the complex metamorphic histories in different domains of the Grenville Province require that more high-precision thermochronology data with modern closure temperature calibration are developed.
+
+% The previous older age estiamtes could be a result of unrecognized crypitc excess argon and incorrect calibration of Pb diffusion rate in accessory minerals in earlier studies such as the rutile closure temperature used by Mezger 1991 which has been revised significantly upwards by the experiemntal data of Cherniak.
+
+% Such large distance between pole positions associated with uncertainties in ages of Grenville magnetizations led to the controversial ``Grenville Problem"---did the Grenville rocks acquire magnetic remanence prior to the continental collision and thus the pole discrepancies are a result of significant crustal shortening (i.e.
 the ``two plate model"; \citeA<e.g.>{Irving1972a, Palmer1973a, Buchan1973a, Halls2015b}), or did they acquire remanence after the continental accretion and thus the southerly poles from the Grenville rocks represent a coherent plate motion through more recent times (i.e.
@@
 
-The two plate model is based on the interpretation that the Grenville rocks acquired magnetizations on allochthonous terranes in the late Mesoproterozoic, prior to the final accretion to Laurentia craton, resulting in their apparent southerly paleomagnetic pole positions today (Fig.
+% The two plate model is based on the interpretation that the Grenville rocks acquired magnetizations on allochthonous terranes in the late Mesoproterozoic, prior to the final accretion to Laurentia craton, resulting in their apparent southerly paleomagnetic pole positions today (Fig.
 \ref{fig:pole_plot}).
@@
 
-Given the new constraints on the age and position of the Jacobsville pole (Fig.
+% Given the new constraints on the age and position of the Jacobsville pole (Fig.
 \ref{fig:pole_plot}), the two plate model would require that the majority of the crustal shortening happened during or after the Rigolet phase of the Grenville Orogeny, as the ca.
@@
 
-Alternatively, the one plate model is much more straightforward in explaining the paleomagnetic pole positions of Grenville rocks as it interprets that those rocks acquired remanence during prolonged exhumation of the orogeny that postdate the peak metamorphism.
+% Alternatively, the one plate model is much more straightforward in explaining the paleomagnetic pole positions of Grenville rocks as it interprets that those rocks acquired remanence during prolonged exhumation of the orogeny that postdate the peak metamorphism.
 In contrast to the two plate model, the one plate model would predict that ages associated with the Grenville poles are much younger than ca.
@@
 However, what has not been fully taken into account in previous studies is the effective lowering of magnetic blocking temperatures of the new remanence.
-Theoretical calculations have shown that during prolonged cooling, magnetite which form above its Curie temperature (580\textdegree C) will subsequently acquire thermoviscous remanence \cite{Pullaiah1975a, Dodson1980a}, whereas hematite which often occur as exsolved lamellae in ilmenite in Grenville rocks acquired thermochemical remanence via magnetic ordering of the hemo-ilmenite below the Ne\'el temperature of hematite \cite{Burton1985a, Harrison2006b}.
+Theoretical calculations have shown that during prolonged cooling, magnetite which form above its Curie temperature (580\textdegree C) will subsequently acquire thermoviscous remanence \cite{Pullaiah1975a, Dodson1980a}, whereas hematite which often occur as exsolved lamellae in ilmenite in Grenville rocks acquired thermochemical remanence via magnetic ordering of the hemo-ilmenite below the N\'eel temperature of hematite \cite{Burton1985a, Harrison2006b}.
 
-Due to the long time scales involved with thermoviscous and thermochemical origin of the magnetic remanence, it is not straightforward to interpret the natural blocking temperatures of Grenville rocks based on observations of unblocking temperatures in the lab where rocks are heated and cooled rapidly.
+% Due to the long time scales involved with thermoviscous and thermochemical origin of the magnetic remanence, it is not straightforward to interpret the natural blocking temperatures of Grenville rocks based on observations of unblocking temperatures in the lab where rocks are heated and cooled rapidly.
 \citeA{Brown2012a} studied Grenville rocks from the Adirondack Mountains.
@@
 
-That the estimated ages associated with the Adirondack paleomagnetic poles are biased old can be attributed to the outdated thermochronometer closure temperature calculations of \citeA{Mezger1991a, Mezger1990a} whose data was used by \citeA{Brown2012a}.
+% That the estimated ages associated with the Adirondack paleomagnetic poles are biased old can be attributed to the outdated thermochronometer closure temperature calculations of \citeA{Mezger1991a, Mezger1990a} whose data was used by \citeA{Brown2012a}.
 In particular, the cooling history for the Adirondack area estimated by \citeA{Mezger1991a, Mezger1990a} show that the orogen had cooled to $\sim$400\textdegree C by the time of U-Pb system closure in rutile crystals.
@@
 Therefore, timing of magnetic remanence acquisition for Grenville rocks in the Adirondack Mountains should be younger than previously thought.
-The interpretation of young Grenville poles is supported by high-precision U-Pb apatite thermochronology data from iron oxide-apatite deposits of the eastern Adirondack Mountains \cite{Krestianinov2021a} and from the Wilberforce pyroxenite of the Wilberforce region, Canada \cite{Paul2021a}.
+
+% The interpretation of young Grenville poles is supported by high-precision U-Pb apatite thermochronology data from iron oxide-apatite deposits of the eastern Adirondack Mountains \cite{Krestianinov2021a} and from the Wilberforce pyroxenite of the Wilberforce region, Canada \cite{Paul2021a}.
 Apatite from both regions yield ages of ca.
@@
 
-Overall, albeit that numerous paleomagnetic data has been developed from rocks of the Grenville Province, few are paired with high-precision thermochronology data that are calibrated by modern diffusion paramters.
-More high-precision thermochronology data and statistical models that properly incorporates uncertainties associated with timing of closure of thermochronometers and with magnetic remanence acquisition temperatures in the slowly cooled orogen are needed in order to further calibrate the age of those poles.
-Further development of paired pole-age data from the rich suite of metamorphic rocks in the Grenville Province can help update Laurentia's APWP \cite{Irving1974a} and the paleogeographic configuration of super continent Rodinia in the Neoproterozoic.
+% However, inclination bias associated with sedimentary paleomagnetic directions need to be corrected when synthesizing poles.
+Rotation of grains during sediment deposition and compaction during early diagenesis can result in the acquisition of a detrital remanent magnetization that is shallower than the local geomagnetic field in which it was acquired \cite{King1955a, Tauxe2005a, Kodama2012a}.
+This effect can be summarized as:
+% \begin{equation*}
+% \textup{tan}(I_o) = f\textup{tan}(I_f)
+% \end{equation*}
+% where $f$ represents the amount of inclination shallowing, $I_o$ represents the observed inclination, and $I_f$ represents the inclination of the field in which the magnetization was acquired.
+Methods for correcting sedimentary inclination bias include laboratory-based measurements of magnetic mineral fabrics \cite<e.g.>{Kodama1990a, Bilardello2015a}, as well as statistical methods based on the evaluation of the deviation of an observed virtual geomagnetic pole (VGP) distribution from that of a paleosecular variation model (the ``elongation/inclination" [E/I] method of \citeA{Tauxe2004b}).
+When a large number of sedimentary paleomagnetic sites (typically n$>$100) capture an extensive period of time such that the paleosecular variation is sufficiently averaged, the E/I method can give efficient estimates with reasonable uncertainty bounds via bootstrapping and is less labor-intensive than methods based on mineral fabric measurements.
 
-\section*{Conclusion}
+% Since the deposition of the Jacobsville Formation span over geomagnetic field reversals, it is of interest to investigate the amount of inclination shallowing associated with the normal (n=130) and the reversed (n=148) polarity respectively (transitional fields and specimens affected by complex tilting regime are excluded).
+Applying the E/I method on the normal polarity directions yields an estimated $f$ value of 0.67 with 95\% bootstrap uncertainty bounds from 0.91 to 0.55 (Fig.
+\ref{fig:in_situ_pmag}B).
+Directions of reversed polarity yields an estimated $f$ factor of 0.52 with uncertainty bounds between 0.97 and 0.29 (Fig.
+\ref{fig:in_situ_pmag}B).
+Although the estimated $f$ value for the reversed-polarity directions has large statistical uncertainty bounds, the uncertainties are likely associated with the overall shallow inclinations and distributions of directions in both upper and lower hemispheres (Fig.
+\ref{fig:in_situ_pmag}B).
+Therefore, inclination corrections do not result in significant changes in the mean reversed direction.
 
-Our new paleomagnetic pole from the ca.
-990 Ma Jacobsville Formation passes paleomagnetic field tests and has robust radiometric age constraints.
-This high-quality pole is a crucial addition to Laurentia's apparent polar wander path in the late Proterozoic.
-It extends the Mesoproterozoic Keweenawan Track (Fig.
-\ref{fig:pole_plot}) by showing that the plate slowed down significantly as the Grenville collisional orogenesis commenced on Laurentia's margin.
-It also sets an anchor for the position of supercontinent Rodinia in early Neoproterozoic due to Laurentia's central position in the configuration.
-Being a robust record from rocks in interior Laurentia, the Jacobsville pole implies that ages associated with the paleomagnetic poles recorded by metamorphic rocks of the Grenville Province might need to be adjusted significantly.
-Further paired studies of high-quality paleomagnetism and high-precision thermochronolgy are needed to illuminate Laurentia's development toward mid Neoproterozoic and its connection with conjugate continents such as Baltica.
+% In addition, directions of the two polarities do not pass a reversal test \cite{McFadden1990a, Tauxe2010a}, regardless of whether inclination correction is applied.
+One interpretation of the asymmetric polarities could be that the deposition of the Jacobsville Formation was protracted such that the sediments at Hungarian Falls which dominate the reversed polarity records were deposited prior to those at Snake Creek tributary which dominate the normal polarity records (Fig.
+\ref{fig:in_situ_pmag}).
+This is consistent with our stratigraphic interpretation that Snake Creek tributary sediments could be younger than those near central Keweenaw Peninsula (i.e.
+Hungarian Falls, Natural Wall, Hammel Creek; Fig.
+\ref{fig:strat_column}).
+Other alternative interpretations could include that the mean poles are biased by transient non-dipolar fields due to insufficient averaging of paleosecular variation or transitional fields during field reversals.
+More data are needed to test these hypotheses.
 
-\section*{Data Availability}
-Paleomagnetic data associated with this study are available within the MagIC database (\url{XXX}; UPDATE WHEN DOI IS GENERATED) and all data are within a Github repository associated with this work (\url{https://github.com/Swanson-Hysell-Group/Jacobsville}) that is also archived on Zenodo (\url{https://doi.org/XXX}).
-This repository also contains Python code that implements all of the calculations, visualizations and statistical tests discussed herein.
+% Overall, both inclination-corrected polarities recorded by the Jacobsville Formation show mean pole positions in the southern hemisphere with respect to today's continental configuration (normal polarity longitude=175.1\textdegree E, latitude=19.1\textdegree S, A95=3.8\textdegree; reversed polarity longitude=191.1\textdegree E, latitude=14.5\textdegree S, A95=3.9\textdegree; Fig.
+\ref{fig:EI_results}).
+The two pole positions are distinct from the youngest poles developed from rocks of the Midcontinent Rift (Fig.
+\ref{fig:EI_results}).
+Despite that the Jacobsville sedimentation span over geomagnetic field reversals, the large arc distance between either polarity of the Jacobsville Formation and the pole of the ca.
+1070 Ma upper Freda Formation \cite{Henry1977a} is consistent with the interpretation that the onset of deposition of the Jacobsville Formation at the five studied localities post-date the youngest member of the Oronto Group sediments by tens of millions of years.
 
-\acknowledgments
-This project is funded by NSF grant EAR-1847277 to Nicholas Swanson-Hysell.
-
-\bibliography{YZ_ref}
-
-\end{document}
+% Near the base of the exposed stratigraphic section at Natural Wall (Fig.
+\ref{fig:strat_column}), the strata are steeply tilted or overturned (Fig.
+\ref{fig:Field_photo}).
+Tilt-corrected paleomagnetic directions from samples collected from these strata show dominantly southeasterly declinations that are distinct from either normal- or reversed-polarity directions higher along the strata or at Hungarian Falls.
+It is most likely that these directions resulted from noncylindrical or multiple episodes of complex folding.
+We hypothesize that these DRM directions were acquired during a reverse polarity episode similar to that recorded near the base of Hammel Creek (Fig.
+\ref{fig:strat_column}), but we will not include these directions in further paleomagnetic pole analyses.

### 2023-03-29 Manuscript/support_info.tex
update manuscript tex files

\documentclass[11pt,letterpaper]{article}
\usepackage{xcolor}
\usepackage{textcomp,marvosym}
\usepackage{amsmath,amssymb}
\usepackage[left]{lineno}
\usepackage{changepage}
\usepackage{rotating}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{graphicx}
\usepackage{sidecap}
\usepackage{pdfpages}
\usepackage{longtable}
\usepackage{url}
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
%\doublespacing
%\usepackage[]{tocbibind}

\newcommand{\GG}[1]{}

\raggedright
\textwidth = 6.5 in
\textheight = 8.5 in
\oddsidemargin = 0.0 in
\evensidemargin = 0.0 in
\topmargin = -0.5 in
\headheight = 0.0 in
\headsep = 0.5 in
\parskip = 0.0 in
\parindent = 0.2 in

\begin{document}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}
\subsection*{Supporting Information for ``Tracking Rodinia in the Neoproterozoic: new paleomagnetic constraint from the Jacobsville Formation"}
Yiming Zhang\textsuperscript{1}, Eben B. Hodgin\textsuperscript{1}, Tadesse Alemu\textsuperscript{1}, James Pierce\textsuperscript{1,2}, Anthony Fuentes\textsuperscript{1}, Nicholas L. Swanson-Hysell\textsuperscript{1}

{\small $^{1}$Department of Earth and Planetary Science, University of California, Berkeley, CA, USA

% $^{2}$Department of Earth, Environmental and Planetary Sciences, Brown University, RI, USA

$^{2}$Department of Earth and Planetary Sciences, Yale University, New Haven, CT, USA}

\listoffigures
\listoftables
\tableofcontents

\begin{figure*}[h!]
\centering
\includegraphics[width=0.5\textwidth]{Jacobsville_pdf_directions.pdf}
\caption{Jacobsville present day local field directions. This present day overprint component is typically removed by $\sim$300\textdegree C.}
\label{fig:Jacobsville_pdf}
\end{figure*}

\begin{figure*}[h!]
\centering
\includegraphics[width=0.9\textwidth]{Manuscript/SI_hct_fold_test.pdf}
\caption{Bootstrap paleomagnetic fold test \citep{Tauxe1994a} of the chemical remanent magnetization directions recorded by specimens from the nearly horizontal beds and moderately tilted beds at the Snake Creek tributary. Complete unfolding does not lie within the 95\% confidence limits of the test, consistent with the magnetization having been acquired syn- to post-folding.}
\label{fig:Jacobsville_hct_fold_test}
\end{figure*}

\begin{figure*}[h!]
\centering
\includegraphics[width=0.9\textwidth]{SI_reversal_test.pdf}
\caption{Jacobsville detrital remanence directions showing dual polarities plotted on an equal area stereonet plot. The directions do not pass a reversal test of \cite{McFadden1990a} as the angle between the mean directions is 15.9\textdegree\, larger than the critical angle of 7.5\textdegree. The directions also do not pass the bootstrap reversal test of \cite{Tauxe1991a}. }
\label{fig:Jacobsville_reversal_test}
\end{figure*}

\begin{figure*}[h!]
\centering
\includegraphics[width=0.9\textwidth]{SI_hct.pdf}
\caption{Jacobsville chemical remanence magnetization directions plotted by section. Samples collected from the Natural Wall section do not have interpretable chemical remanence component that can be fit with least-squares lines. }
\label{fig:Jacobsville_CRM}
\end{figure*}

\begin{figure*}[h!]
\centering
\includegraphics[width=0.9\textwidth]{Manuscript/SI_Jacobsville_Nonesuch_rate.pdf}
\caption{Monte Carlo simulations of apparent polar wander rates implied by paleomagnetic and geochronologic data between the Nonesuch Formation and the Jacobsville Formation. (A) The gold and red points are 10,000 simulated Kent distribution mean paleomagnetic poles for the Nonesuch Formation and Jacobsville Formation, respectively. The simulated poles are connected by gray lines which represent the apparent polar wander great circle paths. (B) The points color-coded in the same way as in (A) show the simulated paleomagnetic pole latitudes plotted against their simulated pole ages using uniform distributions. The points are are connected by gray lines which represent the simulated pole latitudinal motion. The histogram on the right shows all 10,000 of the simulated rates that yield the labeled 2.5 percentile value of 2.5 cm/yr, median value of 3.2 cm/yr and 97.5 percentile value of 4.3 cm/yr.}
\label{fig:Jacobsville_Nonesuch_rate}
\end{figure*}

\begin{sidewaystable}[h!]
\begin{scriptsize}
\caption[Compilation of Laurentian paleomagnetic poles.]{\scriptsize Compilation of up-to-date Keweenawan Track paleomagnetic poles and ca. 780-720 Ma Laurentian paleomagnetic poles. }
\begin{tabular}{p{4cm}lllp{4cm}lllp{4cm}}

\hline
\multicolumn{9}{c}{Compilation of Fisher mean paleomagnetic poles}                                                                                                                                                                                        \\ \hline
Pole                                                       & Pole lat & Pole lon & A95  & Pole reference                                                         & AgeNominal & AgeLower & AgeUpper & Age reference                                       \\
Osler reverse (lower)                                      & 40.9     & 218.6    & 4.8  & \cite{Swanson-Hysell2014a}                                           & 1108       & 1105.15  & 1110     & \cite{Davis1985a, Swanson-Hysell2019a}               \\
Osler reverse (upper)                                      & 42.3     & 203.4    & 3.7  & \cite{Halls1974a, Davis1985a, Swanson-Hysell2014a} & 1105.15    & 1104.82  & 1105.48  & \cite{Swanson-Hysell2019a}                       \\
Mamainse lower reversed 1                                  & 49.5     & 227      & 5.3  & \cite{Swanson-Hysell2014b}                                           & 1109       & 1106     & 1112     & As discussed in \cite{Swanson-Hysell2019a}         \\
Mamainse lower reversed 2                                  & 37.5     & 205.2    & 4.5  & \cite{Swanson-Hysell2014b}                                                  & 1105       & 1100.4   & 1109     & \cite{Swanson-Hysell2014b}                              \\
Mamainse lower normal and upper reversed                   & 36.1     & 189.7    & 4.9  & \cite{Swanson-Hysell2014b}                                                  & 1100.36    & 1100.1   & 1100.61  & \cite{Swanson-Hysell2014b}                               \\
Mamainse upper normal                                      & 31.2     & 183.2    & 2.5  & \cite{Swanson-Hysell2014b}                                                  & 1094       & 1090     & 1100     & As discussed in S\cite{Swanson-Hysell2019a}            \\
Grand Portage Basalts                                      & 46       & 201.7    & 6.8  & \cite{Books1968a, Tauxe2009a}                                    & 1106       & 1105.28  & 1108     & \cite{Swanson-Hysell2019a}                         \\
North Shore Volcanic Group (upper NE   sequence)           & 31.1     & 181.7    & 4.2  & \cite{Books1972a, Tauxe2009a}                                    & 1095       & 1092     & 1098     & \cite{Davis1997a, Fairchild2017a}       \\
North Shore Volcanic Group (upper SW   sequence)           & 36.9     & 179.3    & 2.1  & \cite{Tauxe2009a, Swanson-Hysell2019a}                   & 1096.18    & 1093.94  & 1096.75  & \cite{Swanson-Hysell2019a}                        \\
Schroeder Lutsen Basalts                                   & 28.3     & 187.6    & 2.5  & \cite{Books1972a, Tauxe2009a, Fairchild2017a}            & 1090       & 1085     & 1091.5   & \cite{Fairchild2017a}                               \\
Portage Lake Volcanics                                     & 27.5     & 182.5    & 2.3  & \cite{Books1972a, Hnat2006a}                                         & 1092.51    & 1091.59  & 1093.37  & \cite{Swanson-Hysell2019a}                    \\
Lake Shore Traps                                           & 22.2     & 180.8    & 4.5  & \cite{Diehl1994a}                                                     & 1085.47    & 1084     & 1091     & \cite{Fairchild2017a, Swanson-Hysell2019a}    \\
Siemens Creek Volcanics                                    & 45.8     & 214      & 9.2  & \cite{Palmer1986a}                                                 & 1108       & 1105     & 1111     & \cite{Davis1997a}                               \\
Michipicoten Island Formation                              & 17       & 174.7    & 4.4  & \cite{Palmer1987a, Fairchild2017a}                        & 1083.95    & 1083.52  & 1084.39  & \cite{Fairchild2017a}                           \\
Freda Formation                                            & 2.2      & 179      & 4.2  & \cite{Henry1977a}                                                     & 1070       & 1060     & 1083.5   & As discussed in \cite{Swanson-Hysell2019a}            \\
Haliburton Intrusions                                      & -32.6    & 141.9    & 6.3  & \cite{Warnock2000a}                          & 1015       & 1000     & 1030     & \cite{Warnock2000a}                                          \\
Adirondack Highlands gneiss                                & -19      & 148.7    & 11.2 & \cite{Brown2012a}                                                &            &          &          &                                                     \\
Adirondack Highlands anorthositic rocks                    & -25.2    & 143.4    & 12.9 & \cite{Brown2012a}                                                &            &          &          &                                                     \\
Adirondack Highlands granites                              & -28.5    & 131.7    & 7.1  & \cite{Brown2012a}                                                &            &          &          &                                                     \\
Franklin large igneous province   Victoria/Mainland/Baffin & 6.7      & 162.1    & 3    & \cite{Denyszyn2009a}                                                   & 716.33     & 715.79   & 716.87   &                                                     \\
Combo Carbon Butte–Awatubi                                 & 14.2     & 163.8    & 3.5  & \cite{Eyster2019a}                                                   & 751        & 743.4    & 758.6    &                                                     \\
Carbon Canyon                                              & -0.5     & 166      & 9.7  & compilation by \cite{Eyster2019a}                                    & 757        & 750.2    & 763.8    &                                                     \\
UMG Group 3                                                & 4.9      & 160.6    & 3.2  & compilation by \cite{Eyster2019a}                                    & 755        &          &          & As discussed in \cite{Eyster2019a}                \\
Nankoweap                                                  & -10      & 163      & 4.9  & \cite{Weil2003a}                                                     & 782        &          &          & As discussed in Eyster et al., 2020                 \\
UMG Group 2                                                & -5.8     & 158.7    & 2.7  & compilation by \cite{Eyster2019a}                                    & 760        &          &          & As discussed in Eyster et al., 2021                 \\
UMG Group 1                                                & 3        & 163.5    & 3.2  & compilation by \cite{Eyster2019a}                                    & 766        &          &          & As discussed in Eyster et al., 2022                 \\
Gunbarrel mean                                             & 9.1      & 138.2    & 11.7 & \cite{Eyster2019a}                                                   & 774.93     & 774.39   & 775.47   &                                                    
\end{tabular}
\end{scriptsize}
\label{tab:pole_compilation}
\end{sidewaystable}

\clearpage

\begin{sidewaystable}
\begin{scriptsize}
\begin{tabular}{p{1.5cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{2cm}}
\hline
\multicolumn{14}{c}{Compilation of Kent mean paleomagnetic poles}                                                                                                                                                                                           \\ \hline 
Pole                  & Pole lat & Pole lon & Major axis lat & Major axis lon & Major axis magnitude & Minor axis lat & Minor axis lon & Minor axis magnitude & Pole reference         & Age Nominal & Age Lower & Age Upper & Age reference                                       \\
Nonesuch Formation    & 6.6      & 182.9    & 27.6           & 280.2          & 2.8                  & 41.6           & 87             & 2                    & \cite{Slotznick2023a} & 1080       & 1070     & 1083.5   & As discussed in \cite{Swanson-Hysell2019a}         \\
Jacobsville Formation & 16.9    & 183.4    & 45.2           & 255.5          & 4.1                    & 39.9           & 108.1          & 3.1                  & this study             & 990        & 985      & 992     & \cite{Hodgin2022a}; as discussed in the manuscript
\end{tabular}
\end{scriptsize}
\end{sidewaystable}

\clearpage

\bibliographystyle{gsabull}
\bibliography{YZ_ref}

\end{document}

### 2024-01-23 Manuscript/Manuscript.tex
update manuscript for proof

@@
 \item Siltstones and fine-grained sandstones of the ca. 990 Ma Jacobsville Formation have primary magnetization held by detrital hematite.
-\item Following closure of the Unimos Ocean and onset of the Grenvillian orogeny in the late Mesoproterozoic, Laurentia's motion significantly slowed.
-\item The pole position and age suggests that the Grenville Loop paleomagnetic poles are younger than current interpretations.
+\item Following closure of the Unimos Ocean and onset of the Grenvillian orogeny in the late Mesoproterozoic, Laurentia's motion went from fast to slow.
+\item The pole position and age suggests that the paleomagnetic poles of the Grenville Loop are younger than in previous interpretations.
 \end{keypoints}
@@
 High-resolution thermal demagnetization experiments resolve detrital remanent magnetizations held by hematite.
-These directions were reoriented within siltstone rip-up clasts and pass an intraformational conglomerate test---giving confidence that the magnetization is detrital and primary.
+These directions were reoriented within siltstone intraclasts and pass intraformational conglomerate tests---giving confidence that the magnetization is detrital and primary.
 An inclination-corrected mean paleomagnetic pole position for the Jacobsville Formation indicates that Laurentia's motion slowed down significantly following the onset of the Grenvillian orogeny.
-Prior rapid plate motion associated with closure of the Unimos Ocean transitioned to slow drift of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic.
+Prior rapid plate motion associated with closure of the Unimos Ocean between 1110 and 1090 Ma transitioned to slow drift of Laurentia across the equator in the late Mesoproterozoic to early Neoproterozoic.
 We interpret the distinct position of this well-dated pole from those in the Grenville orogen that have been assigned a similar age to indicate that the ages of the poles associated with the Grenville Loop likely need to be revised to be younger due to prolonged exhumation.
@@
 775 Ma Gunbarrel large igneous province \cite{Harlan2003a, Mackinder2019a, Swanson-Hysell2021c}.
-The lack of Laurentian rocks with primary TRM in the early Neoproterozoic has led paleogeography reconstructions to be reliant on magnetizations acquired by rocks that were metamorphosed during the ca.
-1090 to 980 Ma Grenvillian orogeny \cite{Rivers2012a, Swanson-Hysell2023a}.
-However, rocks within the orogen acquired their magnetizations during exhumation long after peak metamorphic conditions \cite{McWilliams1975a}.
+The lack of Laurentian rocks with primary TRM in the early Neoproterozoic has led paleogeographic reconstructions to be reliant on magnetizations acquired by rocks that were metamorphosed during the ca.
+1090 to 980 Ma Grenvillian orogeny \cite{Rivers2008a, Rivers2012a, Swanson-Hysell2023a}.
+However, rocks within the orogen acquired their magnetizations during exhumation after peak metamorphic conditions \cite{McWilliams1975a}.
 As a result, the timing of remanence acquisition needs to be constrained through the more difficult task of dating cooling associated with exhumation.
@@
 The inset green box shows the extent of panel B.
-(B) Geologic map of northern Michigan and Wisconsin showing Midcontinent Rift igneous and sedimentary rocks, the Jacobsville Formation, sedimentary rocks of the Bayfield Group, other older Proterozoic rocks, and major post-rift thrust faults.
+(B) Geologic map of northern Michigan and Wisconsin showing Midcontinent Rift igneous and sedimentary rocks, sedimentary rocks of the Jacobsville Formation and Bayfield Group, other older Paleoproterozoic and Archean rocks, and major post-rift thrust faults.
 The location of the Jacobsville stratigraphic sections in this study are shown with green diamonds.
@@
 \ref{fig:Geologic_map}; \citeA{Hamblin1958a, Hodgin2022a, DeGraff2022a}).
-A 992.51$\pm$0.64 Ma (2$\sigma$ analytical uncertainty) U-Pb detrital zircon date from the Jacobsville Formation constrains its maximum deposition age \cite{Hodgin2022a}.
+U-Pb detrital zircon dates developed through chemical abrasion–isotope dilution–thermal ionization mass spectrometry (CA-ID-TIMS) of 1003.21 $\pm$ 2.23 Ma and 992.51$\pm$0.64 Ma (2$\sigma$ analytical uncertainty) constrain the maximum deposition age of the Jacobsville Formation \cite{Hodgin2022a}.
 A 985.5$\pm$35.8 Ma (2$\sigma$ analytical uncertainty) U-Pb date from calcite veins within the Keweenaw fault provides a minimum age constraint on Jacobsville deposition, as the Jacobsville Formation is folded in the footwall against Midcontinent rift volcanics in the hanging wall.
@@
 990 Ma \cite{Hodgin2022a}.
-These results constrain the formation to have been deposited within a syn-orogenic basin during the Rigolet phase of the Grenvillian orogen and to have been deformed during the same orogenic phase \cite{Hodgin2022a, Swanson-Hysell2023a}.
+These results support the interpretation that the formation was deposited within a syn-orogenic basin during the Rigolet phase of the Grenvillian orogen and was then deformed during the same orogenic phase.
 
@@
 1020 Ma based on APWP extrapolation \cite<e.g.>{Li2008a}.
-However, the uncertainty in its age combined with a lack of field tests, and no application of inclination correction have resulted in the pole not being included in recent curated compilations \cite<e.g.>{Evans2021a}.
+However, the uncertainty in the age of the magnetization combined with a lack of field tests, and no application of inclination correction have resulted in the pole not being included in recent curated compilations \cite<e.g.>{Evans2021a}.
 Obtaining high-quality paleomagnetic data, including robust field tests, motivates the development of new data from the Jacobsville Formation.
@@
 \ref{fig:Geologic_map} and \ref{fig:strat_column}).
-High-resolution thermal demagnetization isolates primary detrital remanent magnetization (DRM) from secondary chemical remanent magnetization (CRM) within analyzed samples.
-This interpretation is supported by the DRM directions passing an intraformational conglomerate test on siltstone intraclasts that were ripped-up and remobilized within the fluvial depositional environment.
+These sections were chosen given the abundance of fine-grained siliciclastic lithologies where magnetization can be constrained through intraclast conglomerate field tests.
+Additionally, they are within the continuous bedrock belt from which chronostratigraphic constraints have been developed.
+In contrast to fluvial sandstones in other exposures such as in the eastern Lake Superior region near Sault Ste.
+Marie, the maximum and minimum depositional ages of these sections can be more confidently constrained.
+
+Within sampled lithologies, high-resolution thermal demagnetization isolates primary detrital remanent magnetization (DRM) from secondary chemical remanent magnetization (CRM).
+This interpretation is supported by the DRM directions passing intraformational conglomerate tests on siltstone intraclasts that were ripped-up and remobilized within the fluvial depositional environment.
 A paleomagnetic fold test further constrains the timing of DRM acquisition to predate the folding of the Jacobsville Formation in the footwall of the Keweenaw Fault soon after deposition.
@@
 Following the end of active extension in the rift, thermal subsidence resulted in the accumulation of $>$4 km of ca.
-1085-1050 Ma Oronto Group sedimentary rocks that are both intercalated with the youngest volcanics and deposited atop them (Fig.
+1085-1050 Ma Oronto Group strata that are both intercalated with the youngest volcanics and deposited atop them (Fig.
 \ref{fig:Geologic_map}; \citeA{Daniels1982a, Cannon1989a}).
@@
 \ref{fig:Geologic_map}; \citeA{Cannon1993a}).
-Rb-Sr radiometric thermochronology data of \citeA{Cannon1993a} show Archean lithologies within the Montreal River monocline to have been exhumed to mid to shallow crustal temperatures ($\sim$270\textdegree C) by ca.
+Rb-Sr biotite thermochronology data of \citeA{Cannon1993a} show Archean lithologies within the Montreal River monocline to have been exhumed to mid to shallow crustal temperatures ($\sim$270\textdegree C) by ca.
 1050 Ma.
-The Jacobsville Formation overlies an erosional angular unconformity that developed on lithologies that were exhumed through this earlier episode of contractional deformation associated with Grenvillian orogenesis (Fig.
+The Jacobsville Formation overlies an angular unconformity that developed on lithologies that were exhumed through this earlier episode of contractional deformation associated with Grenvillian orogenesis (Fig.
 \ref{fig:Geologic_map}; \citeA{Hamblin1958a, Cannon1993a, Kalliokoski1982a}).
-Conglomeratic facies can occur in the basal Jacobsville Formation, with clasts derived from locally uplifted basement lithologies \cite{Irving1885a, Hamblin1958a, Kalliokoski1982a, Alemu2023a}.
+Conglomeratic facies occur in the basal Jacobsville Formation, with clasts derived from locally uplifted basement lithologies \cite{Irving1885a, Hamblin1958a, Kalliokoski1982a}.
 
 Bedding planes in the Jacobsville Formation typically have shallow dips.
-The exception to these near-horizontal orientations is proximal to the Keweenaw fault (Fig.
+The exception to these near-horizontal orientations is proximal to reverse faults such as the Keweenaw fault (Fig.
 \ref{fig:Geologic_map}) where the Jacobsville Formation was occasionally deformed into monoclinal drag folds resulting in steeply tilted to overturned beds (Fig.
@@
 Recent mapping of the Jacobsville Formation also found that it sometimes onlaps the Midcontinent Rift volcanics within or adjacent to en echelon fault branches of the Keweenaw fault system near the northern end of the Keweenaw Peninsula \cite{Tyrrell2019a, Mueller2021a}.
-Clumped isotope data for vein calcite with a U-Pb date of 985.5$\pm$35.8 Ma in the Keweenaw fault breccia within the hanging wall give temperatures of $\sim$50\textdegree C  \cite{Hodgin2022a}.
-The low temperature indicates that the calcite precipitated at shallow depth and that the majority of contractional inversion along the fault was complete by ca.
-985 Ma.
-Combined with the 992.51$\pm$0.64 Ma maximum depositional age, these data are interpreted to indicate that the Jacobsville is a syn-orogenic unit deposited during the ca.
-1010-980 Ma Rigolet phase of the Grenvillian orogeny \cite{Hodgin2022a, Swanson-Hysell2023a}.
-
-In addition, field observations reveal erosional contacts and development of paleosols between the Jacobsville Formation and underlying volcanic rocks \cite{Hamblin1958a, Kalliokoski1975a}, and seismic reflection data beneath Lake Superior have been interpreted to indicate that the Jacobsville Formation lies in angular unconformity atop sedimentary rocks of the Oronto Group \cite{Cannon1989a}.
-These data indicate that there was an interval of erosion following ca.
-1050 Ma prior to the onset of Jacobsville Formation deposition.
+These observations are consistent with the interpretation that there was deposition of Jacobsville Formation sediments during faulting.
 
@@
 
-The Jacobsville Formation is a fluvial sequence of feldspathic and quartzose conglomerates, sandstones, siltstones, and shales devoid of lava flows or cross-cutting igneous dikes although internal clastic dikes are present \cite{Hamblin1958a}.
-Sandstones of the Jacobsville Formation are more quartz-rich with fewer lithic fragments than sandstones of the Oronto Group such that they can be classified as subarkose to quartz arenites (Fig.
+The Jacobsville Formation is a fluvial succession of feldspathic and quartzose conglomerates, sandstones, siltstones, and shales devoid of lava flows or cross-cutting igneous dikes although internal clastic dikes are present \cite{Hamblin1958a}.
+Sandstones of the Jacobsville Formation are more quartz-rich with generally fewer lithic fragments than sandstones of the Oronto Group such that they can typically be classified as subarkose to quartz arenites (Fig.
 \ref{fig:Field_photo}; \citeA{Hamblin1958a, Ojakangas2002a}).
@@
 \centering
-\includegraphics[width=0.8\textwidth]{Jacobsville_Sections_v6.pdf}
+\includegraphics[width=0.75\textwidth]{Jacobsville_Sections_v6.pdf}
 \caption{Lithostratigraphy and magnetostratigraphy for studied sections of the Jacobsville Formation in northern Michigan, USA.
+GPS locations for the base and top of the sections are noted.
 Paleomagnetic sampling focused on the dark red siltstone to fine-grained sandstone lithofacies.
@@
 No U-Pb zircon dates from these samples were younger than the age of Midcontinent Rift volcanics.
-A correlation of a facies association consisting of volcanic-clast conglomerate, coarse arkosic sandstone, and dark red clay-rich siltstone is inferred between the sections at Hungarian Falls (along Dover Creek and a tributary falls), Hammel Creek, and Natural Wall \cite{Brojanigo1984a}.
+A correlation of a facies association consisting of volcanic-clast conglomerate, coarse arkosic sandstone, and dark red clay-rich siltstone is inferred between the sections at Dover Creek (along Dover Creek and a tributary falls), Hammel Creek, and Natural Wall \cite{Brojanigo1984a}.
 The stratigraphic correlation between Jacobsville sections at Agate Falls and Snake Creek tributary with these correlated sections from the central Keweenaw Peninsula is relatively uncertain.}
@@
 In southern exposures of the Jacobsville Formation, conglomerate facies typically have provenance sourced from Paleoproterozoic lithologies such as vein quartz and iron formation clasts \cite{Hamblin1958a, Kalliokoski1982a}.
-The occurrence of these conglomerates in proximity to the Marenisco fault and Pelton Creek fault is interpreted to represent syn-depositional development of local relief (Fig.
-\ref{fig:Geologic_map}; \citeA{Kalliokoski1982a, Hedgman2014a}).
+The occurrence of these conglomerates in proximity to the Marenisco fault and Pelton Creek fault and the downsection steepening of stratal dips interpreted as growth strata may represent syn-depositional development of local relief (Fig.
+\ref{fig:Geologic_map}; \citeA{Kalliokoski1982a, Hedgman1992a}).
 Close to the Keweenaw fault in the central Keweenaw Peninsula (Fig.
-\ref{fig:Geologic_map}), the Jacobsville Formation consists dominantly of quartz-rich, trough cross-bedded, medium-grained sandstone, with locally abundant clast-supported pebble to cobble conglomerate, and interbeds of red, hematite-bearing, micaceous siltstone to fine-grained sandstone (Hungarian Falls, Hammel Creek, Natural Wall; Figs.
+\ref{fig:Geologic_map}), the Jacobsville Formation consists dominantly of quartz-rich, trough cross-bedded, medium-grained sandstone, with locally abundant clast-supported pebble to cobble conglomerate, and interbeds of red, hematite-bearing, micaceous siltstone to fine-grained sandstone (Dover Creek, Hammel Creek, Natural Wall; Figs.
 \ref{fig:strat_column} and \ref{fig:Field_photo}).
 In this region, the conglomerate contains abundant rounded to angular volcanic clasts as large as boulders that are likely derived from uplifted Midcontinent Rift volcanics \cite{Irving1885a, Brojanigo1984a}.
-Thicker intervals of dark red to brick red siltstone and very fine-grained sandstone are interbeded with coarser-grained sandstone and conglomerate at Agate Falls and Hungarian Falls (Figs.
+Thicker intervals of dark red to brick red siltstone and very fine-grained sandstone are interbedded with coarser-grained sandstone and conglomerate at Agate Falls and Dover Creek (Figs.
 \ref{fig:strat_column} and \ref{fig:Field_photo}).
-Rip-up clasts of the siltstones are found within some channelized interbeds of sandstone and conglomerate (Fig.
+Intraclasts of the siltstones are found within some channelized interbeds of sandstone and conglomerate (Fig.
 \ref{fig:intraclast_pmag}).
@@
 \caption{\scriptsize Field photos of the Jacobsville Formation.
-(A) The red/white medium-grained sandstone with decimeter-scale trough cross-stratification in this image is a wide-spread lithology in the formation, but was not targeted for paleomagnetic sampling in this study.
+(A) The red/white medium-grained sandstone with decimeter-scale trough cross-stratification in this image (taken at the shore near the unincorporated community of Jacobsville 46.9819\textdegree N, 88.4068\textdegree W) is a wide-spread lithofacies in the formation, but was not targeted for paleomagnetic sampling in this study.
 (B, C, D) At the Snake Creek tributary and the Natural Wall ravine in the northeastern Keweenaw Peninsula, intervals of red fine-grained sandstone to siltstone beds can be found through the steeply dipping to overturned basal strata near the Keweenaw fault and also the nearly horizontal upper strata.
-At the Hungarian Falls section (E), relatively thick red siltstone horizons interbedded or interfingering with sandstones and conglomerates are found along a waterfall associated with a stream that is a tributary to Dover Creek.
-Abundant reoriented siltstone rip-up clasts can be present within the conglomeratic layers above siltstone beds (Fig.
+At the Dover Creek section (E), thick red siltstone horizons interbedded or interfingering with sandstones and conglomerates are found along a Dover Creek tributary waterfall (47.1708\textdegree N, 88.4466\textdegree W; this waterfall is distinct, but close to Hungarian Falls).
+Abundant reoriented siltstone intraclasts can be present within the conglomeratic layers above siltstone beds (Fig.
 \ref{fig:intraclast_pmag}).
 (F) At Agate Falls, nearly flat-lying red fissile siltstone beds are exposed.
-Subrounded detrital mica grains deposited parallel to the bedding plane are often present within the siltstone facies.
-Another set of siltstone rip-up clasts were sampled from a conglomeratic layer at Agate Falls for a paleomagnetic conglomerate test (Fig.
+Detrital mica grains deposited parallel to the bedding plane are often present within the siltstone facies.
+Another set of siltstone intraclasts were sampled from a conglomeratic layer at Agate Falls for a paleomagnetic conglomerate test (Fig.
 \ref{fig:intraclast_pmag}).}
@@
 
-The abundance of rift volcanic-clasts in the conglomeratic facies at Hungarian Falls, Hammel Creek, and Natural Wall indicates the presence of uplifted volcanics in proximity to the Keweenaw fault in the central part of the peninsula during Jacobsville deposition (Figs.
+The abundance of clasts of rift volcanics in the conglomeratic facies at Dover Creek, Hammel Creek, and Natural Wall indicates the presence of uplifted volcanics in proximity to the Keweenaw fault in the central part of the peninsula during Jacobsville deposition (Figs.
 \ref{fig:Geologic_map} and \ref{fig:strat_column}).
 Coarse-grained arkosic to lithic sandstone is a facies associated with the volcanic-clast conglomerates and likely derived from mechanical weathering of the exhumed volcanics.
-Chemical weathering of exhumed volcanics from the hanging wall is a likely source of clay within the abundant fine-grained lithologies that are associated with the volcanic-clast conglomeratic facies \cite{Hamblin1958a}.
-Detrital mica grains up to $\sim$5 mm in size commonly found within siltstone to fine-grained sandstone are likely sourced from Paleoproterozoic and Archean lithologies in the region (Fig.
-\ref{fig:Geologic_map}).
+Chemical weathering of exhumed volcanics from the hanging wall is a likely source of clay within the abundant fine-grained lithologies that are associated with the volcanic-clast conglomeratic facies (Fig.
+\ref{fig:strat_column}.
+Detrital zircon provenance data from Jacobsville samples on the Keweenaw Peninsula developed by \citeA{Malone2020a} reveal zircon with dates overlapping with Midcontinent Rift volcanism consistent with this provenance interpretation.
+Detrital mica grains up to $\sim$5 mm in size that are commonly found within siltstone to fine-grained sandstone within the Jacobsville are likely sourced from Paleoproterozoic and Archean lithologies in the region (Fig.
+\ref{fig:Geologic_map}) and preferentially settled in lower energy overbank settings.
 
-The thickness of the Jacobsville strata can vary significantly in northern Michigan \cite{Hamblin1958a, Kalliokoski1982a}, and complete stratigraphic sections are not exposed at the studied localities (Figs.
+The thickness of the Jacobsville strata can vary in northern Michigan \cite{Hamblin1958a, Kalliokoski1982a}, and complete stratigraphic sections are not exposed at the studied localities (Figs.
 \ref{fig:Geologic_map} and \ref{fig:strat_column}).
 Clay-rich red siltstone is more abundant in conjunction with conglomerates (e.g.
-20-120 m in the Hungarian Falls section; Figs.
+20-120 m in the Dover Creek section; Figs.
 \ref{fig:strat_column} and \ref{fig:Field_photo}) than with the cross-stratified sandstone which dominates much of the formation.
@@
 Together, the volcanic conglomerates, red siltstones, and arkosic sandstones make up a unique facies association.
-As previously interpreted by \citeA{Brojanigo1984a}, this facies association was likely deposited proximal to actively uplifting volcanics near the Hungarian Falls, Hammel Creek, and Natural Wall sections (Fig.
+As previously interpreted by \citeA{Brojanigo1984a}, this facies association was likely deposited proximal to actively uplifting volcanic rocks near the Dover Creek, Hammel Creek, and Natural Wall sections (Fig.
 \ref{fig:strat_column}).
@@
 \ref{fig:Geologic_map}; \citeA{Hamblin1958a, Cannon1995a, Cannon1996a, Cannon2001a}).
-In the eastern Lake Superior region near Sault Ste Marie, Ontario, there are similar exposures of sedimentary rocks that are also mapped as the Jacobsville Formation \cite{Hamblin1958a}.
+In the eastern Lake Superior region near Sault Ste.
+Marie, Ontario, there are exposures of sedimentary rocks that are also mapped as the Jacobsville Formation \cite{Hamblin1958a}.
 As observed on the Keweenaw Peninsula, these sandstones are within the footwall of faults where they are overthrusted by Midcontinent Rift volcanics near Mamainse Point \cite{Manson1994a}.
-To the west, the Jacobsville Formation has been interpreted to be correlative with sedimentary rocks of the Bayfield Group in northern Wisconsin and the Fond du Lac Formation and Hinckley Formation in Minnesota \cite{Hamblin1958a, Kalliokoski1982a, Ojakangas2001b}.
+To the west, the Jacobsville Formation has been interpreted to be correlative with sedimentary rocks of the Bayfield Group in northern Wisconsin and the Fond du Lac Formation and Hinckley Formation in Minnesota \cite{Thwaites1912a, Hamblin1958a, Wallace1971a, Kalliokoski1982a, Ojakangas2001b}.
 A challenge associated with these correlations is that the ca.
@@
 \ref{fig:Geologic_map} and \ref{fig:strat_column}).
-Additional oriented cores and block samples of siltstone rip-up clasts within conglomerates and conglomeratic sandstones were sampled within the Hungarian Falls and Agate Falls sections (Figs.
+Additional oriented cores and block samples of siltstone intraclasts within conglomerates and conglomeratic sandstones were sampled within the Dover Creek and Agate Falls sections (Figs.
 \ref{fig:Geologic_map} and \ref{fig:strat_column}, and \ref{fig:intraclast_pmag}).
-To maximize sampling of distinct time snapshots of the geomagnetic field at the time of Jacobsville deposition and average out paleosecular variation, we optimized for vertical stratigraphic coverage.
+To maximize sampling of distinct time snapshots of the geomagnetic field at the time of Jacobsville deposition and average out paleosecular variation, we optimized for vertical stratigraphic coverage \cite{Sapienza2023a}.
 Each sample is a distinct stratigraphic horizon and therefore a paleomagnetic site, representing a unique record of the geomagnetic field.
@@
 Sun compass data were preferentially used when available.
-A total of 379 specimens including 30 rip-up clasts were collected for paleomagnetic study.
-To minimize the visual impact of sampling, rock containing core holes was knocked out of outcrops following samples---readily done for the friable siltstones.
+A total of 379 specimens including 30 intraclasts were collected for paleomagnetic study.
+To minimize the visual impact of sampling, rock containing core holes was knocked out of outcrops following samples---readily done for the friable siltstones and sandstones of the Jacobsville Formation.
 
 Challenges exist in isolating the primary magnetization in sedimentary rocks.
-Primary detrital remanence acquired during deposition can be masked by secondary remanence acquired through precipitation and growth of new minerals within sedimentary rocks syn- to post-deposition (e.g.
+Primary detrital remanence acquired during deposition can be masked by secondary remanence acquired through precipitation and growth of diagenetic minerals within sedimentary rocks syn- to post-deposition (e.g.
 pigmentary hematite formed from precursor phases such as ferrihydrite; \citeA{Jiang2018a}).
@@
 
+In the Jacobsville Formation samples that we collected, there are both detrital hematite grains (10s of micrometers) as well as finer grained (sub-micrometer) pigmentary hematite (Fig.
+S1).
 We adopted a thermal demagnetization protocol with increasingly higher resolution approaching the N\'eel temperature of hematite (5 \textdegree C to 2 \textdegree C; Fig.
-\ref{fig:intraclast_pmag}).
+\ref{fig:intraclast_pmag}, S2).
 The specimens underwent stepwise thermal demagnetization at the UC Berkeley Paleomagnetism Lab using an ASC demagnetizer (residual fields $<$10 nT) with measurements made on a 2G DC-SQUID magnetometer.
-While the very fine-grained sandstone and siltstone lithologies typically do not have a significant present-day local field overprint component, such component can be present in fine-grained sandstones and is removed by $\sim$300\textdegree C during thermal demagnetization (Fig.
-S1).
+Due to the thermal gradient in the oven, we keep the specimens at the same location throughout the demagnetization steps.
+This protocol makes the temperature difference between each step relatively consistent for each specimen compared to if they changed positions within the oven.
+While the very fine-grained sandstone and siltstone lithologies typically do not have an appreciable present-day local field overprint component, such a component can be present in fine-grained sandstones and is removed by $\sim$300\textdegree C during thermal demagnetization (Fig.
+S2, S3).
 A total of 356 specimens yielded detrital remanent magnetizations that can be fit with least-square lines \cite{Kirschvink1980a}.
-These fits were made using PmagPy \cite{Tauxe2016a} and all paleomagnetic data are available to the measurement level in the MagIC database (https://earthref.org/MagIC/19780/df1b7c0c-b1cc-4e1e-b40c-2585f007178a; \textit{this link provides access to reviewers and will be updated with a final version of the link when the manuscript is accepted and assigned a DOI}).
+These fits were made using PmagPy \cite{Tauxe2016a} and all paleomagnetic data are available to the measurement level in the MagIC database (\url{https://10.7288/V4/MAGIC/19780}).
 
@@
 
-Similar to the thermal demagnetization results of the hematite-bearing fluvial intraclasts of the Freda Formation \cite{Swanson-Hysell2019b}, the siltstone rip-up clasts of the Jacobsville Formation typically reveal two distinct magnetization components (Fig.
+Similar to the thermal demagnetization results of the hematite-bearing fluvial intraclasts of the Freda Formation \cite{Swanson-Hysell2019b}, the siltstone intraclasts of the Jacobsville Formation typically reveal two distinct magnetization components (Fig.
 \ref{fig:intraclast_pmag}).
@@
 \ref{fig:intraclast_pmag}).
-The relatively low unblocking temperatures and similarity in directions among the reoriented rip-up clasts indicate that it is chemical remanent magnetization acquired through crystallization of secondary pigmentary hematite following clast redeposition.
+The relatively low unblocking temperatures and similarity in directions among the reoriented intraclasts indicate that it is chemical remanent magnetization acquired through crystallization of secondary pigmentary hematite following clast redeposition.
 After removal of this component, further thermal demagnetization with small step increments at higher temperatures up to $\sim$689 \textdegree C often reveal an origin-trending component (Fig.
@@
 As a result, 27 out of 30 intraclast specimens could be fit with high-temperature least squares lines.
-Distinct from the well-grouped mid-temperature component directions, the high-temperature directions of the intraclasts are dispersed and the null hypothesis of randomness cannot be rejected at the 95\% confidence level (n=5 for intraclasts at Agate Falls and n=22 for intraclasts at Hungarian Falls; Fig.
+Distinct from the well-grouped mid-temperature component directions, the high-temperature directions of the intraclasts are dispersed and the null hypothesis of randomness cannot be rejected at the 95\% confidence level (n=5 for intraclasts at Agate Falls and n=22 for intraclasts at Dover Creek; Fig.
 \ref{fig:intraclast_pmag}; \citeA{Watson1956a}).
@@
 \includegraphics[width=.7\textwidth]{intraclast_pmag.png}
-\caption{\scriptsize(A) Field photos of siltstone to fine-grained sandstone fluvial rip-up clasts within conglomerate and conglomeratic sandstone beds in the Jacobsville Formation.
-(B) Representative step-wise thermal demagnetization data from a rip-up clast from Hungarian Falls (HFC-23) plotted on an orthogonal vector diagram.
+\caption{\scriptsize(A) Field photos of siltstone to fine-grained sandstone fluvial intraclasts within conglomerate and conglomeratic sandstone beds in the Jacobsville Formation.
+(B) Representative step-wise thermal demagnetization data from an intraclast from Dover Creek (HFC-23) plotted on an orthogonal vector diagram.
 The plot reveals a mid-temperature component and an origin-trending high-temperature component.
@@
 In contrast, the high-temperature component directions are dispersed and pass the randomness test of \citeA{Watson1956a}.
-The DRM directions are shown in tilt-corrected bedding coordinates and CRM directions are shown in geographic coordinates.
+Both the DRM and the CRM directions are shown in geographic coordinates.
 NRM = natural remanent magnetization.}
@@
 \ref{fig:Geologic_map} and \ref{fig:in_situ_pmag}).
-The exposures of steeply tilted, moderately tilted, as well as nearly horizontal beds along this section allow us to conduct a paleomagnetic fold test to investigate the timing of detrital remanence acquisition.
-The folding in this section is the result of deformation in the immediate footwall of the Keweenaw fault.
-The chronological constraints on the timing of Keweenaw fault motion indicate that it occurred within $\sim$10 Myr of deposition---making this fold test a more informative constraint on the age of the remanence than typical for such tests.
-Given that the more steeply tilted beds accommodated significant deformation during shortening associated with the Keweenaw fault motion, they could be more prone to non-cylindrical folding and complicate tilt-correction of the paleomagnetic directional data \cite<e.g.>{Pueyo2003a, Nabavi2021a}.
-Therefore, we conduct a bootstrap fold test \cite{Tauxe1994a} using specimens from the nearly horizontal beds and the moderately tilted beds.
-The result shows that the degrees of unfolding that result in the best grouping of the DRM directions have 95\% confidence bounds that overlap with 100\%, consistent with the specimens having acquired their magnetization prior to folding (Fig.
+The stratigraphic section is within a large-scale drag fold within the immediate footwall of the Keweenaw fault.
+As a result, there are exposures of steeply tilted, moderately tilted, as well as nearly horizontal beds along this section that allow us to conduct a paleomagnetic fold test to investigate the timing of magnetization.
+The chronological constraints on the timing of Keweenaw fault motion indicate that the folding occurred on a timescale of $\sim$10 Myr of deposition---making this fold test a more informative constraint on the age of the remanence than typical for such tests.
+Given that the more steeply tilted beds accommodated deformation during shortening associated with the Keweenaw fault motion, they could be more prone to non-cylindrical folding and complicate tilt-correction of the paleomagnetic directional data \cite<e.g.>{Pueyo2003a, Nabavi2021a}.
+Therefore, we conduct a paleomagnetic bootstrap fold test \cite{Tauxe1994a} using specimens from the nearly horizontal beds and the moderately tilted beds.
+The results show that the degrees of untilting that result in the best grouping of the DRM directions have 95\% confidence bounds that overlap with 100\% unfolding, consistent with the specimens having acquired their magnetization prior to tilting (Fig.
 \ref{fig:fold_test}).
-This positive fold test further supports that the Jacobsville red beds acquired a primary detrital remanence during deposition.
-In contrast, the low-temperature component directions from the Snake Creek tributary fails a fold test (Fig.
-S2), consistent with them being acquired as chemical remanence magnetizations that postdate the folding.
+This positive fold test further supports the interpretation that the Jacobsville red beds acquired a primary detrital remanence during deposition.
+In contrast, the mid-temperature component directions that typically unblock up to 650 \textdegree C from the Snake Creek tributary fails a fold test (Fig.
+S4), consistent with them being acquired as chemical remanent magnetizations that postdate the tilting.
+The positive fold test on the detrital remanence in combination with the similarity of the detrital and chemical remanence directions (Fig.
+6B) indicates that the majority of chemical remanence acquisition likely occurred geologically soon after the deformation.
 
@@
 \caption{Bootstrap paleomagnetic fold test \cite{Tauxe1994a} of the detrital remanent magnetization directions recorded by specimens from the nearly horizontal beds and moderately tilted beds at the Snake Creek tributary.
-Complete unfolding lies within the 95\% confidence limits of the test, consistent with the magnetization having been acquired prior to folding.}
+Complete unfolding lies within the 95\% confidence limits of the test, consistent with the magnetization having been acquired prior to tilting.}
 \label{fig:fold_test}
@@
 The paleomagnetic polarity reversals are shown in stratigraphic context in Figure \ref{fig:strat_column} with specimen magnetization declinations plotted against stratigraphic heights.
-The paleogeographic context of Laurentia within Rodinia and into the Paleozoic enables us to interpret the westerly DRM directions as normal geomagnetic polarity and the easterly DRM directions as reversed polarity (Fig.
+The record of Laurentia's paleomagnetic poles from the Proterozoic through the Phanerozoic (see compilations in \citeA{Torsvik2012a} and \citeA{Swanson-Hysell2021c}) gives a continuity of the orientation of the continent that enables the geomagnetic polarity of the directions to be interpreted.
+In this framework, the westerly DRM directions are normal geomagnetic polarity and the easterly DRM directions are reversed polarity (Fig.
 \ref{fig:in_situ_pmag}).
 
-At the Hungarian Falls section along Dover Creek, numerous hematite-rich siltstone to fine sandstone layers commonly occur in association with trough cross-bedded sandstone and conglomeratic facies.
+At the Dover Creek section, which goes along Dover Creek and then up the waterfall of a side tributary (the Hungarian Falls side falls), numerous hematite-rich siltstone to fine sandstone layers commonly occur in association with trough cross-bedded sandstone and conglomeratic facies.
 These lithofacies are particularly well-exposed along a tributary waterfall to Dover Creek such that 158 samples were collected for paleomagnetic study (Fig.
-\ref{fig:Geologic_map}, \ref{fig:strat_column}), making this section the most densely sampled amongst all sections.
+\ref{fig:Geologic_map}, \ref{fig:strat_column}), making this section the most densely sampled amongst all the sections.
 The specimen DRMs show that the geomagnetic field reversed at least nine times during deposition within the section (Fig.
@@
 
-The Natural Wall and Hammel Creek sections are in close proximity to the Hungarian Falls section (Fig.
-\ref{fig:Geologic_map}) and the unique sedimentary facies association of volcanic-clast conglomerate, red siltstone, and coarse arkosic sandstone is similar to the basal $\sim$120 meters of the section at Hungarian Falls (Fig.
+The Natural Wall and Hammel Creek sections are in close proximity to the Dover Creek section (Fig.
+\ref{fig:Geologic_map}) and the unique sedimentary facies association of volcanic-clast conglomerate, red siltstone, and coarse arkosic sandstone is similar to the basal $\sim$120 meters of the section at Dover Creek (Fig.
 \ref{fig:strat_column}).
-Our sampling at these two localities is more limited due to fewer occurrences of exposed siltstone and fine-grained sandstone beds due to more limited exposure throughout the sections (Fig.
+Our sampling at these two localities is more limited due to there being poorer exposures of siltstone and fine-grained sandstone beds in these sections (Fig.
 \ref{fig:strat_column}).
@@
 \centering
-\includegraphics[width=0.95\textwidth]{in_situ_pmag.pdf}
-\caption{(A) Specimen detrital remanence directions (DRM) plotted by locality on equal area plots.
+\includegraphics[width=0.8\textwidth]{in_situ_pmag.pdf}
+\caption{\footnotesize (A) Specimen detrital remanence directions (DRM) plotted by locality on equal area plots.
 During sample collection, we optimized for vertical stratigraphic coverage such that each sample constitutes a single horizon and therefore a paleomagnetic site.
-The measurement-level data are available in the MagIC database (https://earthref.org/MagIC/19780/df1b7c0c-b1cc-4e1e-b40c-2585f007178a; \textit{this link provides access to reviewers and will be updated with a final version of the link when the manuscript is accepted}).
+Specimens from the Snake Creek tributary as a mean direction of dec=256.0\textdegree, inc=-30.5\textdegree, k=9.9, $\alpha_{95}$=5.6\textdegree, n=71; the Natural Wall section has a mean direction of dec=242.4\textdegree, inc=-6.4\textdegree, k=9.0, $\alpha_{95}$=13.5\textdegree, n=15; the Hammel Creek section has a mean direction of dec=94.1\textdegree, inc=8.7\textdegree, k=30.7, $\alpha_{95}$=6.8\textdegree, n=16; the Dover Creek section has a mean direction of dec=73.2\textdegree, inc=5.4\textdegree, k=5.6, $\alpha_{95}$=5.6\textdegree, n=138; the Agate Falls section has a mean direction of dec=83.4\textdegree, inc=13.0\textdegree, k=13.4, $\alpha_{95}$=4.9\textdegree, n=67.
+All reported directions are calculated in tilt-corrected coordinates and after unifying polarity.
+The measurement-level data are available in the MagIC database (https://10.7288/V4/MAGIC/19780).
 (B) Summary equal area plots at the study-level combining directions from all localities.
@@
 The \citeA{McFadden1990a} reversal test results show the angle between the two mean directions (15.9\textdegree; no inclination shallowing correction) exceeds the critical angle (7.5\textdegree) (Fig.
-S3), thereby failing the test.
+S5), thereby failing the test.
 This difference largely arises due to normal (west directed) directions in the Snake Creek tributary section being slightly steeper in their upward inclination.
-There are at least three possible explanations for this behavior: 1) the lower number of normal directions which dominantly come from the Snake Creek tributary section may not have averaged out secular variation; 2) there could have been plate motion during Jacobsville deposition with the normal polarity in the Snake Creek tributary section being acquired when Laurentia was at a slightly higher latitude; 3) there could be a bias from the reversed CRM superimposed on the normal DRM in the Snake Creek tributary specimens given that removal of the CRM is associated with progressively shallowing the inclination in some specimens \cite{Cox1960a, Scott1996a}.
+There are at least four possible explanations for this behavior: 1) the lower number of normal directions which dominantly come from the Snake Creek tributary section may not have averaged out secular variation; 2) there could have been plate motion during Jacobsville deposition with the normal polarity in the Snake Creek tributary section being acquired when Laurentia was at a slightly higher latitude; 3) there could be a bias from the reversed CRM superimposed on the normal DRM in the Snake Creek tributary specimens being incompletely removed during thermal demagnetization; 4) the geomagnetic field in the early Neoproterozoic could have been slightly different than that of more recent times in terms of the dispersion of pole position at distinct time snapshots potentially associated with changes in field intensity.
 Regardless, the positive intraclast conglomerate test gives confidence that the Jacobsville Formation has a primary detrital remanent magnetization that can be isolated through thermal demagnetization.
@@
 
-A total of 186 specimens yielded interpretable chemical remanence directions that are dominantly reversed polarity.
-These directions are plotted by stratigraphic section location in Figure S4 and are all shown in Figure 6B.
+A total of 186 specimens yielded interpretable chemical remanence directions that are of dominantly reversed polarity.
+These directions are plotted by stratigraphic section location in Figure S6 and are all shown in Figure 6B.
 That the CRM directions are a secondary remanence component that postdate the deposition of the Jacobsville Formation is supported by the negative intraformational conglomerate test results and failure in the fold test (Figs.
-\ref{fig:intraclast_pmag}, S2).
-However, the similarity of the CRM directions (declination=96.2\textdegree, inclination=15.1\textdegree, $\alpha_{95}$=4.2\textdegree, n=186) with the primary DRM directions (declination=103.1\textdegree, inclination=13.4\textdegree, $\alpha_{95}$=3.3\textdegree, n=307, no inclination shallowing correction) indicate that growth of the secondary pigmentary hematite which carries the CRM likely occurred relatively soon after deposition when Laurentia was in a similar paleogeographic position.
+\ref{fig:intraclast_pmag}, S4).
+However, the similarity of the CRM directions (declination=96.2\textdegree, inclination=15.1\textdegree, $\alpha_{95}$=4.2\textdegree, n=186) with the primary DRM directions (declination=103.1\textdegree, inclination=13.4\textdegree, $\alpha_{95}$=3.3\textdegree, n=307, no inclination shallowing correction) indicate that growth of the secondary pigmentary hematite which carries the CRM likely occurred soon after deposition when Laurentia was in a similar paleogeographic position.
 
 Overall, the multiple geomagnetic field polarity reversals recorded by the Jacobsville Formation constrain that the end of the Keweenawan normal superchron \cite{Driscoll2016b} (which started ca.
-1099 Ma; \cite{Swanson-Hysell2019a}) had to have ended by the onset of ca.
-990 Ma Jacobsville deposition \cite{Hodgin2022a}.
+1099 Ma; \citeA{Swanson-Hysell2019a}) had to have ended by the onset of Jacobsville deposition.
 
@@
 775-719 Ma poles from western Laurentia extensional basins \cite{Weil2006a, Eyster2019a}.
-Developing a new paleomagnetic pole from the Jacobsville Formation presents an opportunity for a well-constrained early Neoproterozoic pole given that the depositional age of the formation has recently been constrained to ca.
-990 Ma \cite{Hodgin2022a}.
+Developing a new paleomagnetic pole from the Jacobsville Formation presents an opportunity for a well-constrained early Neoproterozoic constraint on Laurentia's paleogeographic position.
 
@@
 The scatter in the DRM directions can be interpreted to reflect variations in the geomagnetic field during the deposition of the sediments \cite{Steiner1983a, Tauxe1984a}.
-The multiple geomagnetic field reversals captured by specimens collected from the Hungarian Falls and Agate Falls sections support that a prolonged period of time is represented by the specimens given that the duration of geomagnetic polarity chrons are typically on the order of tens of Kyr to Myr timescales (Fig.
+The multiple geomagnetic field reversals captured by specimens collected from the Dover Creek and Agate Falls sections support that a prolonged period of time is represented by the specimens given that the duration of geomagnetic polarity chrons are typically on the order of tens of Kyr to Myr timescales (Fig.
 \ref{fig:strat_column}).
@@
 Such a directional distribution is due to rotation of detrital hematite grains during deposition and subsequent compaction, resulting in inclinations recorded by sedimentary rocks having shallower angles than the local field in which they are deposited.
-If uncorrected, shallower inclinations obtained from sedimentary rocks can result in erroneously low estimates of paleolatitudes, biasing paleogeographic reconstructions \cite{Muttoni2003a, Domeier2012a}.
+If uncorrected, shallower inclinations obtained from sedimentary rocks can result in erroneously low estimates of paleolatitudes, biasing paleogeographic reconstructions.
 To estimate the amount of inclination shallowing, we apply the statistical elongation/inclination (E/I) method that evaluates the deviation of the distribution of a large number ($>$100) of observed sedimentary paleomagnetic directional data with the predicted distribution given by a statistical paleosecular variation model \cite{Tauxe2004b}.
@@
 1093 Ma Cut Face Creek Sandstone in the Midcontinent Rift successfully corrects the shallow paleomagnetic inclinations of the sediments to that of the North Shore Volcanic Group lava flows that bracket them.
-\citeA{Pierce2022a} also developed a method to represent uncertainties associated with the amount of inclination shallowing into sedimentary paleomagnetic pole positions which can then be summarized with an elliptical Kent distribution \cite{Kent1982a} rather than a circular Fisher distribution \cite{Fisher1953a}.
+\citeA{Pierce2022a} also developed a method, that we use here, to represent uncertainties associated with the amount of inclination shallowing into sedimentary paleomagnetic pole positions which can then be summarized with an elliptical Kent distribution \cite{Kent1982a} rather than a circular Fisher distribution \cite{Fisher1953a}.
 
@@
 \ref{fig:EI_results}).
-The $f$ factor of 0.64 is close to the value of 0.58 which is a the mean of inclination shallowing factors compiled for hematite-bearing rocks in \citeA{Pierce2022a} (building on compilations of \citeA{Bilardello2016b} and \citeA{Vaes2021a}).
+The $f$ factor of 0.64 is close to the value of 0.58 which is the mean of inclination shallowing factors compiled for hematite-bearing rocks in \citeA{Pierce2022a} (building on compilations of \citeA{Bilardello2016b} and \citeA{Vaes2021a}).
+The Fisher mean inclination-corrected detrital remanence mean direction is dec=76.7\textdegree, inc=18.6\textdegree $\alpha_{95}$=3.9\textdegree.
 The Fisher mean inclination-corrected paleomagnetic pole associated with this nominal $f$ factor is Plon=183.4\textdegree E, Plat=16.9\textdegree S, A$_{95}$=3.1\textdegree.
@@
 Following \citeA{Pierce2022a}, the pole can be represented with a Kent distribution 95\% confidence ellipse which is: mean longitude=183.4\textdegree E, mean latitude=16.9\textdegree S, major axis longitude=255.5\textdegree E, major axis latitude=45.2\textdegree N, major axis magnitude=4.1\textdegree, minor axis longitude=108.1\textdegree E, minor axis latitude=39.9\textdegree N, minor axis magnitude=3.1\textdegree (Fig.
-\ref{fig:EI_results}; Table \ref{tab:Kent_means})
-
-% To further represent the uncertainty associated with the $f$ factor, and take all estimated $f$ values from the bootstrap results of the E/I algorithm and apply each value to correct all Jacobsville DRMs (with the reversed directions flipped to antipodes).
-Then we calculate the Fisher statistics for each of the inclination-corrected mean poles.
-Finally, we perform a Monte Carlo approach to generate 100 resamples of each Fisher mean pole.
-As a result, the distribution of the resampled mean pole positions have an elongated distribution which represents more uncertainty associated with paleolatitudes of Laurentia than the orientation of the continent (Fig.
-\ref{fig:EI_results}).
-The distribution of paleolatitudes of the resampled mean poles can be approximated by a normal distribution (Fig.
-\ref{fig:EI_results}).
-Following the \citeA{Pierce2022a} methodology, we then use the Kent distribution to determine the 95\% confidence ellipse of the distribution of the resampled mean poles (mean longitude=183.4\textdegree E, mean latitude=17.1\textdegree S, major axis longitude=255.3\textdegree E, major axis latitude=45.2\textdegree N, major axis magnitude=4.2\textdegree, minor axis longitude=108.2\textdegree E, minor axis latitude=39.8\textdegree N, minor axis magnitude=3.1\textdegree; Fig.
-\ref{fig:EI_results}).
-As compared to the paleomagnetic pole calculated with Fisher statistics using the nominal inclination shallowing correction factor of $f$=0.66, this Kent ellipse indicates that there is more uncertainty along the great circle path between the mean pole position and the locality of the Jacobsville basin (Fig.
 \ref{fig:EI_results}; Table \ref{tab:Kent_means}).
+This pole position is close to the end of the Keweenawan Track (Fig.\ref{fig:pole_plot}) and far away from Laurentia's pole positions during the Paleozoic when there was orogenesis along Laurentia's eastern margin (Fig.
+S7).
+This pole position is consistent with the geochronology constraints on the deposition of the Jacobsville Formation and the positive paleomagnetic field tests that the detrital remanence of the Jacobsville Formation was acquired during the earliest Neoproterozoic.
 
@@
 \caption{\scriptsize Results of the estimated amount of inclination shallowing of the detrital remanent magnetization of the Jacobsville Formation using the elongation/inclination (E/I) method \cite{Tauxe2004b}.
-(A) The E/I method givens an estimated flattening factor $f$=0.64 (red curve) based on where the elongation/inclination curve of the dataset intersects that predicted by the TK03 paleosecular variation model (black curve; \citeA{Tauxe2004b}).
+(A) The E/I method gives an estimated flattening factor $f$=0.64 (red curve) based on where the elongation/inclination curve of the dataset intersects that predicted by the TK03 paleosecular variation model (black curve; \citeA{Tauxe2004b}).
 The grey lines show 1,000 bootstrap resamples of the DRM directions which provide an estimate of the uncertainty associated with the $f$ factor estimate (95\% interval from 0.83 to 0.51).
@@
 (C) The cumulative distribution of inclinations based on the E/I bootstrap results with the 95\% confidence bounds shown in terms of inclination and $f$ factor.
-(D) The new Kent mean inclination corrected DRM pole (red square) of the Jacobsville Formation is shown in the context of the pole calculated for the CRM (purple triangle) as well as the Jacobsville pole developed by \citeA{Roy1978a}, the Chequamegon Formation pole of \citeA{McCabe1983a}, and poles from the Oronto Group (Nonesuch Formation, \citeA{Slotznick2023a}, Freda Formation, \citeA{Henry1977a}).
+(D) The new Kent mean inclination corrected DRM pole (red square) of the Jacobsville Formation is shown in the context of the pole calculated for the CRM (purple triangle) as well as the Jacobsville pole developed by \citeA{Roy1978a}, the Chequamegon Formation pole of \citeA{McCabe1983a}, and poles from the Oronto Group (Nonesuch Formation, \citeA{Slotznick2023a}; Freda Formation, \citeA{Henry1977a}).
 The mean pole position of the Kent distribution was calculated following the method of \citeA{Pierce2022a}.
@@
 
-\citeA{Roy1978a} also studied the paleomagnetism of sedimentary rocks grouped as the Jacobsville Formation and isolated a characteristic magnetization component via alternating-field, thermal, and chemical demagnetization on a suite of Jacobsville red beds near Keweenaw Peninsula (their area A), the town of Marquette (their area B), and Sault Ste Marie (their area C).
+\citeA{Roy1978a} also studied the paleomagnetism of sedimentary rocks grouped as the Jacobsville Formation and isolated a characteristic magnetization component via alternating-field, thermal, and chemical demagnetization on a suite of Jacobsville red beds near the Keweenaw Peninsula (their area A), the town of Marquette (their area B), and Sault Ste Marie (their area C).
 Although least-squares principal component analyses was not a routine approach in fitting paleomagnetic directions at the time, that study had success in removing the present day local field overprint and was able to resolve dual magnetic polarities at one locality.
@@
 \ref{fig:EI_results}; \citeA{Henry1977a, Slotznick2023a}).
-We hypothesize that the red beds studied by \citeA{Roy1978a} from the Sault Ste Marie area are likely time equivalent to sedimentary rocks of the Oronto Group, instead of being correlative with the Jacobsville Formation.
+As suggested by \citeA{Dubois1962a} and \citeA{Roy1978a}, these data suggest that the fluvial red beds in the Sault Ste Marie area that are often taken to be correlative to the Jacobsville Formation \cite<e.g.>{Malone2020a} are likely time equivalent to sedimentary rocks of the Oronto Group and deposited during post-rift thermal subsidence.
 
@@
 \ref{fig:Geologic_map}) has been hypothesized to be coeval with the Jacobsville Formation \cite{Hamblin1958a, Kalliokoski1982a, Malone2016a}.
-Published low precision detrital zircon U-Pb dates developed by LA-ICP-MS dates are consistent with this interpretation with a maximum depositional age of ca.
+Published low precision detrital zircon U-Pb dates developed by laser ablation–inductively coupled plasma–mass spectrometry (LA-ICP-MS) dates are consistent with this interpretation with a maximum depositional age of ca.
 1035 Ma for the Chequamegon Formation of the Bayfield Group \cite{Craddock2013a}.
@@
 This interpretation can be evaluated with further high-precision geochronology focused on the Bayfield Group.
+
+\subsection*{Age of the Jacobsville Pole}
+
+At the western edge of the Jacobsville bedrock belt in northern Michigan, the formation is in angular unconformity with Midcontinent Rift rocks that were uplifted and tilted prior to Jacobsville deposition (Fig.
+\ref{fig:Geologic_map}; \citeA{Hedgman1992a, Cannon1995a}).
+Rb-Sr thermochronologic data developed from Paleoproterozoic to Archean lithologies that were exhumed along the Marenisco Fault along with Midcontinent Rift strata in this region indicate that this uplift occurred ca.
+1050 Ma associated with the Ottawan phase of the Grenvillian orogeny \citeA{Cannon1993a}.
+Field observations reveal erosional contacts and development of paleosols between the Jacobsville Formation and underlying Midcontinent Rift volcanic rocks at the Sturgeon River within the bedrock belt \cite{Hamblin1958a, Zbinden1988a}.
+Seismic reflection data beneath Lake Superior have been also been interpreted to indicate that the Jacobsville Formation lies in angular unconformity atop sedimentary rocks of the Oronto Group \cite{Cannon1989a}.
+This context requires that there was ca.
+1050 Ma contractional deformation of Midcontinent Rift strata, erosion of these tilted strata, and renewed subsidence prior to the onset of Jacobsville deposition atop them.
+
+Maximum depositional detrital zircon dates are consistent with this history and provide additional constraints.
+\citeA{Hodgin2022a} applied tandem dating where the young population emerging from a large number of low-precision LA-ICP-MS dates are followed by high-precision CA-ID-TIMS analysis.
+This approach has the advantage of combining the high number of low-precision analyses that can be acquired through LA-ICP-MS with dates developed through CA-ID-TIMS where the deleterious effects of Pb-loss can be mitigated, and more accurate and precise dates can be obtained.
+These data reveal the youngest zircon at Sandstone Creek near the Keweenaw Fault (Figure \ref{fig:Geologic_map}) to have a date of 992.51$\pm$0.64 Ma (2$\sigma$ analytical uncertainty) and the youngest zircon at Agate Falls (Figure \ref{fig:Geologic_map}) a date of 1003.21 $\pm$ 2.23 Ma (2$\sigma$).
+These dates are consistent with the age of syntectonic magmatism in the Grenville orogen associated with the Rigolet phase of the orogeny \cite<e.g.>{Bussy1995a, Turlin2019a, Jannin2018a} which is a potential source for the grains.
+Overall, these dates constrain a maximum depositional age of the Jacobsville Formation in the earliest Neoproterozoic.
+
+While there has been recent success in applying tandem detrital zircon data to obtain maximum depositional ages that are close to true depositional ages \cite{Karlstrom2020a}, the possibility exists that deposition could significantly postdate the youngest dated detrital zircon.
+A firm minimum depositional age on the age of the Jacobsville is that it is unconformably overlain by the Cambrian Munising Formation whose base was deposited ca.
+501 to 497 Ma during the Dresbachian Stage (the Guzhangian Stage in the modern Cambrian time scale) \cite{Hamblin1958a, Haddox1990a}.
+This unconformity is erosional as it truncates Jacobsville strata and intraformational clastic dikes and can have slight angularity \cite{Hamblin1958a, Haddox1990a}.
+That the Jacobsville Formation is older than the Munising Formation lends tighter constraints than the Cambrian age itself given the evidence discussed above that Jacobsville deposition was syn-orogenic.
+The only candidate for such orogenesis is the ca.
+1010-980 Ma Rigolet phase of the Grenvillian orogeny given that subsequent Paleozoic orogenesis (e.g.
+the Ordovician Taconic and Carboniferous Alleghanian orogenies) post-dates deposition of the Munising Formation.
+
+That the Jacobsville Formation is folded in the footwall of the Keweenaw fault, beneath the older Midcontinent Rift volcanic rocks in the hanging wall, provides another opportunity to develop a minimum depositional age.
+Seeking to obtain such a constraint, \citeA{Hodgin2022a} targeted calcite within a Keweenaw Fault breccia that cross-cuts a dense network of zeolite veins and slickenslides that yielded a U-Pb date of 985.5$\pm$35.8 Ma (2$\sigma$).
+While this calcite date has large analytical uncertainty, it indicates that the fault breccia developed associated with the Grenvillian orogeny and it overlaps with the ca.
+1010-980 Ma ages of metamorphism associated with the Rigolet phase of the orogeny \cite{Swanson-Hysell2023a}.
+It is during this ca.
+1010-980 Ma Rigolet phase of the orogeny that the Grenville Front developed \cite{Rivers2008a} which indicates that this was a time period when contractional deformation associated with the orogeny was propagating into the Superior craton.
+Together, the detrital zircon, sedimentology, and the fault calcite dates are consistent with the Jacobsville Formation being deposited in a syn-orogenic Grenville backbulge basin and subsequently deformed as contraction associated with the Rigolet phase of the Grenville orogeny propagated into the interior of Laurentia.
+
+Previously, some researchers have interpreted there to have been large-scale regional contractional deformation during the Alleghanian orogeny in the Paleozoic leading to reactivation of reverse faults including the Keweenaw Fault \cite{Craddock2017a}.
+While the Keweenaw Fault is defined quite broadly in \citeA{Craddock2017a}, the evidence put forward that is relevant to the segment in this study region are: 1) an interpreted post-Grenvillian orogeny age for the Jacobsville Formation which would require that the deformation of the formation was associated with subsequent orogenesis; and 2) deformation within two $<$1.5 km diameter outliers of Paleozoic sedimentary rocks that overlie the Jacobsville (Limestone Mountain in Figure \ref{fig:Geologic_map}).
+The data used to argue for a post-Grenvillian depositional age of the Jacobsville Formation were the low-precision LA-ICP-MS detrital zircon dates of \citeA{Malone2016a} that have now been superseded by the tandem dates of \citeA{Hodgin2022a}.
+These dates are consistent with syn-Grenvillian deposition.
+This progress leaves the tilted and/or brecciated Paleozoic strata in the small exposed Paleozoic outliers of the 1.5 x 1 km Limestone Mountain and the neighboring 0.5 x 0.3 km Sherman Hill as evidence of Paleozoic deformation: interpreted to be Paleozoic orogenesis by \citeA{Hamblin1958a, Craddock2017a} and suggested to be an impact structure by others \cite{Milstein1987a}.
+The impact origin interpretation builds on previous interpretations of the structure as being cryptovolcanic \cite{Cannon1981a}--a common interpretation for structures that have subsequently been recognized as impact related.
+The tilt and brecciation of these outliers contrasts with widespread flat-laying Paleozoic sedimentary rocks that overlie the reverse faults that underwent contractional deformation associated with rift inversion in Minnesota \cite{Jirsa2011a}.
+These strata constrain Paleozoic deformation to have been relatively minor on the scale of 10s of meters rather than the kilometers of uplift associated with the rift inversion along these reverse faults \cite{Boerboom2018a}.
+Clumped isotope temperatures of $\sim$50\textdegree C from the Neoproterozoic fault calcite, also developed by \citeA{Hodgin2022a}, reveal that the Keweenaw fault zone was within a couple kilometers of the surface at the time of late Mesoproterozoic to early Neoproterozoic calcite precipitation.
+These data are inconsistent with major Paleozoic uplift and support the interpretation that the majority of contractional inversion along the Keweenaw Fault occurred during the Rigolet Stage of the orogeny.
+
+Our new paleomagnetic data provides additional insights into the timing of Jacobsville deformation in the early Neoproterozoic.
+At the Snake Creek tributary section which is folded against the Keweenaw Fault (Figure \ref{fig:Geologic_map}, \ref{fig:strat_column}), the chemical remanence directions held by pigmentary hematite fail a fold test (Figure S4), indicating that the pigmentary hematite holding the remanence grew after deformation of the strata.
+The mean pole position of the chemical remanent magnetization at the Snake Creek tributary section overlaps with paleomagnetic poles of the Grenville Loop that have been assigned exhumation ages of ca.
+970-960 Ma (Figure S7; \citeA{Brown2012a}).
+This chemical remanence pole position and the mean Jacobsville detrital remanence pole position are both far away from Laurentia's pole positions at the times of Paleozoic orogenesis or any younger time in the Phanerozoic (Figure S7).
+This result indicates that the majority of the tilting of the Jacobsville in the footwall of the Keweenaw Fault could not have occurred during Paleozoic orogenesis such as the Alleghanian orogeny as the chemical remanent magnetization would then be expected to record a direction aligned with the Phanerozoic pole path.
+Instead, the Rigolet phase of Grenvillian orogeny, which ended by ca.
+980 Ma, is the only feasible orogenic interval that could have tilted these strata.
+This result constrains the Jacobsville Formation in the studied outcrop belt to have been deposited prior to ca.
+980 Ma.
+While Paleozoic reactivation could have occurred, and there could be deformation of that age elsewhere such as at Limestone Mountain, deformation at this time on the Keweenaw Fault would have been relatively minor compared to that in the early Neoproterozoic.
+
+In conclusion, our paleomagnetic data and existing chronological data combined with the geological constraints are most consistent with the interpretation that the Jacobsville Formation within the studied bedrock belt was deposited in a syn-orogenic basin associated with the Rigolet Phase of the Grenvillian orogeny.
+This syn-orogenic basinal setting, which is consistent with the sedimentology of the formation (Fig.
+\ref{fig:strat_column}), provides a mechanism for developing accommodation space.
+The total duration of deposition of the Jacobsville is uncertain with these constraints alone particularly given the low precision on the 985.5$\pm$35.8 Ma calcite date.
+However, data from the Grenville orogen itself constrain the end of the Rigolet phase to be ca.
+980 Ma \cite{Swanson-Hysell2023a}.
+As the studied sections proximal to the Keweenaw Fault were deformed in its footwall during this phase of contractional deformation, it is reasonable to interpret deposition as occurring before the ca.
+980 Ma cessation of Grenvillian orogenesis.
+Taken together, we consider the best nominal age constraint for the depositional age of the Jacobsville Formation to pair with the Jacobsville paleomagnetic pole position to be ca.
+990 Ma as the Rigolet phase of the orogeny was ongoing.
+While deposition within peripheral foreland basins typically lasts between 10 and 50 million years \cite{Woodcock2004a}, the backbulge depocenter may not be as stable or long-lived.
+Given the available chronometric constraints from the Jacobsville Formation, the Keweenaw Fault and the Rigolet phase in the Grenville orogen, deposition of the Jacobsville can be considered to have occurred within the 1000 to 980 Ma time interval.
 
@@
 Incorporating temporal and spatial uncertainty on the poles gives APWP rate of $\sim$3.2 cm/yr (95\% interval from 2.5 cm/yr to 4.3 cm/yr; Fig.
-S5) between ca.
+S8) between ca.
 1070 to 990 Ma.
@@
 \includegraphics[width=\textwidth]{Jacobsville_pole_plot.pdf}
-\caption{The inclination-corrected Jacobsville Formation mean pole position with its Kent uncertainty ellipse is plotted in context of the ca.
+\caption{The inclination-corrected Jacobsville Formation detrital remanent magnetization mean pole position with its Kent uncertainty ellipse is plotted in context of the ca.
 1109-1070 Ma Keweenawan Track (poles color-coded by age shown in the colorbar), selected poles from the Grenville Province, and late Tonian poles of Laurentia as compiled by \citeA{Eyster2019a} (orange colored poles).
@@
 990 Ma Jacobsville pole and the poles of the Grenville loop, such as those of the the Haliburton intrusions, we suggest that estimates of their age as $>$1000 Ma \cite{Warnock2000a, Halls2015b, Evans2021b} are likely too old.
+The solid curve with arrow represents a continuous path of the Keweenawan Track, the dashed curves with arrow represent inferred apparent polar wander in the early Neoproterozoic based on data developed and compiled in this study and by \citeA{Eyster2019a}.
 The pole compilation used in this figure is included as Table S1.
@@
 
-This significant slow down in plate motion is geodynamically consistent with the onset of Grenvillian orogenesis.
+This slowdown in plate motion is geodynamically consistent with the onset of Grenvillian orogenesis.
 The earliest records of orogenesis on the leading margin of Laurentia occur ca.
@@
 1050 Ma \cite{Rivers2012a}.
-The large slowdown in plate speed as now constrained by the Jacobsville paleomagnetic pole is associated with this progression to hard continent-continent collision involving increased crustal thickening and progressive transmission of contractional stress that eventually developed into the interior of Laurentia \cite{Cannon1994a}.
+The large slowdown in plate speed as now constrained by the Jacobsville paleomagnetic pole is associated with this progression to hard continent-continent collision involving increased crustal thickening and progressive transmission of contractional stress that eventually progressed into the interior of Laurentia \cite{Cannon1994a}.
 
@@
 \ref{fig:paleogeography}).
-The Jacobsville paleomagnetic pole constrains this motion to have slowed to below average plate speed as orogeny progressed to be a large-scale continent-continent collision.
+The Jacobsville paleomagnetic pole constrains this motion to have slowed as orogeny progressed to be a large-scale continent-continent collision.
 As a result, Laurentia maintained a low-latitude position throughout the duration of the Grenvillian orogeny which sutured continents together into the supercontinent Rodinia (Fig.
@@
 \ref{fig:pole_plot}).
-Determining ages associated with those Grenville Loop poles is crucial for constraining the motion of Laurentia at this time and the configuration between Laurentia and hypothesized conjugate continents such as Baltica within Rodinia \cite{Cawood2017a, Gong2018a, Swanson-Hysell2021c}.
+Determining ages associated with the Grenville Loop poles is crucial for constraining the motion of Laurentia at this time and the configuration between Laurentia and hypothesized conjugate continents such as Baltica within Rodinia \cite{Cawood2017a, Gong2018a, Swanson-Hysell2021c}.
 
@@
 1075 and 990 Ma as the Grenvillian orogeny progressed.
-The inset figure compares Laurentia's position predicted by the new Jacobsville paleomagnetic pole (red) and that constrained by the Haliburton intrusion of the Grenvillian orogeny (light blue) which has been assigned an age of 1015 Ma \cite{Warnock2000a}).
+The paleogeographic positions of conjugate continents to Laurentia long the Grenville margin in the reconstructions follow the interpretation of \citeA{Swanson-Hysell2023a}.
+The inset figure compares Laurentia's position predicted by the new Jacobsville paleomagnetic pole (red) and that constrained by the Haliburton intrusions of the Grenvillian orogeny (light blue) which has been assigned an age of 1015 Ma \cite{Warnock2000a}).
 We suggest that this and other Grenville loop poles are younger and that Laurentia traveled from low latitudes toward higher latitudes following ca.
@@
 \ref{fig:pole_plot}).
-If ages that have been assigned to Grenville Loop poles are taken at face value, those poles represent the position of Laurentia both before and after the position constrained by the 990 Ma Jacobsville pole.
+If ages that have been assigned to Grenville Loop poles are taken at face value, those poles represent the position of Laurentia both before and after the position constrained by the ca.
+990 Ma Jacobsville pole.
 Such an interpretation would imply either rapid plate tectonic motion of Laurentia coeval with the Rigolet phase of the Grenvillian orogeny or rapid oscillatory true polar wander \cite<e.g.>{Evans2003a}.
-Alternatively, a more straightforward model that could explain the distinct pole positions between the Jacobsville pole and the Grenville Loop poles is that the poles from the Grenville orogen are younger than currently interpreted.
-Instead of poles such as that from the Haliburton intrusions having remanence that was acquired during the Grenvillian orogeny, the magnetic remanence could instead have been acquired during protracted exhumation that followed the ca.
+A more straightforward alternative model that could explain the distinct pole positions between the Jacobsville pole and the Grenville Loop poles is that the poles from the Grenville orogen are younger than currently interpreted.
+Instead of having acquired remanence during the orogeny, the Grenville rocks could have acquired magnetic remanence during protracted exhumation that followed the ca.
 980 Ma cessation of contractional deformation associated with the Rigolet stage of the orogeny.
@@
 While numerous paleomagnetic data sets have been developed from rocks of the Grenville Province, few are paired with well-calibrated, high-precision thermochronology data.
-Recent data sets, such as new apatite U-Pb dates of 923$\pm$14 Ma (2$\sigma$ analytical uncertainty) from the Wilberforce pyroxenite \cite{Paul2021a}, give tantalizing hints that temperatures that blocked magnetizations in portions of the orogen were achieved later than the ages assigned to paleomagnetic poles would suggest; given that U-Pb apatite closure temperatures are close to magnetite blocking temperatures.
-To test the hypothesis that the Grenville Loop is younger than in current interpretations, geochronology studies are needed to exploit the resolving power of high-precision whole grain analytical methods (i.e.
+Recent data sets, such as new apatite U-Pb dates of 923$\pm$14 Ma (2$\sigma$ analytical uncertainty) from the Wilberforce pyroxenite \cite{Paul2021a}, give tantalizing hints that temperatures that blocked magnetizations in portions of the orogen were achieved later than the ages assigned to paleomagnetic poles would suggest as U-Pb apatite closure temperatures are close to magnetite blocking temperatures (ca.
+360-570\textdegree C; \citeA{Cherniak1991a}).
+To test the hypothesis that the Grenville Loop is younger than in previous interpretations, geochronology studies are needed to exploit the resolving power of high-precision whole grain analytical methods (i.e.
 ID-TIMS) and high-spatial-resolution in-situ methods (e.g.
@@
 
-The magnetization of hematite-bearing fine-grained siliciclastic sedimentary rocks from the Jacobsville Formation can be shown to be detrital and primary through positive intraformational conglomerate tests.
+The magnetization of hematite-bearing fine-grained siliciclastic sedimentary rocks from the Jacobsville Formation can be shown to be detrital and primary through positive intraformational conglomerate and fold tests.
 We use these data to develop a new inclination-corrected paleomagnetic pole that can be constrained with recently published radiometric age constraints from \citeA{Hodgin2022a} to be ca.
@@
 We thank Madeline Swanson-Hysell for her assistance in the field.
+We thank David Malone, Douglas Elmore, and two anonymous reviewers for their helpful reviews.
 
-\nocite{Swanson-Hysell2014a, Swanson-Hysell2014b, Books1972a, Davis1997a, Davis1985a, Denyszyn2009a, Diehl1994a, Eyster2019a, Fairchild2017a, Halls1974a, Hnat2006a, Palmer1986a, Tauxe2009a, Tauxe1991a, Weil2003a}
+\nocite{Swanson-Hysell2014a, Swanson-Hysell2014b, Books1968a, Books1972a, Davis1997a, Davis1985a, Denyszyn2009a, Diehl1994a, Eyster2019a, Fairchild2017a, Halls1974a, Hnat2006a, Palmer1986a, Tauxe2009a, Tauxe1991a, Weil2003a, Zhang2023b, Zhang2023c, Zhang2023d, Tauxe1994a, McFadden1990a, Torsvik2012a, Brown2012a, Henry1977a, Buchan1976a, Warnock2000a, Brown2012a, Denyszyn2009a, Eyster2019a, Weil2003a, Slotznick2023a, Hodgin2022a}
 \bibliography{YZ_ref}
@@
 \end{document}
-
-%%%EXTRA TEXT%%%
-
-% Evidence for the Grenville Loop being younger than previously estimated is emerging in recent advances of thermochronological studies in the Grenville Province.
-\citeA{Paul2021a} developed a new apatite U-Pb age of 923$\pm$14 Ma (2$\sigma$ analytical uncertainty) from the Wilberforce pyroxenite using isotope dilution thermal ionization mass spectrometry (ID-TIMS).
-In the Adirondack Highlands, \citeA{Krestianinov2021a} also developed U-Pb ID-TIMS ages of 907$\pm$14 Ma and 924$\pm$13 Ma (2$\sigma$ analytical uncertainties) from apatite associated with low-Ti iron ores that are likely cogenetic with Adirondack AMCG units \cite{Postel1952a, Krestianinov2021a}.
-These ages represent the timing of closure of Pb diffusion within apatite grains during exhumation \cite{Dodson1973a}.
-The temperature range that corresponds to this process for apatite is 420-540\textdegree C \cite{Cherniak1991a} which overlaps with magnetic blocking temperatures of these magnetite-bearing rocks, taking into account the effective lowering in blocking temperatures during slow cooling (400-580\textdegree C; \citeA{Palmer1973a, Brown2012a, Halgedahl1980a, Dodson1980a}).
-Therefore, the Grenville Loop pole ages are better assigned to be close to the dates given by apatite thermochronology.
-Given that K-rich minerals are known to be prone to obtain excess Ar \cite{McDougall1999a}, the discrepancy between the younger apatite U-Pb ages and generally older hornblende $^{40}$Ar/$^{39}$Ar ages indicate that potentially widespread excess Ar could have biased previous cooling history interpretations of the Grenville orogen \cite<e.g.>{Warnock2000a, Mezger1991a}.
-
-% Previously, the large-magnitude shifts in Laurentia's apparent polar wander path through the Keweenawan Track and Grenville Loop (Fig.
-\ref{fig:pole_plot}; \citeA{Irving1981a, Weil1998a, Swanson-Hysell2019a}) were hypothesized to represent rapid true polar wander which involves the rotation of the entire solid Earth about the spin axis \cite{Evans2003a, Mitchell2014b}.
-However, The slowdown in Laurentia's plate motion implied by the new Jacobsville paleomagnetic pole is consistent with the interpretation that the onset of the Grenvillian orogeny had dominant kinematic control on plate speeds as it involved continental-scale collisional orogenesis that likely led to the cessation and subsequent inversion of the Midcontinent Rift \cite{Cannon1994a}.
-
-% More high-precision thermochronology data and statistical models that properly incorporates uncertainties associated with timing of closure of thermochronometers and with magnetic remanence acquisition temperatures in the slowly cooled orogen are needed in order to further calibrate the age of those poles.
-Further development of paired pole-age data from the rich suite of metamorphic rocks in the Grenville Province can help update Laurentia's APWP \cite{Irving1974a} and the paleogeographic configuration of super continent Rodinia in the Neoproterozoic.
-
-% The interpretation of young Grenville poles is supported by high-precision U-Pb apatite thermochronology data from iron oxide-apatite deposits of the eastern Adirondack Mountains \cite{Krestianinov2021a} and from the Wilberforce pyroxenite of the Wilberforce region, Canada \cite{Paul2021a}.
-Apatite from both regions yield ages of ca.
-900 Ma with associated closure temperatures $\sim$450\textdegree C \cite{Cherniak1991a, Chamberlain2001a}.
-Considering delayed remanence acquisition of magnetite and hematite, ca.
-900 Ma should be closer to the actual timing when rocks of the Grenville Province finally blocked in their remanence that is preserved today.
-However, the exact age of each paleomagnetic pole recorded by rocks of different domains of the Grenville Province remains largely uncertain.
-Complications such as the origin of the dated iron oxide-apatite ore, crystal-size dependence of closure temperature of accessory minerals including rutile and apatite \cite{Smye2018a, Chamberlain2001a}, and the complex metamorphic histories in different domains of the Grenville Province require that more high-precision thermochronology data with modern closure temperature calibration are developed.
-
-% The previous older age estiamtes could be a result of unrecognized crypitc excess argon and incorrect calibration of Pb diffusion rate in accessory minerals in earlier studies such as the rutile closure temperature used by Mezger 1991 which has been revised significantly upwards by the experiemntal data of Cherniak.
-
-% Such large distance between pole positions associated with uncertainties in ages of Grenville magnetizations led to the controversial ``Grenville Problem"---did the Grenville rocks acquire magnetic remanence prior to the continental collision and thus the pole discrepancies are a result of significant crustal shortening (i.e.
-the ``two plate model"; \citeA<e.g.>{Irving1972a, Palmer1973a, Buchan1973a, Halls2015b}), or did they acquire remanence after the continental accretion and thus the southerly poles from the Grenville rocks represent a coherent plate motion through more recent times (i.e.
-the ``one plate model"; \citeA<e.g.>{Roy1978a, McWilliams1975a, Weil1998a})?
-
-% The two plate model is based on the interpretation that the Grenville rocks acquired magnetizations on allochthonous terranes in the late Mesoproterozoic, prior to the final accretion to Laurentia craton, resulting in their apparent southerly paleomagnetic pole positions today (Fig.
-\ref{fig:pole_plot}).
-Further, the lack of paleomagnetic poles in the southern hemisphere from interior Laurentia at this time (i.e.
-the Keweenawan Track) is further interpreted to support such allochthonous magnetic remanence hypothesis \cite<e.g.>{Warnock2000a, Halls2015b}.
-A corollary of such interpretation is that the Grenville paleomagnetic poles could record a large amount of crustal shortening associated with the Grenvillian Orogeny.
-\citeA{Halls2015b} suggested the lost crust must be up to 4000 $\pm$ 1000 km long.
-
-% Given the new constraints on the age and position of the Jacobsville pole (Fig.
-\ref{fig:pole_plot}), the two plate model would require that the majority of the crustal shortening happened during or after the Rigolet phase of the Grenville Orogeny, as the ca.
-990 Ma Jacobsville pole is still $\sim$45\textdegree\ arc distance ($\sim$5000 km) away from the paleomagnetic pole of the Haliburton intrusions \cite{Buchan1976a} and $\sim$38\textdegree\ ($\sim$4000 km) from the pole of the Wilberforce intrusions \cite{Palmer1973a} (Fig, \ref{fig:pole_plot}).
-However, geologic and geochronologic evidence from both within the Grenville Province \cite{Aleinikoff2021a} and from interior Laurentia \cite{Hodgin2022a} indicate that majority of the deformation in the Grenville Province and the propagation of its compressional stresses into interior Laurentia happened during the ca.
-1090-1020 Ma Ottawan phase orogeny, whereas smaller amount of convergence happened during the ca.
-1010-980 Ma Rigolet phase.
-The lack of geological evidence for a striking amount of crustal shortening is consistent with the slow average plate motion between the deposition of the upper Freda Formation and the deposition of the Jacobsville Formation (average $\sim$0.2\textdegree/Myr; Fig.
-\ref{fig:pole_plot}).
-Therefore, it is kinetically unlikely that the plate of Laurentia suddenly increased its motion near the end of a collisional orogenesis in order to satisfy the Grenville pole positions.
-The plate kinetics interpretation based solely on paleomagnetic data and poorly constrained ages of magnetic remanence put forward by \citeA{Halls2015b} is implausible.
-
-% Alternatively, the one plate model is much more straightforward in explaining the paleomagnetic pole positions of Grenville rocks as it interprets that those rocks acquired remanence during prolonged exhumation of the orogeny that postdate the peak metamorphism.
-In contrast to the two plate model, the one plate model would predict that ages associated with the Grenville poles are much younger than ca.
-980 Ma (Fig.
-\ref{fig:pole_plot}).
-Adopting this model necessitates that the ages previously assigned to many Grenville poles be adjusted to much younger.
-Many previous paleomagnetic studies recognized that a large portion of the Grenville Province experienced up to granulite facies metamorphism (e.g.
-$>$700\textdegree C).
-The high metamorphism temperatures resulted in complete unblocking and overwriting of any previous remanence regardless of whether magnetite or hematite being the remanence carriers \cite<e.g.>{Ueno1975a, Alvarez1998a, Buchan1973a, Buchan1976a, Brown2012a}.
-However, what has not been fully taken into account in previous studies is the effective lowering of magnetic blocking temperatures of the new remanence.
-Theoretical calculations have shown that during prolonged cooling, magnetite which form above its Curie temperature (580\textdegree C) will subsequently acquire thermoviscous remanence \cite{Pullaiah1975a, Dodson1980a}, whereas hematite which often occur as exsolved lamellae in ilmenite in Grenville rocks acquired thermochemical remanence via magnetic ordering of the hemo-ilmenite below the N\'eel temperature of hematite \cite{Burton1985a, Harrison2006b}.
-
-% Due to the long time scales involved with thermoviscous and thermochemical origin of the magnetic remanence, it is not straightforward to interpret the natural blocking temperatures of Grenville rocks based on observations of unblocking temperatures in the lab where rocks are heated and cooled rapidly.
-\citeA{Brown2012a} studied Grenville rocks from the Adirondack Mountains.
-They recognized that the hemo-ilmenite exsolution occur at relatively low temperatures and acquire thermochemical remanence during cooling of the orogen and assigned an age of ca.
-960 Ma to a paleomagnetic pole developed from hemo-ilmenite-bearing microcline gneiss from northwestern Adirondacks (Fig.
-\ref{fig:pole_plot}).
-Using the theoretical models of \citeA{Pullaiah1975a}, that study also extrapolated remanence ages of ca.
-990 and ca.
-970 Ma for the Wanakena granite and Marcy Massif anorthositic rocks whose dominant remanence carriers are magnetite grains that acquired thermoviscous remanence (Fig.
-\ref{fig:pole_plot}).
-However, in light of the ca.
-990 Ma Jacobsville pole position and the slow down of Laurentia's apparent polar wander since ca.
-1070 Ma, the ages assigned to the poles in \citeA{Brown2012a} could be still too old to be compatible with the magnetization being acquired after the major pulse of the orogenic collision.
-
-% That the estimated ages associated with the Adirondack paleomagnetic poles are biased old can be attributed to the outdated thermochronometer closure temperature calculations of \citeA{Mezger1991a, Mezger1990a} whose data was used by \citeA{Brown2012a}.
-In particular, the cooling history for the Adirondack area estimated by \citeA{Mezger1991a, Mezger1990a} show that the orogen had cooled to $\sim$400\textdegree C by the time of U-Pb system closure in rutile crystals.
-However, this estimated U-Pb closure temperatures range for rutile was not based on laboratory Pb diffusion experiment, but instead was based on an assumption that thediffusion rate of Pb in rutile approximate that of Fe and Ti \cite{Mezger1989a}.
-Later experimental results \cite{Cherniak2000a} as well as analytical estimates \cite{Vry2006a, Kooijman2010a} show that diffusion rates of Pb in rutile are in fact much slower, indicating that previous calculations significantly underestimated closure temperatures of the system.
-Therefore, timing of magnetic remanence acquisition for Grenville rocks in the Adirondack Mountains should be younger than previously thought.
-
-% The interpretation of young Grenville poles is supported by high-precision U-Pb apatite thermochronology data from iron oxide-apatite deposits of the eastern Adirondack Mountains \cite{Krestianinov2021a} and from the Wilberforce pyroxenite of the Wilberforce region, Canada \cite{Paul2021a}.
-Apatite from both regions yield ages of ca.
-900 Ma with associated closure temperatures $\sim$450\textdegree C \cite{Cherniak1991a, Chamberlain2001a}.
-Considering delayed remanence acquisition of magnetite and hematite, ca.
-900 Ma should be closer to the actual timing when rocks of the Grenville Province finally blocked in their remanence that is preserved today.
-However, the exact age of each paleomagnetic pole recorded by rocks of different domains of the Grenville Province remains largely uncertain.
-Complications such as the origin of the dated iron oxide-apatite ore, crystal-size dependence of closure temperature of accessory minerals including rutile and apatite \cite{Smye2018a, Chamberlain2001a}, and the complex metamorphic histories in different domains of the Grenville Province require that more high-precision thermochronology data with modern closure temperature calibration are developed.
-
-% However, inclination bias associated with sedimentary paleomagnetic directions need to be corrected when synthesizing poles.
-Rotation of grains during sediment deposition and compaction during early diagenesis can result in the acquisition of a detrital remanent magnetization that is shallower than the local geomagnetic field in which it was acquired \cite{King1955a, Tauxe2005a, Kodama2012a}.
-This effect can be summarized as:
-% \begin{equation*}
-% \textup{tan}(I_o) = f\textup{tan}(I_f)
-% \end{equation*}
-% where $f$ represents the amount of inclination shallowing, $I_o$ represents the observed inclination, and $I_f$ represents the inclination of the field in which the magnetization was acquired.
-Methods for correcting sedimentary inclination bias include laboratory-based measurements of magnetic mineral fabrics \cite<e.g.>{Kodama1990a, Bilardello2015a}, as well as statistical methods based on the evaluation of the deviation of an observed virtual geomagnetic pole (VGP) distribution from that of a paleosecular variation model (the ``elongation/inclination" [E/I] method of \citeA{Tauxe2004b}).
-When a large number of sedimentary paleomagnetic sites (typically n$>$100) capture an extensive period of time such that the paleosecular variation is sufficiently averaged, the E/I method can give efficient estimates with reasonable uncertainty bounds via bootstrapping and is less labor-intensive than methods based on mineral fabric measurements.
-
-% Since the deposition of the Jacobsville Formation span over geomagnetic field reversals, it is of interest to investigate the amount of inclination shallowing associated with the normal (n=130) and the reversed (n=148) polarity respectively (transitional fields and specimens affected by complex tilting regime are excluded).
-Applying the E/I method on the normal polarity directions yields an estimated $f$ value of 0.67 with 95\% bootstrap uncertainty bounds from 0.91 to 0.55 (Fig.
-\ref{fig:in_situ_pmag}B).
-Directions of reversed polarity yields an estimated $f$ factor of 0.52 with uncertainty bounds between 0.97 and 0.29 (Fig.
-\ref{fig:in_situ_pmag}B).
-Although the estimated $f$ value for the reversed-polarity directions has large statistical uncertainty bounds, the uncertainties are likely associated with the overall shallow inclinations and distributions of directions in both upper and lower hemispheres (Fig.
-\ref{fig:in_situ_pmag}B).
-Therefore, inclination corrections do not result in significant changes in the mean reversed direction.
-
-% In addition, directions of the two polarities do not pass a reversal test \cite{McFadden1990a, Tauxe2010a}, regardless of whether inclination correction is applied.
-One interpretation of the asymmetric polarities could be that the deposition of the Jacobsville Formation was protracted such that the sediments at Hungarian Falls which dominate the reversed polarity records were deposited prior to those at Snake Creek tributary which dominate the normal polarity records (Fig.
-\ref{fig:in_situ_pmag}).
-This is consistent with our stratigraphic interpretation that Snake Creek tributary sediments could be younger than those near central Keweenaw Peninsula (i.e.
-Hungarian Falls, Natural Wall, Hammel Creek; Fig.
-\ref{fig:strat_column}).
-Other alternative interpretations could include that the mean poles are biased by transient non-dipolar fields due to insufficient averaging of paleosecular variation or transitional fields during field reversals.
-More data are needed to test these hypotheses.
-
-% Overall, both inclination-corrected polarities recorded by the Jacobsville Formation show mean pole positions in the southern hemisphere with respect to today's continental configuration (normal polarity longitude=175.1\textdegree E, latitude=19.1\textdegree S, A95=3.8\textdegree; reversed polarity longitude=191.1\textdegree E, latitude=14.5\textdegree S, A95=3.9\textdegree; Fig.
-\ref{fig:EI_results}).
-The two pole positions are distinct from the youngest poles developed from rocks of the Midcontinent Rift (Fig.
-\ref{fig:EI_results}).
-Despite that the Jacobsville sedimentation span over geomagnetic field reversals, the large arc distance between either polarity of the Jacobsville Formation and the pole of the ca.
-1070 Ma upper Freda Formation \cite{Henry1977a} is consistent with the interpretation that the onset of deposition of the Jacobsville Formation at the five studied localities post-date the youngest member of the Oronto Group sediments by tens of millions of years.
-
-% Near the base of the exposed stratigraphic section at Natural Wall (Fig.
-\ref{fig:strat_column}), the strata are steeply tilted or overturned (Fig.
-\ref{fig:Field_photo}).
-Tilt-corrected paleomagnetic directions from samples collected from these strata show dominantly southeasterly declinations that are distinct from either normal- or reversed-polarity directions higher along the strata or at Hungarian Falls.
-It is most likely that these directions resulted from noncylindrical or multiple episodes of complex folding.
-We hypothesize that these DRM directions were acquired during a reverse polarity episode similar to that recorded near the base of Hammel Creek (Fig.
-\ref{fig:strat_column}), but we will not include these directions in further paleomagnetic pole analyses.

### 2024-01-23 Manuscript/support_info.tex
update manuscript for proof

@@
 Yiming Zhang\textsuperscript{1}, Eben B.
-Hodgin\textsuperscript{1}, Tadesse Alemu\textsuperscript{1}, James Pierce\textsuperscript{1,2}, Anthony Fuentes\textsuperscript{1}, Nicholas L.
+Hodgin\textsuperscript{1,2}, Tadesse Alemu\textsuperscript{1}, James Pierce\textsuperscript{1,3}, Anthony Fuentes\textsuperscript{1}, Nicholas L.
 Swanson-Hysell\textsuperscript{1}
@@
 
-% $^{2}$Department of Earth, Environmental and Planetary Sciences, Brown University, RI, USA
+$^{2}$Department of Earth, Environmental and Planetary Sciences, Brown University, RI, USA
 
-$^{2}$Department of Earth and Planetary Sciences, Yale University, New Haven, CT, USA}
+$^{3}$Department of Earth and Planetary Sciences, Yale University, New Haven, CT, USA}
 
@@
 \listoftables
-\tableofcontents
+%\tableofcontents
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=0.9\textwidth]{Manuscript/SI_petrography.png}
+\caption{Side-by-side reflected light and oblique illumination images of a thin section of the same field of view of sample NW2-7 collected from Natural Wall ravine.
+The white arrows point towards examples of large sub-milimeter-scale detrital hematite grains that have bright reflectance under reflected light.
+Almost all subangular grains in grey color in the reflected image are quartz grains.
+The red color in the oblique illumination image is a result of diffuse reflectance of pigmentary hematite grains that often coat the boundaries of detrital grains.
+}
+\label{fig:SI_orthogonal}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=0.5\textwidth]{Manuscript/SI_orthogonal.pdf}
+\caption{Representative orthogonal vector diagrams of demagnetization experiments for specimens from different stratigraphic sections of this study.
+An overprint component subparallel to the present day field direction in the study area is typically minimally present in siltstone to very fine-grained sandstone facies.
+This component shown in light green can be fit with a least-squares line in fine-grained facies.
+After the removal of the low-temperature component, a mid-temperature component typically unblocks through a wide range of temperature steps by up to $\sim$650\textdegree C.
+Finally, an origin-trending component with typically a shallower inclination than that of the mid-temperature component sharply unblocks through heating with smaller temperature intervals up toward the N\'eel temperature of hematite ($\sim$690\textdegree C).
+AF-Agate Falls section; HF-Hungarian side falls section near Dover Creek; JK-Hammel Creek section; SC-Snake Creek tributary section.
+M-magnetic moment; NRM-natural remanent magnetic moment.
+All diagrams are shown in tilt-corrected coordinates.}
+\label{fig:SI_orthogonal}
+\end{figure*}
 
@@
 Samples collected from the Natural Wall section do not have interpretable chemical remanence component that can be fit with least-squares lines.
-}
+All directions are plotted in geographic coordinates.}
 \label{fig:Jacobsville_CRM}
+\end{figure*}
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[width=0.75\textwidth]{Manuscript/SI_Jacobsville_vs_Appalachian_poles.pdf}
+\caption{Jacobsville detrital remanence Kent mean pole position and chemical remanence Fisher mean pole position (in geographic coordinates) plotted in context of Laurentia's paleomagnetic poles during Appalachian orogeny (ca.
+460-260 Ma) as compiled in \cite{Torsvik2012a}.
+The mean chemical remanence pole position from the Snake Creek tributary section is plotted in context of the ca.
+970-960 Ma poles developed by \cite{Brown2012a} from the Adirondack Highlands of the Grenville orogen.
+That this chemical remanence component failed a fold test (Figure S4) and yields a pole position that overlaps with the Grenville poles are consistent with the interpretation that the chemical remanence of the Jacobsville Formation at this locality was acquired  during the early Neoproterozoic, soon after Jacobsville deposition and deformation.
+That the Jacobsville DRM and CRM mean poles are distinct and far away from any of Laurentia's mid- to late-Paleozoic poles supports the interpretation that the Jacobsville detrital remanence magnetization was acquired during the Rigolet phase of the Grenvillian orogeny and the chemical remanence overprint was likely acquired soon after deposition as discussed in detail in the manuscript.}
+\label{fig:Jacobsville_Appalachian}
 \end{figure*}
@@
 Freda Formation                                            & 2.2      & 179      & 4.2  & \cite{Henry1977a}                                                     & 1070       & 1060     & 1083.5   & As discussed in \cite{Swanson-Hysell2019a}            \\
-Haliburton Intrusions                                      & -32.6    & 141.9    & 6.3  & \cite{Warnock2000a}                          & 1015       & 1000     & 1030     & \cite{Warnock2000a}                                          \\
+Haliburton Intrusions                                      & -36.2    & 141.7    & 6.0  & \cite{Buchan1976a}                          & 1015       & 1000     & 1030     & \cite{Warnock2000a}                                          \\
 Adirondack Highlands gneiss                                & -19      & 148.7    & 11.2 & \cite{Brown2012a}                                                &            &          &          &                                                     \\

### 2024-01-27 Manuscript/Manuscript.tex
sync notebooks, some figures, data files, and manuscript after proof correction

@@
 \section*{Data Availability}
-Paleomagnetic data associated with this study are available within the MagIC database (\url{10.7288/V4/MAGIC/19780}) and all data are within a Github repository associated with this work (\url{https://github.com/Swanson-Hysell-Group/Jacobsville}) that is also archived on Zenodo (\url{https://doi.org/XXX}; \textit{will be archived in Zenodo upon acceptance}).
+Paleomagnetic data associated with this study are available within the MagIC database (\url{10.7288/V4/MAGIC/19780}) and all data are within a Github repository associated with this work (\url{https://github.com/Swanson-Hysell-Group/Jacobsville}) that is also archived on Zenodo (\url{https://zenodo.org/doi/10.5281/zenodo.10556464}; \textit{will be archived in Zenodo upon acceptance}).
 This repository also contains Python code that implements all of the calculations, visualizations and statistical tests discussed herein.
